---
title: "2025_10_pnnl_s5_EDA"
output: html_document
---

This follows Tyler Sagendorf's Section 5 Exploratory Data Analysis Tutorial from Proteomics Data Analysis in R/Bioconductor (pnnl)

-   oca.set details - <https://rdrr.io/github/vladpetyuk/vp.misc/man/oca.set.html>

    -   Processed iTRAQ data from both PNNL & JHU sites (from CPTAC study)

        -   an MSnID instance - 8103 proteins by 73 features

    -   CPTAC det <https://statomics.github.io/PDA/pda_tutorialPreprocessing.html>

```{r}
# ## Install missing packages
# if (!require("remotes", quietly = T)) install.packages("remotes")
# if (!require("MSnSet.utils", quietly = T))
#   remotes::install_github("PNNL-Comp-Mass-Spec/MSnSet.utils")
```

```{r}
library(MSnSet.utils)
```

```{r}
# Load the example MSnSet
data(cptac_oca) # loads in oca.set (is a Large MsnSet)
```

```{r}

head(pData(oca.set)) # patient details?
head(fData(oca.set)) # feature
head(exprs(oca.set)) # expression
```

```{r}
str(pData(oca.set))
str(fData(oca.set)) # feature
str(exprs(oca.set)) # expression
```

## 5.1 Count Features in Samples 

<https://pnnl-comp-mass-spec.github.io/proteomics-data-analysis-tutorial/count-features-in-samples.html>

```{r}
# Count no. of features ID per sample - using colSums that are not NA in exprs(oca.set)

# Calculate the no. of proteins ID (not NA) in each sample
oca.set$num_proteins <- colSums(!is.na(exprs(oca.set)))
# vector of no. proteins stored in pData(oca.set)
# pData(oca.set) should be in final col
```

```{r}
# summary (5 nums)
summary(oca.set$num_proteins)
```

```{r}
# as horizontal box plot
boxplot(oca.set$num_proteins, horizontal = TRUE,
        xlab = "No. of Proteins Detected")
```

```{r}
# a scatter plot with more info
plot(oca.set$num_proteins, 
     pch = 19, # solid circle shape
     xlab = "Sample Index",
     ylab = "Number of Proteins Detected") 

# chaotic but useful for showing any major outliers
```

```{r}
# shows distribution - density ploy has the benefits of both box & scatter plots
# Kernel density plot 
plot(density(oca.set$num_proteins, na.rm = TRUE),
     xlim = range(oca.set$num_proteins, na.rm = TRUE),
     main = "Density Plot", xlab = "No. of Proteins Detected")

# shows peaks ~6900-7000 & 7300-7500
```

## 5.2 Sample Boxplots

<https://pnnl-comp-mass-spec.github.io/proteomics-data-analysis-tutorial/sample-boxplots.html>

```{r}
# boxplots by sample?
boxplot(exprs(oca.set))
```

5.3 Estimate Blood Contamination

<https://pnnl-comp-mass-spec.github.io/proteomics-data-analysis-tutorial/estimate-blood-contamination.html>

## 5.4 PCA 

-   <https://pnnl-comp-mass-spec.github.io/proteomics-data-analysis-tutorial/pca.html>

-   Add details & note on missing values & methods

```{r fig.height=9, fig.width=11}
# to see details
pcaMethods::pca
```

```{r}
# Default plot
plot_pca(oca.set)

# prints - "Subsetting to 4738 complete rows for PCA"
# so of the total 8103 features, 4738 were present in all 73 samples
# PC1 & PC2 explain less than 1/5 of the variance in the original data
```

```{r fig.height=9, fig.width=11}
# Points colored by Subtype
plot_pca(oca.set,
         phenotype = "SUBTYPE", # color by SUBTYPE group
         legend_title = "Subtype",
         show_ellipse = TRUE) # adds a 50% normal confidence ellipse | def TRUE
```

```{r}
# Points by batch
plot_pca(oca.set, phenotype = "Batch")

# generally overlaps & centred on origin (except X17) but hard to tell batch effect
# should correct for - see section
```

```{r fig.height=9, fig.width=11}
# determine most influential features in PCA with biplot arg

plot_pca(oca.set,
         phenotype = "SUBTYPE",
         biplot = TRUE,
         label_args = list(color = "black"))


# NP_9976.. are druvers of separation of Mesenchmyal & Proliterative samples 
# Also blood proteins (hemoglobin subunits: NP000549.1 and NP_000550.2) drive PC2 sep 
# but PC2 only explains ~ 6% variance?
```

\
