---
title: "2026_03_LCMS_methods_table"
output: html_document
---

```{r}
library(tidyverse) # dplur.readr, stringr, tidyr, ggplot, tibble, purr etc
```

```{r}
# read in methods csvs (exported from run info from Thermo FreeStyle)
LC_raw <- read_csv("input/Steph_TMTpro-16plex_NOV2022_B4_MP_TR2_HpHF_F1_LC.csv")
MS_raw <- read_csv("input/Steph_TMTpro-16plex_NOV2022_B4_MP_TR2_HpHF_F1_MS.csv")
```

-   LC Instrument Settings - code below grabs the basic LC instrument settings (row 10-54), then separates based on common patterns.
-   Everything before the 1st . is the part.
-   Everything at the end after the : is the value
-   Everything in the middle is the setting (after 1st . & b4 : ). this can be split further into
-   The code uses separate_wider_regex to split into the 3 main segments (part, setting, value).
-   Then further spliting using the basic separate_wider_delim to split if there are further dots into settings (if not dot . - keeps as is) , subsettings (if 1 .) & subsubsettings (if . . ) dots. This is enabled by too_few = "align_start"
-   Finally converts NA to "" blank space for those that do not get split further.
-   & export!

```{r}
#Instrument Setup - 	
# from 10 Sampler.LowDispersionMode: Off to 54 ColumnOven.ValveRight: 1_2
LC_setup_raw <- LC_raw[10:54,1] 

LC_setup <- LC_setup_raw %>%
  separate_wider_regex(`Instrument Method`,
                       patterns = c(
                         part = "[^.]+", # everything b4 dot
                         "\\.", # . itself not captured
                         setting ="[^:]+", # everything before : 
                         ": ", #skip : & space
                         value = ".*" # everything after 
                       )) %>%
  separate_wider_delim(setting, delim = ".", 
                       names = c("settings", "subsettings", "subsubsettings"),
                       too_few = "align_start") %>% # deal w/ variable length split
  replace_na(list("subsettings" = "", subsubsettings = "")) # NAs to "" blank
```

```{r}
# write_excel_csv(LC_setup, "output/LC_setup_instrument_settings_split.csv")
```

**LC Method Plot**

```{r}
LC_gradient_1 <- LC_raw[55:101, 1]

LC_gradient_2 <- LC_gradient_1 %>%
  mutate(set_val = if_else(str_detect(`Instrument Method`, "\\[min\\]"),
                            NA,`Instrument Method`)) %>%
  mutate(time_raw = if_else(!str_detect(`Instrument Method`, "\\[min\\]"),
                            NA,`Instrument Method`), .before = set_val) %>%
  fill(time_raw, .direction = "down")  # fill down like excel corner click! 



#print(LC_gradient_3$action) # 2 check
LC_gradient_3 <- LC_gradient_2 %>%
  
  drop_na(set_val) %>% # remove NAs rows in set_val (those prev time rows)
  select(time_raw, set_val) %>%
  
  separate_wider_regex(time_raw,
                       patterns = c(
                         min = ".*", # greedy - gets everything b4 [min]
                         " \\[min\\]", # "delim" 
                         action = ".*"), #greedy - everything after [min]
                       too_few = ("align_start")) %>%
  
  mutate(action = na_if(action, "")) %>% # recodes "" as NA (so below works)
  fill(action, .direction = "down") %>% # autofill NAs with value above
  

  separate_wider_delim(set_val, delim = ":",
                       names = c("setting", "value"), 
                       too_few = "align_start" ) %>%
  
  mutate(across(where(is.character), str_squish)) # remove white spaces
  

LC_gradient_4 <- LC_gradient_3 %>%
  filter(action == "Run") %>%
   separate_wider_delim(setting, delim = ".",
                       names = c("setting", "subsettings",
                                 "subsubsettings", "subsubsubsettings"), 
                       too_few = "align_start")  %>%
  replace_na(list(subsubsettings = "", subsubsubsettings = "")) # NAs to "" blank
  
```

```{r}
# write_excel_csv(LC_gradient_3, "output/LC_gradient_3.csv")
# write_excel_csv(LC_gradient_4, "output/LC_gradient_4.csv")
```
