---
title: "2026_05_QC_25_09_v3.2.3_PD2.4_inspection"
author: "Stephanie Huang"
date: "2026-05-08"
output: github_document
---

```{2026_05_QC_PD}
```

2026_05_08 - Double checking whether the 1st protein protein in Master Protein Accession (MPA) Column is the 1st protein in the Protein Accession (PA) column of the PSMs results file – because fig301 of PD2.4 userguide shows that it sometimes isn't – intially thought Protein Accessions were ranked ordered because it is not alphabettical at least??!?!.

Answer - yes matches if unique peptide. However, if more than 1 accession, then 1st protein in PA does not necessarily match the MPA – although it does most of the time. So still not sure if the order it is listed in is meaningful...

NOTE - this is not a problem if you use MSstatsTMT's defaults since it removes shared peptides.

```{r}
library(tidyverse)
```

doubling checking PD2.4 PSMs Master Protein Accession is the same as Protein Accession - answer - yes it is, but only if it is a unique peptide.

```{r}
# load("output/2025_10_03_QC_v3.2.1_PSMs_4s_96f_PD_db.rda")
fasta <- read_csv("input/anno/UP000002494_10116_Rat_Ref_Proteome_22403_1pspg_2025_04_UPfastadf.csv") 
```

```{r}
dim(PSMs)
str(PSMs)
colnames(PSMs)

count(PSMs, Confidence) # all high confidence
count(PSMs, `PSM Ambiguity`) # all Unambigous
```

```{r}
PSMs_calcs <- PSMs %>%
  
  mutate(Master_PA_1 = ifelse(grepl(";", `Master Protein Accessions`), # match ;
                              sub(";.*", "", `Master Protein Accessions`), # rm ; & aft
                              `Master Protein Accessions`), # if 1 then asis
         .after = `Master Protein Accessions`) %>%
  
  separate_wider_delim(`Protein Accessions`, # split protein accession
                       delim = "; ", # by separate ; 
                       names_sep = "", # make as name cols as needed
                       too_few = "align_start") %>%  # align @ start
  
  mutate(first_match = ifelse(Master_PA_1 == `Protein Accessions1`,
                              TRUE, FALSE), .before = Master_PA_1)
  
```

```{r}
colnames(PSMs_calcs) # see how many new cols Protein Accessions1 - 65!


summary(PSMs_calcs$first_match) # FALSE (258685), TRUE (3040912), NA's (1373)
# interesting there are some NA's no master protein accession but there is protein accession | falses seem mainly driven by isoforms 
```

<https://www.statology.org/dplyr-left-join-select-columns/> - based on this

```{r}
PSMs_calcs_false <- PSMs_calcs %>%
  filter(first_match == FALSE) #%>% # mismatches
  # filter(PD_set %in% c("1pspg_2025", "sp_2025")) %>% # no isoforms
  # filter(PD_set == "1pspg_2025") %>% 
  # left_join(dplyr::select(fasta, uniprot_acc, gene, description), # 2 cols only
  #           by = join_by("Master_PA_1" == "uniprot_acc")) %>%
  # relocate(c(gene,description), .before = Confidence) %>%
  # filter(str_detect(description, "receptor")) 

```

```{r}
# no NAs so basically this is never an issue when there is only one protein accession
# which means defaults in msstatsTMTs avoids any issues!
sum(is.na(PSMs_calcs_false$`Protein Accessions2`))
```
