---
title: "QC_2025_09_v3.2.1_basic"
author: "Stephanie Huang"
date: "2025-09-30"
output: html_document
---

Overview

Purpose & Comparisons -

Data used

Based off of

Links

1.  [Mastering File Manipulation with R’s list.files() Function](https://www.r-bloggers.com/2023/05/mastering-file-manipulation-with-rs-list-files-function/)
2.  [Reading Data from multiple files](https://www.r-bloggers.com/2023/05/mastering-file-manipulation-with-rs-list-files-function/)

Load Packages

```{r}
library(readr) # read csv files read_csv
library(readxl) # read in excel files read_excel
library(tidyr) # pivot longer
library(stringr) # str_detect()
library(dplyr)
library(ggh4x) # facet_nested - extensions of facet_grid for nesting
library(cowplot) # for plot_grid()
library(gt)
library(ggridges)
library(forcats) #fct_rev
library(purrr) # for map()
library(ggvenn) #for ggvenn
```

Load Data

PSMs

```{r}
#set_wd to sourcefile location
#input_dir <- "input/PD/"
# dirs <- list.dirs(input_dir) # lists directories | not needed 

PSMs_2_read <-  list.files("input/PD", # where 2 read from
                       pattern = ".*PSM.*\\.txt$", # pattern PSM & ends .txt
                       full.names = TRUE, # get full name including path
                       recursive = TRUE) # gets files in subdirectories too


PD_PSMs <- read_tsv(PSMs_2_read, # read all files from list above
                    id = "file_path") # keep file path as an id column

# dim(PD_PSMs) #3658257 x 55
```

MS/MS Spectrum Info

```{r}
MSMS_2_read <- list.files("input/PD",
                          pattern = ".*MSMSSpectrumInfo.*\\.txt$",
                          full.names = TRUE,
                          recursive = TRUE)

PD_MSMS <- read_tsv(MSMS_2_read, id = "file_path")
  
dim(PD_MSMS) #16724105 x 55
```

Proteins

```{r}
Proteins_2_read <- list.files("input/PD",
                          pattern = ".*Proteins.*\\.txt$", # make sure Proteins with s
                          full.names = TRUE,
                          recursive = TRUE)

PD_Proteins <- read_tsv(Proteins_2_read, id = "file_path") # error uneven df size?
  
dim(PD_Proteins) #16724105 x 55
```

Results Statistics

```{r}
RS_2_read <- list.files("input/PD",
                        pattern = ".*ResultStatistics.*\\.txt$",
                        full.names = TRUE,
                        recursive = TRUE)

PD_RS <- read_tsv(RS_2_read, id = "file_path")

dim(PD_RS) # 155572 x 12
```

\~ \~ \~

Load Annotation Sheets

Raw File Annotation

```{r}
raw_anno <- read_csv("input/anno/TMT_B1-4_lcms_run_info_anno_20250830.csv")
```

Sample Annotation

\~ \~ \~ \~

Write out data?

Alternatively use left join to save on writing out extra.

```{r}
# checks
# colnames(PD_PSMs)
# str(PD_PSMs)
# length(unique(PD_PSMs$file_path)) # 331 
# length(unique(PD_PSMs$`File ID`)) # 112 (so 96 canonical + 16 extras)

# for relocate(any_of()) function below (can't do c())
raw_cols <- c("Mixture", "TechRepMixture", "Fraction")

# Clean up & modify data 
PSMs <- PD_PSMs %>%
  
  filter(!str_detect(file_path, "extra")) %>% # gets rid of files in the extra folders
  
  # new cols
  mutate(folder = str_extract(file_path, "(?<=input/PD/)[^/]+")) %>%  # folder name
  mutate(PD_result_file = basename(file_path)) %>% # file name 
  mutate(PD_set = recode(folder, # PD_set or analysis | old = new 
                  "2025_08_Steph_TMTpro_PSMs_MSstatsTMT_EEvsSH_for_PhD" =
                           "1pspg_2025",
                  "2025_09_sp_Steph_TMTpro_PSMs_MSstatsTMT_EEvsSH_for_PhD" = 
                           "sp_2025",
                  "2025_09_sp_w_iso_Steph_TMTpro_PSMs_MSstatsTMT_EEvsSH_for_PhD" = 
                           "sp_w_iso_2025",
                  "Steph_01NOV2024_all_EE-SH_M-F_MSstatsTMT" = 
                          "sp_w_iso_2024_rab"),
         .after = file_path) %>%
  
  # join 2 anno file for raw
  full_join(raw_anno, join_by(`Spectrum File` == Run)) %>%
  
  # move columns
  relocate(file_path, .after = last_col()) %>%
  relocate(any_of(raw_cols), .after = PD_set) %>%
  
  # make 1st 4 cols as factor
  # mutate(across((1:4), as.factor)) # error
  mutate(across(c(Mixture, TechRepMixture, Fraction, PD_set), as.factor))

```

```{r}
# bit convoluted...
num_PSMs <- PSMs %>%
  group_by(pick(where(is.factor))) %>%
  mutate(PSMs = n()) %>%
  select(c(1:4, PSMs)) %>%
  unique()


```

```{r}

```
