---
title: "QC_2025_09_v3.2.1_basic"
author: "Stephanie Huang"
date: "2025-09-30"
output: html_document
---

Overview

Purpose & Comparisons -

Data used

Based off of

Links

1.  [Mastering File Manipulation with R’s list.files() Function](https://www.r-bloggers.com/2023/05/mastering-file-manipulation-with-rs-list-files-function/)
2.  [Reading Data from multiple files](https://www.r-bloggers.com/2023/05/mastering-file-manipulation-with-rs-list-files-function/)

Load Packages

```{r}
library(readr) # read csv files read_csv
library(readxl) # read in excel files read_excel
library(tidyr) # pivot longer
library(stringr) # str_detect()
library(dplyr)
library(ggh4x) # facet_nested - extensions of facet_grid for nesting
library(cowplot) # for plot_grid()
library(gt)
library(ggridges)
library(forcats) #fct_rev
library(purrr) # for map()
library(ggvenn) #for ggvenn
library(paletteer) # color palette
library(ggExtra) # fpr ggMarginal
```

Load Data

PSMs

```{r}
#set_wd to sourcefile location
#input_dir <- "input/PD/"
# dirs <- list.dirs(input_dir) # lists directories | not needed 

PSMs_2_read <-  list.files("input/PD", # where 2 read from
                       pattern = ".*PSM.*\\.txt$", # pattern PSM & ends .txt
                       full.names = TRUE, # get full name including path
                       recursive = TRUE) # gets files in subdirectories too


PD_PSMs <- read_tsv(PSMs_2_read, # read all files from list above
                    id = "file_path") # keep file path as an id column

# dim(PD_PSMs) #3658257 x 55
```

MS/MS Spectrum Info

```{r}
MSMS_2_read <- list.files("input/PD",
                          pattern = ".*MSMSSpectrumInfo.*\\.txt$",
                          full.names = TRUE,
                          recursive = TRUE)

PD_MSMS <- read_tsv(MSMS_2_read, id = "file_path")
  
dim(PD_MSMS) #16724105 x 55
```

Proteins

Error: object 'PD_Proteins' not found Error during wrapup: not that many frames on the stack Error: no more error handlers available (recursive errors?); invoking 'abort' restart

```{r}
# # Does not work ? - look into
# Proteins_2_read_ <- list.files("input/PD",
#                           pattern = ".*Proteins.*\\.txt$", # make sure Proteins with s
#                           full.names = TRUE,
#                           recursive = TRUE)
# 
# PD_Proteins <- read_tsv(Proteins_2_read_, id = "file_path") # error uneven df size?
#   
# dim(PD_Proteins) #16724105 x 55
```

Results Statistics

```{r}
RS_2_read <- list.files("input/PD",
                        pattern = ".*ResultStatistics.*\\.txt$",
                        full.names = TRUE,
                        recursive = TRUE)

PD_RS <- read_tsv(RS_2_read, id = "file_path")

dim(PD_RS) # 155572 x 12
```

\~ \~ \~

Load Annotation Sheets

Raw File Annotation

```{r}
raw_anno <- read_csv("input/anno/TMT_B1-4_lcms_run_info_anno_20250830.csv")
```

Sample Annotation

\~ \~ \~ \~

Write out data?

Alternatively use left join to save on writing out extra.

<https://sparkbyexamples.com/r-programming/r-if-else-multiple-conditions/>

```{r}
# checks
# colnames(PD_PSMs)
# str(PD_PSMs)
# length(unique(PD_PSMs$file_path)) # 331 
# length(unique(PD_PSMs$`File ID`)) # 112 (so 96 canonical + 16 extras)

# for relocate(any_of()) function below (can't do c())
raw_cols <- c("Mixture", "TechRepMixture", "Fraction")

# Clean up & modify data 
PSMs <- PD_PSMs %>%
  
  filter(!str_detect(file_path, "extra")) %>% # gets rid of files in the extra folders
  
  # new cols
  mutate(folder = str_extract(file_path, "(?<=input/PD/)[^/]+")) %>%  # folder name
  mutate(PD_result_file = basename(file_path)) %>% # file name 
  mutate(PD_set = recode(folder, # PD_set or analysis | old = new 
                  "2025_08_Steph_TMTpro_PSMs_MSstatsTMT_EEvsSH_for_PhD" =
                           "1pspg_2025",
                  "2025_09_sp_Steph_TMTpro_PSMs_MSstatsTMT_EEvsSH_for_PhD" = 
                           "sp_2025",
                  "2025_09_sp_w_iso_Steph_TMTpro_PSMs_MSstatsTMT_EEvsSH_for_PhD" = 
                           "sp_w_iso_2025",
                  "Steph_01NOV2024_all_EE-SH_M-F_MSstatsTMT" = 
                          "sp_w_iso_2024_rab"),
         .after = file_path) %>%
  
  # join 2 anno file for raw
  full_join(raw_anno, join_by(`Spectrum File` == Run)) %>%
  
  # move columns
  relocate(file_path, .after = last_col()) %>%
  relocate(any_of(raw_cols), .after = PD_set) %>%
  
  mutate(Year = ifelse(Mixture == 'B1' | Mixture == 'B2',
                       '2021',
                       '2022'),
         .after = PD_set) %>%
  
  # make 1st 4 cols as factor
  # mutate(across((1:4), as.factor)) # error
  mutate(across(c(PD_set, Year, Mixture, TechRepMixture, Fraction), as.factor))

```

```{r}
# #Write out mega PSMs file 
# write_csv(PSMs, "output/2025_10_03_QC_v3.2.1_PSMs_4s_96f_PD_db.csv") # too big 2 open
# write_tsv(PSMs, "output/2025_10_03_QC_v3.2.1_PSMs_4s_96f_PD_db.txt")
# save(PSMs, file = "output/2025_10_03_QC_v3.2.1_PSMs_4s_96f_PD_db.rda")
```

```{r}
# bit convoluted...
num_PSMs <- PSMs %>%
  group_by(pick(where(is.factor))) %>%
  mutate(PSMs = n()) %>%
  dplyr::select(c(1:4, PSMs)) %>%
  unique()
```

```{r}
RS <- PD_RS %>%
  
  # new cols
  mutate(folder = str_extract(file_path, "(?<=input/PD/)[^/]+")) %>%  # folder name
  mutate(PD_result_file = basename(file_path)) %>% # file name 
  mutate(PD_set = recode(folder, # PD_set or analysis | old = new 
                  "2025_08_Steph_TMTpro" =
                           "1pspg_2025",
                  "2025_09_sp_Steph_TMTpro" = 
                           "sp_2025",
                  "2025_09_sp_w_iso_Steph_TMTpro" = 
                           "sp_w_iso_2025",
                  "Steph_01NOV2024" = 
                          "sp_w_iso_2024_rab"),
         .after = file_path) %>%

  # can't do previous join - do if else & str_extract
  # mutate(fraction = str_extract(file_path, ))
  mutate(Year = ifelse(str_detect(file_path, 'DEC2021'),
                                  '2021',
                                  '2022'),
         .after = PD_set) %>%
  
  mutate(Mixture = str_extract(file_path, '(?<=_B)[^_\\.]+'),
         .after = Year) %>%
  
  mutate(TechRepMixture = str_extract(file_path, '(?<=_TR)[^_\\.]+'),
         .after = Mixture) %>%
  
  # Fraction - "(?<=HpHF_)[^_\\.]+" - captures after HpHF_ until next _ or .
  mutate(Fraction = str_extract(file_path, '(?<=HpHF_)[^_\\.-]+'),
         .after = TechRepMixture) %>%
  
  
  
  # join 2 anno file for raw | doesn't work no Spectrum File & weird structure df
  # full_join(raw_anno, join_by(`Spectrum File` == Run)) %>%
  
  # move columns
  relocate(file_path, .after = last_col()) %>%
  # relocate(any_of(raw_cols), .after = PD_set) %>%

  # make 1st 5 cols as factor
  # mutate(across((1:4), as.factor)) # error
  mutate(across(c(PD_set, Year, Mixture, TechRepMixture, Fraction), as.factor)) %>%
  
  # noob method | maybe next w above?
  mutate(Mixture = paste0('B', Mixture)) %>% 
  
  # just rid of for now - data structure annoying 
  filter(!(PD_set == 'sp_w_iso_2024_rab')) # data structure annoying 
```

```{r}
# Check assignment of 
RS_check <- RS %>%
  dplyr::select(1:5, folder, PD_result_file, file_path) %>%
  unique() 

# dim  - 288 x 8  - 96 x 3 = 288 (got rid of Nov 2024 batch for now)
```

```{r}
# #write out
# write_tsv(RS, "output/2025_10_03_QC_v3.2.1_RS_3s_96f_PD_db.txt")
# save(RS, file = "output/2025_10_03_QC_v3.2.1_RS_3s_96f_PD_db.rda")
```

# MS/MS

# PSMs

```{r}
# mostly from Sam's Code + my L&H4_11d v2 LCMS R file
PSMs_calc <- PSMs %>%
  
  # ANNOTATED SEQUENCE COLUMNS
  # trimmed sequence - trimmed_seq 
  mutate(trimmed_seq = gsub('^\\[.*?\\].|.\\[.*?\\]$', # from Sam's code 
                            '', # removes [R]. & .[E] at either end
                            `Annotated Sequence`),
         .after = `Annotated Sequence`) %>%
  
  # trimmed sequence w no mods (all upper) - seq_no_mod
  mutate(seq_no_mod = toupper(trimmed_seq), # sam's ver tolower | toupper instead
         .after = trimmed_seq) %>%
  
  mutate(trm_seq_len = str_length(seq_no_mod),
         .after = seq_no_mod) %>%
  
  # SHARED PEPTIDES COUNT 
  # if greater than 1 then shared = yes
  mutate(shared_peptide = ifelse(`Number of Proteins` > 1, 
                                 'yes',
                                 'no'),
         .after = `Number of Proteins`) %>%
  
  
  # TMT LABELLING EFFICIENCY | ALL completely tagged if mod = fixed
  mutate(
    #lysine_count
    lysine_count = str_count(trimmed_seq, 'k'),
    #lysine (K) TMTpro tag count
    lysine_tag_count = str_count(Modifications, 'K\\d{1,2}\\(TMTpro'),
    #N-term TMTpro count
    n_term_tagged = grepl('N-Term(TMTpro)',
                          fixed = TRUE, # no need to escape \\( in this case
                          Modifications),
    .after = Modifications) %>%
  
  mutate( # get no. of Ks in anno seq & see if no. lysine count (w TMT) matches
    lysine_tagged = ifelse(
      lysine_count != 0, # if K tagged TMT more than 0
        ifelse(
          lysine_tag_count == lysine_count, # if match = TRUE 4 full K tagging
          TRUE, # all lysines are tagged
          FALSE # not all lysines are tagged
        ),
      NA), # if 0 TMT give NA
  .after = lysine_tag_count) %>%
  
  mutate( # calculate full TMT tagging - so all available N-term & all K residues
    completely_tagged = n_term_tagged & (lysine_tagged | is.na(`lysine_tagged`)),
    .after = n_term_tagged)
```

```{r}
PSMs_calc_check <- PSMs_calc[1:1000,] # check 1st 100 obvs 
```

```{r}
# # https://r4ds.hadley.nz/numbers.html#sec-counts
# # https://r4ds.hadley.nz/numbers.html#numeric-summaries
# PSMs_stat_calc <- PSMs_calc %>%
#   group_by(pick(where(is.factor))) %>% # group by factors
#   
#   # SUMMARY STATISTICS
#   # count of PSMs 
#   mutate(n_PSMs = n()) %>% #n of rows in that group (so PSM no. of sample)
#   
#   #shared peptides
#   add_count(shared_peptide, name = 'n_shared_pep') %>% # new col - n unq shr pep
#   mutate(p_shared_pep = round((n_shared_pep/n_PSMs) * 100, 1)) %>% # 1 dp | percent
#   
#   add_count(`Number of Proteins`, name = 'n_of_s_proteins') %>%
#   mutate(p_of_s_proteins = round(('n_of_s_proteins'/n_PSMs)))
#   
#   # missed cleavages
#   add_count(`Number of Missed Cleavages`, name = 'n_miss_cleave') %>%
#   mutate(p_miss_cleave = round((n_miss_cleave/n_PSMs)*100), 1) %>%
```

```{r}
# PSMs_calc_long <- PSMs_calc %>%
  
```

PSM plot 1

```{r}
my_colors <- c("#A23C33FF", "#83992AFF",  "#3C9770FF", "#44709DFF",
               "#D97828FF", "#DEB340FF", "#E5AD4FFF" )
```

```{r}
anno_seq <- PSMs %>%
  group_by(pick(where(is.factor))) %>%
  count(`Spectrum File`) 
```

```{r}
anno_seq_plot <- anno_seq %>%
  ggplot(aes(x = Fraction,
             y = n,
             fill = PD_set,
             #color = Year,
             #alpha = TechRepMixture
             )) +
  geom_boxplot(alpha = 0.7, width = 0.5,
               position = position_dodge(width =0.7)) +
  stat_summary(fun = mean, geom="point", shape = 4,
               size = 1, stroke =0.5, color = "grey10",
               position = position_dodge(width =0.7)) +
  geom_point(color = "grey10", shape = 1, size = 0.5,
             position=position_jitterdodge(dodge.width = 0.7,
                                           jitter.width = 0.15)) +
  scale_fill_manual(#name = "Dosage",
                    values=alpha(c("#E5AD4FFF", "#D97828FF",
                                   "#83992AFF", "#A23C33FF", 0.9)),
                    # breaks=c("0.5", "1", "2"),
                    # labels=c("0.5", "1", "2"),
                    #guide = guide_legend(title.position = "top")
                                         ) + 
  labs(y = 'No. of Annotated Sequences (PSMs)',
       title = ) +
  #facet_wrap(vars(Year), ncol = 1) +
  
  theme_bw() +
  theme(legend.position = "bottom")

anno_seq_plot
```

Shared Proteins

```{r}
# # https://dplyr.tidyverse.org/reference/summarise_all.html
# set_stats <- PSMs %>%
#   group_by(PD_set) %>%
#   mutate(n_PSMs = n()) %>% #n of rows in that group (so PSM no. of sample)
#   summarise_if(is.numeric, mean)
  
```

Overall

```{r}
shared_pep_by_set <- PSMs_calc %>%
  group_by(PD_set) %>%
  count(shared_peptide)
```

```{r}
# https://r-graph-gallery.com/48-grouped-barplot-with-ggplot2.html
shared_pep_by_set_plot <- shared_pep_by_set %>%
  ggplot(aes(x = PD_set,
             fill = shared_peptide,
             y = n,
             color = PD_set
             )) +
  geom_bar(position = "fill", stat = "identity",
           color = 'grey10', width = 0.7) + 
  scale_fill_manual(values=c("#E1ECF4", "#87b1d4")) + 
  theme_bw() +
  theme(legend.position = "bottom")
  
shared_pep_by_set_plot
```

By number

```{r}
no_p_shared_pep <- PSMs %>%
  group_by(PD_set) %>%
  count(`Number of Proteins`)
```

```{r}
no_p_shared_pep_plot <- no_p_shared_pep %>%
  # mutate(`Number of Proteins` = as.factor(`Number of Proteins`)) %>%
  ggplot(aes(x = `Number of Proteins`,
             # alpha = `Number of Proteins`,
             fill = PD_set,
             y = n
             )) +
  xlim(0,10) + 
  # geom_bar(position = "fill", stat = "identity",
  #          color = 'grey10', width = 0.7) + 
  geom_col(position = 'dodge', color = 'grey10',
           alpha = .8) +
  # scale_fill_manual(values=c("#E1ECF4", "#87b1d4")) + 
  # facet_wrap(vars(PD_set), ncol = 1) +
  theme_bw() + 
  theme(legend.position = "bottom")

no_p_shared_pep_plot
```

Correlation between sequence length & no. proteins

-   [[https://r-graph-gallery.com/50-51-52-scatter-plot-with-ggplot2.html]{.smallcaps}](https://r-graph-gallery.com/50-51-52-scatter-plot-with-ggplot2.html){.uri}

-   [[https://www.sthda.com/english/articles/32-r-graphics-essentials/131-plot-two-continuous-variables-scatter-graph-and-alternatives/]{.smallcaps}](https://www.sthda.com/english/articles/32-r-graphics-essentials/131-plot-two-continuous-variables-scatter-graph-and-alternatives/){.uri}

    -   Some interesting options e.g. ggMarginal [ggextra] & ggscatterhist() [ggpubr]

```{r fig.height=7, fig.width=7}

# plot(PSMs_calc$trm_seq_len, PSMs_calc$`Number of Proteins`)

seq_by_nop <- PSMs_calc %>%
  ggplot(aes(x = trm_seq_len,
             y =`Number of Proteins`,
             fill = PD_set,
             color = PD_set)) + 
  geom_point(shape = 1, size = 0.5, alpha = 0.5) +
  geom_smooth(method=lm , color="grey50", fill="grey90", se=TRUE) +
  scale_color_manual(values=alpha(c("#E5AD4FFF", "#D97828FF",
                                   "#83992AFF", "#A23C33FF", 0.9))) +
  facet_wrap(vars(PD_set), ncol = 2) +
  theme_bw() +
  theme(legend.position = "bottom")


seq_by_nop
```

```{r fig.height=6, fig.width=6}
# scatter plot + ggMarginal Density plot

seq_by_nop_2 <- PSMs_calc %>%
  ggplot(aes(x = trm_seq_len,
             y =`Number of Proteins`,
             color = PD_set)) + 
  geom_point(shape = 1,  alpha = 0.5) +
  scale_color_manual(values=alpha(c("#E5AD4FFF", "#D97828FF",
                                   "#83992AFF", "#A23C33FF", 0.9))) +
  theme_bw() +
  theme(legend.position = "bottom")

# add density plot via ggextra's ggMarginal function
seq_by_nop_2_ggmar <- ggMarginal(seq_by_nop_2, type = "density")

# looks cooked lol - prob because not quite continuous data...
seq_by_nop_2_ggmar
```

### Abundance/Intensity plots

```{r}
Abundances <- PSMs %>%
  dplyr::select(PD_set:Fraction, `Abundance 126`:`Abundance 134N`) %>%
  pivot_longer(
    cols = starts_with("Abundance"),
    names_to = 'TMT_tag',
    values_to = "intensity",
    names_prefix = "Abundance" # removes "Abundance" prefix from names value
  ) 
  
```

```{r fig.height=10, fig.width=9}
# weird - goes >4000
# also what is up with trailing distribution...
Abundances_density <- Abundances %>%
  filter(PD_set == '1pspg_2025') %>%
  filter(intensity >1 & intensity <1000) %>%
  ggplot(aes(x = intensity,
             y = TMT_tag,
             fill = TMT_tag)) +
  geom_density_ridges(alpha = 0.2) +
  facet_wrap(vars(Mixture), ncol = 4) +
  theme_bw() +
  theme(legend.position = "none")

Abundances_density
  
```

```{r fig.height=5, fig.width=7}
#Sample intensities box plot
intensity_bp <- Abundances %>%
  filter(PD_set == '1pspg_2025') %>%
  ggplot(aes(x = TMT_tag, 
             y = intensity,
             fill = Mixture)) +
  geom_boxplot() +
  scale_fill_manual(values=c("#d5a6bd", "#9fc5e8",
                             "#ea9999", "#f9cb9c")) +
  labs(subtitle = 'Extreme outliers?!?! + note 2024 & others similiar') +
  theme_bw() +
  theme(legend.position = "none")
  
intensity_bp
```

```{r fig.height=7, fig.width=12}
intensity_bp_2 <- Abundances %>%
  filter(PD_set == '1pspg_2025') %>%
  ggplot(aes(x = TMT_tag, 
             y = intensity,
             fill = Mixture)) +
  geom_boxplot(outliers = FALSE) +
  stat_summary(fun = mean, geom = "point", shape = 4, size = 0.5) +
  scale_fill_manual(values=c("#d5a6bd", "#9fc5e8",
                             "#ea9999", "#f9cb9c")) +
  labs(subtitle = 'mean skewed by outliers (also not accounting for NAs)') +
  # facet_wrap(vars(Year), ncol = 2) +
  facet_wrap(vars(Mixture), ncol = 4) +
  theme_bw() +
  theme(legend.position = "bottom",
        panel.spacing = unit(0, "cm"), # Removes space between panel)
        axis.text.x = element_text(size = 7.5, color ="grey10"))  


intensity_bp_2
```

------------------------------------------------------------------------

### 2026_01_26 - AMPA 

```{r}
EExAge_56_anno <- read_csv("input/anno/2026_01_22_mssTMT_EExAGE_anno_HA_56.csv") %>%
  mutate(across(c(Year, Mixture, TechRepMixture, Fraction), as.factor)) #%>%
  #dplyr::select(Mixture:Sex) # rm NAs entries otherwise doesn't join? 
```

```{r}
#Gria 1-4 AMPA 1-4 | rat acc
AMPA_list <- c("P19490", "P19491", "P19492", "P19493")

AMPA_PSMs <- PSMs_calc %>%
  ungroup() %>%
  # filter rows with where prot acc has  any of  the items in AMPA list 
  filter(str_detect(`Protein Accessions`,
                    str_c(AMPA_list, collapse = "|"))) 
  

AMPA_PSMs_lm <- AMPA_PSMs %>%
  pivot_longer(
    cols = starts_with("Abundance"),
    names_to = 'TMT_tag',
    values_to = "intensity",
    names_prefix = "Abundance " # removes "Abundance " prefix from names value
    # NOTE : Hidden space " 126" from "Adundance" is what cause non joining! 
  ) %>% 
  left_join(EExAge_56_anno,
            by = join_by("Year" == "Year",
                         "Mixture" == "Mixture",
                         "TechRepMixture" == "TechRepMixture",
                         "Fraction" == "Fraction",
                         "TMT_tag" == "Channel"),
            relationship = "many-to-many") 


# Checks
AMPA_PSMs_lm %>%
  count(Year, Mixture, TechRepMixture, Fraction, TMT_tag)  # 1536 x 6

EExAge_56_anno %>%
  count(Year, Mixture, TechRepMixture, Fraction, Channel) # 1536 x 6"

# also use antijoin() to check!!! (sees what was left behind in X & not joined)
```

\
\#### Return to late try with normalised data.

```{r}
AMPA_pep_plot <- ggplot(AMPA_PSMs_lm,
                        aes(x = Condition, y = intensity)) +
  
  geom_boxplot(aes(fill = Condition)) +
  scale_fill_manual(values = c("#8CB585", "#66A61EFF", 
                               "grey50",
                               "#D4C07E", "#E6AB02FF")) + # housing colors
  ylim(0,1000) +
  facet_nested(~ shared_peptide + Year + Mixture) +
  theme_bw() +
  theme(legend.position = "bottom")

AMPA_pep_plot

```

```{r}
AMPA_PSMs_seq_sum <- AMPA_PSMs_lm %>%
  group_by(shared_peptide, Mixture, Condition) %>%
  summarise(mean_abund = mean(intensity, na.rm = TRUE),
            med_abund = median(intensity, na.rm = TRUE))

print(AMPA_PSMs_seq_sum)
```

# Result Statistics

#### Random plot

```{r}
mz_tmt <- c(126.127,127.121,127.131,128.128,
            128.134,129.065,129.131,130.086,
            130.141,131.065,131.144,132.148,
            132.141,133.151,133.145,134.148)

tmt_in <- c(124,122,90,116.1,
            105.1,149,88.6,105,
            114.6,57.3,79.5,137,
            99.5,91.7,85.4,89.2)
```

------------------------------------------------------------------------

## 
