---
title: "Wassouf_2018_RNAseq_rawcounts"
author: "Stephanie Huang"
date: "2025-12-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Data was retrieved from: <https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE96961>

### Setup

```{r}
# # https://bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.html
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# 
# BiocManager::install("DESeq2")
```

```{r}
library(tidyverse)
library(readr)
library(readxl)
library(ggplot2)
library(ggrepel) #  for geom_text_repel() for volcano plot annotations
library(DESeq2) # for GLM for transcriptomics raw data processing to get DEGs
# library(UniProt.ws) 
library(pheatmap) # for heat maps to got with DESeq2 
```

```{r}
#read in raw data
df_1 <- read_csv("input/Wassouf_2018_RNAseq_GSE96961_rawcounts.csv")
```

### Sample information.

The data was matched to the samples based on the following details.

Example of matching

- See the following link <https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSM2546841>

  - Sample GSM2546841

  - Title: WT_SE_1

  - Description - Sample name: I14R018b05

  - GSM2546842 \| I14R018b06 \| WT_SE_2

  - GSM2546843 \| I14R018b07 \| WT_SE_3

  - GSM2546844 \| I14R018b08 \| WT_SE_4

  - GSM2546845 \| I14R018b01 \| WT_EE_1

  - GSM2546848 \| I14R018b04 \| WT_EE_4

  - GSM2546849 \| I14R018b13 \| TG_SE_1

  - GSM2546853 \| I14R018b09 \| TG_EE_1

```{r}
summary(df_1)
colnames(df_1)
```

```{r}
# list2check <- c('Dlg4', 'Syp', 'Bdnf', 'Shank1', 'Shank3', 'Psd95', 'Dlgh4',
#                 'Arx')
# 
# list2check_2 <- c('Aldh1a1','Anxa11','Arx','Bdnf','Bok','Camk2n2','Ccdc136',
#                   'Cdh22','Coro6','Dagla','Dpp10','Ebf4','Echdc2','Egr3','Fosb',
#                   'Gng4','Iqgap2','Kif17','Klk8','Lct','Lingo3','Lrrc10b',
#                   'Myoc','Nap1l5','Nptx2','Pdyn','Plk5','Ptgs2','Rasd1','Sbsn',
#                   'Sema5a','Sipa1l3','Stmn1','Syt12','Tanc1','Tle2','Tmsb10',
#                   'Trpc6','Psd95','Shank1','Shank3', # 41 from supplmentary table
#                   'Syp', 'Dlg4', 'Dlgh4')
# 
# pattern_2 <- paste(list2check_2, collapse = "|") # | OR 
```

### df_5 - ADD SOME NOTES ON THIS SECTION!

```{r}
# # match either by primary gene names OR str_detect for multi entries 
# # MUCH BETTER THAN PREVIOUS ! 
# df_5 <- df_3 %>%
#   filter(Gene.Names..primary. %in% list2check_2 | str_detect(Gene.Names, pattern_2)) %>% 
#   tidyr::pivot_longer(cols = WT_EE__01_01:TG_SE__04_16,
#                       names_to = c("Genotype", "Housing",  "Sample_No", "Sample_ID"),
#                       names_pattern = "(.+)_(.+)__(.+)_(.+)", #  .+  -> 1 or more chr
#                       values_to = "raw_read_counts")
#   
```

Interim End - Quick summary note.

------------------------------------------------------------------------

# Testing DESeq2 - EEvsSH (WT only)

Replication attempt of Wassouf et al., 2018 t1/v1 - testing 8 samples only WT_EE vs WT_SE.

- <https://bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.html>

## Count matrix input

<https://bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#countmat>

```{r}
# give new names to columns so matches those provided in GEO db
new_names <- c("ENSEMBL",
               'WT_EE__01_01', 'WT_EE__02_02', 'WT_EE__03_03', 'WT_EE__03_04',
               'WT_SE__01_05', 'WT_SE__02_06', 'WT_SE__03_07', 'WT_SE__04_08',
               'TG_EE__01_09', 'TG_EE__02_10', 'TG_EE__03_11', 'TG_EE__04_12',
               'TG_SE__01_13', 'TG_SE__02_14', 'TG_SE__03_15', 'TG_SE__04_16')

colnames(df_1) <- new_names # if writing over original 
```

### Make cts matrix from raw data & coldata df for sample/condition info

```{r}
# # ALL 4 GROUPS 
# 
# # make matrix from Wassouf et al.'s raw read count csv
# cts <- as.matrix(df_1[, 2:17])
# rownames(cts) <- df_1$ENSEMBL 
# 
# # Make sample data/coldata
# # my method is a bit diff to the DESeq2 userguide since i don't have separate sample df
# coldata <- data.frame(row.names = new_names[2:17],
#                       Genotype = c(rep('WT',8), rep('TG',8)),
#                       Housing = c(rep('EE', 4), rep('SE', 4),
#                                   rep('EE', 4), rep('SE', 4)),
#                       stringsAsFactors = TRUE) #
```

```{r}
# JUST WT SE VS EE!

# make matrix from Wassouf et al.'s raw read count csv
cts <- as.matrix(df_1[, 2:9])
rownames(cts) <- df_1$ENSEMBL 

# Make sample data/coldata
# my method is a bit diff to the DESeq2 userguide since i don't have separate sample df
coldata <- data.frame(row.names = new_names[2:9],
                      Housing = c(rep('EE', 4), rep('SE', 4)),
                      stringsAsFactors = TRUE) 
```

```{r}
# quick check to make sure in the same order as the samples
# if not, see the userguide to fix

head(cts,2) # yes matches
coldata
```

### Construct DESeqDataSet (dds) object in preparation

```{r}
#library("DESeq2")
dds <- DESeqDataSetFromMatrix(countData = cts,
                              colData = coldata,
                              design = ~ Housing)
dds
```

## Pre-filtering

<https://bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#pre-filtering>

```{r}
#keeps genes/rows with at least 4 samples with 10 or more read counts
# (size of 4) count of >= 10 is recommened by DESeq2 - keeps 16925
# but Wassouf did min of 50? - they got 12287.
# if you do smallestGroupsize 4 * counts =>50 then you get 12315 |
# 2026_09 - tested >=10 but still gives same no Nas as before, back to 50 to match
smallestGroupSize <- 4
keep <- rowSums(counts(dds) >= 50) >= smallestGroupSize
dds <- dds[keep,] # index dds with keep to select those that meet condition
```

#### Note on factor levels

```{r}
dds$Housing <- relevel(dds$Housing, ref = "SE") # reference level is SE
```

## **Differential expression analysis**

```{r}
# run D
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients
res <- results(dds, name="Housing_EE_vs_SE")
```

### **Log fold change shrinkage for visualization and ranking**

```{r}
resultsNames(dds)
```

```{r}
# resLFC <- lfcShrink(dds, coef = "Housing_EE_vs_SE", type = "apeglm")
```

```{r}
resOrdered <- res[order(res$padj),]
sum(res$padj < 0.1, na.rm=TRUE) #60
summary(res)
```

```{r}
# write.csv(as.data.frame(resOrdered), 
#           file="output/2025_12_22_Wassouf_2018_DESeq2_renalysis_1_WTonly_EEvsSE.csv")
```

```{r}
resOrdered_df <- as.data.frame(resOrdered)
```

## NEW CONTINUES FROM HERE!!! (2026_09_20)

## Filtered version

```{r}
# NEW 
resOrdered_df_adjp.1 <- resOrdered_df %>%
  filter(padj <=0.1) %>%  #60 
  tibble::rownames_to_column(var = "ENSEMBL")

resOrdered_df_adjp.15 <- resOrdered_df %>%
  filter(padj <=0.15) %>%  #105 results
  tibble::rownames_to_column(var = "ENSEMBL")
```

```{r}
write_excel_csv(resOrdered_df_adjp.15, "output/2026_09_20_Wassouf_2018_DESeq2_renalysis_WTonly_EEvsSE_adjp15_105.csv")
```

```{r}
ensembl_2_uniprot <- read_excel("input/Wassouf_2018_DESeq_test_1_EEvsSE_p01_Ensembl_2_Uniprot_idmapping_2025_12_22.xlsx")

ensembl_2_uniprot <- ensembl_2_uniprot %>%
   mutate(existence_ranking = case_when( # ranking level of evidence
    `Protein existence` == "Evidence at protein level" ~ 5,
    `Protein existence` == "Evidence at transcript level" ~ 4,
    `Protein existence` == "Inferred from homology" ~ 3, 
    `Protein existence` == "Predicted" ~ 2, 
    `Protein existence` == "Uncertain" ~ 1
    ), .after = `Protein existence`) %>%
  mutate(across(where(is.character), ~ na_if(.x, ""))) 
   
# need to run in separate line as mouse ref is input (otheriwise object not found )
ensembl_2_uniprot <- ensembl_2_uniprot %>% 
  mutate(occupied = rowSums(!is.na(ensembl_2_uniprot)))
```

```{r}
resOrdered_df_adjp.15_mapped <- left_join(resOrdered_df_adjp.15, ensembl_2_uniprot,
                                       by = join_by(ENSEMBL == From)) # 397x31
#1 unmapped
```

```{r}
length(unique(resOrdered_df_adjp.15_mapped$ENSEMBL)) # 105

resOrdered_df_adjp.15_mf <- resOrdered_df_adjp.15_mapped %>%  #397
  filter(!is.na(Entry)) %>%  #396
  filter(!is.na(padj)) %>%  #396
  group_by(ENSEMBL) %>% # keeps unique ensembl!
  filter(if(any(Reviewed == "reviewed", na.rm = TRUE)) {
    Reviewed == "reviewed" # keeps reviewed if possible
  } else{TRUE})  %>% #149 
  slice_max(Annotation, n = 1) %>% # 129 - keep highest anno ranking!
  slice_max(`Protein existence`, n = 1) %>%  #129 still
  slice_max(occupied, n = 1, with_ties = FALSE) %>% # 104 - keep highest complete
  mutate(Regulation = ifelse(log2FoldChange > 0, "Up", "Down"),
         .after = lfcSE) %>% # add up or down col
  arrange(Regulation, padj)


# checks
length(unique(resOrdered_df_adjp.15_mf$ENSEMBL)) # 104
length(unique(resOrdered_df_adjp.15_mf$`Gene Names`)) # 104
length(unique(resOrdered_df_adjp.15_mf$`Gene Names (primary)`)) # 104
length(unique(resOrdered_df_adjp.15_mf$Entry)) # 104
colnames(resOrdered_df_adjp.15_mf)
```

```{r}
# write_excel_csv(resOrdered_df_adjp.15_mf,
#    "output/2026_09_20_Wassouf_EEvsSH_WTonly_resOrdered_df_adjp15_mapped104.csv")
```

------------------------------------------------------------------------

### Unfiltered version (but rm no adj.p vals)

```{r}
resOrdered_df_full <- resOrdered_df %>% # 12315
  filter(!is.na(padj)) %>% # 1800
  tibble::rownames_to_column(var = "ENSEMBL")
```

```{r}
resOrdered_df_full_map <- left_join(resOrdered_df_full, ensembl_2_uniprot,
                                       by = join_by(ENSEMBL == From)) #2105
```

```{r}
length(unique(resOrdered_df_full_map$ENSEMBL)) # 2105

resOrdered_df_full_mf <- resOrdered_df_full_map %>%  # 2105
  filter(!is.na(Entry)) %>%  #404 
  filter(!is.na(padj)) %>%  #404 - same since filtered above
  group_by(ENSEMBL) %>% # keeps unique ensembl!
  filter(if(any(Reviewed == "reviewed", na.rm = TRUE)) {
    Reviewed == "reviewed" # keeps reviewed if possible
  } else{TRUE}) %>% #152
  slice_max(Annotation, n = 1) %>% # 134 - keep highest anno ranking!
  slice_max(`Protein existence`, n = 1) %>%  #134 still
  slice_max(occupied, n = 1, with_ties = FALSE) %>% # 109 - keep highest complete
  mutate(Regulation = ifelse(log2FoldChange > 0, "up", "down"),
         .after = lfcSE) %>% # add up or down col
  arrange(Regulation, padj)


# # checks
length(unique(resOrdered_df_full_map$ENSEMBL)) # 1810
length(unique(resOrdered_df_full_map$`Gene Names`)) # 168
length(unique(resOrdered_df_full_map$`Gene Names (primary)`)) # 110
length(unique(resOrdered_df_full_map$Entry)) # 405
colnames(resOrdered_df_full_map)
```

```{r}
# write_excel_csv(resOrdered_df_full_mf,
#    "output/2026_09_20_Wassouf_EEvsSH_WTonly_resOrdered_df_full_noNAs_mapped109.csv")
```

------------------------------------------------------------------------

## Lit Review map (new - 2026_09_20)

Note generally matches directional of results in their paper (+ Supplementary) - e.g. shank1, Psd95 Shank3 up in EE while Nap1l5 down in EE etc -

BUT not entirely consistent with their supp figure 4b (which Perez 2024 compares to) - see below

```{r}
# for lit review table 
lit_review_transcriptomics <- resOrdered_df_full_mf %>%
  mutate(
    space = " ",
    Year = 2018,
    Author = 'Wassouf',
    Animal = 'Mice',
    Condition = 'WT',
    Region = 'Hippocampus',
    Whole = 'Yes',
    Exp_type = 'mRNA',
    Method = 'TRANSCRIPTOMICS_RNAseq',
    Used_name = paste0(`Gene Names (primary)`, ' - ', `ENSEMBL`),
    Mod = " ",
    Mod_type = " ",
    Gene_name = `Gene Names (primary)`, 
    Comparison = 2, # EE vs SH
    In_EE = Regulation,
    Fold_change = log2FoldChange, # see also lfcSE
    Arrows = case_when( # dplyr case when, instead of ifelse
      log2FoldChange > 2.5 ~ '+++',
      log2FoldChange> 1 ~ '++',
      log2FoldChange > .05 ~ '+',
     log2FoldChange > -.05 ~ '=',
      log2FoldChange> -1 ~ '-',
      log2FoldChange > -2.5 ~ '--',
     log2FoldChange <= -2.5 ~ '---',
      ),
    Significant = if_else(padj <= 0.05, "yes", "no"),
    pvalue = padj,
    N_tested = '4', # 4 WT per group 4 EE WT vs 4 SH WT
    Stats_adj = 'adjusted',
    Note = ' ',
    Full_name = `Protein names`,
    Uniprot_rat_accession = Entry,
    Uniprot_entry_name = `Entry Name`,
    Uniprot_link = ifelse(!is.na(Entry),
        paste0('https://www.uniprot.org/uniprotkb/', Entry,'/entry'),' '),
    Wikipedia_link = ifelse(!is.na(Entry),
          paste0('https://en.wikipedia.org/wiki/', `Gene Names (primary)`), ' '),
    
    Curation_method = case_when(
      Reviewed == "reviewed" ~ "auto - mice ENSEMBL 2 uniprot ref & reviewed ( mice primary gene name)",
      Reviewed == "unreviewed"  ~ "auto - mice ENSEMBL 2 uniprot ref & unreviewed ( mice primary gene name)" ),
    
    Class_group = ' ',
    Other_link = ' ',
    Other_comments = 'see Wassouf_2018_RNAseq_rawcounts.Rmd EE lit review script',
    misc = paste0(ENSEMBL, '_', `Gene Names (primary)`, '_(',  `Protein names`,')')
  ) %>%
  arrange(desc(In_EE), Gene_name)

```

```{r}
# # write out lit review df | done 2026_09_20!
# write_excel_csv(lit_review_transcriptomics,
#    "output/2026_09_20_Wassouf_2018_RNAseq_mRNA_lit_review_WT_only_EEvsSH.csv")

```

### Manual curation

Confused - so in supplementary fig 4b there are 23 up and 15 down which is 38, but in their supp fig 4a pic it says 41 DEG (23 up & 18 down.)

Not sure if missing is a matter of mapping because hard 2 tell from ENSEMBL alone (check later) or maybe due to using all mice (transgenic & WT)

```{r}
# Comparison to Wassouf 2018 (from supp figure 4b) 
# perez comp has 22 + 15 - this is 23 + 15 
down <- c('Tmsb10','Stmn1','Tle2','Ebf4','Echdc2','Cdh22','Arx',
          'Camk2n2','Ccdc136','Nap1l5','Dpp10','Gng4','Myoc','Sbsn','Coro6')

up <- c('Nptx2','Klk8','Lingo3','Bok','Lrrc10b','Lct',
        'Aldh1a1','Anxa11','Sipa1l3','Dagla','Syt12',
        'Kif17','Rasd1','Iqgap2','Tanc1','Sema5a', 'Trpc6',
        'Bdnf','Ptgs2','Plk5','Pdyn','Egr3','Fosb')
```

```{r}
wassouf_DEG <- data.frame(Direction_w  = c(rep("up", 23), rep("down", 15)),
                            gene = c(up, down)) 
```

```{r}
down_check <-resOrdered_df_full_mf %>% # 2 only - Nap1l5 & Stmn1
  filter(`Gene Names (primary)` %in% down)

up_check <-resOrdered_df_full_mf %>% # 2 only -  Sipa1l3 & Dagla
  filter(`Gene Names (primary)` %in% up)
```

```{r}
lit_review_manual <- wassouf_DEG %>%
  mutate(
    space = " ",
    Year = 2018,
    Author = 'Wassouf',
    Animal = 'Mice',
    Condition = 'WT & TG',
    Region = 'Hippocampus',
    Whole = 'Yes',
    Exp_type = 'mRNA',
    Method = 'TRANSCRIPTOMICS_RNAseq',
    Used_name = gene,
    Mod = " ",
    Mod_type = " ",
    Gene_name = gene, 
    Comparison = 2, # EE vs SH
    In_EE = Direction_w,
    Fold_change = " ", # see also lfcSE
    Arrows = " ",
    Significant = "yes",
    pvalue = " ",
    N_tested = ' ', # 4 WT per group 4 EE WT vs 4 SH WT
    Stats_adj = ' ',
    Note = 'Manual curation from Wassouf 2018 Supp Fig 4B (38 - 23 up & 15 down) - only (up) Nap1l5 & Stmn1 & (down) Sipa1l3 & Dagla overlapped with the WT EE vs SH re-analysis uniprot automap - not sure if lack of overlap is due to missing data or mapping from ENSEMBL 2 UNIPROT??? possibly due to their analysis actually using all 16 mice including transgenic',
    Full_name = " ",
    Uniprot_rat_accession = " ",
    Uniprot_entry_name = " ",
    Uniprot_link = " ",
    Wikipedia_link = " ",
    Curation_method = "Manually by gene name from Supp Fig 4B",
    Class_group = ' ',
    Other_link = ' ',
    Other_comments = 'see Wassouf_2018_RNAseq_rawcounts.Rmd EE lit review script',
    misc = ' ') %>%
  arrange(desc(In_EE), Gene_name)
```

```{r}
# write out manual lit review df | done 2026_09_20!
# write_excel_csv(lit_review_manual,
#    "output/2026_09_20_Wassouf_2018_RNAseq_mRNA_manual_sfig4B_38_lit_review_WT&TG_EEvsSH.csv")
```

------------------------------------------------------------------------

------------------------------------------------------------------------

# Below here is old - don't use!

```{r}
# OLD 
resOrdered_df_p.01 <- resOrdered_df %>%
  filter(pvalue <=0.01) %>%  #267 results | still lots of NAs
  tibble::rownames_to_column(var = "ENSEMBL")

resOrdered_df_p.015 <- resOrdered_df %>%
  filter(pvalue <=0.015) %>%  #330 results
  tibble::rownames_to_column(var = "ENSEMBL")
```

```{r}
# write_excel_csv(resOrdered_df_p.01, "output/2025_12_22_Wassouf_2018_DESeq2_renalysis_1_WTonly_EEvsSE_p01.csv")
```

Ohhhhhh the reason they have their adjust p-value cutoff at 0.15 is likely due the ones above that give NA! See exported data!

------------------------------------------------------------------------

Map significant ones

```{r}
# mouseUp <- UniProt.ws(10090)
```

```{r}
# keytypes(mouseUp)
# columns(mouseUp)
```

```{r}
# ensembl_to_map <- rownames(resOrdered_df_p.01)
# # pattern_2 <- paste(list2check_2, collapse = "|") # | OR 
```

```{r}
# ensembl_to_up <-  mapUniProt(
#                     from = "xref_ensembl",
#                     to = "UniProtKB",
#                     columns = c("id", "accession", "gene_primary",
#                                 "organism_name", "protein_name", "reviewed"),
#                     query = list(organism_id = 10090, ids = ensembl_to_map))

# Error in `req_perform()`:
# ! HTTP 400 Bad Request.
```

------------------------------------------------------------------------

## Uniprot.ws error so just manually get from uniprot_id mapping

```{r}
# ensembl_2_uniprot <- read_excel("input/Wassouf_2018_DESeq_test_1_EEvsSE_p01_Ensembl_2_Uniprot_idmapping_2025_12_22.xlsx")
# 
# ensembl_2_uniprot <- ensembl_2_uniprot %>%
#    mutate(existence_ranking = case_when( # ranking level of evidence
#     `Protein existence` == "Evidence at protein level" ~ 5,
#     `Protein existence` == "Evidence at transcript level" ~ 4,
#     `Protein existence` == "Inferred from homology" ~ 3, 
#     `Protein existence` == "Predicted" ~ 2, 
#     `Protein existence` == "Uncertain" ~ 1
#     ), .after = `Protein existence`) %>%
#   mutate(across(where(is.character), ~ na_if(.x, ""))) 
#    
# # need to run in separate line as mouse ref is input (otheriwise object not found )
# ensembl_2_uniprot <- ensembl_2_uniprot %>% 
#   mutate(occupied = rowSums(!is.na(ensembl_2_uniprot)))
```

```{r}
resOrdered_df_p.01_mapped <- left_join(resOrdered_df_p.01, ensembl_2_uniprot,
                                       by = join_by(ENSEMBL == From)) # 850 x 29 
# 8 unmapped
```

```{r}
# length(unique(resOrdered_df_p.01_mapped$ENSEMBL)) # 267

resOrdered_df_p.01_mf <- resOrdered_df_p.01_mapped %>% # 850
  filter(!is.na(Entry)) %>% # 842
  filter(!is.na(padj)) %>% # 404 | if KEEP NAs instead -> FINS on 259
  group_by(ENSEMBL) %>% # keeps unique ensembl!
  filter(if(any(Reviewed == "reviewed", na.rm = TRUE)) {
    Reviewed == "reviewed" 
  } else{TRUE})  %>% #154
  slice_max(Annotation, n = 1) %>% # 134 | or 109 no ties
  slice_max(`Protein existence`, n = 1) %>% #134 sill
  slice_max(occupied, n = 1, with_ties = FALSE) %>% # 119 | or 109 with ties
  mutate(Regulation = ifelse(log2FoldChange > 0, "Up", "Down"),
         .after = lfcSE) %>% # add up or down col
  arrange(Regulation, padj)

# checks  || if keep NAs is 267 instead
length(unique(resOrdered_df_p.01_mf$ENSEMBL)) # 109 |109
length(unique(resOrdered_df_p.01_mf$`Gene Names`)) # 111 | 109
length(unique(resOrdered_df_p.01_mf$`Gene Names (primary)`)) # 109 | 109
length(unique(resOrdered_df_p.01_mf$Entry)) # 134 | 109
colnames(resOrdered_df_p.01_mf)
```

```{r}
# write_excel_csv(resOrdered_df_p.01_mf,
#                 "output/2025_12_23_Wassouf_test_1_EEvsSH_WTonly_mapped_109.csv")

# re-export to see
# write_excel_csv(resOrdered_df_p.01_mf,
#                 "output/2025_09_20_Wassouf_test_1_EEvsSH_WTonly_mapped_109.csv")

# write_excel_csv(resOrdered_df_p.01_mf,
#                 "output/2026_09_20_Wassouf_test_1_EEvsSH_WTonly_mapped_259_w_NAs.csv")
```

## Results Plots (old)

```{r}
volcano_adjpv <- ggplot(resOrdered_df_p.01_mf, 
                     aes(x =log2FoldChange, y = -log2(padj), label = `Gene Names (primary)`)) +
 geom_point(aes(color = case_when(
   padj < 0.05 & log2FoldChange > 0 ~ "Upregulated (adj. p<0.05)",
   padj < 0.05 & log2FoldChange < 0 ~ "Downregulated (adj. p<0.05)",
   padj < 0.1 & log2FoldChange > 0 ~ "Upregulated (adj. p<0.1)",
   padj < 0.1 & log2FoldChange < 0 ~ "Downregulated (adj. p<0.1)",
   TRUE ~ "Not Significant")),
   alpha = 1, size = 2.5, shape = 1) +
  # facet_wrap(~ Label) +
  # xlim(-1,1) +
  scale_color_manual(values = c("firebrick3", 
                                "royalblue4", 
                                "lightpink1", 
                                "lightblue3" , 
                                 "grey80"),
                     breaks = c("Upregulated (adj. p<0.05)", 
                                "Downregulated (adj. p<0.05)", 
                                "Upregulated (adj. p<0.1)", 
                                "Downregulated (adj. p<0.1)", 
                                "Not Significant"),
                     labels = c("Upregulated Proteins (up to adj. p<0.05)", 
                                "Downregulated Proteins (up to adj. p<0.05)", 
                                "Upregulated Proteins (up to adj. p<0.1)", 
                                "Downregulated Proteins (up to adj. p<0.1)" , 
                                "Not Significant")) +
  
  geom_text_repel() +
  labs(title = "Wassouf 2018 RNAseq DESeq2 test 1 EE vs SE (WT only 8 mice) - Significant Proteins (adjp<0.05) (2025_12_23)",
       subtitle = "Presented like MSstats and my data log2FC & -log2(adj.pvalue)",
       x = "log2FC",
       y = "-Log2 (adjusted p-value)",
       color = "Protein Regulation") +
  # geom_hline(yintercept = -log2(0.05), linetype = "dashed", linewidth = 0.2,  color = "grey30") +
  # annotate("text", x = .86, y = -log2(0.05), label = "Adj. p-value cutoff (0.05) \n",
  #          fontface = 'italic', size = 4, color = "grey10") +
  # geom_hline(yintercept = -log2(0.1), linetype = "dashed", linewidth = 0.2,  color = "grey50") +
  # annotate("text", x = .86, y = -log2(0.1), label = "Adj. p-value cutoff (0.1) \n",
  #          fontface = 'italic', size = 4, color = "grey30") +
  theme_bw() + 
  theme(legend.position = "bottom")



volcano_adjpv
```

```{r}
# TO DO - FIX LOLLIPOP CHART
# assign colors based on direction. | works by nvm ordering is weird
#label_colors <- ifelse(df_3$`EC log2 fold` > 0 , "#dc322f", "#268bd2")

colnames(resOrdered_df_p.01_mf)

# lollipop - ggplot ver - a big ugly
sig_plot <- resOrdered_df_p.01_mf %>%
  filter(padj < .15) %>%
  ggplot(aes(x = reorder(`Gene Names (primary)`, log2FoldChange), # orders by log fold
                         y = log2FoldChange,
             color = Regulation)) +
  geom_segment(aes(x = `Gene Names (primary)`,
                   yend = log2FoldChange, y = 0)) +
  geom_point(aes(alpha = padj),
             shape = 19, size = 4) +
  scale_alpha(range = c(1,.1), guide = 'none') + # rev alpha so darker = more significant
  # geom_text(aes(label = gene_name), color = "grey10", size = 2.5,
  #           position = position_stack()) +
  geom_hline(yintercept = 0, color = 'grey20', linewidth = 0.1) +
  scale_color_manual(breaks = c("Up", "Down"),
                     values = c("#dc322f", "#268bd2"),
                     guide = 'none') +
  labs(
    title = 'Wassouf 2018 RNAseq DESeq2 test 1 EE vs SE (WT only 8 mice) - Lollipop for padj < .15 (104 DEG) (2025_12_23)',
    subtitle = 'Ordered by Log2FC & then alphabetically by gene name, alpha is significance, blue is down and red is up.',
       y = 'EC Log2FC',
       x = 'Primary Gene Name') +
  #coord_flip() + #fliptheplot
  #facet_wrap(~fraction, space = 'free_x', scale = 'free_x') +
  #theme_void() +
  theme_bw() +
  theme(
    #axis.title = element_blank(),
    #axis.text.x = element_blank()
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, # 90, hj1 t/R, vj.5 cent
                               size = 9, face = 'italic'),
    axis.text.y = element_text(hjust = 1, color = "grey20", size = 10), # rota 90 & hjust=1 (top/right)
        )

sig_plot
```

------------------------------------------------------------------------

# DESeq2 test 2 (factorial 2 x 2 Housing x Genotype)

See **Multi-factor designs** <https://bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#multi-factor-designs>

<https://master.bioconductor.org/packages/release/workflows/vignettes/rnaseqGene/inst/doc/rnaseqGene.html>

See mentions of SVA (Surrogate variable analysis) mentioned in doc & in the paper.

```{r}
# ALL 4 GROUPS

# make matrix from Wassouf et al.'s raw read count csv
cts_t2 <- as.matrix(df_1[, 2:17])
rownames(cts_t2) <- df_1$ENSEMBL

# Make sample data/coldata
# my method is a bit diff to the DESeq2 userguide since i don't have separate sample df
coldata_t2 <- data.frame(row.names = new_names[2:17],
                      Genotype = c(rep('WT',8), rep('TG',8)),
                      Housing = c(rep('EE', 4), rep('SE', 4),
                                  rep('EE', 4), rep('SE', 4)),
                      stringsAsFactors = TRUE) #

# quick check to make sure in the same order as the samples
# if not, see the userguide to fix
head(cts_t2,2) # yes matches
coldata_t2
```

Now construct the DESeqDataSet (dds) object in preparation

```{r}
#library("DESeq2")
dds_t2 <- DESeqDataSetFromMatrix(countData = cts_t2,
                              colData = coldata_t2,
                              design = ~ Housing + Genotype) # 2 x 2?
dds_t2 
```

## Pre-filtering

<https://bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#pre-filtering>

```{r}
#keeps genes/rows with at least 4 samples with 10 or more read counts
# (size of 4) count of >= 10 is recommened by DESeq2 - keeps 16925
# but Wassouf did min of 50? - they got 12287.
# if you do smallestGroupsize 4 * counts =>50 then you get 12315
smallestGroupSize <- 4
keep <- rowSums(counts(dds_t2) >= 50) >= smallestGroupSize 
dds_t2 <- dds_t2[keep,] # index dds with keep to select those that meet condition
```

## **Differential expression analysis**

```{r}
# run DESeq 
dds_t2 <- DESeq(dds_t2)

# outputs
resultsNames(dds_t2) # lists the coefficients
res_t2_housing <- results(dds_t2, name="Housing_EE_vs_SE")
res_t2_genotype <- results(dds_t2, name="Genotype_WT_vs_TG")
```

### **Log fold change shrinkage for visualization and ranking**

```{r}
resultsNames(dds_t2)
```

```{r}
# HOUSING
# 13037 tested | padj <.1 | 178 up & 40 down
resOrdered_t2_housing <- res_t2_housing[order(res_t2_housing$padj),]
sum(res_t2_housing$padj < 0.1, na.rm=TRUE) #60
summary(res_t2_housing)

resOrdered_t2_housing_df <- as.data.frame(resOrdered_t2_housing) # unpack as a df

resOrdered_t2_housing_df <- tibble::rownames_to_column(resOrdered_t2_housing_df, var = "ENSEMBL")
```

```{r}
# write_excel_csv(resOrdered_t2_housing_df, "output/2025_12_23_Wassouf_2018_DESeq2_reanalysis_2_EEvsSH_all_16.csv")
```

```{r}
# GENOTYPE
# 13037 tested | padj <.1 | 61 up & 39 down
resOrdered_t2_genotype <- res_t2_genotype[order(res_t2_genotype$padj),]
sum(res_t2_genotype$padj < 0.1, na.rm=TRUE) #60
summary(res_t2_genotype)

resOrdered_t2_genotype_df <- as.data.frame(resOrdered_t2_genotype)
```

------------------------------------------------------------------------

```{r}
# read in & set up uniprot annotation file for t2 EE vs SH (all samples)
# queried uniprot (349 padj <.159) returned 1235
ensembl_2_uniprot_t2 <- read_excel(
  'input/Wassouf_2018_DESeq_test_2_EEvsSE_all_adjp15_Ensembl_2_Uniprot_idmapping_2025_12_23.xlsx')

ensembl_2_uniprot_t2 <- ensembl_2_uniprot_t2 %>%
  mutate(existence_ranking = case_when ( # ranking level of evidence
    `Protein existence` == "Evidence at protein level" ~ 5,
    `Protein existence` == "Evidence at transcript level" ~ 4,
    `Protein existence` == "Inferred from homology" ~ 3, 
    `Protein existence` == "Predicted" ~ 2, 
    `Protein existence` == "Uncertain" ~ 1), .after = `Protein existence`) %>%
  mutate(across(where(is.character), ~ na_if(.x, "")))

# need to run in separate line as mouse ref is input (otheriwise object not found )
ensembl_2_uniprot_t2 <- ensembl_2_uniprot_t2 %>%
  mutate(occupied = rowSums(!is.na(ensembl_2_uniprot_t2)))
```

```{r}
resOrdered_t2_housing_df_m <- resOrdered_t2_housing_df %>%
  filter(padj <= 0.159) %>% # returns 349 obs
  left_join(ensembl_2_uniprot_t2,
            by = join_by(ENSEMBL == From)) # 3 unmapped (1238 x 31)
```

```{r}
# checks
length(unique(resOrdered_t2_housing_df_m$ENSEMBL)) # 349
length(unique(resOrdered_t2_housing_df_m$`Gene Names`)) # 527
length(unique(resOrdered_t2_housing_df_m$`Gene Names (primary)`)) # 351
length(unique(resOrdered_t2_housing_df_m$Entry)) # 1236

#
resOrdered_t2_housing_df_mf <- resOrdered_t2_housing_df_m %>%
  filter(!is.na(Entry)) %>% # 1235
  group_by(ENSEMBL) %>%
  filter(if(any(Reviewed == "reviewed", na.rm = TRUE)) { 
    Reviewed == "reviewed" # keep reviewed if there
  } else{TRUE}) %>% # 496
  slice_max(Annotation, n = 1) %>% # 407
  slice_max(existence_ranking, n = 1) %>% # 406 
  slice_max(occupied, n = 1, with_ties = FALSE) %>% # 376 | 349 
  mutate(Regulation = ifelse(log2FoldChange > 0, "Up", "Down"), .after = lfcSE) # add up or down column


# checks after
length(unique(resOrdered_t2_housing_df_mf$ENSEMBL)) # 346
length(unique(resOrdered_t2_housing_df_mf$`Gene Names`)) # 346
length(unique(resOrdered_t2_housing_df_mf$`Gene Names (primary)`)) # 346
length(unique(resOrdered_t2_housing_df_mf$Entry)) # 346
colnames(resOrdered_t2_housing_df_mf)
```

```{r}
# write_excel_csv(resOrdered_t2_housing_df_mf,
#                 "output/2025_12_23_Wassouf_2018_test_2_EEvsSH_all16_adjp159_mapped_346.csv")
```

```{r}
volcano_adjpv_2 <- ggplot(resOrdered_t2_housing_df_mf, 
                     aes(x =log2FoldChange, y = -log2(padj), label = `Gene Names (primary)`)) +
 geom_point(aes(color = case_when(
   padj < 0.05 & log2FoldChange > 0 ~ "Upregulated (adj. p<0.05)",
   padj < 0.05 & log2FoldChange < 0 ~ "Downregulated (adj. p<0.05)",
   padj < 0.1 & log2FoldChange > 0 ~ "Upregulated (adj. p<0.1)",
   padj < 0.1 & log2FoldChange < 0 ~ "Downregulated (adj. p<0.1)",
   TRUE ~ "Not Significant")),
   alpha = 1, size = 2.5, shape = 1) +
  # facet_wrap(~ Label) +
  # xlim(-1,1) +
  scale_color_manual(values = c("firebrick3", 
                                "royalblue4", 
                                "lightpink1", 
                                "lightblue3" , 
                                 "grey80"),
                     breaks = c("Upregulated (adj. p<0.05)", 
                                "Downregulated (adj. p<0.05)", 
                                "Upregulated (adj. p<0.1)", 
                                "Downregulated (adj. p<0.1)", 
                                "Not Significant"),
                     labels = c("Upregulated Proteins (up to adj. p<0.05)", 
                                "Downregulated Proteins (up to adj. p<0.05)", 
                                "Upregulated Proteins (up to adj. p<0.1)", 
                                "Downregulated Proteins (up to adj. p<0.1)" , 
                                "Not Significant")) +
  
  geom_text_repel() +
  labs(title = "Wassouf 2018 RNAseq DESeq2 test 2 EE vs SE (all 16 mice) - Significant Proteins (adjp<0.159) (2025_12_23)",
       subtitle = "Presented like MSstats and my data log2FC & -log2(adj.pvalue)",
       x = "log2FC",
       y = "-Log2 (adjusted p-value)",
       color = "Protein Regulation") +
  # geom_hline(yintercept = -log2(0.05), linetype = "dashed", linewidth = 0.2,  color = "grey30") +
  # annotate("text", x = .86, y = -log2(0.05), label = "Adj. p-value cutoff (0.05) \n",
  #          fontface = 'italic', size = 4, color = "grey10") +
  # geom_hline(yintercept = -log2(0.1), linetype = "dashed", linewidth = 0.2,  color = "grey50") +
  # annotate("text", x = .86, y = -log2(0.1), label = "Adj. p-value cutoff (0.1) \n",
  #          fontface = 'italic', size = 4, color = "grey30") +
  theme_bw() + 
  theme(legend.position = "bottom")



volcano_adjpv_2
```

```{r}
# lollipop 2
sig_plot_2 <- resOrdered_t2_housing_df_mf %>%
  # filter(padj < .15) %>% # overplotted...
  #filter(padj <= .159 & log2FoldChange > 0.20 | log2FoldChange < -0.20) %>% #
  # filter(padj <= .10 & log2FoldChange > 0.20 | log2FoldChange < -0.20) %>% #
  filter(padj <= .05) %>% #
  ggplot(aes(x = reorder(`Gene Names`, log2FoldChange), # orders by log fold
                         y = log2FoldChange,
             color = Regulation)) +
  geom_segment(aes(x = `Gene Names`,
                   yend = log2FoldChange, y = 0)) +
  geom_point(aes(alpha = padj),
             shape = 19, size = 4) +
  scale_alpha(range = c(1,.1), guide = 'none') + # rev alpha so darker = more significant
  # geom_text(aes(label = gene_name), color = "grey10", size = 2.5,
  #           position = position_stack()) +
  geom_hline(yintercept = 0, color = 'grey20', linewidth = 0.1) +
  scale_color_manual(breaks = c("Up", "Down"),
                     values = c("#dc322f", "#268bd2"),
                     guide = 'none') +
  labs(
    title = 'Wassouf 2018 RNAseq DESeq2 test 2 EE vs SE (all 16 mice) - Lollipop for padj <=.05 (111 DEG) (2025_12_23)',
    subtitle = 'Ordered by Log2FC & then alphabetically by gene name, alpha is significance, blue is down and red is up.',
       y = 'EC Log2FC',
       x = 'Primary Gene Name') +
  #coord_flip() + #fliptheplot
  #facet_wrap(~fraction, space = 'free_x', scale = 'free_x') +
  #theme_void() +
  theme_bw() +
  theme(
    #axis.title = element_blank(),
    #axis.text.x = element_blank()
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, # 90, hj1 t/R, vj.5 cent
                               size = 9, face = 'italic'),
    axis.text.y = element_text(hjust = 1, color = "grey20", size = 10), # rota 90 & hjust=1 (top/right)
        )

sig_plot_2
```

Trying pheatmaps!

<https://davetang.github.io/muse/pheatmap.html>

```{r}
# davetang example
# example_file <- "https://davetang.org/file/TagSeqExample.tab"
# data <- read.delim(example_file, header = TRUE, row.names = "gene")
# data_subset <- as.matrix(data[rowSums(data)>50000,])
# dim(data_subset)

cts_subset <- cts[rowSums(cts)>50000,]
dim(cts_subset) # 144 X 8 

pheatmap(cts_subset)
```

```{r}
# EE vs SH (WT only) 
# Get ENSEMBLS (that are adjp <= 0.1)
t1_list <- resOrdered_df_p.01_mf$ENSEMBL[resOrdered_df_p.01_mf$padj <= 0.1] #59/109

cts_subset_padj_1 <- cts[t1_list,] # gets the 59 
dim(cts_subset_padj_1) # 59 X 8 

pheatmap(cts_subset_padj_1)
```

```{r}
# Get ENSEMBLS (that are adjp <= 0.05) | keeps 
t2_list <- resOrdered_t2_housing_df_mf$ENSEMBL[resOrdered_t2_housing_df_mf$padj <=0.05] 

cts_t2_padj_05 <- cts_t2[t2_list,]
dim(cts_t2_padj_05) # 111 x 16

pheatmap(cts_t2_padj_05)

# looks about the same for most? (since not processed?)
```

Maybe test heatmaps shown in workflow- <https://master.bioconductor.org/packages/release/workflows/vignettes/rnaseqGene/inst/doc/rnaseqGene.html> OR <https://bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#data-quality-assessment-by-sample-clustering-and-visualization>
