---
title: "BioIn_2026_01_v3.4.2_ClusterProfiler_EExAge_vis_v2"
author: "Stephanie Huang"
date: "2026-01-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# [BioIn_2026_08_v3.4.2_ClusterProfiler_EExAge_vis_v4_Pathways - For KEGG & REACTOME]{.underline}

**2026_08_06** - this branches off from the [BioIn_2026_08_v3.4.2_ClusterProfiler_EExAge_vis_v2.Rmd]{.underline}

```{r}
#colors 
old_lolliplots <- c("#4a6fe3","#9da8e2","#e2e2e2","#e495a5","#d33f6a") #Blue-Red 2 
old_pathview <- c("#3E76E8", "grey85", "#DA3558") 
blended_new <- c("#4473E6", "#DEDEDE", "#D73A61")
blended_new_2 <- c("#6E8FE5", "#DEDEDE", "#DF657F")
blended_new_3 <- c("#5981E6", "#DEDEDE", "#DB5070") # used for new pathview tint bars!  
```

# Setup

### Load Packages

```{r}
set.seed(123) # for reproducibility

# Load Packages 
library(tidyverse) # for dplyr, ggplot, stringr, readr, purrr, tibble, tidyr etc 
library(readxl)
library(clusterProfiler) #GSEA includes gsea, enrichplot etc.
# library(org.Rn.eg.db) # Rat genome anno pack- incl. AnnotationDbi
library(msigdbr) # MSigDB gene sets
library(enrichplot) # for dotplot()
library(ggridges) # for ridgeplot()
# library(pathview) # pathview() - Kegg pathway visualisation

library(paletteer) # colors
library(colorspace) # colors
library(cowplot) # plot_grid

library(readxl)
library(writexl) # for excel tab export
library(ggvenn) # vennys (old ones)
library(scales) #label_wrap text break in lolliplots
library(ggh4x)
library(ggrepel) # geom_text_repel
library(ggrastr) # https://tomsing1.github.io/blog/posts/volcano/
# library(magick) # for overlay tint on pathviews
library(rbioapi) # to get reactome parent categories
```

### Load Data

#### GSEA Results

```{r}
# #READ IN DATA
# KEGG
load("output/gsea_results/2026_01_12_KEGG_gsea_EExAge_mssv3.2.2t1_cpv3.4.2t3.rda")

# # MSIGDB C2 Reactome 
load("output/gsea_results/2026_02_09_gsea_msig_c2_reactome_EExAge_mss3.2.2_t1_cp3.4.2.rda")
```

#### MSstatsTMT Mapped Protein Results

```{r}
# mapped_comp_v3.2.2_EExAge_t1
load("input/protein_lists/2025_12_31_1pspg_v3.2.2_t1_EExAge_mapped_comp.rda")

df <- mapped_comp_v3.2.2_EExAge_t1
```

#### KEGG category annotation

```{r}
# USED PREV TO RETRIEVE PATHWAYS & THEN EXPORTED & MANUALLY ANNOTATED & REIMPORTED!
# kk_gson <- gson_KEGG(species = "rno") # via clusterprofiler
# kk_gson_df <- kk_gson@gsid2name
# write_csv(kk_gson_df, "2026_03_04_KEGG_rno_pathways_kk_gson_df.csv")

# see 2026_03_04_KEGG_rno_pathways_kk_gson_df_w_anno
KEGG_cat_anno <- read_csv(
  "input/db/2026_03_04_KEGG_rno_pathways_kk_gson_df_w_anno.csv")
```

Pull in MSigDB Reactome Annotations

<https://igordot.github.io/msigdbr/articles/msigdbr-intro.html>

```{r}
# msigdbr_collections() #available collections
REACTOME_genesets <- msigdbr(species = "rat", collection = "C2", subcollection = "CP:REACTOME") %>%
  select(gs_name:gs_url) %>%
  distinct(gs_name, .keep_all = TRUE)
```

```{r}
# HALLMARKS_genesets <- msigdbr(species = "rat", collection = "H")
```

```{r}
# KEGG_genesets <- msigdbr(species = "rat", collection = "C2", subcollection = "CP:KEGG_LEGACY")
# 
# GO_genesets <- msigdbr(species = "rat", collection = "C5")
```

### Read in Reactome Pathways (downloaded)

```{r}
REACT_relations_raw <- read_tsv("input/db/2026_08_ReactomePathwaysRelation.txt",
                                col_names = c("parent",
                                              "child")) 

REACT_paths_raw <- read_tsv("input/db/2026_08_ReactomePathways.txt",
                                col_names = c("gs_id",
                                              "description",
                                              "species")) 
```

```{r}
REACT_join <- left_join(REACT_relations_raw, REACT_paths_raw,
                            by = join_by(parent == gs_id)) %>%
  relocate(description, .after = parent) %>%
  rename(parent_description = description) %>%
  left_join(select(REACT_paths_raw,
                   gs_id, description),
                            by = join_by(child == gs_id)) %>%
  relocate(description, .after = child) %>%
  rename(child_description = description)

unique(REACT_join$species)

REACT_join_peeps <- REACT_join %>%
  filter(species == "Homo sapiens")
```

<https://rbioapi.moosa-r.com/reference/rba_reactome_pathways_top.html>

```{r}
top_level_REACT <- rba_reactome_pathways_top(species =  "Homo sapiens")
```

------------------------------------------------------------------------

# Wrangling

### Protein Comparisons with ENTREZ & GENENAME

```{r}
# see BioIn_2025_12_v3.4.2_ClusterProfiler_EExAge_vis_testing_v1.Rmd for more info

protein_comp <- df %>%
  filter(is.na(issue)) %>% # keeps rows with issue == na (so comps that are ok)
  filter(!str_detect(log2FC, "Inf")) %>% # remove infinity/missing log2FC
  mutate(Label_trm = paste0(gsub("^[^_]*_", "", Label)), .after = Label)

# check 
head(protein_comp) 
colnames(protein_comp)


# ENTREZ
protein_comp_uni_gene <- unique(protein_comp$gene) # 8014

bitr_gn_2_entrez <- bitr(protein_comp_uni_gene,
                         fromType = "SYMBOL",
                         toType = "ENTREZID",
                         OrgDb = "org.Rn.eg.db") #3.39% failed to map... |3.72% v2




# SEE MY LEVELS BELOW
my_comp_levels = c("H_EE_vs_SH",
              "HxA_EE_YNG_vs_SH_YNG",
              "HxA_EE_OLD_vs_SH_OLD",
              "A_YNG_vs_OLD",
              "HxA_EE_YNG_vs_EE_OLD",
              "HxA_SH_YNG_vs_SH_OLD",

              "HxA_EE_OLD_vs_SH_YNG", "HxA_EE_YNG_vs_SH_OLD"
)


# MAP ENTREZ TO UNIPROT PROT COMPS
protein_comp_entrez <- left_join(protein_comp, bitr_gn_2_entrez,
                                 by = join_by(gene == SYMBOL)) %>%
  filter(!is.na(ENTREZID))  # remove non mapped - 64384 to 61952 


protein_comp_entrez$Label <- factor(protein_comp_entrez$Label,
                                  levels = my_comp_levels)
```

### Map nested .rda to master dfs & fix names

```{r}
##_KEGG_______
## Map KEGG to a MASTER DF
KEGG_tibs_df <- map_dfr(KEGG_gsea_EExAge_mssv3.2.2t1_cpv3.4.2t3$kegg_results_tib,
                        as.data.frame,
                        .id = "Label")  %>%
  mutate(Label_trm = paste0(gsub("^[^_]*_", "", Label)), .after = Label) %>%
  mutate(Description_2 = str_replace_all(Description, " and ", " & "), # & instead
         .after = Description) %>%
  left_join(KEGG_cat_anno, # join to KEGG subcategory data
            by = join_by(ID == gsid,
                         Description == name)) %>%
  mutate(subcategory_2 = str_replace_all(subcategory, " and ", " & ")) %>%
  mutate(obs_genes = str_count(core_enrichment, "/") + 1)

#Fix some KEGG description names  
#df$col[df$col == "old_val"] <- "new_val"  | SHORTEN NAME! 
KEGG_tibs_df$Description_2[KEGG_tibs_df$Description_2 == "Pathways of neurodegeneration - multiple diseases"] <- "Pathways of neurodegeneration - multiple"

KEGG_tibs_df$Description_2[KEGG_tibs_df$Description_2 == "Chemical carcinogenesis - reactive oxygen species"] <- "Chemical carcinogenesis - ROS"
  


##_REACTOME_______________
## MAP MSIG c2 REACTOME
REACT_tibs_df <- map_dfr(
  gsea_msig_c2_reactome_EExAge_mss3.2.2_t1_cp3.4.2$gsea_results_tib,
  as.data.frame,
  .id = "Label") %>%
  #shorted comparison label
  mutate(Label_trm = paste0(gsub("^[^_]*_", "", Label)), .after = Label) %>%
  #Shorten & mod Reactome Description label
  mutate(Description_2 = str_replace_all(Description, "REACTOME_", ""),
         .after = Description)  %>%
  mutate(Description_2 = str_replace_all(Description_2, "_AND_", "_&_"), # &
         .after = Description) %>%
  mutate(Description_2 = str_replace_all(Description_2, "_", " "), # _ to " "
         .after = Description) %>%
  left_join(select(REACTOME_genesets, # df
                   gs_name, gs_description, gs_exact_source, gs_url), # cols
            by = join_by(ID == gs_name)) %>% # 985
  left_join(REACT_join_peeps, by = join_by(gs_exact_source == child)) # 1001

  
  
  
```

### Filter df for 6 comparisons (+ arrange levels & shorten some names)

```{r}
##_LEVELS________
my_levels_1 = c("H_EE_vs_SH", # OLD LONG LABELS
              "HxA_EE_YNG_vs_SH_YNG",
              "HxA_EE_OLD_vs_SH_OLD",
              "A_YNG_vs_OLD",
              "HxA_EE_YNG_vs_EE_OLD",
              "HxA_SH_YNG_vs_SH_OLD"
              )

my_levels_2 = c("EE_vs_SH", # NEW SHORT LABELS IN Label_trm
              "EE_YNG_vs_SH_YNG",
              "EE_OLD_vs_SH_OLD",
              "YNG_vs_OLD",
              "EE_YNG_vs_EE_OLD",
              "SH_YNG_vs_SH_OLD"
              )


##_KEGG_6_
KEGG_tibs_df_6 <- KEGG_tibs_df %>%
  filter(!(Label %in% c("HxA_EE_OLD_vs_SH_YNG", "HxA_EE_YNG_vs_SH_OLD"))) 

KEGG_tibs_df_6$Label <- factor(KEGG_tibs_df_6$Label,
                                  levels = my_levels_1)
KEGG_tibs_df_6$Label_trm <- factor(KEGG_tibs_df_6$Label_trm,
                                  levels = my_levels_2)


##_REACTOME_6_
REACT_tibs_df_6 <- REACT_tibs_df %>%
   filter(!(Label %in% c("HxA_EE_OLD_vs_SH_YNG", "HxA_EE_YNG_vs_SH_OLD"))) 

REACT_tibs_df_6$Label <- factor(REACT_tibs_df_6$Label,
                                  levels = my_levels_1)
REACT_tibs_df_6$Label_trm <- factor(REACT_tibs_df_6$Label_trm,
                                  levels = my_levels_2)
```

```{r}
unique_REACT_tibs_df_6 <- REACT_tibs_df_6 %>%
  select(gs_exact_source, ID, gs_description) %>%
  unique() #311

REACT_paths_observed <- left_join(REACT_join_peeps, 
                                  unique_REACT_tibs_df_6,
                                  by = join_by(child == gs_exact_source)) %>%
  filter(!is.na(ID) & !is.na(gs_description)) #%>% # 300


f1 <- setdiff(unique_REACT_tibs_df_6$gs_exact_source,
              REACT_paths_observed$child) 
m2 <- intersect(unique_REACT_tibs_df_6$gs_exact_source,
                REACT_paths_observed$parent)

intersect(f1,m2) # 12 missing that are in the parent id
```

```{r}

```

------------------------------------------------------------------------

## Functions

### Lolliplot

- <https://stackoverflow.com/questions/39801041/ggplot2-have-common-facet-bar-in-outer-facet-panel-in-3-way-plot>
- <https://stackoverflow.com/questions/19440069/ggplot2-facet-wrap-strip-color-based-on-variable-in-data-set>
- <https://stackoverflow.com/questions/79755989/changing-facet-strip-colour-by-grouping-variable-when-using-ggh4xfacet-manual>

```{r}
lolliplot <- function(data,
                      x_1 = Description, 
                      x_2 = NES, 
                      y = NES, 
                      fill_1 = NES,
                      # cat_color = NULL, # ONLY NEEDED FOR KEGG CATEGORIES 
                      alpha = p.adjust,
                      facet_formula = ~Label_trm,
                      size = setSize,
                      subtitle = NULL,
                      legend_pos = "none") {
  
  # NEED TO SHUFFLE HERE & THERE BUT WORKS
    strip <- strip_themed( # green for 1st 3 (housing) & Lavender for last 3 (age)
    background_x = elem_list_rect(fill = c(
      "#AFCC9F", "#D4E4CB", "#D4E4CB", 
      "#BFBBD5", "#DDDCE9", "#DDDCE9")))
  
  
  lol <- ggplot(data,
                aes(x = reorder({{x_1}},{{x_2}}),
                    y = {{y}}
                    )) +
    
    geom_segment(aes(x = {{x_1}},
                     yend = {{y}}, y = 0,
                     alpha = {{alpha}}), color = "grey10", linewidth = 0.3) +
    
    geom_point(aes(alpha = {{alpha}},
                  size = {{size}},
                  fill = {{fill_1}}),
                  color ="grey10", shape = 21) +
    
    # scale_alpha(range = c(.95,.3)) + # rev alpha so darker = more sig | all
    scale_alpha(range = c(.99,.49)) + #  GO TRIM VERSION so matches others
    geom_hline(yintercept = 0, color = 'grey20', linewidth = 0.1) +
    scale_color_manual(breaks = c("Up", "Down"),
                     # values = c("#dc322f", "#268bd2"),
                     guide = 'none') + 
    scale_size_continuous(range = c(3,8)) + # map size oflolly (set size)
    # scale_size_continuous(range = c(2,8),
    #      guide = guide_legend(nrow=1)) + # map size oflolly (set size) | GO BP
    

    # scale_fill_continuous_diverging("Blue-Red 2") + # colorspace
    scale_fill_gradient2(low="#4a6fe3", mid="#e2e2e2", high="#d33f6a",
                         midpoint = 0,
                        limits = c(-3.5, 3.5)) +
  
 
    
    # scale_x_discrete(labels = scales::label_wrap(width = 30)) + # break text | KEGG ver
     scale_x_discrete(labels = scales::label_wrap(width = 60)) + # break text |

    ylim(-4,4)+
  
    labs(subtitle = subtitle,
       y = 'NES' ) +
    
    coord_flip() + #fliptheplot
    

   
    # facet_wrap(facet_formula,
    #          space = 'free_x') +
     
    # facet_grid(facet_formula, # KEGG subcategory anno! half working ....
    #              space = "free", scales = "free") +
    
    facet_grid2(facet_formula,
                strip = strip, # custom strip colors
                space = "free", scales = "free", 
                labeller = label_wrap_gen(width = 20), # length of label
                drop = FALSE) + # dropfalse keeps any empty facets

    
    
    theme_bw() +
    theme(
      # strip.background.x = element_rect(fill = "#cbdec1"),
      # strip.background.x = element_rect(fill = my_level_colors),
      strip.text.y = element_text(angle = 0, size = 7.5),
      strip.text.x = element_text(size = 6),
      panel.spacing.y = unit(0, "lines"), # or "cm", "in", etc.
      
    axis.text.x = element_text(angle = 90, hjust = 1, color = "grey10",
                               vjust = 0, # 90, hj1 t/R, vj.5 cent
                               size = 8.5, face = 'italic'), 
    axis.text.y = element_text(hjust = 1, color = "grey10",
                               size = 8, face = 'italic',
                               lineheight = .7), # <1 smaller distances breaks
    axis.title.y = element_blank(),
    axis.title.x = element_text(color = "grey10", size = 8.5),
    axis.ticks = element_line(color = "grey10", linewidth = 0.25),

    
    # legend.margin = margin(-10,-1,-5,-1), # (top, right, bot, left) | bottom ver
    legend.title = element_text(face = 'italic', size =8,
                                hjust = 1),
    legend.key.size = grid::unit(0.3, "cm"), # posit answer to same height
    legend.text = element_text(margin = margin(l = 0)), # marg text & key
    legend.box.just = "right", # right justify legend box
    
    # # #FOR RIGHT LEGEND
    legend.margin = margin(0,0,0,-5), # (top, right, bot, left) | right ver
    legend.justification = "top", # for right ver
    
    legend.position = legend_pos,
    
    strip.background = element_rect(linewidth = 0.25, color = "grey10"),
    panel.border = element_rect(linewidth = 0.25, color = "grey10"),
    panel.spacing.x = unit(.1, "cm"),
    panel.grid.major = element_line(color = "grey90", linewidth = 0.1),
    panel.grid.minor = element_line(color = "grey90", linewidth = 0.1)
    )
  
  
}
```

------------------------------------------------------------------------

# Lolliplots

## KEGG lolliplots (6)

### Full KEGG PLOT with no filtering (appendix plot)

```{r}
# KEGG ALL 6 BASIC
KEGG_lolli_full_6 <- lolliplot(KEGG_tibs_df_6,
                         facet_formula = reorder(subcategory_2, -NES)~Label_trm,
                         x_1 = Description_2,
                         subtitle = "KEGG GSEA", legend_pos = "bottom")

KEGG_lolli_full_6
```

### KEGGs to remove for main thesis lolliplot

```{r}
# rm KEGG human diseases
human_diseases <- c("Non-alcoholic fatty liver disease",
                    "Coronavirus disease - COVID-19",
                    "Chemical carcinogenesis - reactive oxygen species",
                    "Salmonella infection",
                    "Diabetic cardiomyopathy",
                    "Chemical carcinogenesis - receptor activation",
                    "Acute myeloid leukemia",
                    "Hepatocellular carcinoma",
                    "Prostate cancer",
                    "Platinum drug resistance",
                    "Viral life cycle - HIV-1",
                    "Systemic lupus erythematosus",
                    "Staphylococcus aureus infection"
                    )

# organismal systems
organs <- c("Pancreatic secretion", # digestive
            "Cardiac muscle contraction", # circulatory
            
            "Endocrine and other factor-regulated calcium reabsorption", #excretory
            "Proximal tubule bicarbonate reclamation", #excretory
            
            "Salivary secretion", # digestive
            "Bile secretion", # digestive
            "Gastric acid secretion", # digestive
            "Protein digestion and absorption",  # digestive
            "Carbohydrate digestion and absorption",  # digestive
            "Mineral absorption",  # digestive
            
            "Thyroid hormone synthesis", # endocrine
            "Adrenergic signaling in cardiomyocytes",  # circulatory
            
            "Aldosterone synthesis and secretion", # Endocrine System
            "Insulin secretion" # Endocrine system
            
            )

otherKEGGs <- c("Thermogenesis", # Envvironmental Adapation 
                "Folate biosynthesis" # Metabolism of cofactors & vitamins
                )
```

### Trim Combo & subcategory KEGG Lolliplot (for main PhD)

```{r}
# COMBINED KEGG!!! - 2026_08_03
KEGG_tibs_6_trim_PhD <- KEGG_tibs_df_6_anno %>%
  filter(!(Description %in% human_diseases)) %>%
  filter(!(Description %in% organs)) %>%
  filter(!(Description %in% otherKEGGs)) %>%
  lolliplot(
    facet_formula = reorder(subcategory_2, -NES)~Label_trm,
    # facet_formula = category+subcategory_2~Label,
    x_1 = Description_2,
    subtitle = "KEGG GSEA - Housing & Age - (trimmed)", legend_pos = "bottom")

KEGG_tibs_6_trim_PhD
```

### 

## Msigdb C2 Reactome lolliplots

```{r}
# REACTOME GSEA - 770 adjp<0.05(all 6) |  p.adjust <0.02 (550 obs) |p.01 405
REACT_lol_6 <- REACT_tibs_df_6 %>%
  filter(p.adjust <=0.03) %>%
  
  lolliplot(subtitle = "REACTOME (via Msigdbr C2)",
            x_1 = Description_2, legend_pos = "bottom")

REACT_lol_6


# ggsave("2026_02_25_REACTOME_via_Msigdbr_C2_EExAge_6.png", dpi = 200)
```

------------------------------------------------------------------------

------------------------------------------------------------------------

## Pathview for loops (14 02 2026)

Done a nooby lazy way maybe clean up later.

- <https://www.bioconductor.org/packages/devel/bioc/vignettes/pathview/inst/doc/pathview.pdf>
- rat kegg pathways - <https://www.kegg.jp/kegg-bin/show_organism?menu_type=pathway_maps&org=rno>
- <https://igordot.github.io/msigdbr/articles/msigdbr-intro.html>

#### Aes test

```{r}
## Ribosome - rno03010 
pathview(gene.data=KEGG_gsea_EExAge_mssv3.2.2t1_cpv3.4.2t3$geneList_vec$H_EE_vs_SH,
                pathway.id="rno03010",
                species = 'rno',
         out.suffix = "2026_02_14__H_EE_vs_SH_testing",
         kegg.native = T,
         same.layer = T, #slower & on top (out of bounds?)
         res = 600, #400 dpi (def 300 dpi for png) # not sure if changes anything..
         low = list(gene = "#3E76E8", cpd = "#A054A3"), # purple unused 
         mid = list(gene= "grey85", cpd = "grey85"), 
         high = list(gene = "#DA3558", cpd ="#F3AD6AFF"), # org unused
         new.signature = FALSE)

# my_colors <- c("#E22A47", "#DA3558", "#D33F6A", "#317DEC", "#3E76E8", "#4A6FE3" )
```

### Set up master gene vec lists

```{r}
# # get msigdbr kegg pathways into data 
# KEGG_rats_msig <- msigdbr(species = "Rattus norvegicus",
#                           collection = "C2", subcollection = "CP:KEGG_MEDICUS")
# nvm maybe return to later 
```

```{r}
length(unique(KEGG_tibs_df_6$ID)) # 67 

KEGG_multi <- KEGG_tibs_df_6 %>%
  group_by(Description) %>%
  count() %>%
  # filter(n > 1)
  filter(n == 1) # ones 

KEGG_multi_list <- KEGG_multi$Description

# note - & - different?
KEGG_multi_df <- KEGG_tibs_df_6 %>% 
  filter(Description %in% KEGG_multi_list) %>%
  mutate(trm_des = str_replace_all(Description, "[- -()]", "")) %>%  # rm spaces, - & () 
  mutate(trm_desc = str_trunc(trm_des, 15, side = "right", ellipsis = "")) %>%
  filter(!(ID %in% c("rno04723", # rm reterograde endocannbinoid - not found pathbiew err
                "rno01250" # Biosynthesis of nucleotide sugars"
                ))) 
```

```{r}
# Map KEGG to a MASTER genelist DF 
geneLists_df <- map_dfr(KEGG_gsea_EExAge_mssv3.2.2t1_cpv3.4.2t3$geneList_vec,
                        ~enframe(.x),
                        .id = "Label")
```

```{r}
# iterate over... 
labels <- unique(KEGG_tibs_df_6$Label)[1]


kegg_pathways <- unique(KEGG_multi_df$ID)
```

#### Extra Pathways

```{r}
# rm errors - metabolic pathways, fatty acid metabolism (rno01212) - too long
kegg_pathways_2_df <- data.frame("ID" = c("rno00010", # 1
                                          "rno04130", # 
                                          "rno04514", # 
                                          "rno04518", #
                                          "rno04722",
                                          "rno04024",
                                          "rno04151",
                                          "rno04010",
                                          "rno04014",
                                          "rno04020",
                                          "rno04022",
                                          "rno04081",
                                          "rno04150",
                                          "rno04080",
                                          "rno04512",
                                          "rno04060",
                                          "rno04517",
                                          "rno04210",
                                          "rno04110",
                                          "rno04810",
                                          "rno05022",
                                          "rno04211",
                                          "rno04724",
                                          "rno04726",
                                          "rno04720",
                                          "rno04730",
                                          "rno04725",
                                          "rno04728",
                                          "rno04120",
                                          "rno03008",
                                          "rno00970",
                                          "rno04141",
                                          "rno04144",
                                          "rno04145",
                                          "rno04146",
                                          "rno04142"
                                          ),
                                 
                                 "trm_desc" = c("glycolysis_gluc",
                                              "SNARE_int_vt",
                                              "CAM_interac",
                                              "Intergrin_sig",
                                              "Neurotrophic_sig",
                                              "cAMP_sig_path",
                                              "PI3K_Akt_sig_path",
                                              "MAPK_sig_path",
                                              "Ras_sig_path",
                                              "Calcium_sig_path",
                                              "cGMP_PKG_sig_path",
                                              "Hormone_sig",
                                              "mTOR_sig_path",
                                              "Neuract_lig_rep_i",
                                              "ECM_receptor_int",
                                              "Cytokine_rec_int",
                                              "IgSF_CAM_sig",
                                              "Apoptosis",
                                              "Cell_cycle",
                                              "Reg_actin_cytosk",
                                              "Paths_neurodegen",
                                              "longeivity_reg",
                                              "glutamaterg_synp",
                                              "serotonergic_syn",
                                              "LTP",
                                              "LTD",
                                              "cholinergic_syn",
                                              "dopaminergic_syn",
                                              "ubiquit_med_prtly",
                                              "ribosome_biogen",
                                              "Aminoacyl_tRNA_bi",
                                              "prot_proces_in_ER",
                                              "endocytosis",
                                              "phagosome",
                                              "peroxisome",
                                              "lysosome"
                                              ))

kegg_pathways_2 <- unique(kegg_pathways_2_df$ID)
```

#### for loop for KEGG pathviews

```{r}
# # PATHVIEW FOR LOOP
# for (lb in labels) {
# 
#   wdf <- geneLists_df %>%
#     filter(Label == lb) %>%
#     dplyr::select(name, value)
# 
#   wgL <- tibble::deframe(wdf)
# 
#   #for (kp in kegg_pathways) {  # 1 kegg pathways from gsea
#   for (kp in kegg_pathways_2) { # 2 extra kegg paths of interest
# 
#     # c_desc <- KEGG_multi_df %>%  # 1 kegg pathways from gsea
#      c_desc <- kegg_pathways_2_df %>%  # 2 extra kegg paths of interest
#       filter(ID == kp) %>%
#       dplyr::select(trm_desc) %>%
#       slice_head(n = 1)
# 
# 
#     pathview(gene.data = wgL,
#                 pathway.id = kp,
#                 species = 'rno',
#          out.suffix = paste0(lb, "_", c_desc[1]),
#          kegg.native = T,
#          same.layer = T, #slower & on top (out of bounds?)
#          res = 600, #400 dpi (def 300 dpi for png) # nvm doesn't change drawings idk
#          low = list(gene = "#317DEC", cpd = "#A054A3"), # purple unused
#          mid = list(gene= "grey85", cpd = "grey85"),
#          high = list(gene = "#E22A47", cpd ="#F3AD6AFF"), # org unused
#          new.signature = FALSE
#          )
#   }
# }

#noob way manually set working directory as folder you want to output

```

### Kegg Pathviews with tint bar overlay (magick) (2026_08_03)

- <https://cran.r-project.org/web/packages/magick/vignettes/intro.html#Build_from_source>

```{r}
# #PATHVIEW WITH TINT_BAR overlay using magick package (2026_08)

# label_colors <- c(
#   "H_EE_vs_SH" = "#AFCC9F",
#   "HxA_EE_YNG_vs_SH_YNG" = "#D4E4CB",
#   "HxA_EE_OLD_vs_SH_OLD" = "#D4E4CB",
#   "A_YNG_vs_OLD" = "#BFBBD5",
#   "HxA_EE_YNG_vs_EE_OLD" = "#DDDCE9",
#   "HxA_SH_YNG_vs_SH_OLD" = "#DDDCE9"
# )

# for (lb in labels) {
#   wdf <- geneLists_df %>% filter(Label == lb) %>% dplyr::select(name, value)
#   wgL <- tibble::deframe(wdf)
# 
#   my_lab <- geneLists_df$Label[[1]] # my label
# 
#   # for (kp in kegg_pathways) { # v1 observed
#   # c_desc <- KEGG_multi_df %>%
# 
#     for (kp in kegg_pathways_2) { # 2 extra kegg paths of interest
#       c_desc <- kegg_pathways_2_df %>%  # 2 extra kegg paths of interest
# 
#           filter(ID == kp) %>%
#           dplyr::select(trm_desc) %>% slice_head(n = 1)
#         suffix <- paste0(lb, "_", c_desc[[1]])
# 
#     pathview(gene.data = wgL, pathway.id = kp, species = 'rno',
#              out.suffix = suffix, kegg.native = T, same.layer = T, res = 600,
#              low = list(gene = "#5981E6", cpd = "#A054A3"),
#              mid = list(gene = "#DEDEDE", cpd = "#DEDEDE"),
#              high = list(gene = "#DB5070", cpd = "#F3AD6AFF"),
#              new.signature = FALSE)
# 
#     out_png <- paste0(kp, ".", suffix, ".png")  # confirm exact naming once
# 
#     if (file.exists(out_png)) { # uses magick package + grDevices
#       pic <- image_read(out_png) # annotated pathview pic
#       dets <- image_info(pic) # get details of pathview pic
#       overlay <- image_blank(dets$width/43, dets$height, # tinted overlay
#                               color = adjustcolor(label_colors[[lb]], alpha.f = 0.75))
#       tinted <- image_composite(pic, overlay, operator = "over") # composite
#       image_write(tinted, paste0(tools::file_path_sans_ext(out_png), # export
#                                  "_tinted_bar_2026_08.png"))
# 
#     } else { # if none
#       warning("Expected file not found: ", out_png)
#     }
#   }
# }
```

------------------------------------------------------------------------

#### Running Scores

<https://yulab-smu.top/biomedical-knowledge-mining-book/enrichplot.html#running-score-and-preranked-list-of-gsea-result>

```{r}
#head(KEGG_gsea_EExAge_mssv3.2.2t1_cpv3.4.2t3$kegg_results$H_EE_vs_SH$Description, 20)

# 10 = Reterograde Endo cannm, 12 - synaptic vesicle, 20 neuroactive
# gseaplot2(KEGG_gsea_EExAge_mssv3.2.2t1_cpv3.4.2t3$kegg_results$H_EE_vs_SH,
#           geneSetID = c(10,12, 20))


p1 <- gseaplot2(KEGG_gsea_EExAge_mssv3.2.2t1_cpv3.4.2t3$kegg_results$H_EE_vs_SH,
                geneSetID = c(10,12, 20, 6:8, 1), subplots = 1:2,
                color = c("#D3F2A3FF", "#E17C05FF", "#586e75", "#F3AD6AFF",
                          "#F1B6DAFF", "#D43F23" ,"#CF597EFF"),
                title = "KEGG - EE vs SH (all)")

#head(KEGG_gsea_EExAge_mssv3.2.2t1_cpv3.4.2t3$kegg_results$HxA_EE_YNG_vs_SH_YNG$Description, 50)
p2 <- gseaplot2(KEGG_gsea_EExAge_mssv3.2.2t1_cpv3.4.2t3$kegg_results$HxA_EE_YNG_vs_SH_YNG,
          geneSetID = c(1, 9, 10, 22, 27, 31, 32, 46, 47), subplots = 1:2,
          color = c("#2896B5", "#589D4C", "#A054A3", "#E17C05FF",
                   "#F1B6DAFF", "#586e75", "#F3AD6AFF", "#D43F23" ,"#CF597EFF"),
          title = "KEGG - EE vs SH (young)")



p3 <- gseaplot2(KEGG_gsea_EExAge_mssv3.2.2t1_cpv3.4.2t3$kegg_results$HxA_EE_OLD_vs_SH_OLD,
          geneSetID = c(1,3), subplots = 1:2,
          color = c("#586e75", "#CF597EFF"), 
          title = "KEGG - EE vs SH (old)")


plot_list(p1, p2, p3, ncol = 1)
```

------------------------------------------------------------------------

------------------------------------------------------------------------

## MAPPED GSEA VOLCANO PLOTS

### KEGG

- <https://www.kegg.jp/kegg-bin/show_organism?menu_type=pathway_maps&org=rno>
- <https://en.wikipedia.org/wiki/Shades_of_orange>

KEGG_tibs_df_6_split

```{r}
str(KEGG_tibs_df_6) # lil check
#unique(KEGG_tibs_df_6$core_enrichment) 
  

KEGG_tibs_df_6_entrez_split <- KEGG_tibs_df_6 %>%  # 104 x 12s
  mutate(core_enrich_split = str_split(core_enrichment, "/")) %>% # split by /
  tidyr::unnest(core_enrich_split) %>% # unnest | to  3816 x 13
  left_join(bitr_gn_2_entrez, by = join_by(core_enrich_split == ENTREZID)) #%>%

  # group_by(Label, ID) %>%
  # summarise(unique_entrez = n_distinct(core_enrich_split),
  #           unique_genename = n_distinct(SYMBOL)) %>%
  # ungroup() # slight discrepancy between comparisons???

KEGG_tibs_df_6_entrez_split_display <- KEGG_tibs_df_6_entrez_split %>%
  group_by(Label, Description) %>%
  slice_head(n=10) 
```

```{r}
# iterate over... 
#labels <- unique(KEGG_tibs_df_6$Label)

# GET KEGG PATHVWAYS OF INTEREST 
kegg_POI  <- c(
  # nervous system, development, signaling molecules inter, substance depend - 12
  "rno04721", "rno04082", "rno04723", # syanp_ves, neuroa_lig, ret_endoca
  "rno04727", "rno04080", "rno05033", # GABAsyn, neuroacli-Recept, nicotine
  "rno04014", "rno04024", # Ras signalling, cAMP signalling
  "rno04022", "rno04020", # cGMP-PKG & Calcium isngalling
  "rno04360",  # axonguidance
  "rno05033", # nicotine addiction
  
  
  # Genetic Info processing (gene exp & protein syn) + mitophagy - 13:23 (11)
  "rno03040", "rno03082", "rno03050", # spliceo, ATPchromatre, proteasom
  "rno03010", "rno04141", "rno03013", # ribosome, prot proc in ER, nucleocyto transport
  "rno04120", # ubiquitin mediate proteolysis | folding sorting dgrade
  "rno03015", # mRNA surveillance | translation
  "rno03410", # base exicion repair |replication & repair
  "rno03083", # Polycomb repressive complex |chromosome
  "rno04137", #mitophagy |Transport & catabolism
    
  
  # Metabolism - 24:30 (7)
  "rno00190",  # oxidative phos | energy metab
  "rno00900", # terpenoid backbone synthesis | terpenoids 
  "rno00020", "rno00030", "rno00620", #TCA,  pentose phosph path, pyruvate metab |carb
  "rno01250", # Biosynthesis of nucleotide sugars | global 
  "rno00230", #  purine metabolism | nucleotide metab
  
  #Neurodegenerative disease - 31:36 (6)
  "rno05012","rno05016","rno05020", # Parkinsons, Huntingtons, Prion
  "rno05022","rno05010","rno05014", # multi, Alzheimers, ALS
  
  # OTHERS
  #Immune & Endocrine, envr adapt - 37:35 (8) 
  "rno04666", "rno04610", # Fc-gamma phagocytosis, Complement & coagulation
  "rno04910","rno04714", #insulin, # thermogenesis
  "rno04918", "rno04925", # thyroid hormone, aldosterone synthesis
  "rno05207", "rno05208" # chem & ROS carcinogenesis receptor activation | cancer 
  
  ) 

# kegg_POI <- unique(KEGG_tibs_df_6$ID)

# EMPTY PLOT LIST 
kegg_volcano_list <-  vector(mode = "list")

```

```{r}
for(kp in kegg_POI) {
  
# 1 - PREPARE THE DATA
  my_genes <- KEGG_tibs_df_6_entrez_split %>%
    filter(ID == kp) %>%
    dplyr::distinct(SYMBOL) %>%
    tibble::deframe()
  
  my_title <- KEGG_tibs_df_6_entrez_split %>%
    filter(ID == kp) %>%
    dplyr::distinct(Description) %>%
    tibble::deframe()
  
  my_prots <- protein_comp_entrez %>% # from MSstatsTMT protein comps w ENTREZ map
    filter(!(Label %in% c("HxA_EE_OLD_vs_SH_YNG", "HxA_EE_YNG_vs_SH_OLD"))) %>%
    filter(gene %in% my_genes) %>%
    dplyr::group_by(Label) %>%
    mutate(up = sum(log2FC > 0),
           down = sum(log2FC < 0)) 
  
  prot_sums <- my_prots %>% # this is needed to avoid overplotting
    dplyr::select(Label, up, down) %>%
    unique()
  
  mapped_totals <- my_prots %>%  # for mapped total
    ungroup() %>%
    summarise(n_distinct(gene)) %>%
    tibble::deframe()
  
  gs_size <- KEGG_tibs_df_6_entrez_split %>% # for geneset size
    filter(ID == kp) %>%
    dplyr::distinct(setSize) %>%
    tibble::deframe()
  
  
  path_sig <- KEGG_tibs_df_6 %>% # GET DETAILS FOR ANNOTATING SIG 
    filter(!(Label %in% c("HxA_EE_OLD_vs_SH_YNG", "HxA_EE_YNG_vs_SH_OLD"))) %>%
    filter(ID == kp) %>%
    dplyr::group_by(Label) %>% 
    mutate(sig_val = case_when(
      p.adjust <= 0.001 ~ "***",
      p.adjust <= 0.01 ~ "**",
      p.adjust <= 0.05 ~ "*",
      .default = "NA" )) # if bit there then NA (ns)
  
  
  # interesting... this step needed for some reason...
  path_sig_lab <-  path_sig %>% 
    dplyr::select(ID, Label, sig_val, NES) %>%
    unique() 
  
  my_lg_prot <- my_prots %>% #LABEL LEADING GENE???
    arrange(adj.pvalue) %>% # not quite - 6 most sig proteins
    slice_head(n = 6)
  
  
    strip <- strip_themed( # green for 1st 3 (housing) & Lavender for last 3 (age)
    background_x = elem_list_rect(fill = c(
      "#AFCC9F", "#D4E4CB", "#D4E4CB",
      "#BFBBD5", "#DDDCE9", "#DDDCE9")))

    
  # _____________________________________
  # 2 - MAKE THE VOLCANO PLOTS!!!
  vol_plot <- protein_comp_entrez %>%
    filter(!(Label %in% c("HxA_EE_OLD_vs_SH_YNG", "HxA_EE_YNG_vs_SH_OLD"))) %>%
    
    ggplot(aes(x = log2FC, y = -log10(adj.pvalue))) +
    
    
    # limits - rm 17 (but none are leading edge - shrunken for comparisons)
    scale_x_continuous(limits = c(-2,2),
                       expand = c(0, 0)) +
    scale_y_continuous(limits = c(0,6),
                       expand = c(0.04, 0)) +
    
    geom_hline(yintercept = -log10(0.05),
               linetype = "dashed", linewidth = 0.1,
               color = "grey50", alpha = 0.5) +
    geom_hline(yintercept = -log10(0.1),linetype = "dotted",
               linewidth = 0.1, color = "grey60", alpha = 0.5) +
    geom_vline(xintercept = c(0),linetype = "solid",
               linewidth = 0.1, color = "grey35", alpha = 0.5) +
    

    ggrastr::geom_point_rast(aes(fill = case_when( # conditional ass for color mapping
      adj.pvalue < 0.05 & log2FC > 0 ~ "up_adjp.05", # up sig .05
      adj.pvalue < 0.05 & log2FC < 0 ~ "down_adjp.05", # down sig .05
      adj.pvalue < 0.1 & log2FC > 0 ~ "up_adjp.1", # up 0.1
      adj.pvalue < 0.1 & log2FC < 0 ~ "down_adjp.1", # down .1
      TRUE ~ "ns" ),
      alpha = case_when( # conditional ass for ALPHA mapping
        adj.pvalue < 0.05 & log2FC > 0 ~ "up_adjp.05", # up sig .05
        adj.pvalue < 0.05 & log2FC < 0 ~ "down_adjp.05", # down sig .05
        adj.pvalue < 0.1 & log2FC > 0 ~ "up_adjp.1", # up 0.1
        adj.pvalue < 0.1 & log2FC < 0 ~ "down_adjp.1", # down .1
        TRUE ~ "ns" )), # catch the rest
       size = 1.75, shape = 21, stroke = 0.05,
      raster.dpi = getOption("ggrastr.default.dpi", 330)) +
    
    scale_fill_manual(values = c("firebrick", # coloring
                                  "royalblue3",
                                  "lightpink",
                                  "lightblue3",
                                  "grey80"),
                       breaks = c("up_adjp.05",
                                  "down_adjp.05",
                                  "up_adjp.1",
                                  "down_adjp.1",
                                  "ns"),
                       labels = c("\U2191 (adj.p <.05)",
                                  "\U2193 (adj.p <.05)",
                                  "\U2191 (adj.p <.1)",
                                  "\U2193 (adj.p <.1)",
                                  "ns")) +
    scale_alpha_manual( # alpha mapping
      values = c( # darker for more significant!
        up_adjp.05  = 0.3,
        down_adjp.05 = 0.3,
        up_adjp.1   = 0.3,
        down_adjp.1 = 0.3,
        ns = 0.1
      )) +
    
    
    
    geom_point_rast(data = my_prots, # ORGANE POINTS | ggrastr
                    aes(x = log2FC, y = -log10(adj.pvalue)),
                    alpha = .5, size = 1.75, shape = 21,
                    fill = "#E17C05FF", color = "#FCA842FF",
                    raster.dpi = getOption("ggrastr.default.dpi", 330)) + 
    
    geom_text_repel(data =  my_lg_prot, # TOP 6 MOST SIG ANNO
                    aes(x = log2FC, y = -log10(adj.pvalue),
                        label = gene,
                        color = log2FC > 0), # color mapping (see below)
                    max.overlaps = 30, size = 2.1, # gene text on points
                    min.segment.length = .5, segment.alpha = 0.3,
                    segment.size = 0.2, alpha = 0.9) +
                    # nudge_x = 0.155) +

    
    # NUM SIG UP & DOWN! | use sep df, overwise overplots
    geom_text(data = prot_sums, # up 
              aes(label = paste0(up, " \U2191")), 
              x = 1.35, y = 5.6, size = 3.2, color = "firebrick") + 
    geom_text(data = prot_sums, # down
              aes(label = paste0(down, " \U2193")),
              x = -1.4, y = 5.6, size = 3.2, color = "royalblue3") + 
    
    
    geom_text(data = path_sig_lab, # GSEA SIGNIFICANCE LABEL
              aes(label = paste0(sig_val),
                  y = 4.80,
                  x = ifelse(NES > 0, 1.525, -1.525), # side if neg or pos
                  color = NES > 0), # col - TRUE FALSE 
              size = 3.3) +
    scale_color_manual( # based on TRUE FALSE give color! 
      values = c(`TRUE` = "firebrick", `FALSE` = "royalblue3")) +
    
    
    
    
    labs(subtitle = paste0("KEGG - ", my_title,
                           " (", mapped_totals, "/", gs_size, ")")) +
    
    
    # facet_wrap(~Label, nrow = 1) + 
    facet_wrap2(~Label, nrow = 1,
                strip = strip
                ) + 
    
    theme_bw() +
    theme(
      # strip.background.x = element_rect(fill = my_level_colors),
      legend.position = "none",
      plot.subtitle = element_text(size = 8, vjust = 0, color = "grey10"),
      axis.title = element_text(size =7, color = "grey10"),
      axis.text = element_text(size =6.5, color = "grey10"),
      axis.ticks = element_blank(),
      strip.text = element_text(size = 6.2, color = "grey10"),
      panel.grid = element_blank(),
      strip.background = element_rect(linewidth = 0.25, color = "grey10"),
      panel.border = element_rect(linewidth = 0.25, color = "grey10")
    )
  
  print(vol_plot)
  
  # _____________________________________
  # 3 - SAVE THE PLOTS
  kegg_volcano_list[[kp]] <- vol_plot
  
  print(paste0("Done for ", kp, "! :D"))
  
}
```

```{r}
# plot_grid(plotlist = kegg_volcano_list[c(10,11,6:9)], ncol = 2)
# plot_grid(plotlist = kegg_volcano_list[1:12], ncol = 2)
plot_grid(plotlist = kegg_volcano_list[55:67], ncol = 3)
```

```{r}
# save(kegg_volcano_list, file = "output/2026_04_04_KEGG_volcano_67_EExAGE_6.rda")
```

------------------------------------------------------------------------

### REACTOME

```{r}
REACT_tibs_df_6_entrez_split <- REACT_tibs_df_6 %>%
  mutate(core_enrich_split = str_split(core_enrichment, "/")) %>% # split by /
  tidyr::unnest(core_enrich_split) %>% # unnest | to  3816 x 13
  left_join(bitr_gn_2_entrez, by = join_by(core_enrich_split == ENTREZID)) #%>% #map2gene
  # mutate()
```

```{r}
# GET REACTOME PATHVWAYS OF INTEREST 
REACT_POI  <- c(
  #TOP-LEVEL REACTOME
  # "REACTOME_DNA_REPLICATION",
  # "REACTOME_METABOLISM_OF_RNA", 
  # "REACTOME_DNA_REPAIR",
  # "REACTOME_CELL_CYCLE",
  # "REACTOME_PROGRAMMED_CELL_DEATH",
  # "REACTOME_EXTRACELLULAR_MATRIX_ORGANIZATION",
  # "REACTOME_TRANSPORT_OF_SMALL_MOLECULES",
  # "REACTOME_PROTEIN_LOCALIZATION",
  # # "REACTOME_HEMOSTASIS",
  # # "REACTOME_CIRCADIAN_CLOCK",
  # "REACTOME_NEURONAL_SYSTEM",
  # "REACTOME_VESICLE_MEDIATED_TRANSPORT",
  # 
  # 
  # # SIGNALLING PATHWAYS
  # "REACTOME_PROLONGED_ERK_ACTIVATION_EVENTS",
  # "REACTOME_DISEASES_OF_SIGNAL_TRANSDUCTION_BY_GROWTH_FACTOR_RECEPTORS_AND_SECOND_MESSENGERS",
  # "REACTOME_INTRACELLULAR_SIGNALING_BY_SECOND_MESSENGERS",
  # "REACTOME_REGULATION_OF_RAS_BY_GAPS",
  # "REACTOME_PTEN_REGULATION",
  # "REACTOME_MAPK_FAMILY_SIGNALING_CASCADES",
  # "REACTOME_FCERI_MEDIATED_NF_KB_ACTIVATION",
  # "REACTOME_",
  # "REACTOME_",
  # "REACTOME_",
  # "REACTOME_",
  # "REACTOME_",

  
##_L1_NEURONAL_SYSTEM__(づ ◕‿◕ )づ____
'REACTOME_NEURONAL_SYSTEM', # L1 (TOP LEVEL)

##_L2_Transmission_across_the_chemical_synapse_
'REACTOME_TRANSMISSION_ACROSS_CHEMICAL_SYNAPSES', #L2

##_L3_Neurotransmitter_release_cycle_
'REACTOME_NEUROTRANSMITTER_RELEASE_CYCLE', #L3
'REACTOME_GLUTAMATE_NEUROTRANSMITTER_RELEASE_CYCLE', #L4
'REACTOME_GABA_SYNTHESIS_RELEASE_REUPTAKE_AND_DEGRADATION', #L4

##_L3_Neurotransmitter_receptors_&_postsynaptic_signal_transduciton_
'REACTOME_NEUROTRANSMITTER_RECEPTORS_AND_POSTSYNAPTIC_SIGNAL_TRANSMISSION', #L3
##_L4_Glutamate_binding_activation_of_AMPA_&_synaptic_plasticity
'REACTOME_TRAFFICKING_OF_AMPA_RECEPTORS', #L5

##_L4_Activation_of_NMDA_receptors_PSD_events
'REACTOME_UNBLOCKING_OF_NMDA_RECEPTORS_GLUTAMATE_BINDING_AND_ACTIVATION', #L5
'REACTOME_ASSEMBLY_AND_CELL_SURFACE_PRESENTATION_OF_NMDA_RECEPTORS', #L5
##_L5_Post_NMDA_receptor_activation_events_
'REACTOME_CREB1_PHOSPHORYLATION_THROUGH_THE_ACTIVATION_OF_ADENYLATE_CYCLASE', #L6


##_L2_Protein-protein_interactions_at_the_synapse_
'REACTOME_NEUREXINS_AND_NEUROLIGINS', #L3

##_L2_Potassium_Channels_
'REACTOME_POTASSIUM_CHANNELS', #L2
'REACTOME_VOLTAGE_GATED_POTASSIUM_CHANNELS', #L3
  

###_1_DEVELOPMENTAL_BIOLOGY!__(づ ◕‿◕ )づ____
##_L2_NERVOUS_SYSTEM_DEVELOPMENT______
'REACTOME_NERVOUS_SYSTEM_DEVELOPMENT',
##_L3_Axon_Guidance_______
##_L4_L1CAM_interactions___
'REACTOME_L1CAM_INTERACTIONS', #L4 
'REACTOME_INTERACTION_BETWEEN_L1_AND_ANKYRINS', #L5
'REACTOME_RECYCLING_PATHWAY_OF_L1', #L5

##_L4_EPH-Ephrin_signalling___
'REACTOME_EPH_EPHRIN_SIGNALING', #L4
'REACTOME_EPHB_MEDIATED_FORWARD_SIGNALING', #L5

##_L4_Signalling_by_ROBO_receptors___
'REACTOME_SIGNALING_BY_ROBO_RECEPTORS', #L4
'REACTOME_REGULATION_OF_EXPRESSION_OF_SLITS_AND_ROBOS', #L5

##_L4_Semaphorin_interactions___
'REACTOME_SEMAPHORIN_INTERACTIONS', #L4
'REACTOME_CRMPS_IN_SEMA3A_SIGNALING', #L5
'REACTOME_SEMA3A_PAK_DEPENDENT_AXON_REPULSION', #L5
'REACTOME_OTHER_SEMAPHORIN_INTERACTIONS', #L5

##_L2_LGI-ADAM_interactions___
'REACTOME_LGI_ADAM_INTERACTIONS',






##_L1_EXTRACELLUAR_MATRIX_ORGANIZATION_(づ ◕‿◕ )づ_____
'REACTOME_EXTRACELLULAR_MATRIX_ORGANIZATION', # L1 (TOP LEVEL)

##_L2_ECM_Proteoglycans__
"REACTOME_ECM_PROTEOGLYCANS", #L2

##_L2_Non-integrin_membrane-ECM_interactions__
'REACTOME_FORMATION_OF_THE_DYSTROPHIN_GLYCOPROTEIN_COMPLEX_DGC', #L3 

##_L2_Laminin_Interactions__
'REACTOME_LAMININ_INTERACTIONS', #L2



##_L1_CELL-CELL_COMMUNICATION_(づ ◕‿◕ )づ_____
##_L2_Cell_junction_organization__
##_L3_Cell-cell_junction_org__
"REACTOME_ADHERENS_JUNCTIONS_INTERACTIONS" #L4
  

  
  
  
  # "REACTOME_CELL_CYCLE",
  # "REACTOME_SYNTHESIS_OF_DNA",
  # "REACTOME_DNA_REPLICATION",
  # "REACTOME_SWITCHING_OF_ORIGINS_TO_A_POST_REPLICATIVE_STATE",
  # "REACTOME_MITOTIC_G1_PHASE_AND_G1_S_TRANSITION",
  # "REACTOME_S_PHASE",
  # "REACTOME_M_PHASE",
  # "REACTOME_APC_C_MEDIATED_DEGRADATION_OF_CELL_CYCLE_PROTEINS",
  # "REACTOME_CYCLIN_A_CDK2_ASSOCIATED_EVENTS_AT_S_PHASE_ENTRY",
  # "REACTOME_STABILIZATION_OF_P53",
  # 
  # 
  # 
  # 
  # "REACTOME_APOPTOSIS",
  # "REACTOME_REGULATION_OF_APOPTOSIS",
  # "REACTOME_PROGRAMMED_CELL_DEATH",
  # "REACTOME_AGGREPHAGY",
  # "REACTOME_SELECTIVE_AUTOPHAGY",
  # 
  # "REACTOME_MRNA_SPLICING",
  # "REACTOME_CHOLESTEROL_BIOSYNTHESIS",
  
  # ##_IMMUNE_SYSTEM____________________________
  # "REACTOME_CYTOKINE_SIGNALING_IN_IMMUNE_SYSTEM", # CYTOKINE IMMUNE
  # "REACTOME_INTERLEUKIN_1_FAMILY_SIGNALING", 
  # "REACTOME_INTERLEUKIN_12_FAMILY_SIGNALING", 
  # "REACTOME_TNFR2_NON_CANONICAL_NF_KB_PATHWAY", 
  # 
  # "REACTOME_ADAPTIVE_IMMUNE_SYSTEM", # ADATIVE IMMUNE
  # "REACTOME_TCR_SIGNALING",
  # "REACTOME_SIGNALING_BY_THE_B_CELL_RECEPTOR_BCR",
  # 
  # "REACTOME_COMPLEMENT_CASCADE", # INNATE IMMUNE
  # "REACTOME_FC_EPSILON_RECEPTOR_FCERI_SIGNALING",
  # "REACTOME_C_TYPE_LECTIN_RECEPTORS_CLRS",
  # "REACTOME_FCGAMMA_RECEPTOR_FCGR_DEPENDENT_PHAGOCYTOSIS"
  )

# reactome_volcano_top_lvl_list <- reactome_volcano_toplevel_list 

## EMPTY PLOT LIST 
reactome_volcano_list <-  vector(mode = "list")
```

```{r}
for(rct in REACT_POI) {

# 1 - PREPARE THE DATA
  my_genes <- REACT_tibs_df_6_entrez_split %>%
    filter(ID == rct) %>%
    dplyr::distinct(SYMBOL) %>%
    tibble::deframe()
  
  my_title <- REACT_tibs_df_6_entrez_split %>%
    filter(ID == rct) %>%
    dplyr::distinct(Description) %>%
    tibble::deframe()
  
  my_prots <- protein_comp_entrez %>% # from MSstatsTMT protein comps w ENTREZ map
    filter(!(Label %in% c("HxA_EE_OLD_vs_SH_YNG", "HxA_EE_YNG_vs_SH_OLD"))) %>%
    filter(gene %in% my_genes) %>%
    dplyr::group_by(Label) %>%
    mutate(up = sum(log2FC > 0),
           down = sum(log2FC < 0)) 
  
  prot_sums <- my_prots %>% # this is needed to avoid overplotting
    dplyr::select(Label, up, down) %>%
    unique()
  
  
   mapped_totals <- my_prots %>%  # for mapped total
    ungroup() %>%
    summarise(n_distinct(gene)) %>%
    tibble::deframe()
  
  gs_size <- REACT_tibs_df_6_entrez_split %>% # for geneset size
    filter(ID == rct) %>%
    dplyr::distinct(setSize) %>%
    tibble::deframe()
  
  
  path_sig <- REACT_tibs_df_6 %>% # GET DETAILS FOR ANNOTATING SIG 
    filter(!(Label %in% c("HxA_EE_OLD_vs_SH_YNG", "HxA_EE_YNG_vs_SH_OLD"))) %>%
    filter(ID == rct) %>%
    dplyr::group_by(Label) %>% 
    mutate(sig_val = case_when(
      p.adjust <= 0.001 ~ "***",
      p.adjust <= 0.01 ~ "**",
      p.adjust <= 0.05 ~ "*",
      .default = "NA" )) # if bit there then NA (ns)
  
  
  # interesting... this step needed for some reason...
  path_sig_lab <-  path_sig %>% 
    dplyr::select(ID, Label, sig_val, NES) %>%
    unique() 
  
  my_lg_prot <- my_prots %>% #LABEL LEADING GENE???
    arrange(adj.pvalue) %>% # not quite - 6 most sig proteins
    slice_head(n = 6)
  
  
    strip <- strip_themed( # green for 1st 3 (housing) & Lavender for last 3 (age)
    background_x = elem_list_rect(fill = c(
      "#AFCC9F", "#D4E4CB", "#D4E4CB",
      "#BFBBD5", "#DDDCE9", "#DDDCE9")))

    
  # _____________________________________
  # 2 - MAKE THE VOLCANO PLOTS!!!
  vol_plot <- protein_comp_entrez %>%
    filter(!(Label %in% c("HxA_EE_OLD_vs_SH_YNG", "HxA_EE_YNG_vs_SH_OLD"))) %>%
    
    ggplot(aes(x = log2FC, y = -log10(adj.pvalue))) +
    
    
    # limits - rm 17 (but none are leading edge - shrunken for comparisons)
    scale_x_continuous(limits = c(-2,2),
                       expand = c(0, 0)) +
    scale_y_continuous(limits = c(0,6),
                       expand = c(0.04, 0)) +
    
    geom_hline(yintercept = -log10(0.05),
               linetype = "dashed", linewidth = 0.1,
               color = "grey50", alpha = 0.5) +
    geom_hline(yintercept = -log10(0.1),linetype = "dotted",
               linewidth = 0.1, color = "grey60", alpha = 0.5) +
    geom_vline(xintercept = c(0),linetype = "solid",
               linewidth = 0.1, color = "grey35", alpha = 0.5) +


    ggrastr::geom_point_rast(aes(fill = case_when( # conditional ass for color mapping
      adj.pvalue < 0.05 & log2FC > 0 ~ "up_adjp.05", # up sig .05
      adj.pvalue < 0.05 & log2FC < 0 ~ "down_adjp.05", # down sig .05
      adj.pvalue < 0.1 & log2FC > 0 ~ "up_adjp.1", # up 0.1
      adj.pvalue < 0.1 & log2FC < 0 ~ "down_adjp.1", # down .1
      TRUE ~ "ns" ),
      alpha = case_when( # conditional ass for ALPHA mapping
        adj.pvalue < 0.05 & log2FC > 0 ~ "up_adjp.05", # up sig .05
        adj.pvalue < 0.05 & log2FC < 0 ~ "down_adjp.05", # down sig .05
        adj.pvalue < 0.1 & log2FC > 0 ~ "up_adjp.1", # up 0.1
        adj.pvalue < 0.1 & log2FC < 0 ~ "down_adjp.1", # down .1
        TRUE ~ "ns" )), # catch the rest
       size = 1.75, shape = 21, stroke = 0.05,
      raster.dpi = getOption("ggrastr.default.dpi", 330)) +
    
    scale_fill_manual(values = c("firebrick", # coloring
                                  "royalblue3",
                                  "lightpink",
                                  "lightblue3",
                                  "grey80"),
                       breaks = c("up_adjp.05",
                                  "down_adjp.05",
                                  "up_adjp.1",
                                  "down_adjp.1",
                                  "ns"),
                       labels = c("\U2191 (adj.p <.05)",
                                  "\U2193 (adj.p <.05)",
                                  "\U2191 (adj.p <.1)",
                                  "\U2193 (adj.p <.1)",
                                  "ns")) +
    scale_alpha_manual( # alpha mapping
      values = c( # darker for more significant!
        up_adjp.05  = 0.3,
        down_adjp.05 = 0.3,
        up_adjp.1   = 0.3,
        down_adjp.1 = 0.3,
        ns = 0.1)) +
    
    
    geom_point_rast(data = my_prots, # ORGANE POINTS | ggrastr
                    aes(x = log2FC, y = -log10(adj.pvalue)),
                    alpha = .5, size = 1.75, shape = 21,
                    fill = "#E17C05FF", color = "#FCA842FF",
                    raster.dpi = getOption("ggrastr.default.dpi", 330)) + 
    
    geom_text_repel(data =  my_lg_prot, # TOP 6 MOST SIG ANNO
                    aes(x = log2FC, y = -log10(adj.pvalue),
                        label = gene,
                        color = log2FC > 0), # color mapping (see below)
                    max.overlaps = 30, size = 2.1, # gene text on points
                    min.segment.length = .5, segment.alpha = 0.3,
                    segment.size = 0.2, alpha = 0.9) +
                    # nudge_x = 0.155) +

    
    # NUM SIG UP & DOWN! | use sep df, overwise overplots
    geom_text(data = prot_sums, # up 
              aes(label = paste0(up, " \U2191")), 
              x = 1.35, y = 5.6, size = 3.2, color = "firebrick") + 
    geom_text(data = prot_sums, # down
              aes(label = paste0(down, " \U2193")),
              x = -1.4, y = 5.6, size = 3.2, color = "royalblue3") + 
    
    
    geom_text(data = path_sig_lab, # GSEA SIGNIFICANCE LABEL
              aes(label = paste0(sig_val),
                  y = 4.80,
                  x = ifelse(NES > 0, 1.525, -1.525), # side if neg or pos
                  color = NES > 0), # col - TRUE FALSE 
              size = 3.3) +
    scale_color_manual( # based on TRUE FALSE give color! 
      values = c(`TRUE` = "firebrick", `FALSE` = "royalblue3")) +
    

    labs(subtitle = paste0(my_title," (", mapped_totals, "/", gs_size, ")")) +
    
    # facet_wrap(~Label, nrow = 1) + 
    facet_wrap2(~Label, nrow = 1,
                strip = strip) + 
    
    theme_bw() +
    theme(
      # strip.background.x = element_rect(fill = my_level_colors),
      legend.position = "none",
      plot.subtitle = element_text(size = 8, vjust = 0, color = "grey10"),
      axis.title = element_text(size =7, color = "grey10"),
      axis.text = element_text(size =6.5, color = "grey10"),
      axis.ticks = element_blank(),
      strip.text = element_text(size = 6.2, color = "grey10"),
      panel.grid = element_blank(),
      strip.background = element_rect(linewidth = 0.25, color = "grey10"),
      panel.border = element_rect(linewidth = 0.25, color = "grey10")
    )
  
  print(vol_plot)

  # _____________________________________
  # 3 - SAVE THE PLOTS
  reactome_volcano_list[[rct]] <- vol_plot
  
  print(paste0("Done for ", rct))
  
}
```

```{r}
plot_grid(plotlist = reactome_volcano_list[1:12], ncol = 2)

# ggsave("2026_03_23_REACT_21_volcanos_300dpi.png", dpi = 300) # 11.4 x 5.99
```

------------------------------------------------------------------------

### Exporting Mapped Volcano Plots

#### Cholesterol & Isoprenoids/Terpenoid Backbone synthesis

```{r}
#rno00900
plot_grid(kegg_volcano_list$rno00900,
          reactome_volcano_list$REACTOME_CHOLESTEROL_BIOSYNTHESIS,
          ncol = 1, labels = c("a","b"), label_colour = "grey10", label_size = 11)

# ggsave("2026_03_30_Cholesterol_Terpenoid_mapped_volcano_330dpi.png", dpi = 330)
```

```{r}
# REACTOME INTRACELLULAR SIGNALLING
plot_grid(reactome_volcano_list$REACTOME_PROLONGED_ERK_ACTIVATION_EVENTS,
          reactome_volcano_list$REACTOME_REGULATION_OF_RAS_BY_GAPS,
          reactome_volcano_list$REACTOME_DISEASES_OF_SIGNAL_TRANSDUCTION_BY_GROWTH_FACTOR_RECEPTORS_AND_SECOND_MESSENGERS,
          reactome_volcano_list$REACTOME_INTRACELLULAR_SIGNALING_BY_SECOND_MESSENGERS,
  ncol = 1, labels = c("a","b", "c", "d"), label_colour = "grey10", label_size = 12)

#ggsave("2026_03_30_REACT_intracellular_sig_mapped_volcano_330dpi.png", dpi = 330) 
# 8.5 x 7.97 (6 labels x 4 grpahs)
```

```{r}
# # Ras signalling, cAMP signalling # cGMP-PKG & Calcium isngalling
plot_grid(kegg_volcano_list$rno04014,
          kegg_volcano_list$rno04024,
          kegg_volcano_list$rno04022, kegg_volcano_list$rno04020,
          # kegg_volcano_list$rno04360, kegg_volcano_list$rno05033, # axon guidance |nic
    labels = c("a", "b", "c"),
    label_colour = "grey10", label_size = 11,  ncol = 1)

# ggsave("2026_04_04_sig_EEyng_KEGG_intracell_sig_mapped_volcano_300dpi.png", dpi = 300)
# dev.off()
```

#### KEGG DNA, RNA, & PROTEIN

```{r}
# KEGG ATP-dependent chromatin remodelling & Spliceosome
plot_grid(kegg_volcano_list$rno03082, kegg_volcano_list$rno03040,
          ncol = 1,
          labels = c("a","b"),
          label_colour = "grey10", label_size = 11)
```

```{r}
# KEGG Ribosome & nucleocytoplasmic transport
plot_grid(kegg_volcano_list$rno03010, kegg_volcano_list$rno03013,
          ncol = 1,
          labels = c("a","b"),
          label_colour = "grey10", label_size = 11)
```

```{r}
# KEGG Proteasome & Protein Processing in the ER
plot_grid(kegg_volcano_list$rno03050, kegg_volcano_list$rno04141,
          ncol = 1,
          labels = c("a","b"),
          label_colour = "grey10", label_size = 11)
```

```{r}
# KEGG - others - sig in EE_YNG -  gene expression/protein/mitophagy 
# ubiqui med proteoly|mRNA surveillance|base exicion rep| Polycomb repressive|mitophagy 
plot_grid(kegg_volcano_list$rno04120, kegg_volcano_list$rno03015,
          kegg_volcano_list$rno03410, kegg_volcano_list$rno03083,
          kegg_volcano_list$rno04137,
          ncol = 1,
          labels = c("a","b", "c", "d", "e"),
          label_colour = "grey10", label_size = 11) 
```

Synaptic/Neurotransmission

```{r}
# neurotransmitter release & synaptic vesicle cycle
plot_grid(kegg_volcano_list$rno04721, reactome_volcano_list$REACTOME_NEUROTRANSMITTER_RELEASE_CYCLE,
          ncol = 1, labels = c("a","b"), label_colour = "grey10", label_size = 11)
```

```{r}
# axonguidance \& nervous system development
plot_grid(kegg_volcano_list$rno04360,
          reactome_volcano_list$REACTOME_NERVOUS_SYSTEM_DEVELOPMENT,
          ncol = 1, labels = c("a","b"), label_colour = "grey10", label_size = 11)
```

```{r}
plot_grid(kegg_volcano_list$rno04721, kegg_volcano_list$rno04360,
          ncol = 1, labels = c("a","b"), label_colour = "grey10", label_size = 11)
```

```{r}
# GABA, ret_endoca, neuroa_lig x 2,  nicotine
plot_grid(
  kegg_volcano_list$rno04723, kegg_volcano_list$rno04727,
          # kegg_volcano_list$rno04082, kegg_volcano_list$rno04080,
          kegg_volcano_list$rno05033,
          # labels = c("a","b", "c", "d", "e"),
          labels = c("a","b", "c"),
          ncol = 1, label_colour = "grey10", label_size = 11)
```

```{r}
# NERVOUS SYSTEM 
plot_grid(reactome_volcano_list$REACTOME_NEURONAL_SYSTEM, 
          reactome_volcano_list$REACTOME_NERVOUS_SYSTEM_DEVELOPMENT,
          ncol = 1, labels = c("a","b"),
          label_colour = "grey10", label_size = 11)
```

```{r}
# REACTOME BRAAIN 
plot_grid(

  ##_Nervous_System_Development_____        
  # reactome_volcano_list$REACTOME_NERVOUS_SYSTEM_DEVELOPMENT,
  # reactome_volcano_list$REACTOME_SEMAPHORIN_INTERACTIONS,
  # reactome_volcano_list$REACTOME_SIGNALING_BY_ROBO_RECEPTORS,
  # reactome_volcano_list$REACTOME_CRMPS_IN_SEMA3A_SIGNALING,
  # reactome_volcano_list$REACTOME_SEMA3A_PAK_DEPENDENT_AXON_REPULSION,
  # reactome_volcano_list$REACTOME_EPH_EPHRIN_SIGNALING,
  # reactome_volcano_list$REACTOME_LGI_ADAM_INTERACTIONS,
  # reactome_volcano_list$REACTOME_L1CAM_INTERACTIONS,
  # 
  
##______ECM_____________________________ 
# reactome_volcano_list$REACTOME_EXTRACELLULAR_MATRIX_ORGANIZATION, # L1 (TOP) ECM
# reactome_volcano_list$REACTOME_ECM_PROTEOGLYCANS, #L2 ECM
# reactome_volcano_list$REACTOME_FORMATION_OF_THE_DYSTROPHIN_GLYCOPROTEIN_COMPLEX_DGC, #L3
# reactome_volcano_list$REACTOME_LAMININ_INTERACTIONS, #L3

##_Cell2Cell_____Adherins_&_Gap_Junctions_______
# reactome_volcano_list$REACTOME_ADHERENS_JUNCTIONS_INTERACTIONS, #L4 Cell2Cell Coms
# reactome_volcano_list$REACTOME_GAP_JUNCTION_ASSEMBLY,
# reactome_volcano_list$REACTOME_GAP_JUNCTION_TRAFFICKING_AND_REGULATION,


  
  # ##_Neuronal_sytem____(づ ◕‿◕ )づ___________________ 
  # reactome_volcano_list$REACTOME_NEURONAL_SYSTEM, # top level
  # 
  # reactome_volcano_list$REACTOME_TRANSMISSION_ACROSS_CHEMICAL_SYNAPSES, #L2
  # 
  # ##_L3_Neurotransmitter_release_cycle_
  # reactome_volcano_list$REACTOME_NEUROTRANSMITTER_RELEASE_CYCLE, #L3
  # reactome_volcano_list$REACTOME_GLUTAMATE_NEUROTRANSMITTER_RELEASE_CYCLE, #L4
  # reactome_volcano_list$REACTOME_GABA_SYNTHESIS_RELEASE_REUPTAKE_AND_DEGRADATION, #L4
  
  
  # ##_L3_Neurotransmitter_receptors_&_postsynaptic_signal_transduciton_
  # reactome_volcano_list$REACTOME_NEUROTRANSMITTER_RECEPTORS_AND_POSTSYNAPTIC_SIGNAL_TRANSMISSION, #L3
  # 
  # ##_L4_Glutamate_binding_activation_of_AMPA_&_synaptic_plasticity
  # reactome_volcano_list$REACTOME_TRAFFICKING_OF_AMPA_RECEPTORS, # L5 (L4 AMPA)
  # 
  # ##_L4_Activation_of_NMDA_receptors_PSD_events (L5 & L5)
  # reactome_volcano_list$REACTOME_UNBLOCKING_OF_NMDA_RECEPTORS_GLUTAMATE_BINDING_AND_ACTIVATION,
  # reactome_volcano_list$REACTOME_ASSEMBLY_AND_CELL_SURFACE_PRESENTATION_OF_NMDA_RECEPTORS,
  # ##_L5_Post_NMDA_receptor_activation_events_(L6)
  # reactome_volcano_list$REACTOME_CREB1_PHOSPHORYLATION_THROUGH_THE_ACTIVATION_OF_ADENYLATE_CYCLASE, #L6

  # #_L2_Protein-protein_interactions_at_the_synapse_L3
  # reactome_volcano_list$REACTOME_NEUREXINS_AND_NEUROLIGINS, #L3


  # ##_L2_Potassium_Channels_L2&L3
  # reactome_volcano_list$REACTOME_POTASSIUM_CHANNELS,  #L2
  # reactome_volcano_list$REACTOME_VOLTAGE_GATED_POTASSIUM_CHANNELS, #L3

 # ##_IMMUNE_SYSTEM____________________________
# reactome_volcano_list$REACTOME_CYTOKINE_SIGNALING_IN_IMMUNE_SYSTEM, # CYTOKINE IMMUNE
#   reactome_volcano_list$REACTOME_INTERLEUKIN_1_FAMILY_SIGNALING,  #3
#  reactome_volcano_list$REACTOME_INTERLEUKIN_12_FAMILY_SIGNALING,  #3
#  reactome_volcano_list$REACTOME_TNFR2_NON_CANONICAL_NF_KB_PATHWAY,  #2

# reactome_volcano_list$REACTOME_ADAPTIVE_IMMUNE_SYSTEM,  #adaptive immune T1
#  reactome_volcano_list$REACTOME_TCR_SIGNALING,  # TCR #2
#  reactome_volcano_list$REACTOME_SIGNALING_BY_THE_B_CELL_RECEPTOR_BCR, #2

 # reactome_volcano_list$REACTOME_COMPLEMENT_CASCADE, # complement | innate immune?
 # reactome_volcano_list$REACTOME_FC_EPSILON_RECEPTOR_FCERI_SIGNALING,
 # reactome_volcano_list$REACTOME_C_TYPE_LECTIN_RECEPTORS_CLRS,
 # reactome_volcano_list$REACTOME_FCGAMMA_RECEPTOR_FCGR_DEPENDENT_PHAGOCYTOSIS,
  
  
  # #high_level_terms
  # reactome_volcano_list$REACTOME_NEURONAL_SYSTEM, #L1
  # # reactome_volcano_list$REACTOME_NERVOUS_SYSTEM_DEVELOPMENT, #L2
  # reactome_volcano_list$REACTOME_EXTRACELLULAR_MATRIX_ORGANIZATION, #L1
  # reactome_volcano_list$REACTOME_TRANSPORT_OF_SMALL_MOLECULES, #L1
  # reactome_volcano_list$REACTOME_VESICLE_MEDIATED_TRANSPORT, #L1
  # 
  # reactome_volcano_list$REACTOME_HEMOSTASIS, #L1
  # reactome_volcano_list$REACTOME_CIRCADIAN_CLOCK, #L1

  
  reactome_volcano_list$REACTOME_DNA_REPLICATION,
  reactome_volcano_list$REACTOME_METABOLISM_OF_RNA,
  reactome_volcano_list$REACTOME_DNA_REPAIR,
  reactome_volcano_list$REACTOME_CELL_CYCLE,
  reactome_volcano_list$REACTOME_PROGRAMMED_CELL_DEATH,
  reactome_volcano_list$REACTOME_PROTEIN_LOCALIZATION, #L1



  # labels = c("a","b", "c", "d", "e"),    
  # labels = c("a","b", "c", "d"),
  ncol = 1, label_colour = "grey10", label_size = 11

      )

# ggsave("2026_03_31_axonguidance_brain_dev_mapped_volcano_330dpi.png", dpi = 330)

```

```{r}
plot_grid(plotlist = reactome_volcano_list[15:31], ncol = 3)
```

#### KEGG Oxidative Phosphorylation \\& TCA cycle

```{r}
# KEGG Oxidattive Phosphorylation & TCA Cycle
plot_grid(
  kegg_volcano_list$rno00190, kegg_volcano_list$rno00020,
    labels = c("a","b"), 
    ncol = 1, label_colour = "grey10", label_size = 11)
# ggsave("2026_04_04_KEGG_2_Energy_metabolism_mapped_volcano_300dpi.png", dpi = 300)
# dev.off()
```

#### KEGG Neurodegenerative Diseases

```{r}
# KEGG NEURO DEGENERATIVE
plot_grid(kegg_volcano_list$rno05022, kegg_volcano_list$rno05012,
          kegg_volcano_list$rno05016, kegg_volcano_list$rno05020,
          kegg_volcano_list$rno05010, kegg_volcano_list$rno05014,
          ncol = 1,
          labels = c("a","b", "c", "d", "e", "f"),
          label_colour = "grey10", label_size = 11)
```

CONT HERE

```{r}
# TO REDO
# REACTOME - NT systems & transmission
plot_grid(plotlist = reactome_volcano_list[6:11], 
          ncol = 1,
          # labels = c("a","b","c"),
          label_colour = "grey10", label_size = 12)

# ggsave("2026_03_31_REACT_NT_systems_mapped_volcano_330dpi.png", dpi = 330)
```

```{r}
#REDO
# REACTOME - other brain
plot_grid(plotlist = reactome_volcano_list[12:17], 
          ncol = 1,
          # labels = c("a","b","c"),
          label_colour = "grey10", label_size = 12)

# ggsave("2026_03_31_REACT_other_brain_mapped_volcano_330dpi.png", dpi = 330)
```

------------------------------------------------------------------------
