---
title: "BioIn_2026_06_v3.4.2_ClusterProfiler_EExAge_vis_v3"
author: "Stephanie Huang"
date: "2026-06-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## MSigDB Hallmarks - Lolliplots w Categories & Mapped Volcanos

## Load Packages

```{r}
# https://bioconductor.org/install/
set.seed(123) # for reproducibility

# Load Packages 
library(tidyverse) # for dplyr, ggplot, stringr, readr, purrr, tibble, tidyr etc 
library(clusterProfiler) #GSEA includes gsea, enrichplot etc
library(topGO) # needed for clusterProfiler's plotGOgraph() graph
library(msigdbr) # MSigDB gene sets
library(enrichplot) #emapplot, etc

# library(drawCell) # drawing GO CC - SwissBioPics | not in new ver R
library(htmlwidgets) 
library(webshot2)

library(paletteer) # colors
library(colorspace) # colors
library(cowplot) # plot_grid

library(readxl)
library(writexl) # for excel tab export
library(ggvenn) # vennys (old ones)
library(scales) #label_wrap text break in lolliplots
library(ggh4x)
library(ggrepel) # geom_text_repel
library(ggrastr) # for volcano points! 
library(gt)
```

## Load Data

```{r}
# mapped_comp_v3.2.2_EExAge_t1
load("input/protein_lists/2025_12_31_1pspg_v3.2.2_t1_EExAge_mapped_comp.rda")

df <- mapped_comp_v3.2.2_EExAge_t1
```

```{r}
# MSIGDB Hallmarks
load("output/gsea_results/2026_01_12_gsea_msig_Hallmark_EExAge_mssv3.2.2_t1_cpv3.4.2_t2.rda")

# GO 
load("output/gsea_results/2026_01_12_GO_gsea_EExAge_mssv3.2.2t1_cpv3.4.2t2.rda")

# GO SIMPLIFIED
load("output/gsea_results/2026_01_24_GO_gsea_EExAge_cp_v3.4.2_t2_simplified.rda")
```

```{r}
load("input/db/rat_ref_proteome_2025_12_13_51872x295.rda")

rat_ref <- rat_ref_proteome_2025_12_13
```

```{r}
rat_FASTA <- read_csv("input/db/UP000002494_10116_Rat_Ref_Proteome_22403_1pspg_2025_04_UPfastadf.csv")
```

------------------------------------------------------------------------

```{r}
# see BioIn_2025_12_v3.4.2_ClusterProfiler_EExAge_vis_testing_v1.Rmd for more info
# & BioIn_2026_01_v3.4.2_ClusterProfiler_EExAge_vis_v2.Rmd

protein_comp <- df %>%
  filter(is.na(issue)) %>% # keeps rows with issue == na (so comparisons that are ok)
  filter(!str_detect(log2FC, "Inf")) # remove infinity/missing fold change values

# check 
head(protein_comp) 
colnames(protein_comp)


# ENTREZ
protein_comp_uni_gene <- unique(protein_comp$gene) # 8014

bitr_gn_2_entrez <- clusterProfiler::bitr(protein_comp_uni_gene,
                         fromType = "SYMBOL",
                         toType = "ENTREZID",
                         OrgDb = "org.Rn.eg.db") #3.39% failed to map..better | 3.72% v2

# wait R 4.6 -  2026_06 - 3.72% of input gene IDs are fail to map... 

# SEE MY LEVELS BELOW 
my_comp_levels = c("H_EE_vs_SH",
              "HxA_EE_YNG_vs_SH_YNG",
              "HxA_EE_OLD_vs_SH_OLD",
              "A_YNG_vs_OLD", 
              "HxA_EE_YNG_vs_EE_OLD",
              "HxA_SH_YNG_vs_SH_OLD",
              
              "HxA_EE_OLD_vs_SH_YNG", "HxA_EE_YNG_vs_SH_OLD"
)


# MAP ENTREZ TO UNIPROT PROT COMPS
protein_comp_entrez <- left_join(protein_comp, bitr_gn_2_entrez,
                                 by = join_by(gene == SYMBOL)) %>%
  filter(!is.na(ENTREZID))  # remove non mapped - 64384 to 61952 


protein_comp_entrez$Label <- factor(protein_comp_entrez$Label,
                                  levels = my_comp_levels)

```

## Preparing MSigDB Hallmarks

```{r}
# MAP Msigdb Hallmarks to masterdf
HM_tibs_df <- map_dfr(gsea_msig_Hmrk_EExAge_mss3.2.2_t1_cp3.4.2_t2$gsea_results_tib,
                        as.data.frame,
                        .id = "Label")
# unique(HM_tibs_df_6$Description)

# Based on Liberzon 2015 paper process categories (table 1)
HM_tibs_df_6 <- HM_tibs_df %>%
    filter(!(Label %in% c("HxA_EE_OLD_vs_SH_YNG", "HxA_EE_YNG_vs_SH_OLD"))) %>%
  mutate(Category = case_when( # just doing manually from paper 
    Description == "HALLMARK_ADIPOGENESIS" ~ "Development",
    Description == "HALLMARK_ANDROGEN_RESPONSE" ~ "Signaling",
    
    Description %in% c("HALLMARK_COAGULATION",
                       "HALLMARK_INFLAMMATORY_RESPONSE") ~ "Immune",
    
    Description %in% c("HALLMARK_OXIDATIVE_PHOSPHORYLATION",
                       "HALLMARK_CHOLESTEROL_HOMEOSTASIS", 
                       "HALLMARK_FATTY_ACID_METABOLISM") ~ "Metabolic",
    
    grepl("TARGETS", Description) ~ "Proliferation", # MYC & E2F Targets
  ))


# SEE MY LEVELS BELOW 
HM_tibs_df_6_lvls = c("H_EE_vs_SH",
              "HxA_EE_YNG_vs_SH_YNG",
              "A_YNG_vs_OLD", 
              "HxA_EE_YNG_vs_EE_OLD",
              "HxA_SH_YNG_vs_SH_OLD"
)

HM_tibs_df_6$Label <- factor(HM_tibs_df_6$Label,
                             levels = HM_tibs_df_6_lvls)

```

------------------------------------------------------------------------

## MSigDB Lolliplot Redo

### Lolliplot function

```{r}
lolliplot <- function(data,
                      x_1 = Description, 
                      x_2 = NES, 
                      y = NES, 
                      fill_1 = NES,
                      alpha = p.adjust,
                      facet_formula = ~Label,
                      size = setSize,
                      subtitle = NULL,
                      legend_pos = "none") {
  
  # NEED TO SHUFFLE HERE & THERE BUT WORKS
    strip <- strip_themed( # green for 1st 3 (housing) & Lavender for last 3 (age)
    background_x = elem_list_rect(fill = c(
      "#AFCC9F", "#D4E4CB", 
      "#BFBBD5", "#DDDCE9", "#DDDCE9")))
  
  
  lol <- ggplot(data,
                aes(x = reorder({{x_1}},{{x_2}}),
                    y = {{y}}
                    )) +
    
    geom_segment(aes(x = {{x_1}},
                     yend = {{y}}, y = 0,
                     alpha = {{alpha}}
                     )) +
    
    geom_point(aes(alpha = {{alpha}},
                  size = {{size}},
                  fill = {{fill_1}}),
             shape = 21) +
    # scale_alpha(range = c(.95,.3)) + # rev alpha so darker = more sig | all
    scale_alpha(range = c(.95,.45)) + #  GO TRIM VERSION so matches others
    geom_hline(yintercept = 0, color = 'grey20', linewidth = 0.1) +
    scale_color_manual(breaks = c("Up", "Down"),
                     # values = c("#dc322f", "#268bd2"),
                     guide = 'none') + 
    scale_size_continuous(range = c(3,9), # map size oflolly (set size)
                          guide = guide_legend(nrow = 1)) + # 1 row
    # scale_size_continuous(range = c(2,8),
    #      guide = guide_legend(nrow=1)) + # map size oflolly (set size) | GO BP
    
    # scale_fill_continuous_diverging("Blue-Red 3") + # colorspace
    scale_fill_continuous_diverging("Blue-Red 2", mid = 0) +

    
    scale_x_discrete(labels = scales::label_wrap(width = 50)) + # break text |
  
    ylim(-4,4)+
  
    labs(subtitle = subtitle,
       y = 'NES' ) +
    
    coord_flip() + #fliptheplot
    
    # facet_wrap(facet_formula,
    #          space = 'free_x') +
     
    # facet_grid(facet_formula, # KEGG subcategory anno! half working ....
    #              space = "free", scales = "free") +
    
    facet_grid2(facet_formula,
                strip = strip,
                space = "free", scales = "free") +

    
    theme_bw() +
    theme(
      strip.text.y = element_text(angle = 0, size = 8.5),
      strip.text.x = element_text(size = 7),
      strip.background.y = element_rect(fill = fill_alpha("grey80", 0.5 )), 
      panel.spacing.y = unit(0, "lines"), # or "cm", "in", etc.
      
    axis.text.x = element_text(angle = 90, hjust = 1, color = "grey15",
                               vjust = 0, # 90, hj1 t/R, vj.5 cent
                               size = 9.5, face = 'italic'), 
    axis.text.y = element_text(hjust = 1, color = "grey15",
                               size = 8.5, face = 'italic',
                               lineheight = .7), # <1 smaller distances breaks
    axis.title.y = element_blank(),
    
    legend.margin = margin(-10,0,-5,0), # (top, right, bot, left) | bottom ver
    legend.title = element_text(face = 'italic', size =8,
                                hjust = 1),
    legend.key.size = grid::unit(0.3, "cm"), # posit answer to same height
    legend.text = element_text(margin = margin(l = 0)), # marg text & key
    legend.box.just = "right", # right justify legend box,
    legend.direction = "horizontal",
    
    legend.position = legend_pos) 
    
}
```

### Plot MSig Hallmarks Lolliplot with Category Labels

```{r}
HM_lol <- HM_tibs_df_6 %>%
 mutate(Description_2 = gsub("HALLMARK_", "", Description))  %>%
  lolliplot(subtitle = "MSigDB Hallmarks GSEA - Biological Themes",
            facet_formula = reorder(Category, -NES)~Label,
            x_1 = Description_2,
            legend_pos = "bottom")

HM_lol

```

------------------------------------------------------------------------

## MAPPED GSEA VOLCANO PLOTS - MSIGDB HALLMARKS

### Split leading edge in GSEA & map 2 gene names

```{r}
HM_tibs_df_6_entrez_split <- HM_tibs_df_6 %>%  #19x12
  mutate(core_enrich_split = str_split(core_enrichment, "/")) %>% #split by /
  tidyr::unnest(core_enrich_split) %>% # unnest to 760 x 13
  left_join(bitr_gn_2_entrez, # map 2 gene name via ENTREZID
            by = join_by(core_enrich_split == ENTREZID)) 
```

### Set up vectors & labels for loop

```{r}
Msig_HM_POI <- unique(HM_tibs_df_6$ID) # 9 

# EMPTY PLOT LIST 
Msig_HM_volcano_list <-  vector(mode = "list")
```

### Run for loop for plots

```{r}
for(hm in Msig_HM_POI) {
  
  # 1 - PREPARE THE DATA
  my_genes <- HM_tibs_df_6_entrez_split %>%
    filter(ID == hm) %>%
    dplyr::distinct(SYMBOL) %>%
    tibble::deframe()
  
  my_title <- HM_tibs_df_6_entrez_split %>%
    filter(ID == hm) %>%
    dplyr::distinct(Description) %>%
    tibble::deframe()
  
  my_prots <- protein_comp_entrez %>% # from MSstatsTMT protein comps w ENTREZ map
    filter(!(Label %in% c("HxA_EE_OLD_vs_SH_YNG", "HxA_EE_YNG_vs_SH_OLD"))) %>%
    filter(gene %in% my_genes) %>%
    dplyr::group_by(Label) %>%
    mutate(up = sum(log2FC > 0),
           down = sum(log2FC < 0)) 
  
  prot_sums <- my_prots %>% # this is needed to avoid overplotting
    dplyr::select(Label, up, down) %>%
    unique()
  
  
  path_sig <- HM_tibs_df_6 %>% # GET DETAILS FOR ANNOTATING SIG 
    filter(!(Label %in% c("HxA_EE_OLD_vs_SH_YNG", "HxA_EE_YNG_vs_SH_OLD"))) %>%
    filter(ID == hm) %>%
    dplyr::group_by(Label) %>% 
    mutate(sig_val = case_when(
      p.adjust <= 0.001 ~ "***",
      p.adjust <= 0.01 ~ "**",
      p.adjust <= 0.05 ~ "*",
      .default = "NA" )) # if bit there then NA (ns)
  
  
  # interesting... this step needed for some reason...
  path_sig_lab <-  path_sig %>% 
    dplyr::select(ID, Label, sig_val, NES) %>%
    unique() 
  
  my_lg_prot <- my_prots %>% #LABEL LEADING GENE???
    arrange(adj.pvalue) %>% # not quite - 6 most sig proteins
    slice_head(n = 6)
  
  
    strip <- strip_themed( # green for 1st 3 (housing) & Lavender for last 3 (age)
    background_x = elem_list_rect(fill = c(
      "#AFCC9F", "#D4E4CB", "#D4E4CB",
      "#BFBBD5", "#DDDCE9", "#DDDCE9")))

    
  # _____________________________________
  # 2 - MAKE THE VOLCANO PLOTS!!!
  vol_plot <- protein_comp_entrez %>%
    filter(!(Label %in% c("HxA_EE_OLD_vs_SH_YNG", "HxA_EE_YNG_vs_SH_OLD"))) %>%
    
    ggplot(aes(x = log2FC, y = -log10(adj.pvalue))) +
    
    
    # limits - rm 17 (but none are leading edge - shrunken for comparisons)
    scale_x_continuous(limits = c(-2,2),
                       expand = c(0, 0)) +
    scale_y_continuous(limits = c(0,6),
                       expand = c(0.04, 0)) +
    
    geom_hline(yintercept = -log10(0.05),
               linetype = "dashed", linewidth = 0.1,
               color = "grey50", alpha = 0.5) +
    geom_hline(yintercept = -log10(0.1),linetype = "dotted",
               linewidth = 0.1, color = "grey60", alpha = 0.5) +
    geom_vline(xintercept = c(0),linetype = "solid",
               linewidth = 0.1, color = "grey35", alpha = 0.5) +
    

    ggrastr::geom_point_rast(aes(fill = case_when( # conditional ass for color mapping
      adj.pvalue < 0.05 & log2FC > 0 ~ "up_adjp.05", # up sig .05
      adj.pvalue < 0.05 & log2FC < 0 ~ "down_adjp.05", # down sig .05
      adj.pvalue < 0.1 & log2FC > 0 ~ "up_adjp.1", # up 0.1
      adj.pvalue < 0.1 & log2FC < 0 ~ "down_adjp.1", # down .1
      TRUE ~ "ns" ),
      alpha = case_when( # conditional ass for ALPHA mapping
        adj.pvalue < 0.05 & log2FC > 0 ~ "up_adjp.05", # up sig .05
        adj.pvalue < 0.05 & log2FC < 0 ~ "down_adjp.05", # down sig .05
        adj.pvalue < 0.1 & log2FC > 0 ~ "up_adjp.1", # up 0.1
        adj.pvalue < 0.1 & log2FC < 0 ~ "down_adjp.1", # down .1
        TRUE ~ "ns" )), # catch the rest
       size = 1.75, shape = 21, stroke = 0.05,
      raster.dpi = getOption("ggrastr.default.dpi", 330)) +
    
    scale_fill_manual(values = c("firebrick", # coloring
                                  "royalblue3",
                                  "lightpink",
                                  "lightblue3",
                                  "grey80"),
                       breaks = c("up_adjp.05",
                                  "down_adjp.05",
                                  "up_adjp.1",
                                  "down_adjp.1",
                                  "ns"),
                       labels = c("\U2191 (adj.p <.05)",
                                  "\U2193 (adj.p <.05)",
                                  "\U2191 (adj.p <.1)",
                                  "\U2193 (adj.p <.1)",
                                  "ns")) +
    scale_alpha_manual( # alpha mapping
      values = c( # darker for more significant!
        up_adjp.05  = 0.3,
        down_adjp.05 = 0.3,
        up_adjp.1   = 0.3,
        down_adjp.1 = 0.3,
        ns = 0.1
      )) +
    
    
    
    geom_point_rast(data = my_prots, # ORGANE POINTS | ggrastr
                    aes(x = log2FC, y = -log10(adj.pvalue)),
                    alpha = .5, size = 1.75, shape = 21,
                    fill = "#E17C05FF", color = "#FCA842FF",
                    raster.dpi = getOption("ggrastr.default.dpi", 330)) + 
    
    geom_text_repel(data =  my_lg_prot, # TOP 6 MOST SIG ANNO
                    aes(x = log2FC, y = -log10(adj.pvalue),
                        label = gene,
                        color = log2FC > 0), # color mapping (see below)
                    max.overlaps = 30, size = 2.1, # gene text on points
                    min.segment.length = .5, segment.alpha = 0.3,
                    segment.size = 0.2, alpha = 0.9) +
                    # nudge_x = 0.155) +

    
    # NUM SIG UP & DOWN! | use sep df, overwise overplots
    geom_text(data = prot_sums, # up 
              aes(label = paste0(up, " \U2191")), 
              x = 1.35, y = 5.6, size = 3.2, color = "firebrick") + 
    geom_text(data = prot_sums, # down
              aes(label = paste0(down, " \U2193")),
              x = -1.4, y = 5.6, size = 3.2, color = "royalblue3") + 
    
    
    geom_text(data = path_sig_lab, # GSEA SIGNIFICANCE LABEL
              aes(label = paste0(sig_val),
                  y = 4.80,
                  x = ifelse(NES > 0, 1.525, -1.525), # side if neg or pos
                  color = NES > 0), # col - TRUE FALSE 
              size = 3.3) +
    scale_color_manual( # based on TRUE FALSE give color! 
      values = c(`TRUE` = "firebrick", `FALSE` = "royalblue3")) +
    
    
    
    
    labs(subtitle = paste0("MSigDB - ", my_title)) +
    
    
    # facet_wrap(~Label, nrow = 1) + 
    facet_wrap2(~Label, nrow = 1,
                strip = strip
                ) + 
    
    theme_bw() +
    theme(
      # strip.background.x = element_rect(fill = my_level_colors),
      legend.position = "none",
      plot.subtitle = element_text(size = 8, vjust = 0, color = "grey10"),
      axis.title = element_text(size =7, color = "grey10"),
      axis.text = element_text(size =6.5, color = "grey10"),
      axis.ticks = element_blank(),
      strip.text = element_text(size = 6.2, color = "grey10"),
      panel.grid = element_blank(),
      strip.background = element_rect(linewidth = 0.25, color = "grey10"),
      panel.border = element_rect(linewidth = 0.25, color = "grey10")
    )
  
  
  print(vol_plot)
  
  # _____________________________________
  # 3 - SAVE THE PLOTS
  Msig_HM_volcano_list[[hm]] <- vol_plot
  
  print(paste0("Done for ", hm, "! :D"))
  
}
```

### Plot & Export MSigDB Hallmarks Mapped Volcanos

```{r}
# plot_grid(plotlist = Msig_HM_volcano_list, ncol = 1)

# ggsave("2026_06_14_MsigDB_HM_9_volcanos_330dpi.png", dpi = 330) # 8.99 x 18.2
# dev.off()
```

```{r}
plot_grid(plotlist = Msig_HM_volcano_list[1], ncol = 1)
```

```{r}
# Metabolic - 3 - oxidative phos, cholesterol, fatty acid
plot_grid(plotlist = Msig_HM_volcano_list[c(1,3,8)], ncol = 1,
          labels = c("a", "b", "c"),
          label_colour = "grey10", label_size = 11)

# ggsave("2026_06_14_MsigDB_HM_3_volcano_metabolic_330dpi.png", dpi = 330) #8.21x595
# dev.off() 
# 2026_08_01 - 7.65 x 5.65 PDF | 2026_08_02 - 7.65 x 5.65 PDF v2 & v3 w/ ^ font
```

```{r}
# proliferation - 2 - E2F targets & Myc Targets 1 
plot_grid(plotlist = Msig_HM_volcano_list[c(2,5)], ncol = 1,
          labels = c("a", "b"),
          label_colour = "grey10", label_size = 11)

# ggsave("2026_06_14_MsigDB_HM_2_volcano_proliferation_330dpi.png", dpi = 330) #8.21x4
# dev.off()
# 2026_08_01 - 7.65 x 3.75 PDF | 2026_08_02 - 7.65 x 3.75 PDF v2
```

```{r}
# 4 - immume(inflamm, coagulat), development(adipogenesis), & signalling(androgen)
plot_grid(plotlist = Msig_HM_volcano_list[c(4,6,7,9)], ncol = 1,
          labels = c("a", "b", "c", "d"),
          label_colour = "grey10", label_size = 11)

# ggsave("2026_06_14_MsigDB_HM_4_volcano_immune_develop_signal_330dpi.png", dpi = 330) #8.21x7.84
# dev.off()
# 2026_08_01 - 7.65 x 7.5 PDF | 2026_08_02 - 7.65 x 7.5 PDF v2
```

## Hallmarks gt

```{r}
#TOP 10 most leading edge in sig hallmarks pathways 
HM_tibs_df_6_entrez_split_top_10 <- HM_tibs_df_6_entrez_split %>%
  group_by(Label, Description) %>%
  slice_head(n=10) %>%
  left_join(dplyr::select(rat_FASTA,
                          uniprot_acc, entry_name, gene),
            by = join_by(SYMBOL == gene)) %>%
  left_join(dplyr::select(rat_ref,
                   Entry,`Gene.Names..primary.`, `Protein.names`, 
                   `Protein.families`,Keywords, `Subcellular.location..CC.`,
                   Function..CC., Pathway,
       Gene.Ontology..cellular.component.:Gene.Ontology..molecular.function.,
                   `Developmental.stage`, Tissue.specificity),
            by = join_by(uniprot_acc == Entry)) %>%
  ungroup()
```

```{r}
HM_tibs_df_6_entrez_split_top_10_gt <- HM_tibs_df_6_entrez_split_top_10  %>%
  dplyr::select(Label, ID, Description, SYMBOL, setSize, NES, p.adjust, Category,
         uniprot_acc:Tissue.specificity) %>%
  gt(
    rowname_col = "Label", 
    groupname_col = "ID" 
  ) %>%
  fmt_number(columns = c(NES, p.adjust), decimals= 3) %>%
    tab_style(
    style = cell_text(weight = "bold",style = "italic"),
    locations = cells_body(columns = SYMBOL)
  ) %>%
  gtExtras::gt_theme_espn() %>%
  tab_options(table.font.size = px(10)) 


HM_tibs_df_6_entrez_split_top_10_gt
```

------------------------------------------------------------------------

## GO

```{r}
# MAP GO (nested nested purrr's map_dfr) to single df 
GO_tibs_df <- map_dfr(
  GO_gsea_EExAge_msstats_v3.2.2_t1_cp_v3.4.2_t2$GO_results_tib,
  ~map_dfr(.x, as_tibble, .id = "GO"),
  .id = "Label") # 1489 x 13 
```

- <https://blog.djnavarro.net/posts/2024-12-16_regex-backreferences/>
- <https://www.regular-expressions.info/>

```{r}
GO_tibs_df_6 <- GO_tibs_df %>%
  filter(!(Label %in% c("HxA_EE_OLD_vs_SH_YNG", "HxA_EE_YNG_vs_SH_OLD"))) %>% #1180
  mutate(top_core_enrichment =  gsub("^((.*?/){6}.*?)/.*", # trimmed 5 / + b4 6 /
                                     "\\1", # gets top 6 ENTREZ ID of core enrich
                                     core_enrichment))

GO_tibs_df_6$Label <- factor(GO_tibs_df_6$Label,
                             levels = c("H_EE_vs_SH",
                                        "HxA_EE_YNG_vs_SH_YNG",
                                        "HxA_EE_OLD_vs_SH_OLD",
                                        "A_YNG_vs_OLD", 
                                        "HxA_EE_YNG_vs_EE_OLD",
                                        "HxA_SH_YNG_vs_SH_OLD"))

# check for NAs 
colSums(is.na(GO_tibs_df)) # 0 NAs
colSums(is.na(GO_tibs_df_6)) #0  NAs ! 
```

```{r}
# Let's look @ the strongest leading edge genes.
GO_tibs_df_6_split <- GO_tibs_df_6 %>%
  mutate(top_core_enrich_split = str_split(top_core_enrichment, "/")) %>%
  unnest(top_core_enrich_split) %>%
  left_join(bitr_gn_2_entrez, # map 2 gene name via ENTREZID
            by = join_by(top_core_enrich_split == ENTREZID))
```

```{r}
# COUNTS OF TOP DRIVER GENES (LEADING EDGES in GO)
counts <- GO_tibs_df_6_split %>%
  group_by(Label, GO) %>%
  count(SYMBOL) %>%
  slice_max(n, n=10)
```

------------------------------------------------------------------------

- <https://bioconductor.posit.co/packages/devel/bioc/vignettes/topGO/inst/doc/topGO_manual.html#74_Visualising_the_GO_structure>

- <https://github.com/YuLab-SMU/enrichplot/tree/ce94b308d0ce5e835117bacd8858aadf2b2de9bc/R>

- <https://github.com/YuLab-SMU/clusterProfiler/issues/786>

```{r}
# Cluster Profiler's plotGOgraph is like topGO's plot (depends on topGO)
plotGOgraph(GO_gsea_EExAge_msstats_v3.2.2_t1_cp_v3.4.2_t2$GO_results$H_EE_vs_SH$BP,
            firstSigNodes = 10)
```

# TO CONTINUE - look \# cnetplot & ggtangle options to fix!

<https://github.com/YuLab-SMU/enrichplot/issues/322>

geom_edge(aes(linetype = relationship),

arrow = arrow(length = unit(2, 'mm')),

colour = "darkgrey", position = position, linewidth = 1) +

Simplified has more unique but arranges hierarchy a bit weirdly

20 is the sweet spot for full page appendix figure e.g. - HxA_EE_OLD_vs_SH_OLD\$CC

```{r}
# TEST PLOT 
set.seed(123)
# clusterProfiler's version of the plotGOgraph - goplot
# see v2 clusterProfiler script for modifying emapplot - same principles here!
go_p <- goplot(
  GO_gsea_EExAge_msstats_v3.2.2_t1_cp_v3.4.2_t2$GO_results$HxA_EE_OLD_vs_SH_OLD$MF,
    # GO_gsea_EExAge_cp_v3.4.2_t2_simplified$sim_GO_res$HxA_EE_OLD_vs_SH_OLD$CC,
  color = "NES",
  geom = "label",
  showCategory = 20,
  layout = igraph::layout_with_sugiyama, # default for DAG
  ID = "Description") +
  
  scale_fill_continuous_diverging("Blue-Red 2", name = "NES", alpha = .95,
                                  na.value = scales::alpha("grey90", 0.70)) +
  theme(
    legend.position = "bottom",
    legend.key.size = unit(0.3, "cm"),
    legend.title = element_text(size = 8, color = "grey10"),
    legend.text = element_text(size = 7, color = "grey10"),
    legend.margin = margin(-5,0,5,0),
    plot.background = element_rect(fill = "#DDE9D6", color = NA),
    plot.margin = margin(t = 0, r = 5, b = 0, l = 5, unit = "pt") # add space 2 margin
  ) 

# FINETUNING AESTHETICS
# lines & arrows
go_p$layers[[1]]$aes_params$linewidth <- 0.5

# geoms
go_p$layers[[2]]$aes_params$colour <- "grey10"
go_p$layers[[2]]$aes_params$size <- 3 # overwrite layer w/ size
# go_p$layers[[2]]$geom_params$max.overlaps <- 15
go_p$layers[[2]]$geom_params$box.padding <- 0 # gets rid of weird black lines?


# ## COLOR GO MF, BP, CC Label
# # matches node_data with _ e.g Molecular_Function, or Biological_Process
# root_row <- node_data[grepl("_", node_data$Term), ]
# 
# go_p <- go_p + # color separately with orange bg
#   annotate("label",
#            x = root_row$x, y = root_row$y,
#            label = root_row$Term, alpha = 0.9,
#            fill = "orange", size =3.9)

go_p

#args(goplot)
# go_p$layers[[1]] # segments/lines
# go_p$layers[[2]] # geom/box - get settings 
```

### GO Plot Loop

#### set up

```{r}
comp_colors <- c( # comparison colors
  "H_EE_vs_SH"  = "#D3E3CA", # HOUSING
  "HxA_EE_YNG_vs_SH_YNG" = "#E4EEDF",
  "HxA_EE_OLD_vs_SH_OLD" = "#E4EEDF",
  
  "A_YNG_vs_OLD" = "#DBD9E8", # AGE
    "HxA_EE_YNG_vs_EE_OLD" = "#EAE8F1",
    "HxA_SH_YNG_vs_SH_OLD" = "#EAE8F1",
  
  "HxA_EE_YNG_vs_SH_OLD" = "#FFFFFF", # extras
  "HxA_EE_OLD_vs_SH_YNG" = "#FFFFFF"
)

GP_comps <- names(GO_gsea_EExAge_msstats_v3.2.2_t1_cp_v3.4.2_t2$GO_results[1:6]) # list of comparisons 

ontologies <- c("BP", "CC","MF") # GO ontology types 

GO_plots_list <- list() #empty vector for go plots
```

#### loop

```{r}
for (comp in GP_comps) {
  for (ont in ontologies) {
    
    res <- GO_gsea_EExAge_msstats_v3.2.2_t1_cp_v3.4.2_t2$GO_results[[comp]][[ont]] # regular
    
    go_p <- tryCatch({
      p <- goplot(res, # GO PLOT 
        color = "NES", geom = "label", showCategory = 25) +
        
        scale_fill_continuous_diverging("Blue-Red 2", alpha = .95, name = "NES",  # fill color
                                        na.value = scales::alpha("grey90", 0.70)) +
        
      labs(subtitle = sprintf("GO DAG Plot (%s) - Top 25 from %s", ont, comp)) +

        theme(legend.position = "bottom", # THEME
              legend.key.size = unit(0.35, "cm"),
              legend.title = element_text(size = 7.5, color = "grey10"),
              legend.text = element_text(size = 7, color = "grey10"),
              legend.margin = margin(-5,0,5,0),
              plot.subtitle = element_text(size = 9, color = "grey10", face = "italic"),
              plot.background = element_rect(fill = comp_colors[[comp]], # compcol
                                             color = NA),
              plot.margin = margin(t = 0, r = 5, b = 0, l = 5, unit = "pt")) 
        
      # FINETUNING AESTHETICS
      # lines & arrows
      p$layers[[1]]$aes_params$linewidth <- 0.5
      # geoms
      p$layers[[2]]$aes_params$colour <- "grey10"
      p$layers[[2]]$aes_params$size <- 3 # overwrite layer w/ size
      p$layers[[2]]$geom_params$box.padding <- 0 # gets rid of weird blacklines?
      
      p # print plot 
      
      
    }, error = function(e) { # IF failed e.g. no
      message(sprintf("failed plot for %s: %s :(", comp, e$message))
      NULL
    })
    
    
    # SAVE TO LIST
    GO_plots_list[[comp]][[ont]] <- go_p
    
    if (!is.null(go_p)) { # EXPORT
    ggsave( #PDF
      filename = sprintf("%s_GOplot_DAG_GO_%s_%s_top25_8x12.pdf", Sys.Date(), comp, ont),
      plot=go_p, width=8, height=12, units="in")
    
    ggsave( # PNG
      filename = sprintf("%s_GOplot_DAG_GO_%s_%s_top25_8x12.png", Sys.Date(), comp, ont),
      plot=go_p, width=8, height=12, units="in")
    }
    
  }
}

# e.g. error message is failed plot for HxA_EE_YNG_vs_SH_OLD: Some vertex names in `d` are not listed in `vertices` :(
```

```{r}
plot_grid(GO_plots_list$H_EE_vs_SH$CC,GO_plots_list$H_EE_vs_SH$BP)
plot_grid(GO_plots_list$H_EE_YNG_vs_SH_YNG$CC,GO_plots_list$H_EE_YNG_vs_SH_YNG$BP)
plot_grid(GO_plots_list$A_YNG_vs_OLD$CC,GO_plots_list$A_YNG_vs_OLD$BP)
```

## Other plots

```{r}
enrichplot::manhattanplot(GO_gsea_EExAge_msstats_v3.2.2_t1_cp_v3.4.2_t2$GO_results$H_EE_vs_SH$MF)
```

## GO SIM

```{r}
# MAP GO (nested nested purrr's map_dfr) to single df 
GO_sim_tibs_df <- map_dfr(
  GO_gsea_EExAge_cp_v3.4.2_t2_simplified$sim_GO_res_tib, 
  ~map_dfr(.x, as_tibble, .id = "GO"),
  .id = "Label") # 690 x 13
```

```{r}
plotGOgraph(GO_gsea_EExAge_cp_v3.4.2_t2_simplified$sim_GO_res$A_YNG_vs_OLD$CC)
# EE_vs_SH MF one keeps giving errors (regardless of GO normal or simplified?!?!?!)
# YNG_vs_OLD BP gives error - 'snodes' contains nodes not in graph: ‘GO:0043604’
```

```{r}
GO_sim_tibs_df_6 <- GO_sim_tibs_df  %>%
  filter(!(Label %in% c("HxA_EE_OLD_vs_SH_YNG", "HxA_EE_YNG_vs_SH_OLD"))) %>% #1180
  mutate(top_core_enrichment =  gsub("^((.*?/){6}.*?)/.*", # trimmed 5 / + b4 6 /
                                     "\\1", # gets top 6 ENTREZ ID of core enrich
                                     core_enrichment))


GO_sim_tibs_df_6$Label <- factor(GO_sim_tibs_df_6$Label,
                             levels = c("H_EE_vs_SH",
                                        "HxA_EE_YNG_vs_SH_YNG",
                                        "HxA_EE_OLD_vs_SH_OLD",
                                        "A_YNG_vs_OLD", 
                                        "HxA_EE_YNG_vs_EE_OLD",
                                        "HxA_SH_YNG_vs_SH_OLD"))
```

```{r}
GO_sim_tibs_df_6_split  <- GO_sim_tibs_df_6 %>%
  mutate(top_core_enrich_split = str_split(top_core_enrichment, "/")) %>%
  unnest(top_core_enrich_split) %>%
  left_join(bitr_gn_2_entrez, # map 2 gene name via ENTREZID
            by = join_by(top_core_enrich_split == ENTREZID))
```

```{r}
# GO_CC_POI <- unique(GO_tibs_df_6$ID) # 705 omg! nvm too many 


GO_sim_tibs_df_6_split_brain <- GO_sim_tibs_df_6_split %>%
  filter(Label %in%  c("H_EE_vs_SH", "A_YNG_vs_OLD")) %>%
  filter(str_detect(Description, regex(
    # "axon|dendrite|schwann|myelin|glutamate|gaba|synap|dense|bouton|nerve|neuro|vesicle|paranodal",
    "proton-transporting two-sector ATPase complex|monoatomic ion channel complex|proton-transporting ATP synthase complex|proton-transporting two-sector ATPase complex|channel activity",
    ignore_case = TRUE))) #43

GO_ONT <- c("CC", "BP", "MF")
unique_brain <- unique(GO_sim_tibs_df_6_split_brain$ID)

GO_brain_volcano_list <- vector(mode = "list")
```

```{r}
for(go in unique_brain) {
  # HUH? error caused by capital GO?!?!, but go okay? weird! or conflict w/ GO col
  
  
  # 1 - PREPARE THE DATA
  my_genes <- GO_sim_tibs_df_6_split %>%
    filter(ID == go) %>%
    dplyr::distinct(SYMBOL) %>%
    tibble::deframe()
  
  my_title <- GO_sim_tibs_df_6_split %>%
    filter(ID == go) %>%
    dplyr::distinct(Description) %>%
    tibble::deframe()
  
  my_id <- GO_sim_tibs_df_6_split %>%
    filter(ID == go) %>%
    dplyr::distinct(ID) %>%
    tibble::deframe()
  
    my_go <- GO_sim_tibs_df_6_split %>%
    filter(ID == go) %>%
    dplyr::distinct(GO) %>%
    tibble::deframe()
  
  
  my_prots <- protein_comp_entrez %>% # from MSstatsTMT protein comps w ENTREZ map
    filter(!(Label %in% c("HxA_EE_OLD_vs_SH_YNG", "HxA_EE_YNG_vs_SH_OLD"))) %>%
    filter(gene %in% my_genes) %>%
    dplyr::group_by(Label) %>%
    mutate(up = sum(log2FC > 0),
           down = sum(log2FC < 0)) 
  
  prot_sums <- my_prots %>% # this is needed to avoid overplotting
    dplyr::select(Label, up, down) %>%
    unique()
  
  mapped_totals <- my_prots %>%  # for mapped total
    ungroup() %>%
    summarise(n_distinct(gene)) %>%
    tibble::deframe()
  
  gs_size <- GO_sim_tibs_df_6_split %>% # for geneset size
    filter(ID == go) %>%
    dplyr::distinct(setSize) %>%
    tibble::deframe()
  
  
  path_sig <- GO_sim_tibs_df_6 %>% # GET DETAILS FOR ANNOTATING SIG 
    filter(!(Label %in% c("HxA_EE_OLD_vs_SH_YNG", "HxA_EE_YNG_vs_SH_OLD"))) %>%
    filter(ID == go) %>%
    dplyr::group_by(Label) %>% 
    mutate(sig_val = case_when(
      p.adjust <= 0.001 ~ "***",
      p.adjust <= 0.01 ~ "**",
      p.adjust <= 0.05 ~ "*",
      .default = "NA" )) # if bit there then NA (ns)
  
  
  # interesting... this step needed for some reason...
  path_sig_lab <-  path_sig %>% 
    dplyr::select(ID, Label, sig_val, NES) %>%
    unique() 
  
  my_lg_prot <- my_prots %>% #LABEL LEADING GENE???
    arrange(adj.pvalue) %>% # not quite - 6 most sig proteins
    slice_head(n = 6)
  
  
    strip <- strip_themed( # green for 1st 3 (housing) & Lavender for last 3 (age)
    background_x = elem_list_rect(fill = c(
      "#AFCC9F", "#D4E4CB", "#D4E4CB",
      "#BFBBD5", "#DDDCE9", "#DDDCE9")))

    
  # _____________________________________
  # 2 - MAKE THE VOLCANO PLOTS!!!
  vol_plot <- protein_comp_entrez %>%
    filter(!(Label %in% c("HxA_EE_OLD_vs_SH_YNG", "HxA_EE_YNG_vs_SH_OLD"))) %>%
    
    ggplot(aes(x = log2FC, y = -log10(adj.pvalue))) +
    

    geom_hline(yintercept = -log10(0.05),
               linetype = "dashed", linewidth = 0.1,
               color = "grey30") +
    geom_hline(yintercept = -log10(0.1),linetype = "dotted",
               linewidth = 0.1, color = "grey45") +
    geom_vline(xintercept = c(0),linetype = "solid",
               linewidth = 0.1, color = "grey20") +

    
    ggrastr::geom_point_rast(aes(fill = case_when( # condit ass for color mapping
      adj.pvalue < 0.05 & log2FC > 0 ~ "up_adjp.05", # up sig .05
      adj.pvalue < 0.05 & log2FC < 0 ~ "down_adjp.05", # down sig .05
      adj.pvalue < 0.1 & log2FC > 0 ~ "up_adjp.1", # up 0.1
      adj.pvalue < 0.1 & log2FC < 0 ~ "down_adjp.1", # down .1
      TRUE ~ "ns" ),
      alpha = case_when( # conditional ass for ALPHA mapping
        adj.pvalue < 0.05 & log2FC > 0 ~ "up_adjp.05", # up sig .05
        adj.pvalue < 0.05 & log2FC < 0 ~ "down_adjp.05", # down sig .05
        adj.pvalue < 0.1 & log2FC > 0 ~ "up_adjp.1", # up 0.1
        adj.pvalue < 0.1 & log2FC < 0 ~ "down_adjp.1", # down .1
        TRUE ~ "ns" )), # catch the rest
       size = 1.7, shape = 21, stroke = 0.03,
      raster.dpi = getOption("ggrastr.default.dpi", 330)) +
    
    scale_fill_manual(values = c("firebrick", # coloring
                                  "royalblue3",
                                  "lightpink",
                                  "lightblue3",
                                  "grey80"),
                       breaks = c("up_adjp.05",
                                  "down_adjp.05",
                                  "up_adjp.1",
                                  "down_adjp.1",
                                  "ns"),
                       labels = c("\U2191 (adj.p <.05)",
                                  "\U2193 (adj.p <.05)",
                                  "\U2191 (adj.p <.1)",
                                  "\U2193 (adj.p <.1)",
                                  "ns")) +
    scale_alpha_manual( # alpha mapping
      values = c( # darker for more significant!
        up_adjp.05  = 0.3,
        down_adjp.05 = 0.3,
        up_adjp.1   = 0.3,
        down_adjp.1 = 0.3,
        ns = 0.1)) +
    
    
    geom_point_rast(data = my_prots, # ORGANE POINTS | ggrastr
                    aes(x = log2FC, y = -log10(adj.pvalue)),
                    alpha = .5, size = 1.75, shape = 21,
                    fill = "#E17C05FF", color = "#FCA842FF",
                    raster.dpi = getOption("ggrastr.default.dpi", 330)) + 
    
    geom_text_repel(data =  my_lg_prot, # TOP 6 MOST SIG ANNO
                    aes(x = log2FC, y = -log10(adj.pvalue),
                        label = gene,
                        color = log2FC > 0), # color mapping (see below)
                    max.overlaps = 30, size = 2.1, # gene text on points
                    min.segment.length = .5, segment.alpha = 0.3,
                    segment.size = 0.2, alpha = 0.9) +


    # NUM SIG UP & DOWN! | use sep df, overwise overplots
    geom_text(data = prot_sums, # up
              aes(label = paste0(up, " \U2191")),
              x = 1.4, y = 5.6, size = 3.2, color = "firebrick") +
    geom_text(data = prot_sums, # down
              aes(label = paste0(down, " \U2193")),
              x = -1.45, y = 5.6, size = 3.2, color = "royalblue3") +


    geom_text(data = path_sig_lab, # GSEA SIGNIFICANCE LABEL
              aes(label = paste0(sig_val),
                  y = 4.85,
                  x = ifelse(NES > 0, 1.525, -1.525), # side if neg or pos
                  color = NES > 0), # col - TRUE FALSE
              size = 3.3) +
    scale_color_manual( # based on TRUE FALSE give color!
      values = c(`TRUE` = "firebrick", `FALSE` = "royalblue3")) +
    
    
    # limits - rm 17 (but none are leading edge - shrunken for comparisons)
    scale_x_continuous(limits = c(-2,2),
                       expand = c(0, 0)) +
    scale_y_continuous(limits = c(0,6),
                       expand = c(0.04, 0)) +
    
    
    labs(subtitle = paste0("GO ", my_go, " - ", my_title, " (", my_id, ") (",
                           mapped_totals, "/", gs_size, ")")) +
    
    facet_wrap2(~Label, nrow = 1,
                strip = strip) + 
    
    theme_bw() +
    theme(
      legend.position = "none",
      plot.subtitle = element_text(size = 8, vjust = 0, color = "grey10"),
      axis.title = element_text(size =7, color = "grey10"),
      axis.text = element_text(size =6.5, color = "grey10"),
      axis.ticks = element_blank(),
      strip.text = element_text(size = 6.2, color = "grey10"),
      panel.grid = element_blank(),
      strip.background = element_rect(linewidth = 0.25, color = "grey10"),
      panel.border = element_rect(linewidth = 0.25, color = "grey10")
    )
  
  print(vol_plot)
  
  # _____________________________________
  # 3 - SAVE THE PLOTS
  GO_brain_volcano_list[[go]] <- vol_plot
  
  print(paste0("Done for ", go, "! :D"))
  
}
```

```{r}
plot_grid(plotlist = GO_brain_volcano_list[37:43],
          ncol = 2)
```

```{r}
plot_grid(plotlist = GO_brain_volcano_list[1:4],
          ncol = 1)
```
