---
title: "BioIn_2025_09_v3.3.2_ClusterProfiler.Rmd"
author: "Stephanie Huang"
date: "2025-09-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ClusterProfiler (BioIn Test 2)

## Details & Data Used

## Links & Resources

-   

## Setup

Install if needed

```{r}
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# 
# BiocManager::install("clusterProfiler")
# BiocManager::install("msigdbr")
# BiocManager::install("org.Rn.eg.db")
# BiocManager::install("enrichplot")
# BiocManager::install("pathview")
```

Load Packages

```{r message=FALSE, warning=FALSE}
library(tidyverse) # for dplyr, ggplot, stringr, readr etc 
library(clusterProfiler) #GSEA includes gsea, enrichplot etc.
library(org.Rn.eg.db) # Rat genome annotation package - incl. AnnotationDbi
library(msigdbr) # MSigDB gene sets
library(enrichplot) # for dotplot()
library(ggridges) # for ridgeplot()
library(pathview) # pathview() - Kegg pathway visualisation
```

## GSEA

### Prepare Protein List into Ranked Gene list

#### Import Protein List (MSstatsTMT comparison output)

```{r}
# Set working directory to source file location

# read in protein list df from MSstatsTMT comparison
df <- read_csv(
  "input/protein_lists/2025_09_12_rp_1pspg_test_pairwise_msstats_v3.1.5_t8.csv")
```

##### lil checks

```{r}
# lil checks
head(df) # peak
str(df) # structure | 7746 x 9
colnames(df) # column names

# lil checks (bit messy)
unique(df$issue) # if = na in issue column then good. if not exclude
df %>% filter(str_detect(Protein, "Cont_")) # rows with contaminants - to exlude
df %>% filter(str_detect(log2FC, "Inf")) # rows with Infinity /no log fold - to exclude
df %>% filter(is.na(pvalue)) # same observation as the "Inf" in log2FC
df %>% filter(is.na(adj.pvalue)) # same observation as the "Inf" in log2FC
```

#### Clean up df & rename

```{r}
protein_comp <- df %>%
  filter(is.na(issue)) %>% # keeps rows with issue == na (so comparisons that are ok)
  filter(!str_detect(Protein, "Cont_")) %>% # remove contaminant proteins
  filter(!str_detect(log2FC, "Inf")) %>% # remove infinity/missing fold change values
  dplyr::select(Protein:adj.pvalue) # rm weird ...1 col & issue col, skip this? 

# check 
head(protein_comp) 
str(protein_comp) # 7724 x 7
colnames(protein_comp)
```

Convert UniProt Accessions into EntrezID using bitr()

-   [bitr: Biological ID Translator](https://yulab-smu.top/biomedical-knowledge-mining-book/useful-utilities.html)

-   Note Entrez id is a central gene identifier - so want to map uniprot accession with gene ID which then can be mapped to other IDs.

    -   Side note - DO NOT ANNOTATE WITH MULTIPLE TOTYPE e.g. toType = c("ENTREZID", "SYMBOL", "GENENAME"), will result in duplication errors since various gene names or symbols get mapped on?

```{r}
keytypes(org.Rn.eg.db) # key types available

# 7158/7724 mapped (8.05% missing)
entrezID <- bitr(protein_comp$Protein, # input df
             fromType = "UNIPROT", # input type
             toType = "ENTREZID", # output type
             OrgDb = org.Rn.eg.db) # annotation db
             
# UNIPROT -> ENTREZID - 8.05% of input gene failed to map
# UNIPROT -> ENSEMBL 12.91% of input gene IDs are fail to map...
# UNIPROT -> REFSEQ - 8.05% of input gene IDs are fail to map...
# UNIPROT -> SYMBOL -  8.05% of input gene IDs are fail to map...
```

Merge EntrezID onto protein list

```{r}
protein_genes <- merge(entrezID, protein_comp,
                       by.x = "UNIPROT",
                       by.y = "Protein")

head(protein_genes)
dim(protein_genes)
```

```{r}
#write out?
```

Prepare Ranked Gene List

```{r}
# use sign(df$log2fc)*(-log10(df$pval)) + BSS & PNNL

# using 
geneList_df <- protein_genes %>%
  mutate(ranking = sign(protein_genes$log2FC)*(-log10(protein_genes$pvalue))) %>%
  group_by(ENTREZID) %>%
  summarise(ranking = mean(ranking, na.rm = TRUE)) %>%
  arrange(-ranking)

# can export 

dim(geneList_df) # 7153 x 2 - same as prev test! 

# pvalue & log2FC  min = -7.6191 max 4.559
# adj.pvalue * log2FC = max 1.3938 & min -3.7302

# ~ ~ ~ ~ ~ ~ ~ 
# convert into a named vector to supply to fgsea()
geneList <- tibble::deframe(geneList_df) 

head(geneList)
tail(geneList)
plot(geneList)
sum(is.na(geneList)) # 0 | without extra line (not sure why previous version had NA)
```

### Running GSEA via clusterProfiler

### msigbdr

-   [Introduction to msigdbr](https://igordot.github.io/msigdbr/articles/msigdbr-intro.html)

    -   [species & orthologs](https://igordot.github.io/msigdbr/articles/msigdbr-intro.html#species) - has human & mouse but can rat orthologs & use.

        -   TO CHECK/DECIDE - Using mouse or human genesets?

            -   nvm only allows `species =  "Mus musculus"` if you have `db_species = "MM"` ?

    -   [Pathway enrichment analysis](https://igordot.github.io/msigdbr/articles/msigdbr-intro.html#pathway-enrichment-analysis) (uses enricher() from clusterProfiler or fgsea())

-   [Package 'msigdbr' (July 2025)](https://cran.r-project.org/web/packages/msigdbr/index.html)

-   [12 Universal enrichment analysis](https://yulab-smu.top/biomedical-knowledge-mining-book/universal-api.html) (ClusterProfiler)

    -   [12.3 MSigDb analysis](https://yulab-smu.top/biomedical-knowledge-mining-book/universal-api.html#msigdb-analysis)

    -   [**12.3.2 MSigDb gene set enrichment analysis**](https://yulab-smu.top/biomedical-knowledge-mining-book/universal-api.html#msigdb-gsea) **using `GSEA()`**

    -   A bit more detailed <https://uclouvain-cbio.github.io/WSBIM2122/sec-gsea.html#using-the-clusterprofiler-package>

-   [Gene Set Enrichment Analysis with ClusterProfiler (NYU NGS)](https://learn.gencore.bio.nyu.edu/rna-seq-analysis/gene-set-enrichment-analysis/)

    -   lots of useful stuff on plots

```{r}
# # msigdbr helper functions
# msigdbr_species() # 20 species
# msigdbr_collections() # 33730584 bytes?
# unique(msigdbr_df$db_version) # check ver
# citation("msigdbr") # how to cite
```

#### Hallmark Collection (H)

```{r}
# Preps for fgsea()  

# # example
# # Get CGP (chemical and genetic perturbations) gene sets with genes mapped to rat
# orthologs gs <- msigdbr(species = "Rattus norvegicus",
#                         collection = "C2",
#                         subcollection = "CGP")

# pull in hallmark genesets from msigdbr, maps rat orthologs to human geneset 
hallmark_gs_df <- msigdbr(species = "Rattus norvegicus", # my species
                             collection = "H") # Hallmark collection
                             
head(hallmark_gs_df) 
dim(hallmark_gs_df) # 7294 x 23

# ~ ~ ~ ~ ~ ~ ~ 
# Convert in 2 name list format required for fgsea()
# Splits in separate lists containing Entrez grouped by the geneset name (gs_name)
# ENTREZID is ncbi_gene?
hallmark_genesets <-  hallmark_gs_df %>%
  split(x = .$ncbi_gene, # containing - df/vector
        f = .$gs_name) # file - factor

#str(hallmark_genesets) # list of 50
```

```{r}
# Preps TERM2GENE for GSEA() | 1st col = term_id/geneset & 2nd col = mapped gene
msig_h <- msigdbr(species = "Rattus norvegicus",
                  collection = "H") %>%
 dplyr::select(gs_name, ncbi_gene) # get only geneset name & corresponding ENTREZID

head(msig_h)
dim(msig_h) #  7294 x 2
```

Running gsea with GSEA()

```{r}
# using GSEA() from clusterProfiler

# uses fgseaMultilevel() function? | auto 
GSEA_hallmark <- GSEA(geneList = geneList, # ranked gene list
                      TERM2GENE = msig_h, # genesets
                      by = 'fgsea' # method - fgsea or DOSE
                      )

# Output has lots of extra info! results + genesets, Genelist, input params! -> save rda
# can make into tibble for easy viewing 
GSEA_hallmark_tibble <- as_tibble(GSEA_hallmark)
```

Plots

```{r fig.height=6, fig.width=9}
enrichplot::dotplot(GSEA_hallmark,
                    showCategory = 15,
                    title = "Housing - EE vs. SH (t8 - t2)")
```

```{r}
dotplot(GSEA_hallmark,
        split = ".sign", # split by sign - positive vs negative enrichment score
        title = "Housing - EE vs. SH (t8 - t2) - split by sign",
        font.size = 11) +  
  facet_grid(.~.sign) # split plots by sign
```

GSEA plot using gseaplot() from enrichplot

-   Look into making into a function!

```{r fig.height=6, fig.width=9}
# GSEA plot - positive - most sig [1] - Hallmark_oxidative_phosphorylation
gseaplot(GSEA_hallmark, 
         geneSetID = 1, # 1st geneset is 1
         by = "all", # running score or position
         title = GSEA_hallmark$Description[1])
```

```{r fig.height=6, fig.width=9}
# GSEA plot - positive - most sig [2] - Hallmark_MYC_targets_v1
gseaplot(GSEA_hallmark, 
         geneSetID = 2, # 2nd geneset is 2
         by = "all", # running score or position
         title = GSEA_hallmark$Description[2])
```

Ridgeplot

```{r fig.height=6, fig.width=9}
ridgeplot(GSEA_hallmark) + labs(x = "enrichment distribution")
```

#### C5 Ontology gene sets (GO)

```{r}
# Preps TERM2GENE for GSEA() | 1st col = term_id/geneset & 2nd col = mapped gene
msig_C5_CC <- msigdbr(species = "Rattus norvegicus",
                      collection = "C5", # C5 - ontology set
                      subcollection = "CC") %>% # GO cellular component
  dplyr::select(gs_name, ncbi_gene)

dim(msig_C5_CC) # 98011 x 2 for C5 CC!!!
```

```{r}
# using GSEA() from clusterProfiler
# uses fgseaMultilevel() function? | auto 
GSEA_msig_C5_CC <- GSEA(geneList = geneList, # ranked gene list
                      TERM2GENE = msig_C5_CC, # genesets
                      #organism = "rno",
                      by = 'fgsea') # method - fgsea or DOSE

# had to restart R to fix error | worked after - see warnings.                      
                      

# Output has lots of extra info! results + genesets, Genelist, input params! -> save rda
# can make into tibble for easy viewing 
GSEA_msig_C5_CC_tibble <- as_tibble(GSEA_msig_C5_CC)
```

```{r fig.height=27, fig.width=11}
# 102 significant beyond adj pvalue 0.05 
# neuronal & synaptic ones activated in EE. phew! 
dotplot(GSEA_msig_C5_CC,
        split = ".sign", # split by pos vs neg enrichment score
        showCategory = 102,
        font.size = 11,
        label_format = 55,
        title = "Housing - EE vs. SH (t8 - t2) - GSEA_msig_C5_CC") +
  facet_grid(.~.sign) # split plots by sign
```

```{r}
# simplify
sim_msig_go_cc <- pairwise_termsim(GSEA_msig_C5_CC)

```

#### NOTE TO SELF - so many other curated genesets available on MSigDB!!! Look into!

### GO via gseGO

-   [**6 GO enrichment analysis**](https://yulab-smu.top/biomedical-knowledge-mining-book/clusterprofiler-go.html#clusterprofiler-go)

    -   [6.4 GO Gene Set Enrichment Analysis](https://yulab-smu.top/biomedical-knowledge-mining-book/clusterprofiler-go.html#go-gene-set-enrichment-analysis)

-   [17.4 GO utilities](https://yulab-smu.top/biomedical-knowledge-mining-book/useful-utilities.html#go-utilities)

    -   [**17.4.3 reduce redundancy of enriched GO terms via `simplfy()`**](https://yulab-smu.top/biomedical-knowledge-mining-book/useful-utilities.html#reduce-redundancy-of-enriched-go-terms)

-   [**enrichplot pairwise_termsim()**](https://rdrr.io/bioc/enrichplot/man/pairwise_termsim.html)

#### gseGO CC

```{r}
gseGO_cc <- gseGO(geneList = geneList, 
             ont ="CC", # cell component
             keyType = "ENTREZID",
             OrgDb = org.Rn.eg.db, # if you unset it uses bg based on provide geneList?
             by = "fgsea",
             minGSSize = 10, 
             maxGSSize = 500, 
             pvalueCutoff = 0.05, 
             verbose = TRUE, 
             pAdjustMethod = "BH")

# can make into tibble for easy viewing 
gseGO_cc_tibble <- as_tibble(gseGO_cc) 
```

```{r fig.height=18, fig.width=10}
# 102 significant beyond adj pvalue 0.05 
enrichplot::dotplot(gseGO_cc,
                    split  = ".sign", # split by + vs - enrirchment score
                    showCategory = 102,
                    font.size = 11,
                    label_format = 55,
                    title = "Housing - EE vs. SH (t8 - t2) - gseGO_CC"
                    ) +
  facet_grid(.~.sign) # split plots by sign

```

```{r fig.height=9, fig.width=10}
# 6.6 Visualize enriched GO terms as a directed acyclic graph
# goplot requires ggarchery | plot induced GO DAG of significant terms
goplot(gseGO_cc,
       showCategory = 25, # default 10
       layout = "sugiyama") # default
```

```{r}
# Gene-Concept Network with cnetplot()

```

```{r}
# 15.4 Heatmap-like functional classification

```

```{r fig.height=15, fig.width=12}
#15.5 Tree plot
# The treeplot() function performs hierarchical clustering of enriched terms. 
# Relies on the pairwise similarities of the enriched terms calc by the pairwise_termsim() func, which by default using Jaccard’s similarity index (JC).

gseGO_cc_pwts <- pairwise_termsim(gseGO_cc)

treeplot(gseGO_cc_pwts,
         hclust_method = "average", # see hclust () documentation on methods
         showCategory = 70, # default 30
         nCluster = 10, # default 5 | parent term changes a lot & weird assign at times
         )

# works but unstable assignment - look into
```

```{r fig.height=6, fig.width=7}
# 15.6 Enrichment Map 
# also use gseGO_cc_pwts <- pairwise_termsim(gseGO_cc) from above

emapplot(gseGO_cc_pwts, 
         showCategory = 30, # default 30
         # cex_category = 1.5, # can be used to resize nodes | not in ?emapplot?
         # layout = "kk" # change layout | not in ?emapplot? nvm old doesn't accept?
         )
```

```{r fig.height=7, fig.width=7}
# simplfy
sim_gseGO_cc <- simplify(gseGO_cc_pwts,
                         cutoff=0.7,
                         by="p.adjust",
                         select_fun=min)

# plot simplified
emapplot(sim_gseGO_cc)
```

#### NOTE TO SELF - Look at the other plots & what are appropriate presentation guidelines

#### gseGO MF

```{r}
gseGO_mf <- gseGO(geneList = geneList, 
             ont ="MF", # molecular function
             keyType = "ENTREZID",
             OrgDb = org.Rn.eg.db, # if you unset it uses bg based on provide geneList?
             by = "fgsea",
             minGSSize = 10, 
             maxGSSize = 500, 
             pvalueCutoff = 0.05, 
             verbose = TRUE, 
             pAdjustMethod = "BH")

# can make into tibble for easy viewing 
gseGO_mf_tibble <- as_tibble(gseGO_mf) 
```

```{r fig.height=15, fig.width=8}
# 63 significant beyond adj pvalue 0.05 
enrichplot::dotplot(gseGO_mf,
                    split  = ".sign", # split by + vs - enrirchment score
                    showCategory = 65,
                    font.size = 11,
                    label_format = 55,
                    title = "Housing - EE vs. SH (t8 - t2) - gseGO_mf"
                    ) +
  facet_grid(.~.sign) # split plots by sign
```

```{r fig.height=20, fig.width=12}
#15.5 Tree plot
gseGO_mf_pwts <- pairwise_termsim(gseGO_mf)

treeplot(gseGO_mf_pwts,
         hclust_method = "average", # see hclust () documentation on methods
         showCategory = 65, # default 30
         nCluster = 20) # default 5 | parent term changes a lot & weird assign at times
         

# works but unstable assignment - look into
```

```{r fig.height=7, fig.width=7}
sim_gseGO_mf <- simplify(gseGO_mf_pwts,
                         cutoff=0.7,
                         by="p.adjust",
                         select_fun=min)
# plot simplified
emapplot(sim_gseGO_mf)
```

```{r fig.height=8, fig.width=8}
goplot(gseGO_mf,
       showCategory = 25, # default 10
       layout = "sugiyama") # default
```

#### gseGO BP

```{r}
gseGO_bp <- gseGO(geneList = geneList, 
             ont ="BP", # Biological process
             keyType = "ENTREZID",
             OrgDb = org.Rn.eg.db, # if you unset it uses bg based on provide geneList?
             by = "fgsea",
             minGSSize = 10, 
             maxGSSize = 500, 
             pvalueCutoff = 0.05, 
             verbose = TRUE, 
             pAdjustMethod = "BH")

# can make into tibble for easy viewing 
gseGO_bp_tibble <- as_tibble(gseGO_bp) 
```

```{r fig.height=35, fig.width=9}
# 176 significant beyond adj pvalue 0.05 
enrichplot::dotplot(gseGO_bp,
                    split  = ".sign", # split by + vs - enrirchment score
                    showCategory = 176,
                    font.size = 11,
                    label_format = 80,
                    title = "Housing - EE vs. SH (t8 - t2) - gseGO_bp"
                    ) +
  facet_grid(.~.sign) # split plots by sign
```

```{r fig.height=18, fig.width=12}
#15.5 Tree plot
gseGO_bp_pwts <- pairwise_termsim(gseGO_bp)

treeplot(gseGO_bp_pwts,
         hclust_method = "average", # see hclust () documentation on methods
         showCategory = 176, # default 30
         nCluster = 20, # default 5 | parent term changes a lot & weird assign at times
         )

# works but unstable assignment - look into
```

```{r fig.height=7, fig.width=7}
sim_gseGO_bp <- simplify(gseGO_bp_pwts,
                         cutoff=0.7,
                         by="p.adjust",
                         select_fun=min)
# plot simplified
emapplot(sim_gseGO_cc)
```

\~ \~ \~ \~ \~ \~ \~ \~ \~

### KEGG via gseKEGG

-   [7 KEGG enrichment analysis](https://yulab-smu.top/biomedical-knowledge-mining-book/clusterprofiler-kegg.html#clusterprofiler-kegg)

    -   [7.3 KEGG pathway gene set enrichment analysis](https://yulab-smu.top/biomedical-knowledge-mining-book/clusterprofiler-kegg.html#clusterprofiler-kegg-pathway-gsea)

    -   [7.5 KEGG module gene set enrichment analysis](https://yulab-smu.top/biomedical-knowledge-mining-book/clusterprofiler-kegg.html#clusterprofiler-kegg-module-gsea)

    -   [7.6 Visualize enriched KEGG pathways](https://yulab-smu.top/biomedical-knowledge-mining-book/clusterprofiler-kegg.html#visualize-enriched-kegg-pathways) -

```{r}
# look at rat info kegg
rat <- search_kegg_organism('rno', by='kegg_code')
head(rat)
```

```{r}
gseKEGG_r <- gseKEGG(geneList = geneList, 
                     organism = 'rno', #RAT
                     keyType = "kegg", 
                     minGSSize = 10, # min gene set size
                     maxGSSize = 500, # max gene set size
                     pvalueCutoff = 0.05, # pvalue cutoff
                     pAdjustMethod = 'BH', #BejaminiHoc Correciton
                     verbose = TRUE) # print message or not

gse_kegg_tibble <- as_tibble(gseKEGG_r) # 26 sig or 23?? (0.05)| 32 if adjp = 0.1
```

```{r fig.height=9, fig.width=7}
enrichplot::dotplot(gseKEGG_r,
                    split = ".sign",  # split by + vs - enrirchment score
                    showCategory = 30,
                    title = "Housing - EE vs. SH (t8 - t2) - gseKEGG_r") +
  facet_grid(.~.sign) # split plot by seign

```

#### Visualise enriched KEGG pathways via pathview() & browseKEGG()

-   [KEGG Rat Pathways](https://www.kegg.jp/kegg-bin/show_organism?menu_type=pathway_maps&org=rno)

Glutamatergic synapse - Rattus norvegicus (rat)

```{r}
# view browseKEGG from clusterProfiler
# launches to web when function run! 
# browseKEGG(gseKEGG_r,
#            pathID = 'rno04724') # glutamtergic synapse rat
```

KEGG - Organismal Systems \> Nervous System (10 pathways)

```{r}
# via pathview::pathview()

# rno04724 <- pathview(gene.data = geneList,
#                      pathway.id = 'rno04724', # Glutamatergic synapse - Rat
#                      species = 'rno',
#                      limit = list(gene=max(abs(geneList)), cpd=1))

# writes out png & xml files to source file location
# look into how to plot & also save as PDF
```

```{r}
# rno04723 - Retrograde endocannabinoid signaling

# rno04723 <- pathview(gene.data = geneList,
#                      pathway.id = 'rno04723',
#                      species = 'rno',
#                      limit = list(gene=max(abs(geneList)), cpd=1))

# gives error ???
```

```{r}
# rno05016 - Huntington disease | runs okay!
# rno05016 <- pathview(gene.data = geneList,
#                      pathway.id = 'rno05016',
#                      species = 'rno',
#                      limit = list(gene=max(abs(geneList)), cpd=1))
```

```{r}
# rno04722 - neurotrophin_signaling
# rno04722 <- pathview(gene.data = geneList,
#                      pathway.id = 'rno04722',
#                      species = 'rno',
#                      limit = list(gene=max(abs(geneList)), cpd=1))
```

```{r}
# rno04720 - Long Term Potentiation (LTP)
# rno04720 <- pathview(gene.data = geneList,
#                      pathway.id = 'rno04720',
#                      species = 'rno',
#                      limit = list(gene=max(abs(geneList)), cpd=1))
```

```{r}
# rno05012 - Parkinsons
# rno05012 <- pathview(gene.data = geneList,
#                      pathway.id = 'rno05012',
#                      species = 'rno',
#                      limit = list(gene=max(abs(geneList)), cpd=1))
```

```{r}
# rno04082 - Neuroactive ligand signaling | nice holistic view of many neuroligand paths
# rno04082 <- pathview(gene.data = geneList,
#                      pathway.id = 'rno04082',
#                      species = 'rno',
#                      limit = list(gene=max(abs(geneList)), cpd=1))
```

```{r}
# rno05033 - Nicotine Addiction | not much just ampa & gaba
# rno05033 <- pathview(gene.data = geneList,
#                      pathway.id = 'rno05033',
#                      species = 'rno',
#                      limit = list(gene=max(abs(geneList)), cpd=1))
```

\~ \~ \~

```{r}
# rno04721 - Synaptic_vesicle_cycle | most up regulated in EE 
# rno04721 <- pathview(gene.data = geneList,
#                      pathway.id = 'rno04721',
#                      species = 'rno',
#                      limit = list(gene=max(abs(geneList)), cpd=1))
```

```{r}
# rno04727 - GABAergic_synapse_ 
# rno04727 <- pathview(gene.data = geneList,
#                      pathway.id = 'rno04727',
#                      species = 'rno',
#                      limit = list(gene=max(abs(geneList)), cpd=1))
```

```{r}
# rno04728 - Dompaminergic_synapse_
# rno04728 <- pathview(gene.data = geneList,
#                      pathway.id = 'rno04728',
#                      species = 'rno',
#                      limit = list(gene=max(abs(geneList)), cpd=1))
```

```{r}
# rno04726 - Serotonergic_synapse_
# rno04726 <- pathview(gene.data = geneList,
#                      pathway.id = 'rno04726',
#                      species = 'rno',
#                      limit = list(gene=max(abs(geneList)), cpd=1))
```

```{r}
# rno04725 - Cholinergic_synapse_ | Interesting seems 2 be mediating synaptic plasticity
# rno04725 <- pathview(gene.data = geneList,
#                      pathway.id = 'rno04725',
#                      species = 'rno',
#                      limit = list(gene=max(abs(geneList)), cpd=1))
```

```{r}
# rno04730 - LTD_ | cerebellum pathview? | otherwise interesting
# rno04730 <- pathview(gene.data = geneList,
#                      pathway.id = 'rno04730',
#                      species = 'rno',
#                      limit = list(gene=max(abs(geneList)), cpd=1))
```

Environmental Information Processing \> Signal Transduction

```{r}
# rno04010 - MAPK_signaling_ | Interesting - cell proliferation & differentiation
# rno04010 <- pathview(gene.data = geneList,
#                      pathway.id = 'rno04010',
#                      species = 'rno',
#                      limit = list(gene=max(abs(geneList)), cpd=1))
```

```{r}
# rno04014 - Ras_signaling_
# rno04014 <- pathview(gene.data = geneList,
#                      pathway.id = 'rno04014',
#                      species = 'rno',
#                      limit = list(gene=max(abs(geneList)), cpd=1))
```

Most significant Pathways

1 - Oxidative Phosphorylation (rno00190) - Metabolism \>Energy Metabolism \> Ox Pho

```{r}
# rno00190 - Oxidative_phosphorylation_ | several highly upregulated
# rno00190 <- pathview(gene.data = geneList,
#                      pathway.id = 'rno00190',
#                      species = 'rno',
#                      limit = list(gene=max(abs(geneList)), cpd=1))
```

2 rno05415 Diabetic_cardiomyopathy - Human Diseases; Cardiovascular Disease

```{r}
# rno05415 - Diabetic_cardiomyopathy_ | upreg @ mitocondrial - what does this mean?
# rno05415 <- pathview(gene.data = geneList,
#                      pathway.id = 'rno05415',
#                      species = 'rno',
#                      limit = list(gene=max(abs(geneList)), cpd=1))
```

3 - Non-alcoholic fatty liver disease rno04932 - Endocrine & metabolic disease

```{r}
# # rno04932 - Non-alcoholic_fatty_liver_d_ | driven by cx1,3,4 mito resp chain
# rno04932 <- pathview(gene.data = geneList,
#                      pathway.id = 'rno04932',
#                      species = 'rno',
#                      limit = list(gene=max(abs(geneList)), cpd=1))
```

4 - Thermogenesis\_ (rno04714) - adipose tissue & keep warm \| fat SH rats & skinny EE

```{r}
# rno04714 - Thermogenesis_ | mitochondrial + others - look into
# rno04714 <- pathview(gene.data = geneList,
#                      pathway.id = 'rno04714',
#                      species = 'rno',
#                      limit = list(gene=max(abs(geneList)), cpd=1))
```

9 - Pathways of neurodegeneration - multiple diseases (rno05022) (neurodegen disea

```{r}
# rno05022 - Pathways_of_neurodegeneration_multi_ |interesting/confusing
# rno05022 <- pathview(gene.data = geneList,
#                      pathway.id = 'rno05022',
#                      species = 'rno',
#                      limit = list(gene=max(abs(geneList)), cpd=1))
```

Not enriched but interesting

```{r}
# #rno04081 - Hormone_signaling_ | various guanine nucleotide binding pro & PKA down
# # and only SST (somatostatin) & Opioid (Penk) Up | also only brain specific up
# rno04081 <- pathview(gene.data = geneList,
#                      pathway.id = 'rno04081',
#                      species = 'rno',
#                      limit = list(gene=max(abs(geneList)), cpd=1))
```

```{r}
# rno04080 - Neuroactive_ligand_receptor_interaction_ | not many but interesting - look
# rno04080 <- pathview(gene.data = geneList,
#                      pathway.id = 'rno04080',
#                      species = 'rno',
#                      limit = list(gene=max(abs(geneList)), cpd=1))
```

```{r}
# rno04060 - Cytokine_cytokine_receptor_interaction_ | no up or down
# rno04060 <- pathview(gene.data = geneList,
#                      pathway.id = 'rno04060',
#                      species = 'rno',
#                      limit = list(gene=max(abs(geneList)), cpd=1))
```

```{r}
# rno04512 - ECM_receptor_interaction_ | all down 
# rno04512 <- pathview(gene.data = geneList,
#                      pathway.id = 'rno04512',
#                      species = 'rno',
#                      limit = list(gene=max(abs(geneList)), cpd=1))
```

```{r}
# rno04514 - Cell_adhesion_molecules_ | no up or down except CXADR down
# rno04514 <- pathview(gene.data = geneList,
#                      pathway.id = 'rno04514',
#                      species = 'rno',
#                      limit = list(gene=max(abs(geneList)), cpd=1))
```

```{r}
# rno04360 - Axon_guidance_ | most down except Ras
# rno04360 <- pathview(gene.data = geneList,
#                      pathway.id = 'rno04360',
#                      species = 'rno',
#                      limit = list(gene=max(abs(geneList)), cpd=1))
```

\~ \~ \~ \~ \~ \~ \~ \~ \~ \~ \~ \~

#### KEGG Module GSEA via gseMKEGG()

```{r}
gseMKEGG_r <- gseMKEGG(geneList = geneList,
                       organism = 'rno',
                       pvalueCutoff = 1)

head(gseMKEGG_r)
gseMKEGG_r_tibble <- as_tibble(gseMKEGG_r) # 17 sig @ adj-pval .05? | why spec as 1 abv?
```

```{r fig.height=9, fig.width=8}
# im confused - loot at gseMKEGG function later
enrichplot::dotplot(gseMKEGG_r,
                    split = ".sign",  # split by + vs - enrirchment score
                    showCategory = 30,
                    title = "Housing - EE vs. SH (t8 - t2) - gseMKEGG_r") +
  facet_grid(.~.sign) # split plot by seign
```
