---
title: "BioIn_2025_12_v3.4.2_ClusterProfiler"
author: "Stephanie Huang"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ClusterProfiler BioIn - 1pspg v3.2.2 EE x Age t1

*for MSstatsTMT 1pspg v3.2.2 EE x Age t1 -(defaults) with B1 129N removed (TMT channel with too much loaded)*

## Set up

### Packages

```{r}
# # Install if needed 
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# 
# BiocManager::install("clusterProfiler")
# BiocManager::install("msigdbr")
# BiocManager::install("org.Rn.eg.db")
# BiocManager::install("enrichplot")
# BiocManager::install("pathview")
```

```{r}
# Load Packages 
library(tidyverse) # for dplyr, ggplot, stringr, readr etc 
library(clusterProfiler) #GSEA includes gsea, enrichplot etc.
library(org.Rn.eg.db) # Rat genome anno pack- incl. AnnotationDbi
library(msigdbr) # MSigDB gene sets
library(enrichplot) # for dotplot()
library(ggridges) # for ridgeplot()
library(pathview) # pathview() - Kegg pathway visualisation
```

### Import data

Load the mapped comparisons output generated from

```{r}
# mapped_comp_v3.2.2_EExAge_t1
load("input/protein_lists/2025_12_31_1pspg_v3.2.2_t1_EExAge_mapped_comp.rda")

df <- mapped_comp_v3.2.2_EExAge_t1
```

lil checks

```{r}
# lil checks
head(df) # peak
str(df) # structure | 7746 x 9
colnames(df) # column names

# lil checks (bit messy)
unique(df$issue) # if = na in issue column then good. if not exclude
check_1 <- df %>% filter(str_detect(log2FC, "Inf")) # 5 inf 
check_2 <- df %>% filter(str_detect(Protein, "Cont")) # 0 (since already removed)
unique(df$Label) 
length(unique(df$Label)) #8
```

### **Clean up df & rename**

```{r}
protein_comp <- df %>%
  filter(is.na(issue)) %>% # keeps rows with issue == na (so comparisons that are ok)
  filter(!str_detect(log2FC, "Inf")) # remove infinity/missing fold change values

# check 
head(protein_comp) 
colnames(protein_comp)

```

## **KEGG GSEA via gseKEGG with for loops**

based of for loop made for the Lichti 2014 reanalysis

### Setup containers & prepare info for sequences 

```{r}
# check 
unique(df$Label) 
# don't want "HxA_EE_YNG_vs_SH_OLD" "HxA_EE_OLD_vs_SH_YNG", so just do manually 

# VECTOR for labels (contrasts from MSstatsTMT analysis)
label_names <- c("H_EE_vs_SH", "A_YNG_vs_OLD",
                 "HxA_EE_YNG_vs_EE_OLD", "HxA_SH_YNG_vs_SH_OLD",
                 "HxA_EE_YNG_vs_SH_YNG", "HxA_EE_OLD_vs_SH_OLD")

# CONTAINER - EMPTY VECTOR
gsea_KEGG_results_list <- vector(mode = "list")
```

### Operation

```{r}
for (lb in label_names) {
  
  # PART 1
  geneList_working_df <- protein_comp %>%
    filter(Label == lb) %>% 
    mutate(ranking = sign(log2FC)*(-log10(pvalue))) %>%
    group_by(Protein) %>%
    summarise(ranking = mean(ranking, na.rm = TRUE)) %>%
    arrange(-ranking)
  
  # next deframe into a named vector needed to supply 2 fgsea()
  geneList_vec_working <- tibble::deframe(geneList_working_df)
  
  # store in list 
  gsea_KEGG_results_list$geneList_vec[[lb]] <- geneList_vec_working
  

  # PART 2 - KEGG GSEA
  # run KEGG gsea with clusterProfiler
  kegg_res_working <- clusterProfiler::gseKEGG(geneList = geneList_vec_working, # ALL DEFAULTS
                              organism = 'rno',
                              keyType = 'uniprot', # using UNIPROT directly
                              by = 'fgsea',
                              pvalueCutoff = 0.05, 
                              pAdjustMethod = 'BH', #BejaminiHoc Correciton
                              minGSSize = 10, 
                              maxGSSize = 500,
                              eps = 1e-10
                              )
  
  # save kegg result original to list vec
  gsea_KEGG_results_list$kegg_results[[lb]] <- kegg_res_working
  
  # not needed here so compute & save directly? or can do anyway? idk?
  gsea_KEGG_results_list$kegg_results_tib[[lb]] <- tibble::as_tibble(kegg_res_working) 
  
  
  # PART 3 - PLOTS
  # Make dotplots | this one plots & saves properly!
  # 3.1 - regular dot blot
  dp <- enrichplot::dotplot(kegg_res_working,
                            showCategory = 30,
                            font.size = 11,
                            title = paste0(stringr::str_glue(
                              'Housing x Age - {lb} - 1pspg MSstatsTMT (v3.2.2) 2025_12_31 - gseKEGG_r'))) 

  
  print(dp)   # print dot plots 
  gsea_KEGG_results_list$dotplots[[lb]] <- dp # save to list 
  
  # 3.2 - dot plot split by sign
  dp_sign <- enrichplot::dotplot(kegg_res_working,
                            split = '.sign',
                            showCategory = 30,
                            font.size = 11,
                            title = paste0(stringr::str_glue(
                              'Housing x Age - {lb} - 1pspg MSstatsTMT (v3.2.2) 2025_12_31 - gseKEGG_r - by sign'))) +
    facet_grid(.~.sign)
  
  print(dp_sign) # print dot plots 
  gsea_KEGG_results_list$dotplots_sign[[lb]] <- dp_sign # save to list
  
}
```

```{r}
# horrific name but at least clear...
KEGG_gsea_EExAge_cp_v3.4.2_msstats_v3.2.2_t1 <- gsea_KEGG_results_list

# save(KEGG_gsea_EExAge_cp_v3.4.2_msstats_v3.2.2_t1,
#      file = "output/2025_12_31_KEGG_gsea_EExAge_cp_v3.4.2_msstats_v3.2.2_t1.rda")
```

------------------------------------------------------------------------

## GO GSEA via gseGO with nested for loop

based of for loop made for the Lichti 2014 reanalysis

### **Setup containers & prepare info for sequences**

```{r}
# # ALREADY DONE ABOVE 
# # VECTOR for labels (contrasts from MSstatsTMT analysis)
# label_names <- c("H_EE_vs_SH", "A_YNG_vs_OLD",
#                  "HxA_EE_YNG_vs_EE_OLD", "HxA_SH_YNG_vs_SH_OLD",
#                  "HxA_EE_YNG_vs_SH_YNG", "HxA_EE_OLD_vs_SH_OLD")

GO_names <- c('BP', 'CC', 'MF', 'ALL') # my GO NAMES 

# CONTAINER - EMPTY VECTOR
gsea_GO_results_lists <- vector(mode = "list")
```

### **Operation**

```{r}
for (lb in label_names){ 
  
 # PART 1
  geneList_working_df_go <- protein_comp %>%
    filter(Label == lb) %>% 
    mutate(ranking = sign(log2FC)*(-log10(pvalue))) %>%
    group_by(Protein) %>%
    summarise(ranking = mean(ranking, na.rm = TRUE)) %>%
    arrange(-ranking)
  
 
  # Part 1.2
  # next, deframe into a named vector needed to supply to fgsea()
  geneList_vec_working_go <- tibble::deframe(geneList_working_df_go)
  
  geneList_vec_working_go <- geneList_vec_working_go[!is.na(names(geneList_vec_working_go))]
  
  # store in list 
  gsea_GO_results_lists$geneList_vec[[lb]] <- geneList_vec_working_go
  
  # Part 1.3 | plots fine but does not save?
  # Plot ranked gene List
  gL_plot_go <- plot(geneList_vec_working_go, main = paste0(lb, '- Ranked gene list'))
  
  print(gL_plot_go)
  
  gsea_GO_results_lists$geneList_plots[[lb]] <- gL_plot_go
  
  
  
  # PART 2 
  # nested for loop
  # so now for each label, go through all 3 go terms (BP, CC, MF)
  for (go in GO_names){ 
    
    # operation
    # run 
    GO_res_working <- clusterProfiler::gseGO(geneList = geneList_vec_working_go,
                                             OrgDb = org.Rn.eg.db, 
                                             keyType = "UNIPROT",
                                             ont = go,
                                             exponent = 1, # default settings for all 
                                              minGSSize = 10,
                                              maxGSSize = 500,
                                              eps = 1e-10,
                                              pvalueCutoff = 0.05,
                                              pAdjustMethod = "BH",
                                              by = "fgsea"
                                             )
    
  # save GO result original to list vec
  gsea_GO_results_lists$GO_results[[lb]][[go]] <- GO_res_working
    
    
  # extra (not really needed but save anyway
  gsea_GO_results_lists$GO_results_tib[[lb]][[go]] <- tibble::as_tibble(GO_res_working)
    

  # PART 3 - if else for dot plot to deal with any nas
  if (nrow(GO_res_working@result) > 0 ) {
    
    #Make dot plots | this one saves properly
    # Part 3.1 - regular dot plots 
    dp_go <- enrichplot::dotplot(GO_res_working,
                              showCategory = 30,
                              font.size = 11,
                              title = paste0(stringr::str_glue(
                                'Housing x Age - {lb} - {go} - 1pspg MSstatsTMT (v3.2.2) 2025_12_31 - gseGO_r'
                              )))
    print(dp_go) # print dot plots
    gsea_GO_results_lists$dotplots[[lb]][[go]] <- dp_go # save to list
    
    #Make sign dot plot 
    # Part 3.2 - regular dot plots 
    dp_go_sign <- enrichplot::dotplot(GO_res_working,
                              showCategory = 30,
                              font.size = 11,
                              split = '.sign',
                              title = paste0(stringr::str_glue(
                              'Housing x Age - {lb} - {go} - 1pspg MSstatsTMT (v3.2.2) 2025_12_31 - gseGO_r - by sign'
                              )))
    print(dp_go_sign) # print dot plots
    gsea_GO_results_lists$dotplots_sign[[lb]][[go]] <- dp_go_sign # save to list
    
  } else{print(paste0("No GO gsea result for dot plot in ", lb, "-", go, " :("))}

  
  
  } 
  
  
}
```

```{r}
# horrific name but at least clear...
GO_gsea_EExAge_cp_v3.4.2_msstats_v3.2.2_t1 <- gsea_GO_results_lists

save(GO_gsea_EExAge_cp_v3.4.2_msstats_v3.2.2_t1,
     file = "output/2025_12_31_GO_gsea_EExAge_cp_v3.4.2_msstats_v3.2.2_t1.rda")
```

------------------------------------------------------------------------
