---
title: "2026_02_mss_3.1.2_sp_pairwise_vis"
author: "Stephanie Huang"
date: "2026-02-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Quick graphs for pairwise counts & pairwise volcano plots!

2026_02_18 (v1 - drafts) & 2026_07_24-27 /(v2)- for PhD write up

Contents

- **Wrangling/Data**

- 

  ## **Plots**

  - Volcano Plots

***General notes***

1.  Sex/Batch - there are lots of "oneConditionMissing" in the NA column, these are largely from running across years & the gradient issue.
2.  

## Setup & Load

```{r}
library(tidyverse)
library(readr)
library(ggrepel) #  for geom_text_repel()
library(colorspace)
library(cowplot)
library(gt)
library(ggrastr) # https://tomsing1.github.io/blog/posts/volcano/
```

------------------------------------------------------------------------

If loading in raw data from MSstatsTMT

```{r eval=FALSE, include=FALSE}
# fasta used to search (includes sp)
up_fasta <- read_csv("input/db/UP000002494_10116_Rat_Ref_Proteome_22403_1pspg_2025_04_UPfastadf.csv")
```

```{r eval=FALSE, include=FALSE}
# full uniprot 295 cols data
load("input/db/rat_ref_proteome_2025_12_13_51872x295.rda")
```

```{r eval=FALSE, include=FALSE}
rat_ref_trm <- rat_ref_proteome_2025_12_13 %>%
  select(Entry,
         Protein.families, Pathway, Keywords, Developmental.stage,
         Induction, Function..CC., Subcellular.location..CC.)
```

```{r eval=FALSE, include=FALSE}
pw_litter <- read_csv("output/MSstatsTMT/comp/2025_12_31_sp_test_pairwise_msstats_v3.1.2_t6_litter.csv")
pw_sex <- read_csv("output/MSstatsTMT/comp/2025_12_31_sp_test_pairwise_msstats_v3.1.2_t5_sex.csv")
pw_beh <- read_csv("output/MSstatsTMT/comp/2025_12_31_sp_test_pairwise_msstats_v3.1.2_t4_behaviour.csv")
pw_cohort <- read_csv("output/MSstatsTMT/comp/2025_12_31_sp_test_pairwise_msstats_v3.1.2_t3_cohort.csv")
pw_housing <- read_csv("output/MSstatsTMT/comp/2025_12_31_sp_test_pairwise_msstats_v3.1.2_t1_housing.csv")
```

### Merge Data

```{r eval=FALSE, include=FALSE}
# combined df of all conditions
com_df <- bind_rows(list(Housing = pw_housing,
                         Sex = pw_sex,
                         Cohort = pw_cohort,
                         ExpH = pw_beh,
                         Litter = pw_litter),
                    .id = "pairwise.comp") %>%
  left_join(up_fasta,
            by = join_by(Protein == uniprot_acc))
  
```

```{r eval=FALSE, include=FALSE}
# # WRITE OUT MERGED & MAPPED! (note - raw & unfiltered!)
# write_excel_csv(com_df,
#                 "output/2025_12_31_sp_test_pairwise_msstats_v3.1.2_all_merge_map.csv")
# save(com_df,
#      file = "output/2025_12_31_sp_test_pairwise_msstats_v3.1.2_all_merge_map.rda")
```

------------------------------------------------------------------------

### Resume/Start Here (loads in output from above)

```{r}
# Load com_df in here, if needed. 
load("input/MSstatsTMT/comp/2025_12_31_sp_test_pairwise_msstats_v3.1.2_all_merge_map.rda")

```

### com_df_mapped - calculations df for volcano plot annotations & tally tab ano

```{r}
com_df_mapped <- com_df %>%
  
  # Filtering 
  filter(is.na(issue)) %>% # no issues  
  filter(is.numeric(adj.pvalue)) %>%  
  filter(!is.na(adj.pvalue)) %>% # fixes NA issue? 
  select(!c(...1, issue)) %>% # rm extra cols
  
  #___FOR_VOLCANO_ANNOTATION_______________
  # Make Gene Label for Sig
  mutate(sig_g_label = if_else(adj.pvalue <= .05, # text label if up to adj.pvlue 
                               gene, NA)) %>%
  # Make one for no of sig up & down for each comaprison group/label
  group_by(Label) %>%
  mutate(Direction = if_else(log2FC > 0, "Up", "Down")) %>%
  mutate(sig_no_up_lab = sum(adj.pvalue < 0.05 & log2FC > 0), # up
         sig_no_down_lab = sum(adj.pvalue < 0.05 & log2FC < 0)) %>%  #  down
  
  #___FOR_TALLY_TABLE_&_NUMBER_REPORTED______________
  # Make Tally/combined 
  mutate(sig_no = sig_no_up_lab + sig_no_down_lab, # total sig per label
         smodel_no = n(), # no.sucessful model proteins- label -since issue rm abv
         model_no = 4562, # no_modelled (no removals)
         smodel_mis_no = 4562 - n()) %>% 
  ungroup()

colnames(com_df_mapped)
```

### Sig counts summary table (34 rows)

<https://stackoverflow.com/questions/56810473/uniting-many-rows-into-one-r>

```{r}
# sig counts table 
sig_counts <- com_df_mapped %>%
  select(pairwise.comp, Label, sig_no_up_lab:smodel_mis_no) %>%
  group_by(pairwise.comp, Label) %>% # not needed?
  unique() %>%
  mutate(smodel_per = round(smodel_no/model_no*100,2), # per of successful modelled
         smodel_mis_no_per = round(smodel_mis_no/model_no*100,2), # per missing
         sig_per = round(sig_no/model_no*100,2), #
         sig_per_up = round(sig_no_up_lab/model_no*100,2), #
         sig_per_down = round(sig_no_down_lab/model_no*100,2)) #

sig_counts
colnames(sig_counts)

sig_counts$pairwise.comp <- factor(sig_counts$pairwise.comp,
                                   levels = c("Housing", "ExpH", "Cohort",
                                              "Litter", "Sex"))

```

### Sig genes summary table (34 rows)

<https://stackoverflow.com/questions/56810473/uniting-many-rows-into-one-r>

```{r}
# sig genes
sig_genes <- com_df_mapped %>%
  select(pairwise.comp, Label, gene, sig_g_label,
         Direction, log2FC, adj.pvalue) %>%
  group_by(pairwise.comp, Label) %>% # not needed?
  slice_min(adj.pvalue, n = 10) %>%
  ungroup() %>%
  group_by(pairwise.comp, Label, Direction) %>%
  summarize(sum_genes = paste(sig_g_label[!is.na(sig_g_label)], # ignores NAs!
                              collapse = "/")) %>%
  pivot_wider(names_from = Direction, values_from = sum_genes)

```

### sig_genes_count summmary table (combined)

```{r}
sig_genes_count <- left_join(sig_counts, sig_genes,
                             by = join_by(pairwise.comp == pairwise.comp,
                                          Label == Label))

sig_genes_count$pairwise.comp <- factor(sig_genes_count$pairwise.comp,
                                   levels = c("Housing", "ExpH", "Cohort",
                                              "Litter", "Sex"))
```

### gt

<https://gt.rstudio.com/reference/tab_style.html#examples>

```{r}
sig_genes_count_gt <- sig_genes_count %>%
  select(pairwise.comp, Label,
         sig_no, sig_per,
         sig_no_up_lab, sig_per_up, Up,
         sig_no_down_lab, sig_per_down, Down,
         # smodel_no, smodel_per,
         smodel_mis_no) %>% # org cols

  gt(
    groupname_col = "pairwise.comp",
    rowname_col = "Label") %>%
  
  # gtExtras::gt_theme_nytimes() %>%
  opt_vertical_padding(scale = 0.35) %>% # vert padding
  cols_align(align = "left") %>% # align left
  
  # tab_header(title = "Preliminary Pairwise Comparisons Summary Table (MSstatsTMT)",
  #            subtitle = md("`2026_07_27` - Breakdown of significant proteins (adj.p ≤.05) + Top 10 most significant proteins(/genes) split by into \u2191 up & \u2193 down-regulated. | Also includes number of missing (NA#) (e.g. `oneConditionMissing`) out of 4562 modelled proteins (Swiss-Prot) - from MSstatsTMT (2025_12_31_sp_test_pairwise_msstats_v3.1.2).")) %>%

   tab_options(# GENERAL TABLE OPTIONS
    table.border.top.width = px(2.5), # very top of table
    table.border.bottom.width = px(2.5), # very bottom of table |not very visible
    table.border.bottom.color = "#BDBDBD",
    column_labels.border.top.width = px(2.5), # top of column labs |not very vis
    column_labels.border.top.color = "#BDBDBD", 
    column_labels.border.bottom.width = px(2.5), # top of column labs |not very vis
    column_labels.border.bottom.color = "#BDBDBD", 
    heading.border.bottom.width = px(2.5), # more visible than above
    heading.border.bottom.color = "#BDBDBD", # more visible than above
    table.font.size = "11.5px",
     stub.font.size = "12px",
    row_group.font.weight = "bold",
    column_labels.font.weight = "bolder"
  ) %>%
  
  cols_merge( # sig no & percentage
    columns = c(sig_no, sig_per),
    pattern = "{1} ({2}%)") %>%
  
  cols_merge( # ↑ sig 
    columns = c(sig_no_up_lab, sig_per_up),
    pattern = "{1} ({2}%)") %>%
  
  
  cols_merge( # ↓ sig 
    columns = c(sig_no_down_lab, sig_per_down),
    pattern = "{1} ({2}%)") %>%
  
    # cols_merge(# modelled 
    # columns = c(smodel_no, smodel_per),
    # pattern = "{1} ({2}%)") %>%
  
   cols_label( # relabel columns
    sig_no = "# Sig. ",
    sig_no_up_lab = "# Sig.  \u2191",
    sig_no_down_lab = "# Sig. \u2193",
    smodel_mis_no = "NA#") %>%
  
  
 tab_style( # Prevent breaking of value columns!
    style = cell_text(whitespace = "nowrap"),
    locations = list(cells_stub(),
                     cells_body(columns = c(sig_no,sig_no_up_lab,
                                            sig_no_down_lab)))) %>%
  
 tab_style(
    style = list(
      cell_text(color = "#5e5e5e")),
    locations = list(
      cells_body(columns = c(smodel_mis_no)),
      cells_column_labels(columns = c(smodel_mis_no)))) %>%
  
  
  tab_style(style = list(
    cell_text(style = "italic", size = "9.5px")),
  locations = cells_body(
    columns = c(Up, Down))) %>%
  
  tab_style(
    style = list(
      # cell_text(color = "#d33682")),
      cell_text(color = "#aa344d")),
    locations = list(
      cells_body(columns = c(sig_no_up_lab)),
      cells_column_labels(columns = c(Up,sig_no_up_lab)))) %>%
  
  tab_style(
    style = list(
      cell_text(color = "#344daa")),
    locations = list(
      cells_body(columns = c(sig_no_down_lab)),
      cells_column_labels(columns = c(Down,sig_no_down_lab)))) 
  

sig_genes_count_gt
```

### Count of Significant proteins by pairwise comparison

```{r}
sig_counts_labels <- sig_counts %>%
  mutate(pairwise.comp = recode(pairwise.comp, Housing = "Hous")) %>%
  dplyr::select(Label, sig_no) %>%
  unique() 
# hcl_palettes(plot = TRUE)


count_plot <- sig_counts %>%
  pivot_longer(cols = c("sig_no_up_lab", "sig_no_down_lab"),
               names_to = "direction",
               values_to = "no") %>%
  mutate(pairwise.comp = recode(pairwise.comp, Housing = "Hous")) %>%
  
  # filter(sig == 1) %>%
  ggplot() +
   geom_col(aes(x = reorder(Label, sig_no), y = no, fill = direction),
            stat = "identity", # so col sam wid
    width = .75, color = "grey15", linewidth = 0.3) +
  
  geom_bar(aes(x = reorder(Label, sig_no), y = sig_no/2), stat = "identity",
           fill= "grey90", alpha = 0.05) + # just for mapping to anchor text!
  geom_text(data = sig_counts_labels, # needs geom_bar above to anchor
            aes(label = sig_no, x = reorder(Label, sig_no),  y = sig_no),
            vjust = -0.4, size = 2.7) +

  
  scale_fill_manual(name = "Direction (Log2FC)",
                    values = alpha(c("#aa344d","#2753BF"),0.4), 
                    breaks = c("sig_no_up_lab", "sig_no_down_lab"),
                    labels = c("\u2191 Upregulated", "\u2193 Downregulated")) +
  
  facet_wrap(~pairwise.comp, nrow = 1, scales = 'free_x', space = 'free_x') +
  theme_bw() +
  labs(y = "No. of Significant Proteins (adj.pvalue ≤ .05)") +
  theme(
    legend.position = "bottom",
    legend.title = element_text(size = 9.5, color="grey15", face = "italic"),
     legend.text = element_text(size = 9, color="grey15"),
    axis.text.x = element_text(angle = 90, hjust =1, vjust = 0.5, face = "italic"),
     legend.margin = margin(0,0,0,0),  # space between plot & legend
    legend.key.size = unit(0.5, "cm"), # size of legend key icon
    axis.title.x = element_blank(),
    axis.title.y = element_text(color ="grey15", size = 9.5),
    axis.ticks = element_line(linewidth = .3),
    strip.text = element_text(size = 9),
    strip.background = element_rect(linewidth = .3),
    panel.border = element_rect(linewidth = 0.3),
    # panel.grid.major = element_blank(),
    # panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "grey95", linewidth = 0.1),
    panel.grid.minor = element_line(color = "grey97", linewidth = 0.1),
    panel.spacing.x = unit(0.1, "cm"), # horizontal spacing - COHORT 
  )

count_plot
```

------------------------------------------------------------------------

```{r}
# top_10_genes
top_20_genes <- com_df_mapped %>%
  select(pairwise.comp, Label, gene, sig_g_label,
         Direction, log2FC, adj.pvalue) %>%
  group_by(pairwise.comp, Label) %>% # not needed?
  slice_min(adj.pvalue, n = 20, with_ties = FALSE) %>%
  ungroup() %>%

  group_by(pairwise.comp, Label, Direction) %>%
  summarize(sum_genes = paste(gene[!is.na(gene)], # ignores NAs!
                              collapse = "/")) %>%
  pivot_wider(names_from = Direction, values_from = sum_genes) %>%




  gt(groupname_col = "pairwise.comp",
    rowname_col = "Label") %>%
     tab_options(# GENERAL TABLE OPTIONS
    table.border.top.width = px(2.5), # very top of table
    table.border.bottom.width = px(2.5), # very bottom of table |not very visible
    table.border.bottom.color = "#BDBDBD",
    column_labels.border.top.width = px(2.5), # top of column labs |not very vis
    column_labels.border.top.color = "#BDBDBD",
    column_labels.border.bottom.width = px(2.5), # top of column labs |not very vis
    column_labels.border.bottom.color = "#BDBDBD",
    heading.border.bottom.width = px(2.5), # more visible than above
    heading.border.bottom.color = "#BDBDBD", # more visible than above
    table.font.size = "11.5px",
     stub.font.size = "12px",
    row_group.font.weight = "bold",
    column_labels.font.weight = "bolder"
  ) #%>%

top_20_genes

# OLD
# #write_excel_csv(top_200_genes, "output/2026_07_27_test_top_200_prelim_check.csv")

```

### Cohort Comparison Geom Tile Heatmap

<https://rpubs.com/Shefali/ggplot-scale-function>

```{r}
# COHORT COMPARISON HEATMAP TILE (2026_07_28)

# sig_in_3a_vs_3b <- com_df_mapped %>%
#   filter(Label == "3a vs 3b" & adj.pvalue <=0.15) %>% 
#   pull(gene)

sig_in_cohorts <- com_df_mapped %>%
  filter(pairwise.comp == "Cohort" & adj.pvalue <=0.1) %>% # .05 - 223 | .1 - 396
  count(gene) %>%
  filter(n >= 2) %>% # .05 - 22 (in 2 or more)| .1 - 54
  pull(gene)


cohorts_comp_plot <- com_df_mapped %>%
  filter(pairwise.comp == "Cohort") %>%
  filter(gene %in% c(sig_in_cohorts)) %>%
  # filter(gene %in% c(sig_in_3a_vs_3b)) %>%
  # filter(adj.pvalue <= 0.2) %>%
  
  ggplot(aes(x = Label, y = reorder(gene, -adj.pvalue), fill= log2FC,
             alpha = adj.pvalue)) + 
  geom_tile(color = "grey10") +
  geom_text(aes(label = paste0("log2FC = ",round(log2FC,2),
                               # ", se = ",round(SE,2),
                               ", adj.p = ",round(adj.pvalue,3))),
            size = 2, fontface = "italic") +
  scale_x_discrete(expand = expansion(mult = c(0.05,0.05)),
                   labels = c("3a vs 3b" = "3a (~111 vs 3b (~113)", 
                              "4 vs 3a" = "4 (~86) vs 3a (~111)", 
                              "4 vs 3b" = "4 (~86) vs 3b (~113)")) +
  labs(title = "Overlap Across Cohorts", 
       subtitle = "Proteins @ adj.pval ≤ .1 (396) & present in 2 or more cohorts (\u2192 54 unique in 2 or more)") +
  colorspace::scale_fill_continuous_diverging("Blue-Red 2", mid = 0) +
  scale_alpha(range = c(1,0.75), guide = "none") + # more transparent less sig
  theme_void() +
  theme_minimal() +
  theme(
    legend.position = "none",
     axis.title = element_blank(),
    # axis.text.x = element_text("grey10", size = 11),
    axis.text.y = element_text(hjust = 1, face =  "italic")
  )

cohorts_comp_plot
```

```{r}
##Checks between ExpH & Housing
# top_50_pval <- com_df_mapped %>%
#  filter(pairwise.comp %in% c("Housing", "ExpH")) %>%
# group_by(pairwise.comp, Label) %>% # not needed?
# slice_min(pvalue, n = 50, with_ties = FALSE) %>%
# slice_min(log2FC, n = 25, with_ties = FALSE) %>%
# filter(gene %in% c('Hcfc2', 'Tmc4', 'Apoc2	', 'Abcc9', # very non-sig no overlap!
#                    'Gapdhs', 'Mucl3', 'Osm', 'Pnrc2', 'Slc12a3')) #%>%
# ungroup() %>%
# count(gene)
```

------------------------------------------------------------------------

# Volcano Plots

### No Litter volcano plots

```{r}
# make new df for the geom_text bit below.
# prev had from 1 df but over plots repeately & cant index since mapped_comp huge.
sig_volcano_sum_no_litter <- com_df_mapped %>%
  filter(pairwise.comp == "Cohort") %>%
  # filter(pairwise.comp %in% c("Housing", "ExpH")) %>%
  # filter(pairwise.comp == "Sex") %>%
  dplyr::select(Label, sig_no_up_lab, sig_no_down_lab) %>%
  unique() 
  
#big_volcano_no_lit <- com_df_mapped %>%
# big_volcano_no_lit_EE <- com_df_mapped %>%
big_volcano_no_lit_C <- com_df_mapped %>%
  
  
  filter(pairwise.comp == "Cohort") %>%
  # filter(pairwise.comp %in% c("Housing", "ExpH")) %>%
  # filter(pairwise.comp == "Sex") %>%
  ggplot(aes(x = log2FC, y = -log10(adj.pvalue), label = sig_g_label)) +
  
  
  #vline & hlines
  geom_hline(yintercept = -log10(0.05),
             linetype = "dashed", linewidth = 0.1,
             color = "grey30") +
  annotate("text", x = 1.80, y = -log10(0.05),
           label = "adj.p <.05 \n",
           fontface = 'italic', size = 1.5, color = "grey10") +
  geom_hline(yintercept = -log10(0.1),linetype = "dotted",
             linewidth = 0.15, color = "grey50") +
  annotate("text", x = 1.75, y = -log10(0.1),
           label = " adj.p <.1 \n",
           fontface = 'italic', size = 1.5, color = "grey30") +
  geom_vline(xintercept = c(0),linetype = "solid",
             linewidth = 0.15, color = "grey85") +
  
  # POINTS
  ggrastr::geom_point_rast(aes(color = case_when( # conditional ass for color mapping
    adj.pvalue < 0.05 & log2FC > 0 ~ "up_adjp.05", # up sig .05
    adj.pvalue < 0.05 & log2FC < 0 ~ "down_adjp.05", # down sig .05
    adj.pvalue < 0.1 & log2FC > 0 ~ "up_adjp.1", # up 0.1
    adj.pvalue < 0.1 & log2FC < 0 ~ "down_adjp.1", # down .1
    TRUE ~ "ns" ), alpha = adj.pvalue), # catch the rest
    size = 1.7,
    raster.dpi = getOption("ggrastr.default.dpi", 330)) +
  
  scale_color_manual(values = c("firebrick", # coloring
                                "royalblue3",
                                "lightpink",
                                "lightblue3",
                                "grey80"),
                     breaks = c("up_adjp.05",
                                "down_adjp.05",
                                "up_adjp.1",
                                "down_adjp.1"),
                     labels = c("\U2191 (adj.p <.05)",
                                "\U2193 (adj.p <.05)",
                                "\U2191 (adj.p <.1)",
                                "\U2193 (adj.p <.1)")) +
  

  scale_alpha(range = c(0.55,0.05), guide = "none") + # more transparent less sig
  
  geom_text_repel(max.overlaps = 20, size = 3, # gene text on points
                  color = "grey10",  segment.size	= 0.2,
                  segment.alpha = 0.5, min.segment.length = 0.5) + 

               
  
  # NUM SIG UP & DOWN! | use sep df, overwise overplots
  geom_text(data = sig_volcano_sum_no_litter, # up 
            aes(label = paste0(sig_no_up_lab, " \U2191")),
            # x = 1.78, y = 9.7, size = 4, color = "firebrick") +  # sex
            # x = 1.8, y = 4.35, size = 4.2, color = "firebrick") + # Hous & ExpH
            x = 1.8, y = 6.65, size = 4.2, color = "firebrick") + # Hous & ExpH
  
  geom_text(data = sig_volcano_sum_no_litter, # down
            aes(label = paste0(sig_no_down_lab, " \U2193")),
            # x = -1.78, y = 9.7, size = 4, color = "royalblue3") +  # sex
            # x = -1.8, y = 4.35, size = 4.2, color = "royalblue3") + # Hous & ExpH
             x = -1.8, y = 6.65, size = 4.2, color = "royalblue3") + # cohort
  

  xlim(-2,2) +
  # ylim(0,10) + # sex
  # ylim(0,4.5) + # Housing & ExpH
  # ylim(0,7.6) + # Cohort
  scale_y_continuous(breaks=seq(0,8,2.5)) + # cohort
  facet_wrap(~Label, ncol = 3) + 
  facet_wrap(~Label, ncol = 3) + 
  labs(x = "log2FC",
       y = "-Log10(adj.p value)",
       color = "Protein Regulation") +

  geom_hline(yintercept = -log10(0.05),
             linetype = "dotted", linewidth = 0.1,
             color = "grey30") +
  
  geom_hline(yintercept = -log10(0.1),linetype = "dotted",
             linewidth = 0.1, color = "grey50") +

  geom_vline(xintercept = c(0),linetype = "solid",
             linewidth = 0.1, color = "grey80") +
  
  guides(color=guide_legend(nrow=1, byrow=TRUE)) +
  
  theme_bw() +
  theme(
    # legend.position = "none", # sex & Housing/ExpH
    legend.position = "bottom", # cohort
    axis.title = element_text(color = "grey10", size = 9.5),
    legend.title = element_text(color = "grey10", size = 9.5),
    strip.text = element_text(size = 10.5, color = "grey10"),
    panel.grid.major = element_line(color = "grey90", linewidth = 0.1),
    panel.grid.minor = element_line(color = "grey97", linewidth = 0.1),
    panel.spacing.y = unit(0.5, "cm"), # vert spacing
    panel.spacing.x = unit(0.5, "cm"), # horizontal spacing - COHORT 
    legend.margin = margin(-5,0,0,0),  # space between plot & legend
    legend.key.size = unit(0.7, "cm"), # size of legend key icon
  )

# big_volcano_no_lit
# big_volcano_no_lit_EE 
big_volcano_no_lit_C
```

```{r}
#cohort_volcano <- big_volcano_no_lit
#Housing_ExpH_volcano <- big_volcano_no_lit
#Sex_volcano <- big_volcano_no_lit

# Housing_ExpH_volcano_redo <- big_volcano_no_lit_EE # 2026_07_26
cohort_volcano_redo <- big_volcano_no_lit_C # 2026_07_26
```

### Litter Volcano Plots

```{r}
# make new df for the geom_text bit below.
# prev had from 1 df but over plots repeately & cant index since mapped_comp huge.
sig_volcano_sum_lit <- com_df_mapped %>%
  # filter(pairwise_comp == "litter") %>%
  filter(Label %in% c("S328 vs S329", "S328 vs S324",
                      "S324 vs S422", "S426 vs S420",
                      "S422 vs S423", "S329 vs S423",
                      "S426 vs S423", "S420 vs S422")) %>%
  dplyr::select(Label, sig_no_up_lab, sig_no_down_lab) %>%
  unique() 


# VOLCANO 
big_volcano_lit <- com_df_mapped %>%
  filter(Label %in% c("S328 vs S329", "S328 vs S324",
                      "S324 vs S422", "S426 vs S420",
                      "S422 vs S423", "S329 vs S423",
                      "S426 vs S423", "S420 vs S422")) %>%
  ggplot(aes(x = log2FC, y = -log10(adj.pvalue), label = sig_g_label)) +
   #vline & hlines
  geom_hline(yintercept = -log10(0.05),
             linetype = "dashed", linewidth = 0.1,
             color = "grey30") +
  annotate("text", x = 1.80, y = -log10(0.05),
           label = "adj.p <.05 \n",
           fontface = 'italic', size = 1.5, color = "grey10") +
  geom_hline(yintercept = -log10(0.1),linetype = "dotted",
             linewidth = 0.15, color = "grey50") +
  annotate("text", x = 1.75, y = -log10(0.1),
           label = " adj.p <.1 \n",
           fontface = 'italic', size = 1.5, color = "grey30") +
  geom_vline(xintercept = c(0),linetype = "solid",
             linewidth = 0.15, color = "grey85") +
  
  # POINTS
  ggrastr::geom_point_rast(aes(color = case_when( # conditional ass for color mapping
    adj.pvalue < 0.05 & log2FC > 0 ~ "up_adjp.05", # up sig .05
    adj.pvalue < 0.05 & log2FC < 0 ~ "down_adjp.05", # down sig .05
    adj.pvalue < 0.1 & log2FC > 0 ~ "up_adjp.1", # up 0.1
    adj.pvalue < 0.1 & log2FC < 0 ~ "down_adjp.1", # down .1
    TRUE ~ "ns" ), alpha = adj.pvalue), # catch the rest
    size = 1.5,
    raster.dpi = getOption("ggrastr.default.dpi", 330)) +
  
  scale_color_manual(values = c("firebrick", # coloring
                                "royalblue3",
                                "lightpink",
                                "lightblue3",
                                "grey80"),
                     breaks = c("up_adjp.05",
                                "down_adjp.05",
                                "up_adjp.1",
                                "down_adjp.1"),
                     labels = c("\U2191 (adj.p <.05)",
                                "\U2193 (adj.p <.05)",
                                "\U2191 (adj.p <.1)",
                                "\U2193 (adj.p <.1)")) +
  

  scale_alpha(range = c(0.55,0.05), guide = "none") + # more transparent less sig
  
  geom_text_repel(max.overlaps = 20, size = 3, # gene text on points
                  color = "grey10",  segment.size	= 0.2,
                  segment.alpha = 0.5, min.segment.length = 0.5) + 

               
  
  # NUM SIG UP & DOWN! | use sep df, overwise overplots
  geom_text(data = sig_volcano_sum_lit, # up 
            aes(label = paste0(sig_no_up_lab, " \U2191")),
            x = 1.65, y = 9.65, size = 4, color = "firebrick") +
  
  geom_text(data = sig_volcano_sum_lit, # down
            aes(label = paste0(sig_no_down_lab, " \U2193")),
            x = -1.65, y = 9.65, size = 4, color = "royalblue3") +
  

  xlim(-2,2) +
  ylim(0,10) +
  # ylim(0,4) +
  facet_wrap(~Label, ncol = 4) + 
  # facet_wrap(~Label, ncol = 4) + 
  
  labs(x = "log2FC",
       y = "-Log10(adj.p value)",
       color = "Protein Regulation") +
  
  geom_hline(yintercept = -log10(0.05),
             linetype = "dotted", linewidth = 0.1,
             color = "grey30") +
  
  geom_hline(yintercept = -log10(0.1),linetype = "dotted",
             linewidth = 0.1, color = "grey50") +

  geom_vline(xintercept = c(0),linetype = "solid",
             linewidth = 0.1, color = "grey80") +
  
  guides(color=guide_legend(nrow=1, byrow=TRUE)) +
  
  theme_bw() +
  theme(legend.position = "none",
        axis.title = element_text(color = "grey10", size = 9.5),
        legend.title = element_text(color = "grey10", size = 9.5),
        strip.text = element_text(size = 10.5, color = "grey10"),
        panel.grid.major = element_line(color = "grey90", linewidth = 0.1),
        panel.grid.minor = element_line(color = "grey97", linewidth = 0.1),
        panel.spacing.y = unit(0.5, "cm"), # vert spacing
        #legend.margin = margin(-10,2,-7,0),  # space between plot & legend
        #legend.key.size = unit(0.5, "cm"), # size of legend key icon
        )

big_volcano_lit
```

### Cowplots & Messy Saves

2026_07_25

```{r}
# top_line <- plot_grid(Housing_ExpH_volcano, Sex_volcano,
#                       ncol=2, rel_widths = c(1,0.6), labels = c("A","B"),
#                       label_size = 11)
# top_line
# 
# mid_line <- plot_grid(cohort_volcano, labels = c("C"), label_size = 11)
# mid_line
# 
# bottom_line <- plot_grid(big_volcano_lit, labels = c("D"), label_size = 11)
# 
# bottom_line
# 
# prelim_pairwise_volcanos_cowplot <- c(top_line,mid_line,bottom_line)
```

```{r}
# just manually save -2026_07_25
# prelim_pairwise_volcanos <- c(cohort_volcano, Housing_ExpH_volcano, Sex_volcano,
#                               big_volcano_lit)
# prelim_pairwise_volcanos_cowplot <- c(top_line,mid_line,bottom_line)
# 
# prelim_pairwise_volcanos_2026_07_25 <- c(prelim_pairwise_volcanos,
#                                          prelim_pairwise_volcanos_cowplot)
```

2026_07_26 - redo - used for PhD!

```{r}
# top_line <- plot_grid(Housing_ExpH_volcano_redo, Sex_volcano,
#                       ncol=2, rel_widths = c(1,0.57), labels = c("A","B"),
#                       label_size = 11)
# top_line

# mid_line <- plot_grid(cohort_volcano_redo, labels = c("C"), label_size = 11)
# mid_line

# bottom_line <- plot_grid(big_volcano_lit, labels = c("D"), label_size = 11)
# 
# bottom_line
# 
# prelim_pairwise_volcanos_cowplot <- c(top_line,mid_line,bottom_line)
```

```{r}
# just manually save -2026_07_26 + added cohort 3a
# prelim_pairwise_volcanos_2026_07_26 <- c(cohort_volcano_redo, Housing_ExpH_volcano_redo, Sex_volcano, big_volcano_lit)
```
