---
title: "2026_08_27_SynDive_Schuman_v1.Rmd"
output: html_document
---

***(van Oostrum et al., 2023)** The proteomic landscape of synaptic diversity across brain regions and cell types (Cell Resource)*

- DOI: [10.1016/j.cell.2023.09.028](https://doi.org/10.1016/j.cell.2023.09.028)

- <https://syndive.org/>

- <https://brain.mpg.de/schuman>

- <https://r4ds.hadley.nz/arrow.html> (for Apache Parquet format - arrow package)

- <https://jolars.github.io/eulerr/articles/visualization.html#labeling>

## Setup

```{r}
library(readxl)
library(tidyverse)
library(arrow) # for parquet
library(eulerr)
library(UpSetR)
library(cowplot)
library(ggh4x)
library(ggrepel)
```

```{r}
# Differential Synaptic Protein Abundance across Synapse Types
df1 <- read_parquet("input/db/syndiveDataset_vanoostrum_mm_proteomics_SynapticProteomeFoldChangeFractions.parquet")

# Expressed Synaptic Proteins across Brain Regions and Cell Types
df2 <- read_parquet("input/db/syndiveDataset_vanoostrum_mm_proteomics_SynapticProteomeExpression.parquet")

# Synapse-Type Enriched Proteins across Brain Regions
df3 <- read_parquet("input/db/syndiveDataset_vanoostrum_mm_proteomics_SynapticProteomeEnrichedType.parquet")
```

```{r}
#comp
load("input/MSstatsTMT/comp/2025_12_31_1pspg_v3.2.2_t1_EExAge_mapped_comp.rda") 
#psum
load("input/MSstatsTMT/psum/2025_12_31_1pspg_v3.2.2_t1_EExAge_mapped_psum.rda") # 
#anno
anno <- read_csv("input/anno/EEv&SH_F&M_TMT_anno_v3_df_b_check_sorted_2025_12_31.csv") 
```

```{r}
comp <- mapped_comp_v3.2.2_EExAge_t1 

comp <-  comp %>%
  filter(is.na(issue)) %>% # keeps rows with issue == na (so comps that are ok)
  filter(!str_detect(log2FC, "Inf")) %>% # remove infinity/missing FC values
  filter(!(Label %in% c("HxA_EE_YNG_vs_SH_OLD","HxA_EE_OLD_vs_SH_YNG"))) %>%
  mutate(Direction = if_else(log2FC > 0, "Up", "Down")) 


comp$Label <- factor(comp$Label,
                            levels = c( "H_EE_vs_SH",
                                        "HxA_EE_YNG_vs_SH_YNG",
                                        "HxA_EE_OLD_vs_SH_OLD",
                                        "A_YNG_vs_OLD", 
                                        "HxA_EE_YNG_vs_EE_OLD",
                                        "HxA_SH_YNG_vs_SH_OLD"))   
```

```{r}
# Map details & shorten name
psum <- left_join(mapped_psum_v3.2.2_EExAge,
                                       anno,
                                       by = join_by(
                                         BioReplicate == BioReplicate,
                                         Mixture == Mixture,
                                         Channel == Channel)) %>%
  dplyr::rename(Condition = Condition.x , Housing_c_anno = Condition.y) %>%
  mutate(Age = ifelse(PND > 100, "OLD", "YNG")) %>%
  mutate(Age = factor(Age, levels = c("YNG", "OLD"))) %>% # so YNG is 1st
  mutate(across(c(Housing, Sex, Year, Cohort, Litter, exp_helper),
         as.factor)) # auto alphabetical order (not really needed?)

psum$Condition <- factor(psum$Condition,
                         levels = c("EE_YNG", "SH_YNG",
                                    "EE_OLD", "SH_OLD"))


#Check 
colnames(psum)
```

------------------------------------------------------------------------

# Part 1 - Data from SynDive

## df1

```{r}
str(df1)
summary(df1)
unique(df1$value_sample) # olf, hip, str, cereb, cortex + diff markers.


df1_cortex <- df1 %>%
  filter(str_detect(value_sample, regex("cortex", ignore_case = TRUE))) 

df1_cortex_topg <- df1_cortex %>%
  group_by(value_sample) %>%
  filter(
    min_rank(desc(value_numeric)) <= 25 | # most positive
      min_rank(value_numeric) <= 25 # most negaitve
  ) %>%
  pull(gene_symbol)
  
# plot(df1_cortex$value_numeric) # holy middle after filtering
```

```{r}
tile_1 <- df1_cortex %>%
  filter(gene_symbol %in% df1_cortex_topg) %>%
  ggplot(aes(x = value_sample, y = reorder(gene_symbol, value_numeric),
             fill = value_numeric )) +
  geom_tile(color = "grey10") +
  colorspace::scale_fill_continuous_diverging("Blue-Red 2", mid = 0) +
  theme(
    # axis.text.x = element_text(angle = 90, hjust = 1),
        axis.text.y = element_text(hjust = 1, face =  "italic"),
         legend.key.size = unit(0.5, "cm")) # size of legend key icon)

tile_1
```

------------------------------------------------------------------------

## df3 + Venn prep

```{r}
str(df3)
summary(df3)
unique(df3$value_sample) # olf, hip, str, cereb, cortex + diff markers.

df3_cortex <- df3 %>%
  filter(str_detect(value_sample, regex("cortex", ignore_case = TRUE))) %>%
  mutate(value_sample = case_when(
    value_sample == "cortex_Camk2a" ~ "Camk2a (cortex)",
    value_sample == "cortex_Syn1" ~ "Syn1 (cortex)",
    value_sample == "cortex_Gad2" ~ "Gad2 (cortex)",
    value_sample == "cortex-inhibitory_Gad2" ~ "Gad2 (cortex-inhib)",
    value_sample == "cortex-inhibitory_PV" ~ "Pvalb (cortex-inhib)",
    value_sample == "cortex-inhibitory_SST" ~ "Sst (cortex-inhib)",
    value_sample == "cortex-inhibitory_VIP" ~ "Vip (cortex-inhib)"
  )) %>%
  filter(value_numeric == 1) %>% # gets presence
  select(value_sample, gene_symbol) # select only 2 cols

df3_cortex_list <- df3_cortex %>% # make list for eullerr
  split(.$value_sample) %>%
  map(~pull(., gene_symbol)) # gets 7 gene lists from each
```

### Eulerr of shared & unique makers

```{r fig.height=9, fig.width=12}
#eulerr plot 
# set.seed(123)
plot(euler(df3_cortex_list,
           # shape="ellipse"
           ),
     # quantities = TRUE,
     quantities = list(type = c("percent", "counts")),
      fill = list(fill = c("#2aa198", "#b58900", "#859900", "#cb4b16",
                           "#dc322f", "#d33682", "#6c71c4")),
     alpha = 0.5
     # labels = list(placement = "elbow"),
     # main = "van Ossumm et al., 2023 - df3"
     )

# plot(euler(df3_cortex_list))
```

```{r}
df3_gene_counts <- df3_cortex %>%
  group_by(gene_symbol) %>%
  summarise(n_types = n_distinct(value_sample), #no. unique brainareas gene is in
            types = paste(sort(unique(value_sample)), collapse = " | "), # concat the cell types
            .groups = "drop") # rm group to collapse

df3_unique_genes <- df3_gene_counts %>% filter(n_types == 1) # lists genes that are only inone group/type
df3_unique_genes %>% count(types, sort = TRUE) # tally of unique genes | matches eulerr percentages 


df3_genes_all <- df3_gene_counts %>% filter(n_types == 7) # genes in all 7 

df3_genes_2 <- df3_gene_counts %>% filter(n_types == 2) # genes in 2
```

### mapped tile plots of unique marker genes

```{r}
comp_joined_1 <- left_join(comp, df3_unique_genes,
                         by = join_by(gene == gene_symbol)) %>% # 64392 x 20 
  drop_na(types) %>% # remove blank type -> 3328 x 20 
  mutate(ranking = sign(log2FC)*(-log10(pvalue))) %>%
  group_by(Label) %>%
  filter(
    min_rank(desc(ranking)) <= 50 | # most positive
      min_rank(ranking) <=50# most negaitve
  )

unique(comp_joined_1$gene) # >400
unique(comp_joined_1$types) # all 7 seen! 
```

```{r}
type_colors <- c( # type
  "Camk2a (cortex)" = "#2aa198",
  "Syn1 (cortex)" = "#859900",
  "Gad2 (cortex)" = "#b58900",
  "Gad2 (cortex-inhib)" = "#dc322f",
  "Pvalb (cortex-inhib)" = "#cb4b16",
  "Sst (cortex-inhib)" = "#d33682" ,
  "Vip (cortex-inhib)" = "#6c71c4" 
)


comp_colors <- c( # comparison colors
  "H_EE_vs_SH"  = "#AFCC9F", # HOUSING
  "HxA_EE_YNG_vs_SH_YNG" = "#D4E4CB",
  "HxA_EE_OLD_vs_SH_OLD" = "#D4E4CB",
  "A_YNG_vs_OLD" = "#BFBBD5", # AGE
  "HxA_EE_YNG_vs_EE_OLD" = "#DDDCE9",
  "HxA_SH_YNG_vs_SH_OLD" = "#DDDCE9")


labels4tile <- unique(comp_joined_1$Label)
tile_plots_list <- list()
tile_plots_list_w_labs <- list()
```

```{r}
for (lb in labels4tile) {
  
  
  tile_df <- comp_joined_1 %>%
    filter(Label == lb) %>% 
    arrange(types, gene) %>%
    mutate(gene = fct_inorder(gene)) 
  
  strip_color <- comp_colors[[lb]]
  
  
  tileplot <- ggplot(tile_df, aes(x = Label, y = gene,
                                  fill = types, color = Direction,
                                  alpha = adj.pvalue)) +
    geom_tile() +
    
    scale_color_manual(values = c("#2380c1","#d33682"))+
    scale_fill_manual(values = type_colors,
                      guide = guide_legend(nrow = 2, byrow = TRUE)) +
  scale_alpha_continuous(limits = c(0, 0.9), # fixed - so consistnet across
                         range = c(1,0.5), # value range
                         # guide = "none",
                        guide = guide_legend(nrow = 2, byrow = TRUE),
                         oob = scales::squish) + # so doesn't NA if out of range
    
  scale_x_discrete(expand = expansion(mult = c(0.0,0.0))) +
  scale_y_discrete(expand = expansion(mult = c(0.0,0.0))) +
    
  facet_wrap(Label~Direction, scales = "free_y", space = "free_x")+
      # facet_grid(types~Label+Direction, scales = "free", space = "free")+


    theme_bw() +
    theme(
      # legend.position = "none",
      legend.position  = "bottom",
      strip.background = element_rect(fill =strip_color), 
      axis.text.y = element_text(hjust = 1, face =  "italic"),
      axis.title = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      strip.text = element_text(size =7, color = "grey10"),
      # legend.title = element_text(size =8, color = "grey10"),
      legend.title = element_blank(),
      legend.text = element_text(size =7, color = "grey10",
                                 margin = margin(l = 0.05, unit = "cm")),
      legend.key.size = unit(0.3, "cm")
    )
  
  
  #print(tileplot)
  
  # tile_plots_list[[lb]] <-  tileplot # save
  tile_plots_list_w_labs[[lb]] <-  tileplot # save
}
```

```{r fig.height=14, fig.width=10}
plot_grid(plotlist= tile_plots_list_w_labs[c(3,5,2,6,4,1)])
```

```{r fig.height=36, fig.width=6}
list_dist_df3_tile <- df3_unique_genes %>%
    arrange(types, gene_symbol) %>%
    mutate(gene = fct_inorder(gene_symbol)) %>%
    ggplot(aes(n_types, y=gene, fill=types)) +
    geom_tile(alpha = 0.9, color="grey10") +

    scale_fill_manual(values = type_colors,
                      guide = guide_legend(nrow = 1, byrow = TRUE)) +  
      
  scale_x_discrete(expand = expansion(mult = c(0.0,0.0))) +
  scale_y_discrete(expand = expansion(mult = c(0.0,0.0))) +
  labs(subtitle = "Unique Cortex Synapse Marker Gene (433) (df3 - van Oostrum et al., 2023)") +
    theme_bw() +
    theme(
      # legend.position = "none",
      legend.position  = "bottom",
      strip.background = element_rect(fill =strip_color), 
      axis.text.y = element_text(hjust = 1, face =  "italic"),
      axis.title = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      strip.text = element_text(size =7, color = "grey10"),
      # legend.title = element_text(size =8, color = "grey10"),
      legend.title = element_blank(),
      legend.text = element_text(size =7, color = "grey10",
                                 margin = margin(l = 0.05, unit = "cm")),
      legend.key.size = unit(0.3, "cm")
    )
  
list_dist_df3_tile      
```

------------------------------------------------------------------------

# Part 2 - Data from paper/suppdatatables!

## 2.1 Gad2-SypTOM vs Camk2a-SypTOM 

```{r}
s_df1 <- read_excel("input/db/mmc1_(vanOostrum_2023).xlsx",
                    sheet = "C) Gad2-vs-Camk2a") 

s_df1 <- s_df1 %>%
  mutate(gene = sub(" .*", "", GeneName), .after = GeneName) %>%  #keep 1st (primary)
  filter(!is.na(gene)) %>% # removes SypTOM
  mutate(type = if_else(log2FC > 0, "Gad2", "Camk2a"))


s_df1_sig_list <- s_df1 %>%
  filter(FC_adj.pvalue < 0.05 & log2FC > 1 | log2FC < -1) 

str(s_df1)
colnames(s_df1)
```

```{r fig.height=7, fig.width=12}
gad2_v_camk2a_dots <- ggplot(s_df1, aes(x = log2FC, y = -log10(FC_adj.pvalue),
                                        label = gene, color = type)) +
  geom_point(shape = 1) +
    geom_text_repel(max.overlaps = 10, size = 2.5, # gene text on points
                  color = "grey10",
                 segment.size	= 0.2, segment.alpha = 0.5, min.segment.length = 0.5) +
  scale_color_manual(values = c("#2896A6","#C65A4A"))+
  theme_void()

gad2_v_camk2a_dots
```

```{r fig.height=6, fig.width=12}
gad2_v_camk2a_bars <- s_df1 %>%
  filter(!is.na(gene)) %>%
  filter(FC_adj.pvalue < 0.05 & log2FC > 1 | log2FC < -1) %>%
  ggplot(aes(x = reorder(gene, log2FC), y = log2FC, fill = type)) +
  geom_bar(stat = "identity", color = "grey10",
           linewidth = 0.3, alpha =0.7) +

  scale_fill_manual(values = c("#2896A6","#C65A4A"), name = "Enriched")+
  labs(y = "Gad2-SynTOM vs Camk2a-SynTOM") +
  theme_bw() +
  theme(
    legend.position = "bottom",
    legend.title =  element_text(size = 9),
    axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5,
                               face = "italic"),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 9),
    panel.border = element_rect(color = "grey10", linewidth = 0.3),
    panel.grid = element_blank()
  ) 

gad2_v_camk2a_bars 
                               
                              
```

```{r}
# lattice::barchart(log2FC~gene, data = s_df1, 
#                   origin = 0,
#          main = "Bar Chart", 
#          xlab = "Type", 
#          ylab = "Values")
```

```{r}
psum_gad2_v_camk2a <- left_join(comp, s_df1_sig_list[, c("gene", 'type')],
                               by =  join_by("gene" == "gene")) %>%
  drop_na(type)

EI_tile_plots_list <- list()

```

```{r}
for(lb in labels4tile) {
  
  tile_df <- psum_gad2_v_camk2a %>%
    filter(Label == lb) %>% 
    arrange(Direction, gene) %>%
    mutate(gene = fct_inorder(gene)) 
  
  strip_color <- comp_colors[[lb]]
  
  
  tileplot <- ggplot(tile_df, aes(x = Label, y = gene,
                                  fill = Direction, color = type,
                                  alpha = adj.pvalue)) +
    geom_tile() +
    
    scale_fill_manual(values = c("#2380c1","#d33682"))+
    scale_color_manual(values =c("#2aa198","#b58900"),
                      guide = guide_legend(nrow = 1, byrow = TRUE)) +
  scale_alpha_continuous(limits = c(0, 0.9), # fixed - so consistnet across
                         range = c(1,0.5), # value range
                         # guide = "none",
                        guide = guide_legend(nrow = 1, byrow = TRUE),
                         oob = scales::squish) + # so doesn't NA if out of range
    
  scale_x_discrete(expand = expansion(mult = c(0.0,0.0))) +
  scale_y_discrete(expand = expansion(mult = c(0.0,0.0))) +
    
  # facet_wrap(TypeLabel~Direction, scales = "free_y", space = "free_x")+
      facet_grid(type~Label, scales = "free", space = "free")+

    
    theme_bw() +
    theme(
      legend.position = "none",
      strip.background.x = element_rect(fill =strip_color), 
      strip.background.y = element_rect(color = 'grey10', fill = 'grey90'), 
      strip.text.y = element_text(size = 7.5, face = "italic",
                                 margin = margin(t = 1, b = 1, l = 1.5, r = 1.5)),
      strip.text.x = element_text(size = 6),
      
      axis.text.y = element_text(hjust = 1, face =  "italic", size = 8.5),
      axis.title = element_blank(),
      axis.text.x = element_blank(),
      axis.ticks = element_blank(),
      # axis.text.y = element_blank(),
      panel.border = element_rect(fill = NA,color = "grey10"), #plotbor
      # axis.ticks = element_line(linewidth = .1, colour = "grey15"),
      panel.spacing.y = unit(0.25, "cm"), # vert spacing
      panel.spacing.x = unit(0, "cm") # horizontal spacing 
    )
  
  # 
  # print(tileplot)

  EI_tile_plots_list[[lb]] <-  tileplot # save
  
}
```

```{r fig.height=15, fig.width=10}
plot_grid(plotlist = EI_tile_plots_list[c(3,5,2,6,4,1)], nrow = 1)
```

## 2.2 - s_df3 - anti-correlation pairs

```{r}
s_df3 <- read_excel("input/db/mmc3_(vanOostrum_2023).xlsx",
                    sheet = "D) negative family correlations") 
```

```{r}
s_df3_l <- s_df3 %>%
  arrange(GeneNameA) %>% # arr alpha
  mutate(set = row_number()) %>% # add pair no.,
  pivot_longer(cols = c("GeneNameA", "GeneNameB"),
               names_to = "marker",
               names_prefix = "GeneName",
               values_to = "gene") %>%
  mutate(marker = if_else(marker == "A", "Camk2a", "Gad2")) %>%
  select(set:gene) # keep 3


psum_s_df3_sets <- left_join(psum, s_df3_l,
                              by = join_by(gene == gene)) %>%
  drop_na(set)
```

```{r fig.height=7, fig.width=12}
psum_s_df3_sets_boxplot <- psum_s_df3_sets %>%
  filter((set %in% c(12:22))) %>%
  ggplot(aes(x = interaction(marker,gene, lex.order = TRUE,sep = "\n"),
             y = Abundance, fill = Housing)) +
  geom_boxplot(alpha = 0.75, color = "grey10") +
      stat_summary(fun=mean, geom="point", position = position_dodge(width = 0.7),
               shape=4, size = 1, stroke = 0.5, color ="grey10") + #displays mean
  facet_wrap(~set, space = "free_x", scale="free")+
      # scale_color_manual(values = c("#2aa198", "#b58900"))+
  scale_fill_manual(values = c("#66A61EFF", "#E6AB02FF")) + # housing colors 
  # # facet_nested(~set+marker, scale = "free") +
  # facet_grid(set~marker, space="free") +
  theme_bw() +
  theme(
    legend.position = "none",
    axis.title.x = element_blank()
  )
  
psum_s_df3_sets_boxplot

```

### Figure_6g

```{r}
in_markers <- data.frame(in_marker = c(rep("Pvalb", 9), rep("Vip", 9),
                                    rep("Sst", 5), rep("Gad2_In", 4)),
                         gene = c(
                           'Pvalb', 'Kcnc1', 'Kcnc2', 'Kcnc3', 'Hcn1', 'Hcn2', 'Vamp1', 'Inpp5j', 'Timm13',
                           'Vip', 'Cnr1', 'Calb2', 'Grm7', 'Grm1', 'Dag1', 'Adcy2', 'Traf3', 'Nrxn2', # Vip
                           'Calb1', 'Rab3b', 'Icam5', 'Atp2b4', 'Nos1', # Sst
                           'Lamp5', 'Sncg', 'Ppp1r1b', 'Cplx3' #Gad2
                                )
                        )

psum_in_markers <- left_join(psum, in_markers,
                             by = join_by("gene" == "gene")) %>%
  drop_na(in_marker)


```

```{r fig.height=6, fig.width=12}

psum_in_marker_bp <- psum_in_markers %>%
  # filter((set %in% c(12:22))) %>%
  ggplot(aes(x =gene, y = Abundance, fill = Condition, color = Age)) +
  geom_boxplot(alpha = 0.75, linewidth = 0.2, outlier.shape = 1) +
  stat_summary(fun=mean, geom="point", position = position_dodge(width = 0.7),
               shape=4, size = 1, stroke = 0.5) + #displays mean
  
  scale_color_manual(values = c("#803300","#654A87")) + # housing colors 
  # scale_color_manual(values = c("#2aa198", "#b58900"))+
  scale_fill_manual(values = c("#66A61EFF", "#E6AB02FF",
                               "#66A61EFF","#E6AB02FF")) + # housing colors 
  
  facet_wrap(~in_marker, space = "free_x", scale="free_x")+
  ylim(c(4.5,9)) +
  
  theme_bw() +
  theme(
    legend.position = "none",
    axis.title.x = element_blank()
  )

psum_in_marker_bp
```

## 2.3 Inhibitory Cortex Table S5A

```{r}
s_df5a <- read_excel("input/db/mmc5_(vanOostrum_2023).xlsx")

str(s_df5a)
```

```{r}
s_df5a_l <- s_df5a %>%
  pivot_longer(
    cols = Gad2_PV_Label:VIP_SST_FC_adjpvalue,
    names_to = c("comparison", ".value"), # value is spec below - so label, log2fc, fc etc
    names_pattern = "(.+)_(Label|log2FC|FC_adjpvalue)" # capture end pattern |deals w/_
  ) %>%
  rename(adjp_in = FC_adjpvalue,
         lab_in = Label,
         subtype_prot = `Subtype Proteome`,
         gene = GeneName,
         log2FC_in = log2FC) %>%
  filter(adjp_in <= 0.05)

colnames(s_df5a_l)

s_df5a_list <- s_df5a_l %>%
  split(.$subtype_prot) %>%
  map(~pull(., gene)) # gets 7 gene lists from each
```

### 
