---
title: "MssTMT_Vis_2026_08_v3.2.2_1pspg_EExAge_v2"
output: html_document
---

MssTMT_Vis_2026_08_v3.2.2_1pspg_EExAge_v2

new version for thesis - removed some old stuff & added some new stuff

# Setup

```{r}
#
#load packages
library(tidyverse)
library(readxl)
library(pheatmap)
# library(pcaMethods) # for various pca()

# library(msigdbr)

library(cowplot)
library(ggh4x) # scalenested

library(gt)
library(ggrepel) #  for geom_text_repel()
library(ggrastr) # https://tomsing1.github.io/blog/posts/volcano/

library(colorspace)
library(RColorBrewer)
library(paletteer)
library(ggnewscale) # additional color/fill scale new_scale_fill()
library(ggvenn) # vennys (old ones)

library(htmlwidgets) #
library(webshot2)
library(ggpmisc) # for line of best fit equation on ggplot
```

### Load in data

```{r}
#psum
load("input/MSstatsTMT/psum/2025_12_31_1pspg_v3.2.2_t1_EExAge_mapped_psum.rda") # 

#comp
load("input/MSstatsTMT/comp/2025_12_31_1pspg_v3.2.2_t1_EExAge_mapped_comp.rda") # 

#anno
anno <- read_csv("input/anno/EEv&SH_F&M_TMT_anno_v3_df_b_check_sorted_2025_12_31.csv") 
# double check later!
```

## Map/Fix names - aka + bunch of extra annotation info!

### For Protein Summarisation Data

```{r}
# Map details & shorten name
mapped_psum <- left_join(mapped_psum_v3.2.2_EExAge,
                                       anno,
                                       by = join_by(
                                         BioReplicate == BioReplicate,
                                         Mixture == Mixture,
                                         Channel == Channel)) %>%
  dplyr::rename(Condition = Condition.x , Housing_c_anno = Condition.y) %>%
  mutate(Age = ifelse(PND > 100, "OLD", "YNG")) %>%
  mutate(Age = factor(Age, levels = c("YNG", "OLD"))) %>% # so YNG is 1st
  mutate(across(c(Housing, Sex, Year, Cohort, Litter, exp_helper),
         as.factor)) # auto alphabetical order (not really needed?)

# mapped_psum$Age <- factor(mapped_psum$Age, # WAIT NOT NEEDED?
#                           levels = c(YNG))
# sig_genes_count$pairwise.comp <- factor(sig_genes_count$pairwise.comp,
#                                    levels = c("Housing", "ExpH", "Cohort",
#                                               "Litter", "Sex"))
  
#Check 
colnames(mapped_psum)
```

##### lil checks

```{r}
# lil checks
head(mapped_comp_v3.2.2_EExAge_t1) # peak
str(mapped_comp_v3.2.2_EExAge_t1) # structure 
colnames(mapped_comp_v3.2.2_EExAge_t1) # column names
dim(mapped_comp_v3.2.2_EExAge_t1) # 64392 x 8

# lil checks (bit messy)
unique(mapped_comp_v3.2.2_EExAge_t1$issue) # na in issue col = good. if not excl
check_1 <- mapped_comp_v3.2.2_EExAge_t1 %>%
  filter(str_detect(log2FC, "Inf")) # 5 inf 
check_2 <- mapped_comp_v3.2.2_EExAge_t1 %>%
  filter(str_detect(Protein, "Cont")) # 0 (since already removed)
unique(mapped_comp_v3.2.2_EExAge_t1$Label) 
length(unique(mapped_comp_v3.2.2_EExAge_t1$Label)) #8
```

### For Group Comparison Data

```{r}
# shorten name & clean up! 
mapped_comp <- mapped_comp_v3.2.2_EExAge_t1 %>%
  
  filter(is.na(issue)) %>% # keeps rows with issue == na (so comps that are ok)
  filter(!str_detect(log2FC, "Inf")) %>% # remove infinity/missing FC values
  
  # make gene label for significant! 
  mutate(sig_g_label = if_else(adj.pvalue <= .1, gene, NA)) %>%
  
  # make one for no. of sig up & down for each comaprison group/label
  group_by(Label) %>%
  mutate(sig_no_up_lab = sum(adj.pvalue < 0.05 & log2FC > 0), # up
         sig_no_down_lab = sum(adj.pvalue < 0.05 & log2FC < 0)) %>%  #  down
  
  mutate(Direction = ifelse(log2FC > 0, "Up", "Down"))

# check 
head(mapped_comp) 
colnames(mapped_comp)
dim(mapped_comp) # 64384 x 22

unique_check_mapcomp <- mapped_comp %>% 
  dplyr::select(Label, sig_no_down_lab, sig_no_up_lab) %>%
  unique() # yep matches up! 
```

#### df - significant comparisons

```{r}
mapped_comp_sig <- mapped_comp %>%
  filter(adj.pvalue < 0.05)

dim(mapped_comp_sig) # 458 x 21
```

# Plotting Comp

#### Theme

```{r}
my_theme <- theme(
    axis.line = element_blank(),
    axis.title = element_text(size = 9, color = "grey10"), 
    # axis.title.x = element_blank(),
    axis.text = element_text(size = 7.5, color ="grey10"),
    axis.ticks = element_line(size=0.5, colour = "grey15"),
    axis.ticks.length.y.left = unit(-0.1, 'cm'), # inwards ticks (L)
    axis.ticks.length.y.right = unit(-0.07, 'cm'), # inwards ticks (R)
    panel.background = element_rect(fill = "white"), #bg of plot (inner bit)
    panel.border = element_rect(fill = NA,
                                color = "grey10"), #plot borders
    plot.background = element_rect(fill = "white"), # bg of plot (outer)
    panel.grid = element_blank(),
    panel.spacing = unit(0, "cm"),  # Removes space between panel 
    legend.position = 'bottom',
    legend.title = element_text(size = 9, face="italic", color = "grey10"),
    legend.margin = margin(-5,0,0,0), # space between plot & legend
    #legend.title.align = (0.5), # centres legend title
    legend.text = element_text(size = 8.5, color = "grey10"),
    legend.box = "horizontal", 
    legend.direction = "horizontal", 
    legend.key.size = unit(0.5, "cm"), # size of legend key icon
    strip.background = element_rect(color = 'grey10', fill = 'grey90'), 
    strip.text.x = element_text(size = 9, color = 'grey10',
                                margin = margin(b = 3, t = 2)) #vert spacing
  )

```

# Volcano

Bit messy atm

- Big Volcano - When doing all -\> Error in UseMethod("depth") : no applicable method for 'depth' applied to an object of class "NULL (takes 10 years on laptop) ➝ due to geom_text_repel() trying to plot the ns ones ➝ fixed if make a just sig gene/protein column

- Should have just made a function..... oh well do later.

```{r}
# 
strip_colors <- strip_themed( # green for 1st 3 (housing) & Lavender for last 3 (age)
    background_x = elem_list_rect(fill = c(
      "#AFCC9F", "#D4E4CB",  "#D4E4CB",
      "#BFBBD5", "#DDDCE9", "#DDDCE9"
      # "#D9DFDD","#D9DFDD"
      )))
  
```

```{r}
# make new df for the geom_text bit below.
# prev had from 1 df but over plots repeately & cant index since mapped_comp huge.
sig_volcano_sum <- mapped_comp %>%
  dplyr::select(Label, sig_no_up_lab, sig_no_down_lab) %>%
  filter(!(Label %in% c("HxA_EE_YNG_vs_SH_OLD", "HxA_EE_OLD_vs_SH_YNG"))) %>% # 6 plot
  # filter(!(Label %in% c("A_YNG_vs_OLD",
  #                       "HxA_EE_YNG_vs_EE_OLD", "HxA_SH_YNG_vs_SH_OLD"))) %>%
  # filter(!(Label %in% c("H_EE_vs_SH",
  #                       "HxA_EE_YNG_vs_SH_YNG", "HxA_EE_OLD_vs_SH_OLD"))) %>%
  unique()

# reoder factor levels for plotting
mapped_comp$Label <- factor(
  mapped_comp$Label,
  levels = c( "H_EE_vs_SH", "HxA_EE_YNG_vs_SH_YNG", "HxA_EE_OLD_vs_SH_OLD",
              "A_YNG_vs_OLD",  "HxA_EE_YNG_vs_EE_OLD", "HxA_SH_YNG_vs_SH_OLD",
              "HxA_EE_YNG_vs_SH_OLD", "HxA_EE_OLD_vs_SH_YNG"))
```

```{r}
# unique(mapped_comp$Label)
# unicodes for up \U2191 & down \U2193
big_volcano <- mapped_comp %>%
  filter(!(Label %in% c("HxA_EE_YNG_vs_SH_OLD", "HxA_EE_OLD_vs_SH_YNG"))) %>% # 6 plot
  #   # filter(!(Label %in% c("A_YNG_vs_OLD",
  #   #                     "HxA_EE_YNG_vs_EE_OLD", "HxA_SH_YNG_vs_SH_OLD"))) %>%
  #   filter(!(Label %in% c("H_EE_vs_SH",
  #                       "HxA_EE_YNG_vs_SH_YNG", "HxA_EE_OLD_vs_SH_OLD"))) %>%

  ggplot(aes(x = log2FC, y = -log10(adj.pvalue), label = sig_g_label)) +
  
  
  #vline & hlines
  geom_hline(yintercept = -log10(0.05),
             linetype = "dashed", linewidth = 0.1,
             color = "grey20") +
  annotate("text", x = 1.80, y = -log10(0.05),
           label = "adj.p <.05 \n",
           fontface = 'italic', size = 1.5, color = "grey10") +
  geom_hline(yintercept = -log10(0.1),linetype = "dotted",
             linewidth = 0.15, color = "grey50") +
  annotate("text", x = 1.75, y = -log10(0.1),
           label = " adj.p <.1 \n",
           fontface = 'italic', size = 1.5, color = "grey30") +
  geom_vline(xintercept = c(0),linetype = "solid",
             linewidth = 0.15, color = "grey80") +
  

  
    # POINTS
  ggrastr::geom_point_rast(aes(color = case_when( # conditional ass for color mapping
    adj.pvalue < 0.05 & log2FC > 0 ~ "up_adjp.05", # up sig .05
    adj.pvalue < 0.05 & log2FC < 0 ~ "down_adjp.05", # down sig .05
    adj.pvalue < 0.1 & log2FC > 0 ~ "up_adjp.1", # up 0.1
    adj.pvalue < 0.1 & log2FC < 0 ~ "down_adjp.1", # down .1
    TRUE ~ "ns" ), alpha = adj.pvalue), # catch the rest
    size = 1.7,
    raster.dpi = getOption("ggrastr.default.dpi", 330)) +
  
  scale_color_manual(values = c("firebrick", # coloring
                                "royalblue3",
                                "lightpink",
                                "lightblue3",
                                "grey80"),
                     breaks = c("up_adjp.05",
                                "down_adjp.05",
                                "up_adjp.1",
                                "down_adjp.1"),
                     labels = c("\U2191 (adj.p <.05)",
                                "\U2193 (adj.p <.05)",
                                "\U2191 (adj.p <.1)",
                                "\U2193 (adj.p <.1)")) +
    scale_alpha(range = c(0.55,0.05), guide = "none") + # more transparent less sig
  
  
  geom_text_repel(max.overlaps = 35, size = 2.2, # gene text on points
                  color = "grey10",
                 segment.size	= 0.2, segment.alpha = 0.5, min.segment.length = 0.5) +
  
  

  
  
  # NUM SIG UP & DOWN! | use sep df, overwise overplots
  geom_text(data = sig_volcano_sum, # up 
            aes(label = paste0(sig_no_up_lab, " \U2191")),
            x = 1.62,
             y = 7.3,
            size = 3.6, color = "firebrick"
            ) + 
  geom_text(data = sig_volcano_sum, # down
            aes(label = paste0(sig_no_down_lab, " \U2193")),
            x = -1.62,
             y = 7.3,
            size = 3.6, color = "royalblue3"
            ) + 

  xlim(-2,2) +

  # facet_wrap(~Label, ncol = 4) + 
  facet_wrap2(~Label,
                strip = strip_colors,
                # space = "free",
               # scales = "free",
                ncol = 3) +

  labs(x = "log2FC",
       y = "-Log10(adj.p value)",
       color = "Protein Regulation") +
  
  
  guides(color=guide_legend(nrow=1, byrow=TRUE)) +
  
  theme_bw() +
  theme(
    # legend.position = "none", # sex & Housing/ExpH
    legend.position = "bottom", # cohort
    axis.title = element_text(color = "grey10", size = 9.5),
    legend.title = element_text(color = "grey10", size = 9.5),
    strip.text = element_text(size = 10.5, color = "grey10"),
    panel.grid.major = element_line(color = "grey90", linewidth = 0.1),
    panel.grid.minor = element_line(color = "grey97", linewidth = 0.1),
    panel.spacing.y = unit(0.5, "cm"), # vert spacing
    panel.spacing.x = unit(0.5, "cm"), # horizontal spacing - COHORT 
    legend.margin = margin(-5,0,0,0),  # space between plot & legend
    legend.key.size = unit(0.7, "cm"), # size of legend key icon
  )

big_volcano

# big_volcano_EE <- big_volcano
# big_volcano_AGE <- big_volcano
```

```{r}
big_volcano
  
# exported 18 June 2026
# ggsave("2026_06_18_EExAge_8_volcanos_2x4_v1_330dpi.png", dpi = 330) #6.72x7.38
# dev.off()
ggsave("2026_06_18_EExAge_6_volcanos_2x3_v1_330dpi.png", dpi = 330) # 6.43 x 4.87
dev.off()




# exported 4 x 2 mini version just for overview   10.5 x 6 inches pdf 
# 2026_01_17_EExAge_8_volcanos_4x2_mini_v2 | 10.5 x 6 inches pdf 
# 2026_01_17_EExAge_8_volcanos_4x2_mini_v3 | 10.5 x 6 inches pdf (fixed name)

# ggsave("2026_02_27_EExAge_8_volcanos_4x2_mini_v4_dpi330.png", dpi = 330) #png versions
# ggsave("2026_02_27_EExAge_6_volcanos_3x2_mini_v4_330dpi.png", dpi = 330)
# ggsave("2026_02_27_EE_3_volcano_330dpi.png", dpi = 330) # EE 7.87 x 3.24 
# ggsave("2026_02_27_Age_3_volcano_330dpi.png", dpi = 330) 
```

------------------------------------------------------------------------

# psumplots

EE vs SH

- <https://stats.oarc.ucla.edu/r/modules/subsetting-data/>
- <https://en.wikipedia.org/wiki/List_of_Crayola_crayon_colors>

```{r}
# IGNORE FOR NOW - OLD
temp_colors <- c("#8CB585","#A8D9A0", "#D4C07E", "#FEE697", "#E6AB02FF",
            "#66A61EFF" )
# Banana Mania (#FBE7B2), Sunglow (#FFCC33), 	Dandelion (#FED85D)
# Maximum Green (#5E8C31)| Asparagus #7BA05B, Middle Green Yellow (#ACBF60)
```

### For Loops & Functions

Function for boxplots

- <https://github.com/eliocamp/ggnewscale?tab=readme-ov-file>

```{r}
boxplot_function <- function(data, x , y, # function inputs
                             fill = Housing,
                             # alpha = NULL,
                             shape_color = TechRepMixture,
                             shape_fill = NULL,
                             facet_formula = ~Age+Cohort+Litter+Sex , # default
                             # facet_formula = ~Sex+Age+Cohort+Litter, # default
                             legend_pos = "bottom", 
                             title = NULL, # need to set as NULL?
                             subtitle = NULL) {
  
  #set up polot ^ assign 2 p
  p <- ggplot(data, aes({{x}},{{y}},
              # color = {{color}},
              # shape = {{shape}},
              # alpha = {{alpha}}
              )) +
    
    # make box plot
    geom_boxplot(aes(fill = {{fill}}),
                 width = 0.55, size = 0.7, median.linewidth = .4, linewidth = 0.3,
                 position = position_dodge(width = 0.7),
                 outlier.shape = NA,
                 color = "grey10", alpha = 0.7) +
    
    stat_summary(fun=mean, geom="point", position = position_dodge(width = 0.7),
               shape=4, size = .8, stroke = 0.3, color ="grey10") + #displays mean 
    
    # scale_alpha_discrete(guide = 'none', range = c(0.65, 0.75)) +
    scale_fill_manual(values = c("#66A61EFF", "#E6AB02FF")) + # housing colors 
    # scale_shape_manual(values = c(1, 3, 4)) + # techreplicate (3)
    

    # New SCALE with NEW GEOM 
    ggnewscale::new_scale_fill() +
      geom_point(aes(fill = {{shape_fill}}, color = {{shape_color}}),
      size = 1, shape = 21, stroke = .2, 
      position = position_jitterdodge(dodge.width = 0.4, jitter.width = 0.02)) +
    #2nd scale
    scale_color_manual(values = c("#003a6d","#31135e","#510224")) + # IBM
    scale_fill_manual(values = alpha(c("#eee8d5", "#b58900", "#C06A0B", "#cb4b16",
                                       "#D43F23", "#dc322f", "#D83459", "#d33682",
                                       "#A054A3", "#6c71c4", "#497ECB", "#268bd2",
                                       "#2896B5", "#2aa198", "#589D4C", "#859900",
                                       "#586e75"), 0.5)) + # solarised mix
    
    
    # FACET
    ggh4x::facet_nested(facet_formula) +
    
    labs(title = title, subtitle = subtitle) + # TITLE & SUBTITLE
    
    guides(color = "none",
           fill = guide_legend(nrow=1, byrow=TRUE)) +
    
    # theme (maybe just add theme from above or beeter self contained like this?)
    theme(
    plot.title = element_text(size = 9, color = "grey10",
                              margin = margin(b = 0.5)),
    plot.subtitle = element_text(size = 8, color = "grey10", face = "italic",
                                 margin = margin(b = 0.5)),
    axis.line = element_blank(),
    axis.title.y = element_text(size = 9, color = "grey10"), 
    axis.title.x = element_text(size = 8, color = "grey10"),
    axis.text.x = element_text(size = 7, color = "grey10"),
    axis.text.y = element_text(size = 7.5, color ="grey10"),
    axis.ticks = element_line(size=0.1, colour = "grey15", linewidth = .2),
    
    panel.background = element_rect(fill = "grey97"), #bg of plot (inner)
    plot.margin = margin(t = 1, r = 2, b = 2, l = 2, unit = "pt"),
    panel.border = element_rect(fill = NA, linewidth = .2,
                                color = "grey10"), #plot borders
    # plot.background = element_rect(fill = "white"), # bg of plot (outer)
    # panel.grid = element_blank(),
    panel.spacing = unit(0, "cm"),  # Removes space between panel 
    
    legend.position = legend_pos, # LEGEND POSITON
    
    legend.title = element_text(size = 9, face="italic", color = "grey10"),
    legend.margin = margin(-5,0,0,0), # space between plot & legend
    #legend.title.align = (0.5), # centres legend title
    legend.text = element_text(size = 8.5, color = "grey10"),
    legend.box = "horizontal", 
    legend.direction = "horizontal", 
    legend.key.size = unit(0.5, "cm"), # size of legend key icon
    strip.background = element_rect(color = 'grey10', fill = 'grey92',
                                    linewidth = .2), 
    strip.text.x = element_text(size = 6.5, color = 'grey10',
                                margin = margin(b = 1, t = 1)) #vert spacing
  )
    
}
```

Testing Function! - Yep works!

```{r}
# Synaptopodin
Synpo <- mapped_psum %>%
  filter(Protein %in% c("Q9Z327")) # 

Synpo_plot <- boxplot_function(Synpo, 
                       x = Housing, 
                       y = Abundance,
                       fill = Housing,  # by default
                       # color = interaction(BioReplicate, TechRepMixture),
                       shape_fill = Channel,
                       shape_color = TechRepMixture,
                       # alpha = Age,
                       legend_pos = "none",
                       title = 'Q9Z327 - Synaptopodin',
                       subtitle = '(Synpo) - In all 55 Samples')

Synpo_plot

unique(Synpo$BioReplicate)
unique(Synpo$`Gene Names (primary)`) #Synpo
length(unique(Synpo$BioReplicate)) # 56 observations! 55

Synpo %>% 
  group_by(TechRepMixture) %>%
  count(TechRepMixture)
```

------------------------------------------------------------------------

### For Housing per protein

```{r}
EE_main_sig_df <- mapped_comp %>%
  filter(Label == "H_EE_vs_SH" & adj.pvalue < 0.05) %>%
  group_by(Direction) %>%
  arrange(adj.pvalue)


EE_main_p_sig_list <- unique(EE_main_sig_df$Protein)

EE_main_psum_plot_list <- vector(mode = "list")
```

```{r}
for (p in EE_main_p_sig_list){
  
  #SET UP/CALC
  bp_df <- mapped_psum %>% #from psums
  filter(Protein == p) 
  
  bp_labs <- EE_main_sig_df %>% # using the filter mapped_comp list 
    filter(Protein == p) %>%
    dplyr::slice(1) 
  
  obs_in <- mapped_psum %>% #get no. biorep observed
  filter(Protein == p) %>%
    summarise(n_unique = n_distinct(BioReplicate)) %>%
    pull(n_unique)
  
  #________________________
  # MAKE PLOT
  bp <- boxplot_function( 
       data = bp_df,
       x = Housing, 
       y = Abundance,
       fill = Housing,  # by default
       shape_fill = Channel, #de
       shape_color = TechRepMixture,
       legend_pos = "none", # default
       title = paste0(bp_labs$gene, " - ", bp_labs$description),
       subtitle = paste0(bp_labs$Direction, "regulated in EE (",
                         bp_labs$uniprot_acc, " ", bp_labs$database,
                         " - observed in ", obs_in, " samples)")
       )
  
  print(bp)

  #________________
  # SAVIHG PLOTS - separates up vs down regulated
    if (bp_labs$Direction[1] == "Up") {
      EE_main_psum_plot_list$Up[[p]] <- bp

    } else{
      EE_main_psum_plot_list$Down[[p]] <- bp
    }
  
}
```

#### EE psum cowplots

saved everything for EE sig (16 prots)

```{r}
# get 5 at a time
plot_grid(plotlist = EE_main_psum_plot_list$Down[6:10], ncol = 1)
```

```{r}
plot_grid(plotlist = EE_main_psum_plot_list$Up, ncol = 1,
          rel_heights = c(1,0.75,1,1,1,1) )
# maybe need to export again, a bit squished | 7.5 x 10.5 in
```

------------------------------------------------------------------------

### For Age

```{r}
top_20_sig_Age_df <- mapped_comp %>%
  filter(Label == "A_YNG_vs_OLD" & adj.pvalue <0.05) %>%
  group_by(Direction) %>%
  arrange(adj.pvalue)%>%
  slice_min(adj.pvalue, n = 10)


Age_main_p_sig_list <- unique(top_20_sig_Age_df$Protein)

Age_main_20_psum_plot_list <- vector(mode = "list")
```

```{r}
for (p in Age_main_p_sig_list){
  
  # SET UP/ CALC
  bp_df <- mapped_psum %>% #from psums
  filter(Protein == p) 
  
  bp_labs <- top_20_sig_Age_df %>% # using the filter mapped_comp list 
    filter(Protein == p) %>%
    slice(1) 
  
  obs_in <- mapped_psum %>% #get no. biorep observed
  filter(Protein == p) %>%
    summarise(n_unique = n_distinct(BioReplicate)) %>%
    pull(n_unique)
  
  #_______________
  # MAKE PLOT
  bp <- boxplot_function( 
       data = bp_df,
       x = Housing, 
       y = Abundance,
       fill = Housing,  # by default
       shape_fill = Channel, #de
       shape_color = TechRepMixture,
       legend_pos = "none", # default
       title = paste0(bp_labs$gene, " - ", bp_labs$description),
       subtitle = paste0(bp_labs$Direction, "regulated in YNG (",
                         bp_labs$uniprot_acc, " ", bp_labs$database,
                         " - observed in ", obs_in, " samples)")
       )
  
  print(bp)

  #_______________
  # SAVIHG PLOTS
    if (bp_labs$Direction[1] == "Up") {
      Age_main_20_psum_plot_list$Up[[p]] <- bp

    } else{
      Age_main_20_psum_plot_list$Down[[p]] <- bp
    }
  
}
```

```{r}
# cowplot = DOWN IN AGE (YNG)
plot_grid(plotlist = Age_main_20_psum_plot_list$Down, ncol = 2)
```

```{r}
# cowplot = UP IN AGE (YNG)
plot_grid(plotlist = Age_main_20_psum_plot_list$Up, ncol = 2)
```

```{r}
plot_grid(plotlist = Age_main_20_psum_plot_list$Down[6:10], ncol = 1) # 7.5 x 10.5
```

```{r}
plot_grid(plotlist = Age_main_20_psum_plot_list$Up[6:10], ncol = 1) # 7.5 x 10.5

```

------------------------------------------------------------------------

------------------------------------------------------------------------

## EE FACET

- <https://forum.posit.co/t/reordering-plot-based-on-another-dataframe/26467/3>

```{r}
my_psum_facet_theme <- my_theme +
  theme(
    axis.text.y = element_text(size = 7.5, color ="grey10"),
    axis.text.x = element_text(size = 7, color ="grey10"),
    axis.title.x = element_blank(),
    panel.spacing = unit(0.2, "cm"),  # Removes space between panel 
    legend.margin = margin(0,0,10,0), # space between plot & legend
    axis.ticks = element_line(linewidth = .1, colour = "grey15"),
    panel.border = element_rect(fill = NA,color = "grey10",linewidth = 0.3), #plot bord
    strip.background = element_rect(color = 'grey10', fill = 'grey90',linewidth = .3),
    strip.text = element_text(face = "italic"),
    plot.title = element_text(  size = 11.5,  color = "grey10",face = "italic"),
    panel.grid.major = element_line(color = "grey90", linewidth = 0.1),
    panel.grid.minor = element_line(color = "grey97", linewidth = 0.1),
  )
```

```{r}
# EE UP
EE_facet_up_prot_list <- mapped_comp %>%
  filter(Label == "H_EE_vs_SH" & adj.pvalue < 0.05 & Direction == "Up") %>%
  pull(Protein)


EE_facet_up_plot <- mapped_psum %>%
  
  # mutate(Condition = forcats::fct_reorder(Condition, PND)) %>% # reorder by PND 
  
  filter(Protein %in% EE_facet_up_prot_list) %>%
  
  ggplot(aes(x = interaction(Age, Housing, 
                             sep = "\n", # space
                             lex.order = TRUE),
             y = Abundance,
             fill = Housing, alpha = Age, color = Sex)) +
 
  geom_boxplot(width = 0.6, size = 0.7, linewidth = 0.3,
               position = position_dodge(width = 0.7),
               outlier.shape = NA,
               color = "grey10") +
  stat_summary(fun=mean, geom="point", position = position_dodge(width = 0.7),
               shape=4, size = 1, stroke = 0.5, color ="grey10") +
   ggrastr::geom_point_rast( size = .7, shape = 1, stroke = .3, 
             position = position_jitterdodge(dodge.width = 0.4, jitter.width = 0.2),
             raster.dpi = getOption("ggrastr.default.dpi", 330)) +
  
  
  
  
  scale_fill_manual(values = c("#66A61EFF", "#E6AB02FF")) + # housing colors
  scale_alpha_manual(values = c(.65,.8)) +
  scale_color_manual(values = c("#c90076","#125CCC")) +
  # guides(y.sec = 'axis') + # adds 2nd y axis (right side)
  
  labs(title = "\U2191 Upregulated in EE ") +
  
  facet_wrap(~gene, ncol = 6, scale = "free_y") +
  my_psum_facet_theme 

EE_facet_up_plot
```

```{r}
# EE down 
EE_facet_down_prot_list <- mapped_comp %>%
  filter(Label == "H_EE_vs_SH" & adj.pvalue < 0.05 & Direction == "Down") %>%
  arrange(adj.pvalue) %>%
  pull(Protein)

# PLOT ______________
EE_facet_down_plot <- mapped_psum %>%
  
  filter(Protein %in% EE_facet_down_prot_list) %>%
  
  mutate(gene = ifelse(Protein == "Q5U2Q3", "NA (Ester hydrolase ...)",
                       gene)) %>% # because no name available :(
  
  ggplot(aes(x = interaction(Age, Housing, 
                             sep = "\n", # space
                             lex.order = TRUE),
             y = Abundance,
             fill = Housing, alpha = Age, color = Sex)) +
 
  geom_boxplot(width = 0.6, size = 0.7, linewidth = 0.3,
               position = position_dodge(width = 0.7),
               outlier.shape = NA,
               color = "grey10") +
  stat_summary(fun=mean, geom="point", position = position_dodge(width = 0.7),
               shape=4, size = 1, stroke = 0.5, color ="grey10") +
   ggrastr::geom_point_rast( size = .7, shape = 1, stroke = .3, 
             position = position_jitterdodge(dodge.width = 0.4, jitter.width = 0.2),
             raster.dpi = getOption("ggrastr.default.dpi", 330)) +
  
  
  scale_fill_manual(values = c("#66A61EFF", "#E6AB02FF")) + # housing colors
  scale_alpha_manual(values = c(.65,.8)) +
  scale_color_manual(values = c("#c90076","#125CCC")) +

  # guides(y.sec = 'axis') + # adds 2nd y axis (right side)
  
    labs(title = "\U2193 Downregulated in EE ") +
  
  facet_wrap(~gene, ncol = 6, scales = "free_y") +
  my_psum_facet_theme
  # theme_bw()

EE_facet_down_plot
```

```{r}
plot_grid(EE_facet_up_plot,EE_facet_down_plot, ncol = 1,
          labels = c("a", "b"), rel_heights = c(1.2,2),
          label_size = 11, label_colour = "grey10")

# ggsave("2026_02_27_EE_facet_all_16_sig_psum_v6.png", dpi = 330) # 9 x 8.5
```

## AGE FACET

```{r}
# AGE UP
Age_facet_up_prot_list <- mapped_comp %>%
  filter(Label == "A_YNG_vs_OLD" & adj.pvalue < 0.05 & Direction == "Up") %>%
    slice_min(adj.pvalue, n = 18) %>%
    pull(Protein)

  
Age_facet_up_plot <- mapped_psum %>%
  
  mutate(Condition = forcats::fct_reorder(Condition, PND)) %>% # reoder by PND (no)
  
  
  filter(Protein %in% Age_facet_up_prot_list) %>%
  
  
  ggplot(aes(x = interaction(Age, Housing, 
                             sep = "\n", # space
                             lex.order = TRUE),
             y = Abundance,
             fill = Housing, alpha = Age, color = Sex)) +
  
  
  geom_boxplot(width = 0.6, size = 0.7, linewidth = 0.3,
               position = position_dodge(width = 0.7),
               outlier.shape = NA,
               color = "grey10") +
  stat_summary(fun=mean, geom="point", position = position_dodge(width = 0.7),
               shape=4, size = 1, stroke = 0.5, color ="grey10") +
   geom_point( size = .7, shape = 1, stroke = .3, 
             position = position_jitterdodge(dodge.width = 0.4, jitter.width = 0.2)) +

  
  scale_fill_manual(values = c("#66A61EFF", "#E6AB02FF")) + # housing colors
  scale_alpha_manual(values = c(.65,.8)) +
  scale_color_manual(values = c("#c90076","#125CCC")) +

  # guides(y.sec = 'axis') + # adds 2nd y axis (right side)
  
    labs(title = "\U2191 Upregulated in YNG ") +
  
  facet_wrap(~gene, ncol = 6, scale = "free_y") +
  my_theme +
  theme(
    axis.text.y = element_text(size = 7.5, color ="grey10"),
    axis.text.x = element_text(size = 7, color ="grey10"),
    axis.title.x = element_blank(),
    panel.spacing = unit(0.2, "cm"),  # Removes space between panel 
    legend.margin = margin(0,0,0,0), # space between plot & legend
    axis.ticks = element_line(size=0.1, colour = "grey15"),
    panel.border = element_rect(fill = NA, color = "grey10"), #plot borders
    strip.background = element_rect(color = 'grey10', fill = 'grey90'), 
          panel.grid.major = element_line(color = "grey95"),
        panel.grid.minor = element_line(color = "grey99"),
  )
  

Age_facet_up_plot
```

```{r}
# AGE Down
Age_facet_down_prot_list <- mapped_comp %>%
  filter(Label == "A_YNG_vs_OLD" & adj.pvalue < 0.05 & Direction == "Down") %>%
    slice_min(adj.pvalue, n = 18) %>%
    pull(Protein)


Age_facet_down_plot <- mapped_psum %>%
  
  mutate(Condition = forcats::fct_reorder(Condition, PND)) %>% # reoder by PND (no)
  
  
  filter(Protein %in% Age_facet_down_prot_list) %>%

  
  
  
  ggplot(aes(x = interaction(Age, Housing, 
                             sep = "\n", # space
                             lex.order = TRUE),
             y = Abundance,
             fill = Housing, alpha = Age, color = Sex)) +
  
  
  geom_boxplot(width = 0.6, size = 0.7, linewidth = 0.3,
               position = position_dodge(width = 0.7),
               outlier.shape = NA,
               color = "grey10") +
  stat_summary(fun=mean, geom="point", position = position_dodge(width = 0.7),
               shape=4, size = 1, stroke = 0.5, color ="grey10") +
   geom_point( size = .7, shape = 1, stroke = .3, 
             position = position_jitterdodge(dodge.width = 0.4, jitter.width = 0.2)) +

  
  scale_fill_manual(values = c("#66A61EFF", "#E6AB02FF")) + # housing colors
  scale_alpha_manual(values = c(.65,.8)) +
  scale_color_manual(values = c("#c90076","#125CCC")) +

  # guides(y.sec = 'axis') + # adds 2nd y axis (right side)
  
    labs(title = "\U2193 Downregulated in YNG ") +
  
  facet_wrap(~gene, ncol = 6, scale = "free_y") +
  my_theme +
  theme(
    axis.text.y = element_text(size = 7.5, color ="grey10"),
    axis.text.x = element_text(size = 7, color ="grey10"),
    axis.title.x = element_blank(),
    panel.spacing = unit(0.2, "cm"),  # Removes space between panel 
    legend.margin = margin(0,0,0,0), # space between plot & legend
    axis.ticks = element_line(size=0.1, colour = "grey15"),
    panel.border = element_rect(fill = NA, color = "grey10"), #plot borders
    strip.background = element_rect(color = 'grey10', fill = 'grey90'), 
          panel.grid.major = element_line(color = "grey95"),
        panel.grid.minor = element_line(color = "grey99"),
  )
  

Age_facet_down_plot
```

```{r}
plot_grid(Age_facet_up_plot, Age_facet_down_plot, ncol = 1,
          labels = c("A", "B"))

# ggsave("2026_02_27_Age_facet_36_sig_psum_v3_330dpi.png", dpi = 330) # 9.79 x 12.6
```



```{r}
# Synaptopodin
TrkB <- mapped_psum %>%
  filter(Protein %in% c("Q63604")) # 

TrkB_plot <- boxplot_function(TrkB, 
                       x = Housing, 
                       y = Abundance,
                       fill = Housing,  # by default
                       # color = interaction(BioReplicate, TechRepMixture),
                       shape_fill = Channel,
                       shape_color = TechRepMixture,
                       # alpha = Age,
                       legend_pos = "none",
                       title = 'Q63604 - TrkB/Ntrk2 - ',
                       subtitle = '(TrKB/Ntrk2) - In all 55 Samples')

TrkB_plot

unique(TrkB$BioReplicate)
unique(TrkB$`Gene Names (primary)`) #Synpo
length(unique(TrkB$BioReplicate)) # 56 observations! 55

TrkB %>% 
  group_by(TechRepMixture) %>%
  count(TechRepMixture)
```

```{r}
ribo <- EE_main_only %>%
  filter(stringr::str_detect(description, pattern = "ribosom|elong"))
```

------------------------------------------------------------------------

TO DO LATER!!!

```{r}
UP_Paban_list <- c('Bdnf', 'Gmfb', 'Gdnf', 'Gfra3',
                   'Mt3', 'Ngf', 'Tac1', 'Npffr2', 'Fgf2',
                   'Ptn', 'Cxcr4', 'Myc', 'Stat5a')

Paban_pat <- paste(UP_Paban_list, collapse = "|") # so sep by OR |

me_and_paban <- EE_main_only %>%
  # match any case insenstitive
  filter(str_detect(gene, regex(Paban_pat, ignore_case = TRUE)))

# matched 8, all non sig, closest is Protachykinin-1 but mapping weird Tac1?
# old paper so need to remap genes! 
# NVM DOUBLE CHECK THIS IM VERY CONFUSED!!! 
```

------------------------------------------------------------------------



```{r}
list_SynDiv <- c("Camk2a", "Dlg4", 'Gphn', 'Gad2')
```

```{r}
SynDiv <- mapped_comp %>%
  # filter(Label == "H_EE_vs_SH") %>%
    filter(Label == "HxA_EE_YNG_vs_SH_YNG") %>%
  filter(gene %in% c("Camk2a", "Dlg4", 'Gphn', 'Gad2', 'Mbp', 'Gfap', 'Syp',
                    'Syn2', 'Gria1', 'Gria2', 'Gria3', 'Gria4' ))
```

```{r}

#  'Slc17a7' - vGlut1, 'Slc32a1' - vGat

EE_syn_check <- mapped_psum %>%
  
  mutate(Condition = forcats::fct_reorder(Condition, PND)) %>% # reoder by PND (no)
  filter(gene %in% c(
    # Neuropeptides
    'Penk', "Npy", "Sst", 'Pvalb', 'Cck', "Tac1", "Vip"
    
    # "Dlg4", 'Syp', 'Gria1', 'Gria2', 'Gria3', 'Gria4', 'Fscn1', 'Nrn1',
    # 'Shank1', 'Shank2', 'Shank3', 'Gphn', 'Gad2', 'Mbp', 'Gfap',
    # "Camk2a", 'Syn2', 'Syn1',
    # 'Syt12', 'Ntrk2', 'Homer1', 'Syngap1', 'Faah', 'Abhd6',
    # 'Cnr1', 'Synpo', 'Arhgap39', 'Cck', 'Pvalb', 'Vip',
    # 'Slc17a7', 'Slc32a1', 'Syt2', 'Syngap1' , 'Grin1', 'Gkap1',
    # 
    # 'Mecp2', 'Nptx2' ,'Arc'
  )) %>%
  
  ggplot(aes(x = interaction(Age, Housing, 
                             sep = "\n", # space
                             lex.order = TRUE),
             y = Abundance,
             fill = Housing, alpha = Age, color = Sex)) +
  geom_boxplot(width = 0.5, size = 0.7, median.linewidth = .4, linewidth = 0.2,
               position = position_dodge(width = 0.7),
               outlier.shape = NA,
               color = "grey10") +
  stat_summary(fun=mean, geom="point", position = position_dodge(width = 0.7),
               shape=4, size = 1, stroke = 0.2, color ="grey10") +
  
  ggrastr::geom_point_rast( size = .6, shape = 1, stroke = .3, 
                            position = position_jitterdodge(dodge.width = 0.4, jitter.width = 0.2),
                            alpha = 0.5,
                            raster.dpi = getOption("ggrastr.default.dpi", 330)) +
  
  scale_fill_manual(values = c("#66A61EFF", "#E6AB02FF")) + # housing colors
  scale_alpha_manual(values = c(.7,.75)) +
  scale_color_manual(values = c("#d33682", "#268bd2")) +
  # guides(y.sec = 'axis') + # adds 2nd y axis (right side)
  facet_wrap(~gene, ncol = 7, scale = "free_y") +
  my_psum_facet_theme

EE_syn_check

```

**AGE CHANGES**

```{r}
EE_check_down <- mapped_psum %>%
  
  mutate(Condition = forcats::fct_reorder(Condition, PND)) %>% # reoder by PND (no)
  filter(gene %in% c('Tle2', 'Coro6', 'Echdc2', 'Ebf4', 'Tmsb10', 'Stmn1',
                     'Camk2n2', 'Sbsn', 'Arx', 'Nap1l5', 'Cdh22', 'Ccdc136',
                     'Gng4', 'Myoc' , 'Dpp10')) %>%

  ggplot(aes(x = interaction(Age, Housing,
                             sep = "\n", # space
                             lex.order = TRUE),
             y = Abundance,
             fill = Housing, alpha = Age)) +
  geom_boxplot(width = 0.45, size = 0.7, median.linewidth = .2, linewidth = 0.2,
                 position = position_dodge(width = 0.7),
                 outlier.shape = NA,
                 color = "grey10") +
  stat_summary(fun=mean, geom="point", position = position_dodge(width = 0.7),
               shape=4, size = .8, stroke = 0.2, color ="grey10") +
  geom_point(color = "grey10", size = .5, shape = 1, stroke = .5,
      position = position_jitterdodge(dodge.width = 0.6, jitter.width = 0.3)) +

  scale_fill_manual(values = c("#66A61EFF", "#E6AB02FF")) + # housing colors
  scale_alpha_manual(values = c(.6,.75)) +
  scale_color_manual(values = c("#003a6d", "#31135e")) +
  # guides(y.sec = 'axis') + # adds 2nd y axis (right side)
  facet_wrap(~gene, ncol = 6, scale = "free_y") +
  # my_theme +
  theme(
     axis.text.y = element_text(size = 9, color ="grey10"),
     axis.text.x = element_text(size = 9, color ="grey10"),
     axis.title.x = element_blank(),
     panel.spacing = unit(0.2, "cm"),  # Removes space between panel
     legend.margin = margin(,0,0,0), # space between plot & legend
     axis.ticks = element_line(size=0.1, colour = "grey15", linewidth = .2),
         panel.border = element_rect(fill = NA, linewidth = .2,
                                color = "grey10"), #plot borders
         strip.background = element_rect(color = 'grey10', fill = 'grey90',
                                    linewidth = .2),
  )


EE_check_down
```

```{r}
EE_check_down <- mapped_psum %>%
  
  mutate(Condition = forcats::fct_reorder(Condition, PND)) %>% # reoder by PND (no)
  filter(gene %in% c('Trpc6','Plk5','Lct','Aldh1a1','Bdnf','Iqgap2','Egr3',
                     'Sema5a','Sipa1l3','Pdyn','Lrrc10b','Dagla','Rasd1','Lingo3',
                     'Tanc1','Nptx2','Kif17','Klk8','Ptgs2','Bok','Fosb','Syt12')) %>%

  ggplot(aes(x = interaction(Age, Housing,
                             sep = "\n", # space
                             lex.order = TRUE),
             y = Abundance,
             fill = Housing, alpha = Age)) +
  geom_boxplot(width = 0.45, size = 0.7, median.linewidth = .2, linewidth = 0.2,
                 position = position_dodge(width = 0.7),
                 outlier.shape = NA,
                 color = "grey10") +
  stat_summary(fun=mean, geom="point", position = position_dodge(width = 0.7),
               shape=4, size = .8, stroke = 0.2, color ="grey10") +
  geom_point(color = "grey10", size = .5, shape = 1, stroke = .5,
      position = position_jitterdodge(dodge.width = 0.6, jitter.width = 0.3)) +

  scale_fill_manual(values = c("#66A61EFF", "#E6AB02FF")) + # housing colors
  scale_alpha_manual(values = c(.6,.75)) +
  scale_color_manual(values = c("#003a6d", "#31135e")) +
  # guides(y.sec = 'axis') + # adds 2nd y axis (right side)
  facet_wrap(~gene, ncol = 6, scale = "free_y") +
  my_theme +
  theme(
     axis.text.y = element_text(size = 9, color ="grey10"),
     axis.text.x = element_text(size = 9, color ="grey10"),
     axis.title.x = element_blank(),
     panel.spacing = unit(0.2, "cm"),  # Removes space between panel
     legend.margin = margin(,0,0,0), # space between plot & legend
     axis.ticks = element_line(size=0.1, colour = "grey15", linewidth = .2),
         panel.border = element_rect(fill = NA, linewidth = .2,
                                color = "grey10"), #plot borders
         strip.background = element_rect(color = 'grey10', fill = 'grey90',
                                    linewidth = .2),
  )


EE_check_down
```

```{r}
plot_grid(EE_check_up, EE_check_down, ncol = 1)
```

```{r}
# PErez
'Tle2',
'Coro6',
'Echdc2',
'Ebf4',
'Tmsb10',
'Stmn1',
'Camk2n2',
'Sbsn',
'Arx',
'Nap1l5',
'Cdh22',
'Ccdc136',
'Gng4',
'Myoc' ,
'Dpp10',

'Trpc6',
'Plk5',
'Lct',
'Aldh1a1',
'Bdnf',
'Iqgap2',
'Egr3',
'Sema5a',
'Sipa1l3',
'Pdyn',
'Lrrc10b',
'Dagla',
'Rasd1',
'Lingo3',
'Tanc1',
'Nptx2',
'Kif17',
'Klk8',
'Ptgs2',
'Bok'.
'Fosb',
'Syt12',
```

```{r}
# import 
Perez_prot <- read_csv("input/lit_review/2025_12_27_Perez_2024_proteomics_map_rat_combined_v1.csv")
```

```{r}
Perez_prot_sig <- Perez_prot %>%
  filter(adj.P.Val < 0.05) %>%
  filter(proteomics_id == "all") # 945 x 57


# 945 comp to > 8000 - observe => 8000 present 
intersect(Perez_prot_sig$Entry, EE_main_only$Protein)


# 36 of Perez's EE_all sig adp <0.05 seen in out top 250 
intersect(Perez_prot_sig$Entry, Housing_top_250$Protein)


# 6 of Perez's EE_all sig adp <0.05 seen in our 16 sig adj p<0.05! 
intersect(Perez_prot_sig$Entry, EE_main_sig_df$Protein)


# 7 of Perez's EE_all sig adp <0.05 seen in our 16 sig adj p<0.05! 
intersect(Perez_prot_sig$Entry, EE_main_sig_df_adjp01$Protein)
```

------------------------------------------------------------------------

------------------------------------------------------------------------

# Abundance over Development

Look into further - of those developmental proteins are EE rats reaching them more quickly?

Not sure – look into math more

Age proteins - Proteins that start lower or are the same = do they catch up at higher ages ?

- <https://www.sthda.com/english/wiki/ggplot2-line-plot-quick-start-guide-r-software-and-data-visualization>
- <https://stackoverflow.com/questions/7549694/add-regression-line-equation-and-r2-on-graph>
- <https://forum.posit.co/t/stat-poly-eq-change-the-position-of-the-equation/193755>

```{r}
# EE UP
EE_up_prot_list <- mapped_comp %>%
  filter(Label == "A_YNG_vs_OLD" & adj.pvalue < 0.05 & Direction == "Down") %>%
      slice_min(adj.pvalue, n = 8) %>%
  pull(Protein)

EE_sig_by_PND_line_df <- mapped_psum %>%
   filter(Protein %in% EE_up_prot_list)



EE_sig_by_PND_line_1 <- ggplot(EE_sig_by_PND_line_df,
                               aes(x = PND, y = Abundance,
                                   color = Housing))


EE_sig_by_PND_line_2 <- EE_sig_by_PND_line_1 +
  geom_point(shape=1, size = 0.1) +  
  #geom_point(shape=1, aes(color = read)) +  #colored points
  stat_poly_line(method = "lm", se = TRUE, 
                 size = .5) +
  stat_poly_eq(use_label(c("eq", "R2")),
               size = 3, label.x = 0.95, parse = TRUE) + #label.x = position

  
  scale_color_manual(values = c("#66A61EFF", "#E6AB02FF")) + # housing colors
  scale_alpha_manual(values = c(.65,.8)) +
  # scale_color_manual(values = c("#c90076","#125CCC")) +
  # guides(y.sec = 'axis') + # adds 2nd y axis (right side)
  
  
  facet_grid(Protein~Mixture, scales = "free_y") +
  my_theme +
  theme(
    axis.text.y = element_text(size = 7.5, color ="grey10"),
    axis.text.x = element_text(size = 7, color ="grey10"),
    axis.title.x = element_blank(),
    panel.spacing = unit(0.2, "cm"),  # Removes space between panel 
    legend.margin = margin(0,0,10,0), # space between plot & legend
    axis.ticks = element_line(size=0.1, colour = "grey15"),
    panel.border = element_rect(fill = NA,color = "grey10"), #plot borders
    strip.background = element_rect(color = 'grey10', fill = 'grey90'),
      panel.grid.major = element_line(color = "grey95"),
        panel.grid.minor = element_line(color = "grey99"),
    legend.position = "none"
  )


EE_sig_by_PND_line_2
```

```{r}
myelin_plot <- mapped_psum %>%
  
  mutate(Condition = forcats::fct_reorder(Condition, PND)) %>% # reoder by PND (no)
  
  
  # filter(Protein %in% EE_facet_up_prot_list) %>%
  
  filter(gene %in% c(
    # "Mbp", "Apoe", "Nfasc", "Plp1",
    # "Srebf2", "Cyp46a1", "Abca1", "Abca2",
    # "Mag", "Mog", "Caspr1", "Kcna1", "Kcna2",
    # "Cntnap1", "Cntnap2", "Ank3", "Scn8a",
    # "Gfap", "Mobp", "Omg",
    # "Npy", "Penk", "Vip", "Cck" , "Sst",
    # "Slc17a7", "Slc17a6", "Gad1", "Gad2",
    # "Shank3", "Gria1" , "Gria2", "Gria3", "Gria4",
    # "Dlg4", "Gphn", "Hcn1", "Hcn2", "Pvalb", "Nrxn2",
    # "Ngfr", "Ntf3", "Bdnf", "Sort1", "Ntrk2",
    "Vgf","Egf", "Fgf1", "Nrn1", "Igf1",
    "Furin", "Tnfrsf16", "Plat", "Rtn4r", "Rtn4", "Plg",
    "Ntrk1", "Ntrk3", "Ngf", "Mmp3", "Mmp7", "Tnfrsf1b"
    # "Camk2a", "Syn1", "Fscn1", "Creb1"
  )) %>%

  
  # filter(gene %in% BDNF_STRING) %>%
  # filter(gene %in% unique_growth_factors) %>%
  # filter(gene %in% unique_neurotrophin) %>%

  
  # filter(str_detect(gene, "Syt|Stxbp|Sort1")) %>%
  
  
  
  ggplot(aes(x = interaction(Age, Housing, 
                             sep = "\n", # space
                             lex.order = TRUE),
             y = Abundance,
             fill = Housing, alpha = Age, color = Sex)) +
 
  geom_boxplot(width = 0.65, size = 0.7, linewidth = 0.3,
               position = position_dodge(width = 0.7),
               outlier.shape = NA,
               color = "grey10"
               ) +
  stat_summary(fun=mean, geom="point", position = position_dodge(width = 0.7),
               shape=4, size = 1.5, stroke = 0.5,
               color ="grey10"
               ) +
   geom_point( size = .7, shape = 1, stroke = .3, 
             position = position_jitterdodge(dodge.width = 0.7, jitter.width = 0.15)) +
  
  scale_fill_manual(values = c("#66A61EFF", "#E6AB02FF")) + # housing colors
  scale_alpha_manual(values = c(.65,.8)) +
  scale_color_manual(values = c("#c90076","#125CCC")) +
  # guides(y.sec = 'axis') + # adds 2nd y axis (right side)
  
  # labs(title = "\U2191 Upregulated in EE ") +
  # 
  facet_wrap(~gene, ncol = 11, scale = "free_y") +
  # facet_wrap2(~gene+Sex, ncol = 4) +
  my_theme +
  theme(
    axis.text.y = element_text(size = 7.5, color ="grey10"),
    axis.text.x = element_text(size = 7, color ="grey10"),
    axis.title.x = element_blank(),
    panel.spacing = unit(0.2, "cm"),  # Removes space between panel 
    legend.margin = margin(0,0,10,0), # space between plot & legend
    axis.ticks = element_line(size=0.1, colour = "grey15"),
    panel.border = element_rect(fill = NA,color = "grey10"), #plot borders
    strip.background = element_rect(color = 'grey10', fill = 'grey90'),
      panel.grid.major = element_line(color = "grey95"),
        panel.grid.minor = element_line(color = "grey99"),
  )


myelin_plot
```
