---
title: "MssTMT_Vis_2026_03_lit_review_merge"
author: "Stephanie Huang"
date: "2026-02-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- [x] 2026_07_02 - check why no p-value for perez_transcriptomics? (protein sheet has everything but not the transcriptomics sheet, although significant & other metrics there - check lit_reivew exp sheet!

- [ ] 2026_07_02 - Clean up cases from manual mice sheet

- [ ] 2026_07_02 - Add extras Paban, Novati, etc? Wassouf?

- [ ] 2026_07_02 - Delete redundant - mapped_comp vs mapped_comp_sym

Log

- 2026_07_03 - Hu et al., 2013 done!

```{r}
#load packages
library(tidyverse)
library(gt)
library(readxl)
library(purrr) # for map()
library(tidyselect) # for contains()
library(msigdbr) # genesets
```

## Load Files

You can also embed plots, for example:

```{r}
#psum
load("input/MSstatsTMT/psum/2025_12_31_1pspg_v3.2.2_t1_EExAge_mapped_psum.rda") 

#comp
load("input/MSstatsTMT/comp/2025_12_31_1pspg_v3.2.2_t1_EExAge_mapped_comp.rda") 

# uniprot rat ref proteome anno
load("input/db/rat_ref_proteome_2025_12_13_51872x295.rda")

#anno
anno <- read_csv("input/anno/EEv&SH_F&M_TMT_anno_v3_df_b_check_sorted_2025_12_31.csv") 
# double check later!

# import 
mice_lit_re <- read_excel(
  "input/lit_review/2026_01_13_EE_omics_lit_review_mRNA_Proteins.xlsx",
           sheet = 'mice_proteins_mRNA')

rat_lit_re <- read_excel(
  "input/lit_review/2026_01_13_EE_omics_lit_review_mRNA_Proteins.xlsx",
           sheet = 'rat_proteins_mRNA')
```

## Wrangle

```{r}
# trim uniprot ann
colnames(rat_ref_proteome_2025_12_13)

rat_ref_trm <- rat_ref_proteome_2025_12_13 %>% # trim ver - 18 cols
  dplyr::select(Entry, Protein.families, Keywords,
  Subcellular.location..CC.,
  Activity.regulation, Catalytic.activity, Cofactor, Developmental.stage:Pathway,
  Gene.Names)

rat_ref_trm_trm <- rat_ref_trm  %>% # mini-er ver - 7 cols
  dplyr::select(Entry, Protein.families, Keywords,
  Subcellular.location..CC., Developmental.stage, Function..CC., Pathway)
```

```{r}
# see lil checks in other vis folder - 
# shorten name & clean up! 
mapped_comp <- mapped_comp_v3.2.2_EExAge_t1 %>%
  
  filter(is.na(issue)) %>% # keeps rows with issue == na (so comps that are ok)
  filter(!str_detect(log2FC, "Inf")) %>% # remove infinity/missing FC values
  
  # make gene label for significant! 
  mutate(sig_g_label = if_else(adj.pvalue <= .1, gene, NA)) %>%
  
  # make one for no of sig up & down for each comaprison group/label
  group_by(Label) %>%
  mutate(sig_no_up_lab = sum(adj.pvalue < 0.05 & log2FC > 0), # up
         sig_no_down_lab = sum(adj.pvalue < 0.05 & log2FC < 0)) %>%  #  down
  
  mutate(Direction = if_else(log2FC > 0, "Up", "Down")) %>%
  mutate(Direction_2 = if_else(Direction == "Up", "\u2191", # up arrow
                               "\u2193")) %>% # else down arrow
  
  mutate(protein_existence = as.character(protein_existence)) # as char so no decimal

mapped_comp$Label <- factor(mapped_comp$Label,
                            levels = c( "H_EE_vs_SH",
                                        "HxA_EE_YNG_vs_SH_YNG",
                                        "HxA_EE_OLD_vs_SH_OLD",
                                        "A_YNG_vs_OLD", 
                                        "HxA_EE_YNG_vs_EE_OLD",
                                        "HxA_SH_YNG_vs_SH_OLD",
                                        "HxA_EE_YNG_vs_SH_OLD",
                                        "HxA_EE_OLD_vs_SH_YNG"))   

#____________________________________
# # check 
# head(mapped_comp) 
# colnames(mapped_comp)
# dim(mapped_comp) # 64384 x 22
# 
# unique_check <- mapped_comp %>% 
#   dplyr::select(Label, sig_no_down_lab, sig_no_up_lab) %>%
#   unique() # yep matches up! 
```

```{r}
#___________________________________________
#Quick Counts
direction_df <- mapped_comp %>%
  group_by(Label) %>%
  count(Direction) %>%
  mutate(percent = round(n/8049*100,2))

direction_df
```

## Significant Proteins Tables

```{r}
mapped_comp_2 <- mapped_comp %>%
  left_join(rat_ref_trm,
             by = join_by(Protein == Entry))
```

------------------------------------------------------------------------

## MSigDBR Genesets

Ehhh moved to MssTMT_Vis_2026_07_geneset_gt_tables_wombocombos script instead

<https://igordot.github.io/msigdbr/articles/msigdbr-intro.html>

```{r}
#msigdbr_collections()
genesets <- msigdbr(species = "rat", collection = "C2")
genesets_KEGG <- msigdbr(species = "rat", collection = "C2", subcollection = "CP:KEGG_LEGACY")
```

```{r}
gene_checker <- genesets %>%
  filter(gene_symbol == "Rras2") 

# RASAL1, RASAL, & Rasal1 not found in any genesets of C2 
```

```{r}
geneset_checker <- genesets %>%
  filter(str_detect(gs_description, regex("Neurotrophi", ignore_case = TRUE)))
```

```{r}
KEGG_neurotrophins <- genesets %>%
  # filter(gs_description == "Neurotrophin signaling pathway") %>%
  filter(str_detect(gs_description,
                    "Neurotrophin")) %>%
  # filter(gs_description == "MAPK signaling pathway") %>%
  # filter(gs_description == "Phosphatidylinositol signaling system") %>% # no PI3K-Akt in msig?
  pull(gene_symbol)


ALL_neurotrophins <- genesets %>%
  # filter(str_detect(gs_description, regex("Neurotrophi", ignore_case = TRUE))) 
    filter(str_detect(gs_name, regex("WP_BDNF_SIGNALING_IN_NEURODEVELOPMENT", ignore_case = TRUE))) 

ALL_neurotrophins_genes <- ALL_neurotrophins %>% pull(gene_symbol)
```

------------------------------------------------------------------------

### WOMBO COMBO tables (02 July 2026)

```{r}
mapped_comp_sym <- mapped_comp %>%
  select(Protein:log2FC, adj.pvalue, description, gene) %>% # ! for pivot_wider
  # filter(Label %in% c("H_EE_vs_SH", "A_YNG_vs_OLD")) %>%
  # filter(!(Label %in% c("HxA_EE_OLD_vs_SH_YNG", "HxA_EE_YNG_vs_SH_OLD"))) %>%
      mutate(sym = case_when(
    adj.pvalue < 0.05 & log2FC > 0 ~ "\u2191 *", # up sig .05
    adj.pvalue < 0.05 & log2FC < 0 ~ "\u2193 *", # down sig .05
    adj.pvalue <= 0.11 & log2FC > 0 ~ "\u2191 .", # up sig .1
    adj.pvalue <= 0.11 & log2FC  < 0 ~ "\u2193 .", # down sig .1
    adj.pvalue <= 0.51 & log2FC > 0 ~ "\u2191", # up (~p0.03-05)
    adj.pvalue <= 0.51 & log2FC  < 0 ~ "\u2193", # down 
    .default = " ")) %>%
  rename(adjp = adj.pvalue) %>%
  pivot_wider(names_from = Label,
              # names_sep = ".",
              values_from = c(sym, log2FC, adjp),
              names_glue = "{Label}.{.value}", # need this order for gt
              names_vary = "slowest", # clusters by columns by order! ez4gt!
              names_sort = TRUE # sorts by order of levels in label factor!
              ) %>%
  left_join(rat_ref_trm_trm, 
            by = join_by("Protein" == "Entry"))
```

Switch in or out depending on comparison!

```{r}
# conditional color mapping for gt (based off of Claude example some w/mods)
# labels of interest - comment in & out 
# 6 - EITHER
# my_labels <- c("H_EE_vs_SH", "HxA_EE_YNG_vs_SH_YNG", "HxA_EE_OLD_vs_SH_OLD",
#               "A_YNG_vs_OLD", "HxA_EE_YNG_vs_EE_OLD","HxA_SH_YNG_vs_SH_OLD")

# 2 - OR 
my_labels <- my_labels <- c("H_EE_vs_SH", "A_YNG_vs_OLD")

#______________________________
#___# WOMBO COMBO FUNCTION SET UP
# using purrr's map function 
cell_locations <- function(my_labels, my_pattern) {
  purrr::map(my_labels, function(prefix) { 
    sym_col <- paste0(prefix, ".sym") # so H_EE_vs_SH.sym etc
    fc_col <- paste0(prefix, ".log2FC")
    adjp_col <- paste0(prefix, ".adjp") 
    
    row_condition <- if (is.function(my_pattern)) { # if given my_pattern is func
      my_pattern(sym_col, fc_col, adjp_col)
    } else { # ELSE - my_pattern is not a function but a provided string (symbol)
      rlang::expr(.data[[!!sym_col]] == !!my_pattern) 
    } # need rlang::expr! & !! when using .data - evals later in gt when given df

    cells_body(
      rows = !!row_condition,
      columns = all_of(c(sym_col, fc_col, adjp_col))) # gt cols 2 app 
  })
}


#__CODING_FOR_CONDITIONAL_COLORS
# UP - from most sig to least
up_sig_05_loc <- cell_locations(my_labels, "\u2191 *")
up_sig_1_loc <- cell_locations(my_labels, "\u2191 .")
up_5_loc <- cell_locations(my_labels, "\u2191")
up_fc_loc <- cell_locations(my_labels, function(sym_col, fc_col, adjp_col) {
  rlang::expr(.data[[!!sym_col]] == " " & .data[[!!fc_col]] > 0.04 & .data[[!!adjp_col]] < .99)
})

# DOWN - from most sig to least
down_sig_05_loc <- cell_locations(my_labels, "\u2193 *")
down_sig_1_loc <- cell_locations(my_labels, "\u2193 .")
down_5_loc <- cell_locations(my_labels, "\u2193")
down_fc_loc <- cell_locations(my_labels, function(sym_col, fc_col, adjp_col) {
  rlang::expr(.data[[!!sym_col]] == " " & .data[[!!fc_col]] < -0.04 & .data[[!!adjp_col]] < .99)
})
```

#### wombocombogt

Comment in & out depending on data/comparisons of interest!

```{r}
wombo_combo_tab <- mapped_comp_sym %>%

  select(c(gene, description, H_EE_vs_SH.sym:HxA_SH_YNG_vs_SH_OLD.adjp,
           Protein.families, Function..CC.)) %>%
  
  # filter(gene %in%  ALL_neurotrophins_genes) %>%
  
  # ____COMMENT IN OUR OUT DEPENDING ON TABLE! ____________________
  #_______v1_EE_main___
  mutate(EE_dir = if_else(H_EE_vs_SH.log2FC > 0, "Up \u2191 in EE",
                          "Down \u2193 in EE")) %>%
  arrange(H_EE_vs_SH.adjp) %>%
  
 # #_______v2_YNG_main___
 #   mutate(YNG_dir = if_else(A_YNG_vs_OLD.log2FC > 0, "Up \u2191 in YNG",
 #                          "Down \u2193 in YNG")) %>%
 #  arrange(A_YNG_vs_OLD.adjp) %>%
  # 
  #_____
  # add row number after organisation
  mutate(row_id = as.character(row_number())) %>%

  head(250) %>%
  

 gt(
    # groupname_col = "EE_dir", # v1 -EE main
    groupname_col = "YNG_dir", # v2 - YNG main
    rowname_col = "row_id"
    # rowname_col = "gene"
  ) %>%

  
 #____STYLING____SAME_FOR_ALLL_____-no_touchy!________________
  fmt_number(decimals = 2)  %>%
  
    tab_spanner_delim( delim = ".", # SPLIT 2 SPANNER HEADS
                    split = "first", limit = 1) %>%
  
   cols_label(
    description = "",
    gene = "",
    H_EE_vs_SH.sym = "EE",
    HxA_EE_YNG_vs_SH_YNG.sym = "EE Y",
    HxA_EE_OLD_vs_SH_OLD.sym = "EE O",
    A_YNG_vs_OLD.sym = "YNG",
    HxA_EE_YNG_vs_EE_OLD.sym = "EE Y",
    HxA_SH_YNG_vs_SH_OLD.sym = "SH Y",
    ) %>%

  
  
  tab_style( # STUB STYLE
    style = list(cell_text(size = px(9), color = "#B3B3B3")),
    locations = cells_stub()) %>%

  
  tab_options(# GENERAL TABLE OPTIONS
    table.font.size=11.5, # GENERAL TABLE OPTIONS
    row_group.font.size = 13,


    # ehhh not sure worth changing....
    # table_body.hlines.width = px(0.3), # horizontal rules 
    # row_group.border.top.width = px(0.5), # below colum labs above grouper hrule
    # row_group.border.bottom.width = px(0.5), # below grouper hrule
    # row_group.border.top.color = "#d33682" # use 2 check where
  ) %>%
  opt_vertical_padding(scale = 0.3) %>% # vert padding
  
   # opt_interactive() %>% # do later for scrollable display! 

  cols_align(
    align = "center",
    columns = c(!1:2, 23:24)) %>%
  
  tab_style( # Spanner styling 
    style = cell_text(size = "90%"),
    locations = cells_column_spanners()) %>%
  
  
  tab_style( # Cell column styling! 
    style = cell_text(size = "90%", style = "italic", color = "#808080"),
    locations = cells_column_labels()) %>%
  
    tab_style( # Prevent breaking of value columns!
    style = cell_text(whitespace = "nowrap"),
    locations = cells_body(columns = c(!1:2, 23:24))) %>%
  
  
  tab_style( # gene name column styling
    style = cell_text(size = px(12.5), style = "italic"),
    locations = cells_body(columns = gene)) %>%
  
  
  #______UP_____________
  tab_style( # UP & ADJ P SIG!
    style = list(cell_fill(color = "#fccad8"), cell_text(color = "#FF0099")),
    locations = up_sig_05_loc) %>%
  
  tab_style( 
    style = list(cell_fill(color = "#fddde6"), cell_text(color = "#d33682")),
    locations = up_sig_1_loc ) %>%
  
  tab_style( 
    style = list(cell_fill(color = "#FEEEF3"), cell_text(color = "#d33682")),
    locations = up_5_loc) %>%
  
  tab_style(style = list(cell_text(color = "#d33682")),
            locations = up_fc_loc) %>%
  
  #______DOWN_____________
  tab_style(
    style = list(cell_fill(color = "#C0DDF9"), cell_text(color = "#0080FF")),
    locations = down_sig_05_loc) %>%
  
  tab_style(
    style = list(cell_fill(color = "#DBECFC"),
                 cell_text(color = "#2380c1")),
    locations = down_sig_1_loc) %>%
  
  tab_style( 
    style = list(cell_fill(color = "#EFF6FD"), cell_text(color = "#2380c1")),
    locations = down_5_loc) %>%
  
    tab_style(style = list(cell_text(color = "#2380c1")),
            locations = down_fc_loc) %>%
  
  
  
  #___SPANNER_COLORS____________
  tab_style(style = list(cell_fill(color = "#AFCC9F")),
      locations = cells_column_spanners(spanners = contains("H_EE_vs_SH"))) %>%
  tab_style(style = list(cell_fill(color = "#D4E4CB"),
                         cell_borders(sides = c("left","right"),
                                      color = "#D3D3D3")),
        locations = cells_column_spanners(
          spanners = matches("HxA_EE_YNG_vs_SH_YNG|HxA_EE_OLD_vs_SH_OLD"))) %>%
  
  tab_style(style = cell_fill(color = "#BFBBD5"),
      locations = cells_column_spanners(spanners = contains("A_YNG_vs_OLD"))) %>%
  tab_style(style = list(cell_fill(color = "#DDDCE9"),
                         cell_borders(sides = "left", color = "#D3D3D3")),
        locations = cells_column_spanners(
          spanners = matches("HxA_EE_YNG_vs_EE_OLD|HxA_SH_YNG_vs_SH_OLD"))) 
  

  

wombo_combo_tab #print
```

------------------------------------------------------------------------

# Lit tables

```{r}
EE_mapped_comp <- mapped_comp %>%
  filter(Label == "H_EE_vs_SH")
```

```{r}
EE_mapped_comp_adjp1 <- EE_mapped_comp %>%
  filter(adj.pvalue <= 0.1) %>%
  pull(gene)
```

```{r}
Age_mapped_comp_trm <-  mapped_comp %>%
  filter(Label == "A_YNG_vs_OLD") %>%
  select(Label, gene, log2FC, adj.pvalue) %>%
  rename_with( ~ paste0("A_", .x), everything()) # add prefix onto col names A_
```

# MANUAL CURATION

### RAT

```{r}
# drop spacing rows & convery *** to pvalue numbers instead. 
unique(rat_lit_re$pvalue)

rat_lit_EE <- rat_lit_re %>%
  tidyr::drop_na(Year, Author) %>% # rm empty spacing rowsw
  rename(pvalue_old = pvalue) %>%
  mutate(pvalue =  as.numeric(case_when(
    pvalue_old == "*" ~ "0.049",
    pvalue_old == "**" ~ "0.00999",
    pvalue_old == "***" ~ "0.00099",
    .default = pvalue_old
  )), .after = pvalue_old) %>%
  relocate(pvalue_old, .before = Year) %>% # checked okay 
  mutate(misc = as.character(misc)) %>%
  select(Year:misc)  # checked okay & can drop 
```

### MICE

```{r}
# drop spacing rows & convery *** to pvalue numbers instead. 
unique(mice_lit_re$pvalue)
unique(mice_lit_re$Author) # get names -> filter for complete cases for now
unique(mice_lit_re$N_tested) #

mice_lit_EE <- mice_lit_re %>%
  filter(Author %in% c("Nithianantharajah", "Naka", "Hu")) %>%
  tidyr::drop_na(Year, Author) %>% # rm empty spacing rows
  rename(pvalue_old = pvalue) %>%
  mutate(pvalue =  as.numeric(case_when(
    pvalue_old == "*" ~ "0.049",
    pvalue_old == "**" ~ "0.00999",
    pvalue_old == "***" ~ "0.00099",
    .default = pvalue_old # NA as NA 
  )), .after = pvalue_old) %>%
  relocate(pvalue_old, .before = Year) %>% # checked okay 
  mutate(misc = as.character(misc)) %>%
  mutate(N_tested = case_when( # convert from chr 2 num - for bind_rows later
    N_tested == "6-8" ~ "7", 
    N_tested == "10-11" ~ "10.5",
    N_tested == "7-8" ~ "7.5",
    .default = N_tested)) %>%
  mutate(N_tested = as.numeric(N_tested)) %>%
  select(Year:misc)  # checked okay & can drop 
```

### Rampon et al., 2000

```{r}
rampon <- read_csv(
  "input/lit_review/2026_08_11_Rampon_2024_mRNA_microarray_lit_review_v2.csv") %>%
    select(Year:misc) # 31/62
```

### Zhang et al., 2024

```{r}
zhang <- read_csv(
  "input/lit_review/2025_12_13_Zhang_2024_mPFC_lit_re_mapped_w_rat_ID_v1.csv") %>%
    select(Year:misc) # 31/62
```

# AUTO MAP

## BOTH

### Perez et al., 2024

```{r}
perez_p <- read_excel("input/lit_review/2025_12_27_Perez_2024_proteomics_lit_review_all_v1.xlsx",
                    sheet = "combined") %>%
    select(Year:misc)  #31/74 x 6849


perez_p <- perez_p %>%
  filter(!is.na(Gene_name)) %>%
  filter(!is.na(Uniprot_rat_accession)) #6483 
```

```{r}
perez_t <- read_excel("input/lit_review/2026_07_04_Perez_2024_transcriptomics_lit_review_all_v2.xlsx",
                      sheet = "combined") %>%
  select(Year:misc) #31/74
```

## Proteomics

Lichti et al., 2014

```{r}
lichti <- read_csv(
  "input/lit_review/2025_12_12_Lichti_2014_rats_NAc_data_lit_review_v1.csv") %>%
  select(Year:misc) #31/55
```

### Arroyo et al., 2020

```{r}
arroyo <- read_csv("input/lit_review/2025_12_29_Arroyo_2020_pigs_proteomics_lit_review_sig_v1.csv") %>%
  select(Year:misc) # 31/55
```

### Borgmeyer et al 2021

```{r}
borg <- read_csv("input/lit_review/2025_12_15_Borgmeyer_2021_Hip_SJ_proteomics_lit_re_rats_v1.csv") %>%
  relocate(pvalue, .after = Significant) %>%
  select(Year:misc) # 31/76
```

## Transcriptomics

### Park et al., 2015

```{r}
park <- read_csv("input/lit_review/2025_12_17_Park_2015_mRNA_lit_review_EExAging_summed_v2.csv") %>%
   select(Year:misc) #31/54
```

------------------------------------------------------------------------

------------------------------------------------------------------------

# MERGE

```{r}
merged <- bind_rows(rat_lit_EE, mice_lit_EE,
                    lichti, arroyo, borg, park, perez_p, perez_t, rampon,zhang) %>%
  filter(!is.na(Gene_name))  %>% # removes empty gene name rows (unmapped ones)
  mutate(across(c(Significant, In_EE), tolower)) %>%
  mutate(sym = case_when(
    Significant == "yes" & In_EE == "up" ~ "\u2191 *",
    Significant == "yes" & In_EE == "down" ~ "\u2193 *",
    Significant == "no" & In_EE == "up" ~ "\u2191",
    Significant == "no" & In_EE == "down" ~ "\u2193",
    In_EE == "same" ~ "-",
  )) %>%
  mutate(Author_Year = paste0(Author, "_", Year), .before = Year) %>%
  mutate(Author_Year_Condition = paste0(Author, "_", Year, "_", Condition),
         .before = Year)

length(unique(merged$Author))
unique(merged$Author)
```

```{r}
gene_counts <- merged %>% # wait double check duplicates in Borgmeyer!
  group_by(Gene_name) %>%
  count(Author) %>%
  count(Gene_name)

adjp_overlap <- merged %>%
  filter(Gene_name %in% EE_mapped_comp_adjp1)
```

```{r}
merged_proteins <- merged %>% 
  filter(Exp_type == "Protein")
```

```{r}
# calculation accounts for multi per paper - so no. is based on unique papers
gene_counts_proteins <- merged_proteins %>%
  group_by(Gene_name) %>%
  count(Author) %>% # 1st step
  count(Gene_name) # 2nd step -> gets no. by papers
```

```{r}
# Perez 2024's comparison to Wassouf 2018 (from figure )
down <- c('Tle2', 'Coro6', 'Echdc2', 'Ebf4', 'Tmsb10', 'Stmn1', 'Camk2n2', 'Sbsn',
           'Arx', 'Nap1l5', 'Cdh22', 'Ccdc136', 'Gng4', 'Myoc', 'Dpp10')

up <- c('Trpc6', 'Plk5', 'Lct', 'Aldh1a1', 'Bdnf', 'Iqgap2', 'Egr3',
        'Sema5a', 'Sipa1l3', 'Pdyn', 'Lrrc10b', 'Dagla', 'Rasd1', 'Lingo3',
        'Tanc1', 'Nptx2', 'Kif17', 'Klk8', 'Ptgs2', 'Bok', 'Fosb', 'Syt12')
```

```{r}

wassouf_DEG <- data.frame(Direction_w  = c(rep("Wassouf 2018 - Up \u2191 in EE (Hip - mRNA)", 22), rep("Wassouf 2018 - Down \u2193 in EE (Hip - mRNA)", 15)),
                            gene = c(up, down)) 

was <- EE_mapped_comp %>%
  filter(gene %in% c(up, down))

perez_was_p <- perez_p %>%
  filter(Gene_name %in% c(up, down)) %>%
  filter(Condition == "all") %>%
    mutate(protein_all =  case_when(
    # Significant == "yes" & In_EE == "up" ~ "\u2191 *", # OLD VERSION
    # Significant == "yes" & In_EE == "down" ~ "\u2193 *",
    # Significant == "no" & In_EE == "up" ~ "\u2191",
    # Significant == "no" & In_EE == "down" ~ "\u2193",
    # In_EE == "Same" ~ "-",
    pvalue < 0.05 & Fold_change > 0 ~ "\u2191 *", # up sig .05
    pvalue < 0.05 & Fold_change < 0 ~ "\u2193 *", # down sig .05
    pvalue <= 0.11 & Fold_change > 0 ~ "\u2191 .", # up sig .1
    pvalue <= 0.11 & Fold_change < 0 ~ "\u2193 .", # down sig .1
    pvalue <= 0.51 & Fold_change > 0 ~ "\u2191", # up (~p0.03-05)
    pvalue <= 0.51 & Fold_change < 0 ~ "\u2193", # down 
    In_EE == "Same" ~ "-"
    # .default = " "
   )) %>%
  select(Gene_name, protein_all)

perez_was_t <- perez_t %>%
  filter(Gene_name %in% c(up, down)) %>%
  filter(Condition == "all") %>%
  mutate(mRNA_all =  case_when( # why no p-value for perez_t?
    Significant == "yes" & In_EE == "up" ~ "\u2191 *",
    Significant == "yes" & In_EE == "down" ~ "\u2193 *",
    Significant == "no" & In_EE == "up" ~ "\u2191",
    Significant == "no" & In_EE == "down" ~ "\u2193",
    In_EE == "Same" ~ "-",
    # .default = " "
   )) %>%
  select(Gene_name, mRNA_all)
```

```{r}

perez_was_me <- left_join(wassouf_DEG,  perez_was_p, 
                          by = join_by(gene == Gene_name)) %>%
  left_join(perez_was_t, by = join_by(gene == Gene_name)) %>%
  
  
  left_join(was, by = join_by(gene == gene)) %>% # EE vs SH
    mutate(EE_vs_SH = case_when(
    adj.pvalue < 0.05 & log2FC > 0 ~ "\u2191 *", # up sig .05
    adj.pvalue < 0.05 & log2FC < 0 ~ "\u2193 *", # down sig .05
    adj.pvalue <= 0.5 & log2FC > 0 ~ "\u2191", # up sig .05
    adj.pvalue <= 0.5 & log2FC  < 0 ~ "\u2193", # down sig .05
    .default = " " # ‐
  )) %>%
  
  left_join(Age_mapped_comp_trm, by = join_by(gene == A_gene)) %>% # YNG vs OLD
      mutate(YNG_vs_OLD = case_when(
    A_adj.pvalue < 0.05 & A_log2FC > 0 ~ "\u2191 *", # up sig .05
    A_adj.pvalue < 0.05 & A_log2FC < 0 ~ "\u2193 *", # down sig .05
    A_adj.pvalue <= 0.5 & A_log2FC > 0 ~ "\u2191", # up sig .05
    A_adj.pvalue <= 0.5 & A_log2FC  < 0 ~ "\u2193", # down sig .05
    .default = " " # ‐
  )) %>%
  
  # mutate(Wassouf_direction = if_else(str_detect()))
  arrange(adj.pvalue) 
```

```{r}
perez_was_me_gt <- perez_was_me %>%
  dplyr::select(c(Direction_w, gene,
                  description,
                  mRNA_all, protein_all,
                 EE_vs_SH, log2FC, adj.pvalue, # EE vs SH data
                 YNG_vs_OLD, A_log2FC , A_adj.pvalue
                 )) %>%
  
 gt(
    groupname_col = "Direction_w",
    rowname_col = "gene"
  ) %>%

  fmt_number(decimals = 2)  %>%
  
    cols_label(
    # description = "Protein Description",
    EE_vs_SH = "EE",
    YNG_vs_OLD = "YNG",
    adj.pvalue = "adj.p") %>%
  
  tab_source_note(source_note = "Comparison to Park et al. 2015 - EE x Age SAGEseq Transcriptomics (Hippocampus; male mice). Of the 469 (59 in EE young & 410 in EE old) significant genes identified, 263 were successfully mapped onto proteins in the current dataset, with the top 30 most significant in the H_EE_vs_SH comparison shown here. ") %>%

  tab_style(
    style = list(
      cell_text(style = "italic",  size = px(12.5))
      ),
    locations = cells_stub()) %>%

    tab_options(table.font.size=11.5) %>%
    opt_vertical_padding(scale = 0.3) %>% # vert padding

    # tab_style(
    # style = list(
    #   cell_text(weight = "bold", size = px(14), # pass through px to work! 
    #             align = "center")  
    #   ),
    # locations = list(
    #   cells_body(columns = c(mRNA_all, protein_all, EE_vs_SH))
    #   )) %>%
  
  tab_style(
    style = list(
      cell_text(size = px(11.5), align = "center")),
    locations = list(
      cells_column_labels(columns = c(mRNA_all, protein_all, EE_vs_SH))
      )) %>%
  
  
  tab_spanner(label = "Perez_2024",
              columns = c(mRNA_all, protein_all))  %>%
  tab_spanner(label = "H_EE_vs_SH", columns = c(EE_vs_SH, log2FC, adj.pvalue)) %>%
  tab_spanner(label = "A_YNG_vs_OLD", columns = c(YNG_vs_OLD, A_log2FC,
                                                A_adj.pvalue)) %>%

  
tab_style( # ARROYO
    style = list(cell_fill(color = "#fddde6"),
                 cell_text(color = "#d33682")),
    locations = cells_body(rows = str_detect(mRNA_all,  "\u2191 *"),
                           columns = c(mRNA_all))) %>%
  tab_style( 
    style = list(cell_fill(color = "#DBECFC"),
                 cell_text(color = "#2380c1")),
    locations = cells_body(rows = str_detect(mRNA_all, "\u2193 *"),
                           columns = c(mRNA_all))) %>%
  
  
  tab_style( # ARROYO
    style = list(cell_fill(color = "#fddde6"),
                 cell_text(color = "#d33682")),
    locations = cells_body(rows = str_detect(protein_all,  "\u2191 *"),
                           columns = c(protein_all))) %>%
  tab_style( 
    style = list(cell_fill(color = "#DBECFC"),
                 cell_text(color = "#2380c1")),
    locations = cells_body(rows = str_detect(protein_all, "\u2193 *"),
                           columns = c(protein_all))) %>%
  
  
  
tab_style( # My COLUMN
    style = list(cell_fill(color = "#fddde6"),
                 cell_text(color = "#d33682")),
    locations = cells_body(rows =  adj.pvalue < 0.35 & log2FC > 0 ,
                           columns = c(EE_vs_SH, log2FC, adj.pvalue))) %>%
  tab_style( 
    style = list(cell_fill(color = "#DBECFC"),
                 cell_text(color = "#2380c1")),
    locations = cells_body(rows = adj.pvalue < 0.35 & log2FC < 0,
                           columns = c(EE_vs_SH, log2FC, adj.pvalue))) %>%
  
  tab_style( # My COLUMN
    style = list(cell_fill(color = "#FEEEF3"),
                 cell_text(color = "#d33682")),
 locations = cells_body(rows = adj.pvalue < 0.7 & adj.pvalue > 0.35 & log2FC > 0,
                           columns = c(EE_vs_SH, log2FC, adj.pvalue))) %>%
  tab_style( 
    style = list(cell_fill(color = "#EFF6FD"),
                 cell_text(color = "#2380c1")),
 locations = cells_body(rows = adj.pvalue < 0.7 & adj.pvalue > 0.35 & log2FC < 0,
                           columns = c(EE_vs_SH, log2FC, adj.pvalue))) %>%
  
 tab_style( # My COLUMN - very NON SIG - DIRECTIONAL COLOR
    style = list(cell_text(color = "#d33682")),
  locations = cells_body(rows = adj.pvalue > 0.7 & log2FC > 0 ,
                           columns = c(EE_vs_SH, log2FC, adj.pvalue))) %>%
  tab_style( 
    style = list(cell_text(color = "#2380c1")),
  locations = cells_body(rows = adj.pvalue > 0.7 & log2FC < 0,
                           columns = c(EE_vs_SH, log2FC, adj.pvalue)))

  
  
print(perez_was_me_gt)
```

------------------------------------------------------------------------

### Perez & Wassouf - Better meth? - perez_was_me_2

(used this for PhD write up 2026_08)

- <https://tidyr.tidyverse.org/reference/pivot_wider.html#ref-examples>
- <https://gt.rstudio.com/reference/tab_spanner_delim.html>
- <https://stackoverflow.com/questions/79559857/conditionally-color-cells-based-on-different-column-in-gt>
- <https://thegraphcourses.org/courses/epirep/topics/enhancing-gt-tables/>
- <https://stackoverflow.com/questions/61435048/conditional-formatting-of-multiple-columns-in-gt-table?rq=3>
- <https://gt.rstudio.com/reference/cells_column_spanners.html>

```{r}
was_2 <- mapped_comp %>%
  ungroup() %>% 
  filter(Label %in% c("H_EE_vs_SH", "A_YNG_vs_OLD")) %>%
  select(Protein:log2FC, adj.pvalue, description, gene) %>% # ! for pivot_wider
  filter(gene %in% c(up, down)) %>%
      mutate(Symbols = case_when(
    adj.pvalue < 0.05 & log2FC > 0 ~ "\u2191 *", # up sig .05
    adj.pvalue < 0.05 & log2FC < 0 ~ "\u2193 *", # down sig .05
    adj.pvalue <= 0.11 & log2FC > 0 ~ "\u2191 .", # up sig .1
    adj.pvalue <= 0.11 & log2FC  < 0 ~ "\u2193 .", # down sig .1
    adj.pvalue <= 0.51 & log2FC > 0 ~ "\u2191", # up (~p0.03-05)
    adj.pvalue <= 0.51 & log2FC  < 0 ~ "\u2193", # down 
    .default = " ")) %>%
  pivot_wider(names_from = Label,
              # names_sep = ".",
              values_from = c(Symbols, log2FC, adj.pvalue),
              names_glue = "{Label}.{.value}") # need this order for gt

perez_was_me_2 <- left_join(wassouf_DEG,  perez_was_p,
                          by = join_by(gene == Gene_name)) %>%
  left_join(perez_was_t, by = join_by(gene == Gene_name)) %>%

  left_join(was_2, by = join_by(gene == gene)) %>%

  arrange(H_EE_vs_SH.adj.pvalue)

colnames(perez_was_me_2)
```

```{r}
# temp - my colors
# new is Azure & Barbie pink
up_colors <- c("#d33682", "#fddde6", "#FEEEF3")
new_up_colors <- c("#d33682", "#DA1884", "#fddde6", "#FEEEF3", "#fccad8")
down_colors <- c("#2380c1", "#DBECFC" )
new_down_colors <- c("#0080FF", "#c9e2fa", "#C0DDF9", "#b7d8f8")
```

```{r}
perez_was_me_gt_2 <- perez_was_me_2 %>%
  dplyr::select(c(Direction_w, gene,
                  description,
                  mRNA_all, protein_all,
                # selection of order matters for tab_spanner_delim
                 H_EE_vs_SH.Symbols, H_EE_vs_SH.log2FC, H_EE_vs_SH.adj.pvalue, 
                 A_YNG_vs_OLD.Symbols, A_YNG_vs_OLD.log2FC, A_YNG_vs_OLD.adj.pvalue
                 )) %>%
  
 gt(
    groupname_col = "Direction_w",
    rowname_col = "gene"
  ) %>%

  fmt_number(decimals = 2)  %>%
  
    cols_label(
    description = "Protein Description",
    H_EE_vs_SH.Symbols = "EE",
    A_YNG_vs_OLD.Symbols = "YNG",
    # adj.pvalue = "adj.p"
    ) %>%
  

  tab_style(
    style = list(
      cell_text(style = "italic",  size = px(12))),
    locations = cells_stub()) %>%

    tab_style( # Cell column styling! 
    style = cell_text(size = "90%", style = "italic", color = "#808080"),
    locations = cells_column_labels()) %>%
  
  tab_spanner_delim( delim = ".",
                    split = "first", limit = 1) %>%


  tab_spanner(label = "Perez 2024 - EE (Hip)",
              columns = c(mRNA_all, protein_all))  %>%
  
  
   tab_options(# GENERAL TABLE OPTIONS
    table.border.top.width = px(2), # very top of table
    table.border.bottom.width = px(2), # very bottom of table |not very visible
    table.border.bottom.color = "#BDBDBD",
    column_labels.border.top.width = px(2), # top of column labs |not very vis
    column_labels.border.top.color = "#BDBDBD", 
    column_labels.border.bottom.width = px(2), # top of column labs |not very vis
    column_labels.border.bottom.color = "#BDBDBD", 
    heading.border.bottom.width = px(2), # more visible than above
    heading.border.bottom.color = "#BDBDBD", # more visible than above
    table.font.size = "11.5px",
    stub.font.size = "12px",
    row_group.font.size = "12.5px"
    ) %>%
  
  opt_vertical_padding(scale = 0.3) %>% # vert padding
  

 #_________UP___________________________
  tab_style( # UP & ADJ P SIG!
    style = list(cell_fill(color = "#fccad8"),
                 cell_text(color = "#FF0099")),
    locations = list(
      cells_body(rows = mRNA_all == "\u2191 *",
                 columns = c(mRNA_all)), 
      cells_body(rows = protein_all == "\u2191 *",
                 columns = c(protein_all)),
      cells_body(rows =  H_EE_vs_SH.Symbols == "\u2191 *",
                 columns = c(H_EE_vs_SH.Symbols, H_EE_vs_SH.log2FC,
                             H_EE_vs_SH.adj.pvalue)),
      cells_body(rows =  A_YNG_vs_OLD.Symbols == "\u2191 *",
                 columns = c(A_YNG_vs_OLD.Symbols, A_YNG_vs_OLD.log2FC,
                             A_YNG_vs_OLD.adj.pvalue))
    )
  ) %>%

    tab_style( #_Up_adj.p_0.1
    style = list(cell_fill(color = "#fddde6"),
                 cell_text(color = "#d33682")),
    locations = list(
      cells_body(rows = mRNA_all == "\u2191 .",
                 columns = c(mRNA_all)), 
      cells_body(rows = protein_all == "\u2191 .",
                 columns = c(protein_all)),
      cells_body(rows =  H_EE_vs_SH.Symbols == "\u2191 .",
                 columns = c(H_EE_vs_SH.Symbols, H_EE_vs_SH.log2FC,
                             H_EE_vs_SH.adj.pvalue)),
      cells_body(rows =  A_YNG_vs_OLD.Symbols == "\u2191 .",
                 columns = c(A_YNG_vs_OLD.Symbols, A_YNG_vs_OLD.log2FC,
                             A_YNG_vs_OLD.adj.pvalue))
    )
  ) %>%
  
    tab_style( #_Up___non_sig_but_non_adjp_0.05ish
    style = list(cell_fill(color = "#FEEEF3"),
                 cell_text(color = "#d33682")),
    locations = list(
      cells_body(rows = mRNA_all == "\u2191",
                 columns = c(mRNA_all)), 
      cells_body(rows = protein_all == "\u2191",
                 columns = c(protein_all)),
      cells_body(
        rows =  H_EE_vs_SH.adj.pvalue > 0.11 & H_EE_vs_SH.adj.pvalue < 0.51 &  H_EE_vs_SH.log2FC > 0,
        columns = c(H_EE_vs_SH.Symbols, H_EE_vs_SH.log2FC,
                    H_EE_vs_SH.adj.pvalue)),
      cells_body(rows =  A_YNG_vs_OLD.adj.pvalue > 0.11 & A_YNG_vs_OLD.adj.pvalue < 0.51 & A_YNG_vs_OLD.log2FC > 0,
                 columns = c(A_YNG_vs_OLD.Symbols, A_YNG_vs_OLD.log2FC,
                             A_YNG_vs_OLD.adj.pvalue)))
  ) %>%
  
  
  tab_style( # 
    style = list(cell_text(color = "#d33682")),
    locations = list(
      cells_body(rows =  H_EE_vs_SH.adj.pvalue > 0.51 & H_EE_vs_SH.log2FC > 0.04,
        columns = c(H_EE_vs_SH.Symbols, H_EE_vs_SH.log2FC,
                    H_EE_vs_SH.adj.pvalue)),
  cells_body(rows = A_YNG_vs_OLD.adj.pvalue > 0.51 & A_YNG_vs_OLD.log2FC > 0.04,
                 columns = c(A_YNG_vs_OLD.Symbols, A_YNG_vs_OLD.log2FC,
                             A_YNG_vs_OLD.adj.pvalue))
    )
  ) %>%

  #_________UP___________________________
  tab_style( # Down & ADJ P SIG!
    style = list(cell_fill(color = "#C0DDF9"),
                 cell_text(color = "#0080FF")),
    locations = list(
      cells_body(rows = mRNA_all == "\u2193 *",
                 columns = c(mRNA_all)), 
      cells_body(rows = protein_all == "\u2193 *",
                 columns = c(protein_all)),
      cells_body(rows =  H_EE_vs_SH.Symbols == "\u2193 *",
                 columns = c(H_EE_vs_SH.Symbols, H_EE_vs_SH.log2FC,
                             H_EE_vs_SH.adj.pvalue)),
      cells_body(rows =  A_YNG_vs_OLD.Symbols == "\u2193 *",
                 columns = c(A_YNG_vs_OLD.Symbols, A_YNG_vs_OLD.log2FC,
                             A_YNG_vs_OLD.adj.pvalue))
    )
  ) %>%
  
      tab_style( #__Down_adj.p_0.1
    style = list(cell_fill(color = "#DBECFC"),
                 cell_text(color = "#2380c1")),
    locations = list(
      cells_body(rows = mRNA_all == "\u2193 .",
                 columns = c(mRNA_all)), 
      cells_body(rows = protein_all == "\u2193 .",
                 columns = c(protein_all)),
      cells_body(rows =  H_EE_vs_SH.Symbols == "\u2193 .",
                 columns = c(H_EE_vs_SH.Symbols, H_EE_vs_SH.log2FC,
                             H_EE_vs_SH.adj.pvalue)),
      cells_body(rows =  A_YNG_vs_OLD.Symbols == "\u2193 .",
                 columns = c(A_YNG_vs_OLD.Symbols, A_YNG_vs_OLD.log2FC,
                             A_YNG_vs_OLD.adj.pvalue))
    )
  ) %>%
  
  tab_style( #___Down__but__non_sigish
    style = list(cell_fill(color = "#EFF6FD"),
                 cell_text(color = "#2380c1")),
    locations = list(
       cells_body(rows = mRNA_all == "\u2193",
                 columns = c(mRNA_all)), 
      cells_body(rows = protein_all == "\u2193",
                 columns = c(protein_all)),
      cells_body(
        rows =  H_EE_vs_SH.adj.pvalue > 0.11 & H_EE_vs_SH.adj.pvalue < 0.51 &  H_EE_vs_SH.log2FC < 0,
        columns = c(H_EE_vs_SH.Symbols, H_EE_vs_SH.log2FC,
                    H_EE_vs_SH.adj.pvalue)),
      cells_body(rows =  A_YNG_vs_OLD.adj.pvalue > 0.11 & A_YNG_vs_OLD.adj.pvalue < 0.51 & A_YNG_vs_OLD.log2FC < 0,
                 columns = c(A_YNG_vs_OLD.Symbols, A_YNG_vs_OLD.log2FC,
                             A_YNG_vs_OLD.adj.pvalue)))
  ) %>%
  
  
  tab_style( #___Down__but_very_non_sig
    style = list(cell_text(color = "#2380c1")),
    locations = list(
      cells_body(rows =  H_EE_vs_SH.adj.pvalue > 0.51 & H_EE_vs_SH.log2FC < -0.04,
        columns = c(H_EE_vs_SH.Symbols, H_EE_vs_SH.log2FC,
                    H_EE_vs_SH.adj.pvalue)),
  cells_body(rows = A_YNG_vs_OLD.adj.pvalue > 0.51 & A_YNG_vs_OLD.log2FC < -0.04,
                 columns = c(A_YNG_vs_OLD.Symbols, A_YNG_vs_OLD.log2FC,
                             A_YNG_vs_OLD.adj.pvalue))
    )
  ) %>%
  
   #___SPANNER_COLORS____________
  tab_style(style = list(cell_fill(color = "#AFCC9F")),
      locations = cells_column_spanners(spanners = contains("H_EE_vs_SH"))) %>%
  
  tab_style(style = cell_fill(color = "#BFBBD5"),
      locations = cells_column_spanners(spanners = contains("A_YNG_vs_OLD"))) %>%
  
   tab_style(style = list(cell_fill(color = "#F9F7ED"), 
                          cell_borders(sides = "right", color = "#D3D3D3")),
      locations = cells_column_spanners(spanners = contains("Perez 2024 - EE (Hip)"))) %>%
  
  #____NAs____Grey_out_NA_values________
    tab_style( # Down & ADJ P SIG!
    style = list(cell_text(color = "grey70", size = px(10))),
    locations = list(
      cells_body(rows = is.na(description),
        columns = c(description, H_EE_vs_SH.Symbols,
                    H_EE_vs_SH.log2FC, H_EE_vs_SH.adj.pvalue, 
                    A_YNG_vs_OLD.Symbols, A_YNG_vs_OLD.log2FC,
                    A_YNG_vs_OLD.adj.pvalue)),
      cells_body(rows = is.na(protein_all),
        columns = c(protein_all))
    )
  )


print(perez_was_me_gt_2)
```

#### ~~Best meth? - perez_was_me_3 - just used version above (2026_08_08)~~

------------------------------------------------------------------------

# 1 - Neurotrophins & Intracellular

### Hu 2013

```{r}
# NOTE PROTEIN MORE ACCURATE THAN GENES
Hu_2013 <- merged %>%
  filter(Author == "Hu") %>%
  select(Year:Uniprot_entry_name, sym) %>%
  filter(is.na(Mod)) # rm phos PTMs


# pull protein list
Hu_list <- merged %>%
  filter(Author == "Hu") %>%
  pull(Uniprot_rat_accession) %>% #21
  unique() # 16 

# pull mapped computation list 
me_and_hu <- mapped_comp_sym %>%
  filter(Protein %in% Hu_list) %>%
  # right_join(Hu_2013, by = join_by("gene" == "Gene_name"))
  right_join(Hu_2013, by = join_by("Protein" == "Uniprot_rat_accession")) %>%
  mutate(observed = if_else(is.na(gene), "Not Observed", "Observed"))

colnames(me_and_hu)
```

```{r}
me_and_hu_gt <- me_and_hu %>%

  # select(c(gene, Used_name, description, Region, Exp_type, Method, sym,
  #          H_EE_vs_SH.sym:HxA_SH_YNG_vs_SH_OLD.adjp, observed)) %>% # full 6

    select(c(gene, Used_name, description, Region, Exp_type, Method, sym,
           H_EE_vs_SH.sym:H_EE_vs_SH.adjp, A_YNG_vs_OLD.sym:A_YNG_vs_OLD.adjp,
           observed)) %>% # 2

 gt(
    groupname_col = "observed", # 
    rowname_col = "gene"
  ) %>%

  # gtExtras::gt_theme_nytimes() %>%
  
  fmt_number(decimals = 2)  %>%
  
    tab_spanner_delim( delim = ".", # SPLIT 2 SPANNER HEADS
                    split = "first", limit = 1) %>%
  
   cols_label(
    description = "Protein Description",
    gene = "",
    sym = "EE", 
    H_EE_vs_SH.sym = "EE",
    # HxA_EE_YNG_vs_SH_YNG.sym = "EE Y",
    # HxA_EE_OLD_vs_SH_OLD.sym = "EE O",
    A_YNG_vs_OLD.sym = "YNG",
    # HxA_EE_YNG_vs_EE_OLD.sym = "EE Y",
    # HxA_SH_YNG_vs_SH_OLD.sym = "SH Y",
    ) %>%
  
    tab_header(
    title = md("Comparison to **Hu et al., 2013** across Housing & Age"),
    subtitle = md("`2026_07_06` - `Does not include Phosphoproteins from study.`")) %>%
  

  tab_style( # STUB STYLE
    style = list(cell_text(size = px(10), color = "#B3B3B3", style = "italic")),
    locations = cells_stub()) %>%

  
  
    tab_options(# GENERAL TABLE OPTIONS
    table.border.top.width = px(2), # very top of table
    table.border.bottom.width = px(2), # very bottom of table |not very visible
    table.border.bottom.color = "#BDBDBD",
    column_labels.border.top.width = px(2), # top of column labs |not very vis
    column_labels.border.top.color = "#BDBDBD", 
    column_labels.border.bottom.width = px(2), # top of column labs |not very vis
    column_labels.border.bottom.color = "#BDBDBD", 
    heading.border.bottom.width = px(2), # more visible than above
    heading.border.bottom.color = "#BDBDBD", # more visible than above
    table.font.size = "11.5px",
     stub.font.size = "12px",
    row_group.font.weight = "bold",
    # column_labels.font.weight = "bolder"
  ) %>%
  
  opt_vertical_padding(scale = 0.3) %>% # vert padding
  
   # opt_interactive() %>% # do later for scrollable display! 

  cols_align(
    align = "center",
    columns = !c(1:7)) %>%
  
  tab_style( # Spanner styling 
    style = cell_text(size = "90%"),
    locations = cells_column_spanners()) %>%
  
    tab_spanner(label = "Hu 2013",
              columns = c(Region, Exp_type, Method, sym))  %>%
  
  
  tab_style( # Cell column styling! 
    style = cell_text(size = "90%", style = "italic", color = "#808080"),
    locations = cells_column_labels()) %>%
  
    tab_style( # Prevent breaking of value columns!
    style = cell_text(whitespace = "nowrap"),
    locations = cells_body(columns = !1:3)) %>%
  
  
  tab_style( # gene name column styling
    style = cell_text(size = px(12.5), style = "italic"),
    locations = cells_body(columns = gene)) %>%
  
   tab_style( # region, method, & exp font smaller
    style = cell_text(size = px(10)), 
    locations = cells_body(columns = c(Region, Exp_type, Method))
  ) %>%
  
  
  
  #______UP_____________
  tab_style( # UP & ADJ P SIG!
    style = list(cell_fill(color = "#fccad8"), cell_text(color = "#FF0099")),
    locations = up_sig_05_loc) %>%
  
  tab_style( 
    style = list(cell_fill(color = "#fddde6"), cell_text(color = "#d33682")),
    locations = up_sig_1_loc ) %>%
  
  tab_style( 
    style = list(cell_fill(color = "#FEEEF3"), cell_text(color = "#d33682")),
    locations = up_5_loc) %>%
  
  tab_style(style = list(cell_text(color = "#d33682")),
            locations = up_fc_loc) %>%
  
  #______DOWN_____________
  tab_style(
    style = list(cell_fill(color = "#C0DDF9"), cell_text(color = "#0080FF")),
    locations = down_sig_05_loc) %>%
  
  tab_style(
    style = list(cell_fill(color = "#DBECFC"),cell_text(color = "#2380c1")),
    locations = down_sig_1_loc) %>%
  
  tab_style( 
    style = list(cell_fill(color = "#EFF6FD"), cell_text(color = "#2380c1")),
    locations = down_5_loc) %>%
  
    tab_style(style = list(cell_text(color = "#2380c1")),
            locations = down_fc_loc) %>%
  
  
  #_______PAPER_ARROWS_________
  tab_style( # UP & ADJ P SIG!
    style = list(cell_fill(color = "#fddde6"), cell_text(color = "#d33682")),
    locations = list(cells_body(rows = sym == "\u2191 *", columns = c(sym)))) %>%
  
  tab_style( # Up!
        style = list(cell_fill(color = "#FEEEF3"), cell_text(color = "#d33682")),
    locations = list(cells_body(rows = sym == "\u2191", columns = c(sym)))) %>%
  
  tab_style( # Down & ADJ P SIG!
    style = list(cell_fill(color = "#DBECFC"),cell_text(color = "#2380c1")),
    locations = list(cells_body(rows = sym == "\u2193 *", columns = c(sym)))) %>%
  
  tab_style( # Down!
    style = list(cell_fill(color = "#EFF6FD"), cell_text(color = "#2380c1")),
    locations = list(cells_body(rows = sym == "\u2193", columns = c(sym)))) %>%
  
  
  #___SPANNER_COLORS____________
  tab_style(style = list(cell_fill(color = "#AFCC9F")),
      locations = cells_column_spanners(spanners = contains("H_EE_vs_SH"))) %>%
  # tab_style(style = list(cell_fill(color = "#D4E4CB"),
  #                        cell_borders(sides = c("left","right"),
  #                                     color = "#D3D3D3")),
  #       locations = cells_column_spanners(
  #         spanners = matches("HxA_EE_YNG_vs_SH_YNG|HxA_EE_OLD_vs_SH_OLD"))) %>%
  
  tab_style(style = cell_fill(color = "#BFBBD5"),
      locations = cells_column_spanners(spanners = contains("A_YNG_vs_OLD"))) %>%
  # tab_style(style = list(cell_fill(color = "#DDDCE9"),
  #                        cell_borders(sides = "left", color = "#D3D3D3")),
  #       locations = cells_column_spanners(
  #         spanners = matches("HxA_EE_YNG_vs_EE_OLD|HxA_SH_YNG_vs_SH_OLD"))) %>%

  
  
   tab_style(style = list(cell_fill(color = "#F9F7ED"), 
                          cell_borders(sides = "right", color = "#D3D3D3")),
      locations = cells_column_spanners(spanners = contains("Hu 2013"))) %>%
  
 #____NAs____Grey_out_NA_values________
    tab_style( # Down & ADJ P SIG!
    style = list(cell_text(color = "grey70", size = px(10))),
    locations = list(
      cells_body(rows = is.na(description),
        columns = c(description, sym,
           # H_EE_vs_SH.sym:HxA_SH_YNG_vs_SH_OLD.adjp,
            H_EE_vs_SH.sym:H_EE_vs_SH.adjp, A_YNG_vs_OLD.sym:A_YNG_vs_OLD.adjp
           ))
    )
  )
  

me_and_hu_gt  #print
```

### Checking Hu & I against others

```{r}
Hu_gene_list <- merged %>%
  filter(Author == "Hu") %>%
  pull(Gene_name) %>% #21
  unique() # 16 

# along with Hu_list from above (proteins acc)

Hu_and_others <- merged %>%
  filter(Uniprot_rat_accession %in% Hu_list | Gene_name %in% Hu_gene_list | Gene_name %in% c("Mapk3", "Rasal1", "Nras", "Camk2b", "Rras")) %>%
  filter(Author %in% c("Lichti", "Perez", "Park")) %>%
  select(Author_Year_Condition, Exp_type, Gene_name, Fold_change, Significant,
  pvalue, Uniprot_rat_accession, sym) %>%
   pivot_wider(names_from = Author_Year_Condition,
              # names_sep = ".",
              values_from = c(sym, Fold_change, pvalue),
              names_glue = "{Author_Year_Condition}.{.value}", # need this order for gt
              names_vary = "slowest", # clusters by columns by order! ez4gt!
              names_sort = TRUE # sorts by order of levels in label factor!
              ) 
```

```{r}
Hu_and_others_gt <- Hu_and_others %>%

  # select(c(gene, Used_name, description, Region, Exp_type, Method, sym,
  #          H_EE_vs_SH.sym:HxA_SH_YNG_vs_SH_OLD.adjp, observed)) %>% # full 6

    # select(c(gene, Used_name, description, Region, Exp_type, Method, sym,
    #        H_EE_vs_SH.sym:H_EE_vs_SH.adjp, A_YNG_vs_OLD.sym:A_YNG_vs_OLD.adjp,
    #        observed)) %>% # 2

 gt(
    groupname_col = "Exp_type", # 
    rowname_col = "Gene_name"
  ) %>%

  # gtExtras::gt_theme_nytimes() %>%
  
  fmt_number(decimals = 2)  %>%
  
    tab_spanner_delim( delim = ".", # SPLIT 2 SPANNER HEADS
                    split = "first", limit = 1) %>%
  
    tab_options(# GENERAL TABLE OPTIONS
    table.font.size=11, # GENERAL TABLE OPTIONS
    row_group.font.size = 13,

    table.border.top.width = px(1.5),
    table.border.bottom.width = px(2)
  ) 


Hu_and_others_gt 
```

### Protein Kinases & Phosphatases \| Ras Related

```{r}
smol_GTPase_fam <- mapped_comp_sym %>%
  filter(str_detect(Protein.families, regex("Small GTPase superfamily")) |str_detect(description, "RAS protein")) %>%
    # filter(str_detect(Protein.families, regex("Ras family", ignore_case = TRUE)) | str_detect(description, "RAS protein")) %>%
  pull(gene)


Kinase_Phos_fam <- mapped_comp_sym %>%
    filter(str_detect(Protein.families, regex("Kinase|Phosphatase|Phosphodiesterase", ignore_case = TRUE))) %>%
  pull(gene)


park_protein_phosphatases <-  merged %>%
  filter(Author == "Park") %>%
  filter(str_detect(Full_name, regex("Protein Phosphatase|Calcineurin",
                                     ignore_case = TRUE))) %>%
  pull(Gene_name) %>% #87
  unique() # 7
```

```{r}
NT_ras_kin_pho_gt <- mapped_comp_sym %>%
  
  mutate(Type = case_when( # KINASE & PHOSPHATASE
    str_detect(Protein.families, regex("Kinase", ignore_case = TRUE)) ~ "Kinase",
    str_detect(Protein.families, regex("Phosphodiesterase", ignore_case = TRUE)) ~ "Phosphodiesterase",
    str_detect(Protein.families, regex("Phosphatase", ignore_case = TRUE)) ~
      "Phosphatase"
  )) %>%
  
  
    # mutate(Type = case_when( # Small GTPase Superfamily
    # str_detect(Protein.families, regex("Ras family")) ~ "Ras family", # incl Di-Ras
    # str_detect(description, "RAS protein") ~ "Ras family", # Rasal1 & Rasal2
    # str_detect(Protein.families, regex("Arf family")) ~ "Arf family",
    # str_detect(Protein.families, regex("Rab family")) ~ "Rab family",
    # str_detect(Protein.families, regex("Rho family")) ~ "Rho family",
    # .default = "Ran, Rheb, RGK, & SAR1 family" # remaining less frequent ones
    # )) %>%
  # mutate(Protein.families = str_replace(Protein.families, "Small GTPase superfamily, ", "")) %>% 
  
  select(c(gene, description, Type,  H_EE_vs_SH.sym:HxA_SH_YNG_vs_SH_OLD.adjp)) %>%
  # select(c(gene, description, Type,  H_EE_vs_SH.sym:H_EE_vs_SH.adjp,
  #          A_YNG_vs_OLD.sym:A_YNG_vs_OLD.adjp)) %>%

  filter(gene %in%  Kinase_Phos_fam) %>%
  # filter(gene %in% smol_GTPase_fam) %>% 
  filter(!(is.na(gene))) %>%
  
  
  # ____COMMENT IN OUR OUT DEPENDING ON TABLE! ____________________
  # mutate(Type = case_when(
  #   str_detect(description, regex("Kinase", ignore_case = TRUE)) ~ "Kinase",
  #   str_detect(description, regex("Phosphatase", ignore_case = TRUE)) ~ "Phosphatase"
  # )) %>%

  
  # #_______v1_EE_main___
  # mutate(EE_dir = if_else(H_EE_vs_SH.log2FC > 0, "Up \u2191 in EE",
  #                         "Down \u2193 in EE")) %>%
  
  # arrange(H_EE_vs_SH.adjp) %>%

 #_______v2_YNG_main___
   # mutate(YNG_dir = if_else(A_YNG_vs_OLD.log2FC > 0, "Up \u2191 in YNG",
   #                        "Down \u2193 in YNG")) %>%
  # arrange(A_YNG_vs_OLD.adjp) %>%
  arrange(H_EE_vs_SH.adjp) %>%
  
  #_____
  # add row number after organisation
  mutate(row_id = as.character(row_number())) %>%

  head(250) %>%
  

 gt(
    # groupname_col = "EE_dir", # v1 -EE main
    # groupname_col = "YNG_dir", # v2 - YNG main
   groupname_col = "Type", # Protein Type
    rowname_col = "row_id"
    # rowname_col = "gene"
  ) %>%

  
 #____STYLING____SAME_FOR_ALLL_____-no_touchy!________________
  fmt_number(decimals = 2)  %>%
  
    tab_spanner_delim( delim = ".", # SPLIT 2 SPANNER HEADS
                    split = "first", limit = 1) %>%
  
   cols_label(
    description = "",
    gene = "",
    H_EE_vs_SH.sym = "EE",
    HxA_EE_YNG_vs_SH_YNG.sym = "EE Y",
    HxA_EE_OLD_vs_SH_OLD.sym = "EE O",
    A_YNG_vs_OLD.sym = "YNG",
    HxA_EE_YNG_vs_EE_OLD.sym = "EE Y",
    HxA_SH_YNG_vs_SH_OLD.sym = "SH Y",
    
    ) %>%
  
  # tab_header(
  #   title = md("Comparison of **Small GTPase Superfamily Proteins** across Housing & Age Comparisons"),
  #   subtitle = md("`2026_07_06` - `Done by Filtering Protein.families - Small GTPase superfamily | description - RAS protein`")) %>%
  
    tab_header(
    title = md("Comparison of **Kinases, Phosphatases, & Phosphodiesterases** across Housing & Age Comparisons"),
    subtitle = md("`2026_08_11` - `Sorted by Significance in EE (Top 250 out of 465)`")) %>%
   
  tab_style( # STUB STYLE
    style = list(cell_text(size = px(9), color = "#B3B3B3")),
    locations = cells_stub()) %>%

  

  
   tab_options(# GENERAL TABLE OPTIONS
    table.border.top.width = px(2), # very top of table
    table.border.bottom.width = px(2), # very bottom of table |not very visible
    table.border.bottom.color = "#BDBDBD",
    column_labels.border.top.width = px(2), # top of column labs |not very vis
    column_labels.border.top.color = "#BDBDBD", 
    column_labels.border.bottom.width = px(2), # top of column labs |not very vis
    column_labels.border.bottom.color = "#BDBDBD", 
    heading.border.bottom.width = px(2), # more visible than above
    heading.border.bottom.color = "#BDBDBD", # more visible than above
    table.font.size = "10px",
     stub.font.size = "11px",
      row_group.font.size = "12px",
    row_group.font.weight = "bold",
    # column_labels.font.weight = "bolder"
  ) %>%
  opt_vertical_padding(scale = 0.25) %>% # vert padding


  cols_align(
    align = "center",
    columns = !1:2) %>%
  
  tab_style( # Spanner styling 
    style = cell_text(size = "90%"),
    locations = cells_column_spanners()) %>%
  
  
  tab_style( # Cell column styling! 
    style = cell_text(size = "90%", style = "italic", color = "#808080"),
    locations = cells_column_labels()) %>%
  
    tab_style( # Prevent breaking of value columns!
    style = cell_text(whitespace = "nowrap"),
    locations = cells_body(columns = !1:2)) %>%
  
  
  tab_style( # gene name column styling
    style = cell_text(size = px(11), style = "italic"),
    locations = cells_body(columns = gene)) %>%
  
    tab_style(
    style = cell_text(weight = "bolder"),
    locations = cells_body(columns = gene, 
                           rows = grepl("Ppp", gene))) %>%
  
    tab_style(
    style = cell_fill(color = "#F9F7ED"),
    locations = cells_body(columns = gene, 
                           rows = gene %in% park_protein_phosphatases)) %>%
  
  
  #______UP_____________
  tab_style( # UP & ADJ P SIG!
    style = list(cell_fill(color = "#fccad8"), cell_text(color = "#FF0099")),
    locations = up_sig_05_loc) %>%
  
  tab_style( 
    style = list(cell_fill(color = "#fddde6"), cell_text(color = "#d33682")),
    locations = up_sig_1_loc ) %>%
  
  tab_style( 
    style = list(cell_fill(color = "#FEEEF3"), cell_text(color = "#d33682")),
    locations = up_5_loc) %>%
  
  tab_style(style = list(cell_text(color = "#d33682")),
            locations = up_fc_loc) %>%
  
  #______DOWN_____________
  tab_style(
    style = list(cell_fill(color = "#C0DDF9"), cell_text(color = "#0080FF")),
    locations = down_sig_05_loc) %>%
  
  tab_style(
    style = list(cell_fill(color = "#DBECFC"),
                 cell_text(color = "#2380c1")),
    locations = down_sig_1_loc) %>%
  
  tab_style( 
    style = list(cell_fill(color = "#EFF6FD"), cell_text(color = "#2380c1")),
    locations = down_5_loc) %>%
  
    tab_style(style = list(cell_text(color = "#2380c1")),
            locations = down_fc_loc) %>%
  
  
  
  #___SPANNER_COLORS____________
  tab_style(style = list(cell_fill(color = "#AFCC9F")),
      locations = cells_column_spanners(spanners = contains("H_EE_vs_SH"))) %>%
  tab_style(style = list(cell_fill(color = "#D4E4CB"),
                         cell_borders(sides = c("left","right"),
                                      color = "#D3D3D3")),
        locations = cells_column_spanners(
          spanners = matches("HxA_EE_YNG_vs_SH_YNG|HxA_EE_OLD_vs_SH_OLD"))) %>%
  
  tab_style(style = cell_fill(color = "#BFBBD5"),
      locations = cells_column_spanners(spanners = contains("A_YNG_vs_OLD"))) %>%
  tab_style(style = list(cell_fill(color = "#DDDCE9"),
                         cell_borders(sides = "left", color = "#D3D3D3")),
        locations = cells_column_spanners(
          spanners = matches("HxA_EE_YNG_vs_EE_OLD|HxA_SH_YNG_vs_SH_OLD")))


NT_ras_kin_pho_gt #print
# NT_ras_kin_pho_RAS
```

```{r}

```

------------------------------------------------------------------------

# 2 - Transcription & Translation

### Rampon et al., 2000

```{r}
Rampon_2000 <- merged %>%
  filter(Author == "Rampon") %>%
  # select(Year:Uniprot_entry_name, sym)  %>%
  select(Year:Gene_name, sym, Fold_change, Significant, Class_group)  %>%
  mutate(Phase = sub(" .*", "", Condition)) %>%
  # mutate(Duration = case_when(
  #   Condition == "Early - 3_h" ~ "03_h",
  #   Condition == "Early - 6_h" ~ "06_h",
  #   Condition == "Late - 2_days" ~ "02_d",
  #   Condition == "Late - 14_days" ~ "14_d"
  # )) %>%
    mutate(Time = case_when(
    Condition == "Early - 3_h" ~ "T1",
    Condition == "Early - 6_h" ~ "T2",
    Condition == "Late - 2_days" ~ "T1",
    Condition == "Late - 14_days" ~ "T2"
  )) %>%
    select(!Condition) %>% # otherwise expanded/redundancies
  pivot_wider(names_from = Time,
              # names_sep = ".",
              values_from = c(sym, Fold_change, Significant),
              names_glue = "{Time}.{.value}") %>% # need this order for gt
  
  group_by(Phase, Gene_name) %>%
  slice_head(n =1)  %>% # just to rm Dnpep	Aspartyl aminopeptidase duplicate
  ungroup()

# pull protein list
Rampon_list <- merged %>%
  filter(Author == "Rampon") %>%
  pull(Gene_name) %>% #110 
  unique() # 51 unique 

# pull mapped computation list
me_and_rampon <- mapped_comp_sym %>%
  filter(gene %in% Rampon_list) %>%
  right_join(Rampon_2000, by = join_by("gene" == "Gene_name")) %>%
  mutate(observed = if_else(is.na(Protein), "Not Observed", "Observed")) %>%
  arrange(H_EE_vs_SH.adjp, A_YNG_vs_OLD.adjp) 

colnames(me_and_rampon)
```

```{r}
me_and_rampon_gt <- me_and_rampon %>%

  # DO IT THE NOOBY WAY - either filter in AGE
  # filter(Phase == "Early") %>%
  # filter(Phase == "Late") %>%
  filter(!is.na(description)) %>% # v3 no NAs
  
  mutate(Phase  = if_else(Phase == "Early", "Early - 3 & 6 hours",
                          "Late - 2 & 14 days")) %>%
  
   select(c(Phase, gene, Class_group, description, 
          T1.sym, T2.sym,
          H_EE_vs_SH.sym:H_EE_vs_SH.adjp, A_YNG_vs_OLD.sym:A_YNG_vs_OLD.adjp)) %>% 
  
 gt(
    groupname_col = "Phase", #
      # groupname_col = c("Class_group", "Phase"), #
    # groupname_col = "Class_group", #
    rowname_col = "gene",
    # row_group_as_column = TRUE
  ) %>%

  row_group_order(groups = c("Early - 3 & 6 hours", "Late - 2 & 14 days")) %>%
  
  fmt_number(decimals = 2)  %>%
  
    tab_spanner_delim( delim = ".", # SPLIT 2 SPANNER HEADS
                    split = "first", limit = 1,
                    columns = !c(T1.sym, T2.sym)) %>%
  
   cols_label(
    description = "Protein Description",
    gene = "",
    Class_group = "Class/Group",
    # T1.sym = "3 h", # Early - 3h
    # T2.sym = "6 h", # Early - 6h
    # T1.sym = "2 d", # Late - 2d
    # T2.sym = "14 d", # Late - 14d
    T1.sym = "T1",
    T2.sym = "T2", 
    H_EE_vs_SH.sym = "EE",
    A_YNG_vs_OLD.sym = "YNG"
    ) %>%
  

    #   tab_header(
    # title = md("Comparison to **Rampon et al., 2000** across Housing & Age"),
    # subtitle = md("`2026_08_11` - ` 35 mapped of 51/54 of curated/given | (mRNA cortex mice) | T1 & T2 -Early (3 & 6 h) & Late (2 & 14 d)`")) %>%

  tab_style( # STUB STYLE
    style = list(cell_text(size = px(11),  style = "italic")),
    locations = cells_stub()) %>%

  
    tab_options(# GENERAL TABLE OPTIONS
    table.border.top.width = px(2), # very top of table
    table.border.bottom.width = px(2), # very bottom of table |not very visible
    table.border.bottom.color = "#BDBDBD",
    column_labels.border.top.width = px(2), # top of column labs |not very vis
    column_labels.border.top.color = "#BDBDBD", 
    column_labels.border.bottom.width = px(2), # top of column labs |not very vis
    column_labels.border.bottom.color = "#BDBDBD", 
    heading.border.bottom.width = px(2), # more visible than above
    heading.border.bottom.color = "#BDBDBD", # more visible than above
    table.font.size = "11.5px",
    stub.font.size = "12px",
    # row_group.font.weight = "bolder",
     row_group.font.size = "12.5px",
  ) %>%
  
  opt_vertical_padding(scale = 0.3) %>% # vert padding
  
  cols_align(
    align = "center",
    columns = !c(1:4)) %>%
  
  tab_style( # Spanner styling 
    style = cell_text(size = "90%"),
    locations = cells_column_spanners()) %>%

    tab_spanner(label = "Rampon_2000",
              columns = c(T1.sym, T2.sym))  %>%
  
  
  tab_style( # Cell column styling! 
    style = cell_text(size = "99%", style = "italic"),
    locations = cells_column_labels()) %>%
  
    tab_style( # Prevent breaking of value columns!
    style = cell_text(whitespace = "nowrap"),
    locations = cells_body(columns = !1:2)) %>%
  
  
  tab_style( # gene name column styling
    style = cell_text(size = px(12.5), style = "italic"),
    locations = cells_body(columns = gene)) %>%
  
  tab_style( # CLASS_GROUP - comment in & out as needed 
    style = cell_text(size = px(10), weight = "lighter"),
    locations = cells_body(columns = Class_group)) %>%
  
  
  
  #______UP_____________
  tab_style( # UP & ADJ P SIG!
    style = list(cell_fill(color = "#fccad8"), cell_text(color = "#FF0099")),
    locations = up_sig_05_loc) %>%
  
  tab_style( 
    style = list(cell_fill(color = "#fddde6"), cell_text(color = "#d33682")),
    locations = up_sig_1_loc ) %>%
  
  tab_style( 
    style = list(cell_fill(color = "#FEEEF3"), cell_text(color = "#d33682")),
    locations = up_5_loc) %>%
  
  tab_style(style = list(cell_text(color = "#d33682")),
            locations = up_fc_loc) %>%
  
  #______DOWN_____________
  tab_style(
    style = list(cell_fill(color = "#C0DDF9"), cell_text(color = "#0080FF")),
    locations = down_sig_05_loc) %>%
  
  tab_style(
    style = list(cell_fill(color = "#DBECFC"),cell_text(color = "#2380c1")),
    locations = down_sig_1_loc) %>%
  
  tab_style( 
    style = list(cell_fill(color = "#EFF6FD"), cell_text(color = "#2380c1")),
    locations = down_5_loc) %>%
  
    tab_style(style = list(cell_text(color = "#2380c1")),
            locations = down_fc_loc) %>%
  
  
  ##_______PAPER_ARROWS_________
  ##_T1_______
  tab_style( # UP & ADJ P SIG!
    style = list(cell_fill(color = "#fddde6"), cell_text(color = "#d33682")),
    locations = list(cells_body(rows = T1.sym == "\u2191 *", columns = T1.sym))
    ) %>%
  
  tab_style( # Up!
        style = list(cell_fill(color = "#FEEEF3"), cell_text(color = "#d33682")),
    locations = list(cells_body(rows = T1.sym == "\u2191", columns = T1.sym))) %>%
  
  tab_style( # Down & ADJ P SIG!
    style = list(cell_fill(color = "#DBECFC"),cell_text(color = "#2380c1")),
    locations = list(cells_body(rows = T1.sym == "\u2193 *", columns = T1.sym))
    ) %>%
  
  tab_style( # Down!
    style = list(cell_fill(color = "#EFF6FD"), cell_text(color = "#2380c1")),
    locations = list(cells_body(rows = T1.sym == "\u2193", columns = T1.sym))) %>%
  
  ##_T2__________
    tab_style( # UP & ADJ P SIG!
    style = list(cell_fill(color = "#fddde6"), cell_text(color = "#d33682")),
    locations = list(cells_body(rows = T2.sym == "\u2191 *", columns = T2.sym))
    ) %>%
  
  tab_style( # Up!
        style = list(cell_fill(color = "#FEEEF3"), cell_text(color = "#d33682")),
    locations = list(cells_body(rows = T2.sym == "\u2191", columns = T2.sym))) %>%
  
  tab_style( # Down & ADJ P SIG!
    style = list(cell_fill(color = "#DBECFC"),cell_text(color = "#2380c1")),
    locations = list(cells_body(rows = T2.sym == "\u2193 *", columns = T2.sym))
    ) %>%
  
  tab_style( # Down!
    style = list(cell_fill(color = "#EFF6FD"), cell_text(color = "#2380c1")),
    locations = list(cells_body(rows = T2.sym == "\u2193", columns = T2.sym))) %>%
  
  
  
  ##__________________________________________________________
  #___SPANNER_COLORS____________
  tab_style(style = list(cell_fill(color = "#AFCC9F")),
      locations = cells_column_spanners(spanners = contains("H_EE_vs_SH"))) %>%
  
  tab_style(style = cell_fill(color = "#BFBBD5"),
      locations = cells_column_spanners(spanners = contains("A_YNG_vs_OLD"))) %>%

  
  
   tab_style(style = list(cell_fill(color = "#F9F7ED"), 
                          cell_borders(sides = "right", color = "#D3D3D3")),
      locations = cells_column_spanners(spanners = contains("Rampon_2000"))) %>%
  
 #____NAs____Grey_out_NA_values________
    tab_style( # Down & ADJ P SIG!
    style = list(cell_text(color = "grey70", size = px(10))),
    locations = list(
      cells_body(rows = is.na(description),
        columns = c(description,
                   T1.sym,
           # H_EE_vs_SH.sym:HxA_SH_YNG_vs_SH_OLD.adjp,
            H_EE_vs_SH.sym:H_EE_vs_SH.adjp, A_YNG_vs_OLD.sym:A_YNG_vs_OLD.adjp
           ))
    )
  )

me_and_rampon_gt  
```

### Park et al., 2015

```{r}
park_2015 <- merged %>%
  filter(Author == "Park") %>%
  select(Year:Uniprot_entry_name, sym) 

# pull protein list
park_list <- merged %>%
  filter(Author == "Park") %>%
  pull(Gene_name) %>% #870
  unique() # 870

# pull mapped computation list
me_and_park <- mapped_comp_sym %>%
  filter(gene %in% park_list) %>%
  right_join(park_2015, by = join_by("gene" == "Gene_name")) %>%
  mutate(observed = if_else(is.na(gene), "Not Observed", "Observed")) %>%
  arrange( A_YNG_vs_OLD.adjp, H_EE_vs_SH.adjp) 

colnames(me_and_park)

park_pp_calcineurin <-  merged %>%
  filter(Author == "Park") %>%
  filter(str_detect(Full_name, regex("Protein Phosphatase|Calcineurin",
                                     ignore_case = TRUE))) %>%
  pull(Gene_name) %>% #87
  unique() # 7 - Ppp3ca(calcineruin), Ppp1r1a(prot phosphatase inhib 1 subunit)
```

```{r}
me_and_park_gt <- head(me_and_park, 40) %>%

    select(c(gene, description, sym, 
           H_EE_vs_SH.sym:H_EE_vs_SH.adjp, A_YNG_vs_OLD.sym:A_YNG_vs_OLD.adjp,
           observed)) %>% # 2

 gt(
    groupname_col = "observed", # 
    rowname_col = "gene"
  ) %>%

  fmt_number(decimals = 2)  %>%
  
    tab_spanner_delim( delim = ".", # SPLIT 2 SPANNER HEADS
                    split = "first", limit = 1) %>%
  
   cols_label(
    description = "Protein Description",
    gene = "",
    sym = "EE", 
    H_EE_vs_SH.sym = "EE",
    A_YNG_vs_OLD.sym = "YNG",
    ) %>%
  
    tab_header(
    title = md("Comparison to **Park et al., 2015** across Housing & Age"),
    subtitle = md("`2026_08_11` - `Top 40 overlapping | 870 mapped & 575 NA | (mRNA Hip Mice)`")) %>%
  

  tab_style( # STUB STYLE
    style = list(cell_text(size = px(11),  style = "italic")),
    locations = cells_stub()) %>%

  
    tab_options(# GENERAL TABLE OPTIONS
    table.border.top.width = px(2), # very top of table
    table.border.bottom.width = px(2), # very bottom of table |not very visible
    table.border.bottom.color = "#BDBDBD",
    column_labels.border.top.width = px(2), # top of column labs |not very vis
    column_labels.border.top.color = "#BDBDBD", 
    column_labels.border.bottom.width = px(2), # top of column labs |not very vis
    column_labels.border.bottom.color = "#BDBDBD", 
    heading.border.bottom.width = px(2), # more visible than above
    heading.border.bottom.color = "#BDBDBD", # more visible than above
    table.font.size = "11.5px",
     stub.font.size = "12px",
    #row_group.font.weight = "bold",
  ) %>%
  
  opt_vertical_padding(scale = 0.3) %>% # vert padding
  
  cols_align(
    align = "center",
    columns = !c(1:2)) %>%
  
  tab_style( # Spanner styling 
    style = cell_text(size = "90%"),
    locations = cells_column_spanners()) %>%

    tab_spanner(label = "Park_2015",
              columns = c(sym))  %>%
  
  
  tab_style( # Cell column styling! 
    style = cell_text(size = "99%", style = "italic"),
    locations = cells_column_labels()) %>%
  
    tab_style( # Prevent breaking of value columns!
    style = cell_text(whitespace = "nowrap"),
    locations = cells_body(columns = !1:2)) %>%
  
  
  tab_style( # gene name column styling
    style = cell_text(size = px(12.5), style = "italic"),
    locations = cells_body(columns = gene)) %>%
  
  #  tab_style( # region, method, & exp font smaller
  #   style = cell_text(size = px(10)), 
  #   locations = cells_body(columns = c(Region, Exp_type, Method))
  # ) %>%
  # 
  
  
  #______UP_____________
  tab_style( # UP & ADJ P SIG!
    style = list(cell_fill(color = "#fccad8"), cell_text(color = "#FF0099")),
    locations = up_sig_05_loc) %>%
  
  tab_style( 
    style = list(cell_fill(color = "#fddde6"), cell_text(color = "#d33682")),
    locations = up_sig_1_loc ) %>%
  
  tab_style( 
    style = list(cell_fill(color = "#FEEEF3"), cell_text(color = "#d33682")),
    locations = up_5_loc) %>%
  
  tab_style(style = list(cell_text(color = "#d33682")),
            locations = up_fc_loc) %>%
  
  #______DOWN_____________
  tab_style(
    style = list(cell_fill(color = "#C0DDF9"), cell_text(color = "#0080FF")),
    locations = down_sig_05_loc) %>%
  
  tab_style(
    style = list(cell_fill(color = "#DBECFC"),cell_text(color = "#2380c1")),
    locations = down_sig_1_loc) %>%
  
  tab_style( 
    style = list(cell_fill(color = "#EFF6FD"), cell_text(color = "#2380c1")),
    locations = down_5_loc) %>%
  
    tab_style(style = list(cell_text(color = "#2380c1")),
            locations = down_fc_loc) %>%
  
  
  #_______PAPER_ARROWS_________
  tab_style( # UP & ADJ P SIG!
    style = list(cell_fill(color = "#fddde6"), cell_text(color = "#d33682")),
    locations = list(cells_body(rows = sym == "\u2191 *", columns = c(sym)))) %>%
  
  tab_style( # Up!
        style = list(cell_fill(color = "#FEEEF3"), cell_text(color = "#d33682")),
    locations = list(cells_body(rows = sym == "\u2191", columns = c(sym)))) %>%
  
  tab_style( # Down & ADJ P SIG!
    style = list(cell_fill(color = "#DBECFC"),cell_text(color = "#2380c1")),
    locations = list(cells_body(rows = sym == "\u2193 *", columns = c(sym)))) %>%
  
  tab_style( # Down!
    style = list(cell_fill(color = "#EFF6FD"), cell_text(color = "#2380c1")),
    locations = list(cells_body(rows = sym == "\u2193", columns = c(sym)))) %>%
  
  
  #___SPANNER_COLORS____________
  tab_style(style = list(cell_fill(color = "#AFCC9F")),
      locations = cells_column_spanners(spanners = contains("H_EE_vs_SH"))) %>%
  # tab_style(style = list(cell_fill(color = "#D4E4CB"),
  #                        cell_borders(sides = c("left","right"),
  #                                     color = "#D3D3D3")),
  #       locations = cells_column_spanners(
  #         spanners = matches("HxA_EE_YNG_vs_SH_YNG|HxA_EE_OLD_vs_SH_OLD"))) %>%
  
  tab_style(style = cell_fill(color = "#BFBBD5"),
      locations = cells_column_spanners(spanners = contains("A_YNG_vs_OLD"))) %>%
  # tab_style(style = list(cell_fill(color = "#DDDCE9"),
  #                        cell_borders(sides = "left", color = "#D3D3D3")),
  #       locations = cells_column_spanners(
  #         spanners = matches("HxA_EE_YNG_vs_EE_OLD|HxA_SH_YNG_vs_SH_OLD"))) %>%

  
  
   tab_style(style = list(cell_fill(color = "#F9F7ED"), 
                          cell_borders(sides = "right", color = "#D3D3D3")),
      locations = cells_column_spanners(spanners = contains("Park_2015"))) %>%
  
 #____NAs____Grey_out_NA_values________
    tab_style( # Down & ADJ P SIG!
    style = list(cell_text(color = "grey70", size = px(10))),
    locations = list(
      cells_body(rows = is.na(description),
        columns = c(description, sym,
           # H_EE_vs_SH.sym:HxA_SH_YNG_vs_SH_OLD.adjp,
            H_EE_vs_SH.sym:H_EE_vs_SH.adjp, A_YNG_vs_OLD.sym:A_YNG_vs_OLD.adjp
           ))
    )
  )
  

me_and_park_gt  #print
```

### Arroyo et al., 2020

```{r}
Arroyo_2020 <- merged %>%
  filter(Author == "Arroyo") %>%
  select(Year:Uniprot_entry_name, sym) 

# pull protein list
Arroyo_list <- merged %>%
  filter(Author == "Arroyo") %>%
  pull(Gene_name) %>% #56
  unique() # 56

# pull mapped computation list
me_and_arroyo <- mapped_comp_sym %>%
  filter(gene %in% Arroyo_list) %>%
  right_join(Arroyo_2020, by = join_by("gene" == "Gene_name")) %>%
  mutate(observed = if_else(is.na(gene), "Not Observed", "Observed")) %>%
  # arrange(H_EE_vs_SH.adjp, A_YNG_vs_OLD.adjp) # EE 1st
   arrange( A_YNG_vs_OLD.adjp, H_EE_vs_SH.adjp) # Age 1st

colnames(me_and_arroyo)
```

```{r}
me_and_arroyo_gt <- me_and_arroyo %>%

    select(c(gene, description, sym, 
           H_EE_vs_SH.sym:H_EE_vs_SH.adjp, A_YNG_vs_OLD.sym:A_YNG_vs_OLD.adjp,
           observed)) %>% # 2
  
  head(40) %>%

 gt(
    groupname_col = "observed", # 
    rowname_col = "gene"
  ) %>%

  fmt_number(decimals = 2)  %>%
  
    tab_spanner_delim( delim = ".", # SPLIT 2 SPANNER HEADS
                    split = "first", limit = 1) %>%
  
   cols_label(
    description = "Protein Description",
    gene = "",
    sym = "EE", 
    H_EE_vs_SH.sym = "EE",
    A_YNG_vs_OLD.sym = "YNG",
    ) %>%
  

       tab_header(
    title = md("Comparison to **Arroyo et al., 2020** across Housing & Age"),
    subtitle = md("`2026_08_08` - `Top 40 overlap | 56 sig given & mapped | (Prot Hip Pigs)`")) %>%

  tab_style( # STUB STYLE
    style = list(cell_text(size = px(11),  style = "italic")),
    locations = cells_stub()) %>%

  
    tab_options(# GENERAL TABLE OPTIONS
    table.border.top.width = px(2), # very top of table
    table.border.bottom.width = px(2), # very bottom of table |not very visible
    table.border.bottom.color = "#BDBDBD",
    column_labels.border.top.width = px(2), # top of column labs |not very vis
    column_labels.border.top.color = "#BDBDBD", 
    column_labels.border.bottom.width = px(2), # top of column labs |not very vis
    column_labels.border.bottom.color = "#BDBDBD", 
    heading.border.bottom.width = px(2), # more visible than above
    heading.border.bottom.color = "#BDBDBD", # more visible than above
    table.font.size = "11.5px",
     stub.font.size = "12px",
    #row_group.font.weight = "bold",
  ) %>%
  
  opt_vertical_padding(scale = 0.3) %>% # vert padding
  
  cols_align(
    align = "center",
    columns = !c(1:2)) %>%
  
  tab_style( # Spanner styling 
    style = cell_text(size = "90%"),
    locations = cells_column_spanners()) %>%

    tab_spanner(label = "Arroyo_2020",
              columns = c(sym))  %>%
  
  
  tab_style( # Cell column styling! 
    style = cell_text(size = "99%", style = "italic"),
    locations = cells_column_labels()) %>%
  
    tab_style( # Prevent breaking of value columns!
    style = cell_text(whitespace = "nowrap"),
    locations = cells_body(columns = !1:2)) %>%
  
  
  tab_style( # gene name column styling
    style = cell_text(size = px(12.5), style = "italic"),
    locations = cells_body(columns = gene)) %>%
  
  #  tab_style( # region, method, & exp font smaller
  #   style = cell_text(size = px(10)), 
  #   locations = cells_body(columns = c(Region, Exp_type, Method))
  # ) %>%
  # 
  
  
  #______UP_____________
  tab_style( # UP & ADJ P SIG!
    style = list(cell_fill(color = "#fccad8"), cell_text(color = "#FF0099")),
    locations = up_sig_05_loc) %>%
  
  tab_style( 
    style = list(cell_fill(color = "#fddde6"), cell_text(color = "#d33682")),
    locations = up_sig_1_loc ) %>%
  
  tab_style( 
    style = list(cell_fill(color = "#FEEEF3"), cell_text(color = "#d33682")),
    locations = up_5_loc) %>%
  
  tab_style(style = list(cell_text(color = "#d33682")),
            locations = up_fc_loc) %>%
  
  #______DOWN_____________
  tab_style(
    style = list(cell_fill(color = "#C0DDF9"), cell_text(color = "#0080FF")),
    locations = down_sig_05_loc) %>%
  
  tab_style(
    style = list(cell_fill(color = "#DBECFC"),cell_text(color = "#2380c1")),
    locations = down_sig_1_loc) %>%
  
  tab_style( 
    style = list(cell_fill(color = "#EFF6FD"), cell_text(color = "#2380c1")),
    locations = down_5_loc) %>%
  
    tab_style(style = list(cell_text(color = "#2380c1")),
            locations = down_fc_loc) %>%
  
  
  #_______PAPER_ARROWS_________
  tab_style( # UP & ADJ P SIG!
    style = list(cell_fill(color = "#fddde6"), cell_text(color = "#d33682")),
    locations = list(cells_body(rows = sym == "\u2191 *", columns = c(sym)))) %>%
  
  tab_style( # Up!
        style = list(cell_fill(color = "#FEEEF3"), cell_text(color = "#d33682")),
    locations = list(cells_body(rows = sym == "\u2191", columns = c(sym)))) %>%
  
  tab_style( # Down & ADJ P SIG!
    style = list(cell_fill(color = "#DBECFC"),cell_text(color = "#2380c1")),
    locations = list(cells_body(rows = sym == "\u2193 *", columns = c(sym)))) %>%
  
  tab_style( # Down!
    style = list(cell_fill(color = "#EFF6FD"), cell_text(color = "#2380c1")),
    locations = list(cells_body(rows = sym == "\u2193", columns = c(sym)))) %>%
  
  
  #___SPANNER_COLORS____________
  tab_style(style = list(cell_fill(color = "#AFCC9F")),
      locations = cells_column_spanners(spanners = contains("H_EE_vs_SH"))) %>%
  # tab_style(style = list(cell_fill(color = "#D4E4CB"),
  #                        cell_borders(sides = c("left","right"),
  #                                     color = "#D3D3D3")),
  #       locations = cells_column_spanners(
  #         spanners = matches("HxA_EE_YNG_vs_SH_YNG|HxA_EE_OLD_vs_SH_OLD"))) %>%
  
  tab_style(style = cell_fill(color = "#BFBBD5"),
      locations = cells_column_spanners(spanners = contains("A_YNG_vs_OLD"))) %>%
  # tab_style(style = list(cell_fill(color = "#DDDCE9"),
  #                        cell_borders(sides = "left", color = "#D3D3D3")),
  #       locations = cells_column_spanners(
  #         spanners = matches("HxA_EE_YNG_vs_EE_OLD|HxA_SH_YNG_vs_SH_OLD"))) %>%

  
  
   tab_style(style = list(cell_fill(color = "#F9F7ED"), 
                          cell_borders(sides = "right", color = "#D3D3D3")),
      locations = cells_column_spanners(spanners = contains("Arroyo_2020"))) %>%
  
 #____NAs____Grey_out_NA_values________
    tab_style( # Down & ADJ P SIG!
    style = list(cell_text(color = "grey70", size = px(10))),
    locations = list(
      cells_body(rows = is.na(description),
        columns = c(description, sym,
           # H_EE_vs_SH.sym:HxA_SH_YNG_vs_SH_OLD.adjp,
            H_EE_vs_SH.sym:H_EE_vs_SH.adjp, A_YNG_vs_OLD.sym:A_YNG_vs_OLD.adjp
           ))
    )
  )
  

me_and_arroyo_gt  #print
```

### Zhang et al., 2024

```{r}
Zhang_2024 <- merged %>%
  filter(Author == "Zhang") %>%
  select(Year:Uniprot_entry_name, sym) 

# pull protein list
Zhang_list <- merged %>%
  filter(Author == "Zhang") %>%
  pull(Gene_name) %>% #56
  unique() # 56

# pull mapped computation list
me_and_zhang<- mapped_comp_sym %>%
  filter(gene %in% Zhang_list) %>%
  right_join(Zhang_2024, by = join_by("gene" == "Gene_name")) %>%
  mutate(observed = if_else(is.na(Protein), "Not Observed", "Observed")) %>%
  arrange(H_EE_vs_SH.adjp, A_YNG_vs_OLD.adjp)

colnames(me_and_zhang)
```

```{r}
me_and_zhang_gt <- me_and_zhang %>%

    select(c(gene, description, sym, 
           H_EE_vs_SH.sym:H_EE_vs_SH.adjp, A_YNG_vs_OLD.sym:A_YNG_vs_OLD.adjp,
           observed)) %>% # 2
  
  head(33) %>%

 gt(
    groupname_col = "observed", # 
    rowname_col = "gene"
  ) %>%

  fmt_number(decimals = 2)  %>%
  
    tab_spanner_delim( delim = ".", # SPLIT 2 SPANNER HEADS
                    split = "first", limit = 1) %>%
  
   cols_label(
    description = "Protein Description",
    gene = "",
    sym = "EE", 
    H_EE_vs_SH.sym = "EE",
    A_YNG_vs_OLD.sym = "YNG",
    ) %>%
  
    # 
    #    tab_header(
    # title = md("Comparison to **Zhang et al., 2024** across Housing & Age"),
    # subtitle = md("`2026_08_10` - ` 28 mapped of 32 sig given | (Prot mPFC mice)`")) %>%

  tab_style( # STUB STYLE
    style = list(cell_text(size = px(11),  style = "italic")),
    locations = cells_stub()) %>%

  
    tab_options(# GENERAL TABLE OPTIONS
    table.border.top.width = px(2), # very top of table
    table.border.bottom.width = px(2), # very bottom of table |not very visible
    table.border.bottom.color = "#BDBDBD",
    column_labels.border.top.width = px(2), # top of column labs |not very vis
    column_labels.border.top.color = "#BDBDBD", 
    column_labels.border.bottom.width = px(2), # top of column labs |not very vis
    column_labels.border.bottom.color = "#BDBDBD", 
    heading.border.bottom.width = px(2), # more visible than above
    heading.border.bottom.color = "#BDBDBD", # more visible than above
    table.font.size = "11.5px",
     stub.font.size = "12px",
    #row_group.font.weight = "bold",
  ) %>%
  
  opt_vertical_padding(scale = 0.3) %>% # vert padding
  
  cols_align(
    align = "center",
    columns = !c(1:2)) %>%
  
  tab_style( # Spanner styling 
    style = cell_text(size = "90%"),
    locations = cells_column_spanners()) %>%

    tab_spanner(label = "Zhang_2024",
              columns = c(sym))  %>%
  
  
  tab_style( # Cell column styling! 
    style = cell_text(size = "99%", style = "italic"),
    locations = cells_column_labels()) %>%
  
    tab_style( # Prevent breaking of value columns!
    style = cell_text(whitespace = "nowrap"),
    locations = cells_body(columns = !1:2)) %>%
  
  
  tab_style( # gene name column styling
    style = cell_text(size = px(12.5), style = "italic"),
    locations = cells_body(columns = gene)) %>%
  
  #  tab_style( # region, method, & exp font smaller
  #   style = cell_text(size = px(10)), 
  #   locations = cells_body(columns = c(Region, Exp_type, Method))
  # ) %>%
  # 
  
  
  #______UP_____________
  tab_style( # UP & ADJ P SIG!
    style = list(cell_fill(color = "#fccad8"), cell_text(color = "#FF0099")),
    locations = up_sig_05_loc) %>%
  
  tab_style( 
    style = list(cell_fill(color = "#fddde6"), cell_text(color = "#d33682")),
    locations = up_sig_1_loc ) %>%
  
  tab_style( 
    style = list(cell_fill(color = "#FEEEF3"), cell_text(color = "#d33682")),
    locations = up_5_loc) %>%
  
  tab_style(style = list(cell_text(color = "#d33682")),
            locations = up_fc_loc) %>%
  
  #______DOWN_____________
  tab_style(
    style = list(cell_fill(color = "#C0DDF9"), cell_text(color = "#0080FF")),
    locations = down_sig_05_loc) %>%
  
  tab_style(
    style = list(cell_fill(color = "#DBECFC"),cell_text(color = "#2380c1")),
    locations = down_sig_1_loc) %>%
  
  tab_style( 
    style = list(cell_fill(color = "#EFF6FD"), cell_text(color = "#2380c1")),
    locations = down_5_loc) %>%
  
    tab_style(style = list(cell_text(color = "#2380c1")),
            locations = down_fc_loc) %>%
  
  
  #_______PAPER_ARROWS_________
  tab_style( # UP & ADJ P SIG!
    style = list(cell_fill(color = "#fddde6"), cell_text(color = "#d33682")),
    locations = list(cells_body(rows = sym == "\u2191 *", columns = c(sym)))) %>%
  
  tab_style( # Up!
        style = list(cell_fill(color = "#FEEEF3"), cell_text(color = "#d33682")),
    locations = list(cells_body(rows = sym == "\u2191", columns = c(sym)))) %>%
  
  tab_style( # Down & ADJ P SIG!
    style = list(cell_fill(color = "#DBECFC"),cell_text(color = "#2380c1")),
    locations = list(cells_body(rows = sym == "\u2193 *", columns = c(sym)))) %>%
  
  tab_style( # Down!
    style = list(cell_fill(color = "#EFF6FD"), cell_text(color = "#2380c1")),
    locations = list(cells_body(rows = sym == "\u2193", columns = c(sym)))) %>%
  
  
  #___SPANNER_COLORS____________
  tab_style(style = list(cell_fill(color = "#AFCC9F")),
      locations = cells_column_spanners(spanners = contains("H_EE_vs_SH"))) %>%
  
  tab_style(style = cell_fill(color = "#BFBBD5"),
      locations = cells_column_spanners(spanners = contains("A_YNG_vs_OLD"))) %>%

  
  
   tab_style(style = list(cell_fill(color = "#F9F7ED"), 
                          cell_borders(sides = "right", color = "#D3D3D3")),
      locations = cells_column_spanners(spanners = contains("Zhang_2024"))) %>%
  
 #____NAs____Grey_out_NA_values________
    tab_style( # Down & ADJ P SIG!
    style = list(cell_text(color = "grey70", size = px(10))),
    locations = list(
      cells_body(rows = is.na(description),
        columns = c(description, sym,
           # H_EE_vs_SH.sym:HxA_SH_YNG_vs_SH_OLD.adjp,
            H_EE_vs_SH.sym:H_EE_vs_SH.adjp, A_YNG_vs_OLD.sym:A_YNG_vs_OLD.adjp
           ))
    )
  )
  

me_and_zhang_gt  #print
```

------------------------------------------------------------------------

# 3 - Synapse

### Borgmeyer et al., 2021

```{r}
Borgmeyer_2021 <- merged %>%
  filter(Author == "Borgmeyer") %>%
  select(Year:Uniprot_entry_name, sym)  %>%
  distinct(Gene_name, .keep_all = TRUE)
  

# pull protein list
Borgmeyer_list <- merged %>%
  filter(Author == "Borgmeyer") %>%
  pull(Gene_name) %>% #209
  unique() # 209

# pull mapped computation list
me_and_borgmeyer <- mapped_comp_sym %>%
  filter(gene %in% Borgmeyer_list) %>%
  right_join(Borgmeyer_2021, by = join_by("gene" == "Gene_name")) %>%
  mutate(observed = if_else(is.na(gene), "Not Observed", "Observed")) %>%
  
  filter(H_EE_vs_SH.adjp < 0.35 | A_YNG_vs_OLD.adjp < 0.3) %>%
  arrange(H_EE_vs_SH.adjp)
  
  # arrange(H_EE_vs_SH.adjp,  A_YNG_vs_OLD.adjp) # H_EE_vs_SH.adjp,

colnames(me_and_borgmeyer)
```

```{r}
me_and_borgmeyer_gt <- head(me_and_borgmeyer, 178) %>%

    select(c(gene, description, sym,
           H_EE_vs_SH.sym:H_EE_vs_SH.adjp, A_YNG_vs_OLD.sym:A_YNG_vs_OLD.adjp,
           observed)) %>% # 2
  
  # select(c(gene, description, Fold_change, pvalue, sym, 
  #          H_EE_vs_SH.sym:HxA_SH_YNG_vs_SH_OLD.adjp, observed)) %>% # full 6

 gt(
    groupname_col = "observed", # 
    rowname_col = "gene"
  ) %>%

  fmt_number(decimals = 2)  %>%
  
    tab_spanner_delim( delim = ".", # SPLIT 2 SPANNER HEADS
                    split = "first", limit = 1) %>%
  
   cols_label(
    description = "Protein Description",
    gene = "",
    sym = "EE", 
    H_EE_vs_SH.sym = "EE",
    A_YNG_vs_OLD.sym = "YNG",
    # Fold_change = "FC",
    # pvalue = "adj.p"
    ) %>%
  

    # tab_header(
    # title = md("Comparison to **Borgmeyer et al., 2021** across Housing & Age"),
    # subtitle = md("`2026_08_10` - `| Top 44 (either or adj.p <0.3 in EE or AGE) sorted by EE - 170/178 sig given mapped | (Prot Hip (Syn) Mice)`")) %>%

  tab_style( # STUB STYLE
    style = list(cell_text(size = px(11),  style = "italic")),
    locations = cells_stub()) %>%

  
    tab_options(# GENERAL TABLE OPTIONS
    table.border.top.width = px(2), # very top of table
    table.border.bottom.width = px(2), # very bottom of table |not very visible
    table.border.bottom.color = "#BDBDBD",
    column_labels.border.top.width = px(2), # top of column labs |not very vis
    column_labels.border.top.color = "#BDBDBD", 
    column_labels.border.bottom.width = px(2), # top of column labs |not very vis
    column_labels.border.bottom.color = "#BDBDBD", 
    heading.border.bottom.width = px(2), # more visible than above
    heading.border.bottom.color = "#BDBDBD", # more visible than above
    table.font.size = "11.5px",
     stub.font.size = "12px",
  ) %>%
  
  opt_vertical_padding(scale = 0.3) %>% # vert padding
  
  cols_align(
    align = "center",
    columns = !c(1:2)) %>%
  
  tab_style( # Spanner styling 
    style = cell_text(size = "90%"),
    locations = cells_column_spanners()) %>%

    tab_spanner(label = "Borgmeyer_2021",
              columns = c(sym))  %>%
  
  
  tab_style( # Cell column styling! 
    style = cell_text(size = "99%", style = "italic"),
    locations = cells_column_labels()) %>%
  
    tab_style( # Prevent breaking of value columns!
    style = cell_text(whitespace = "nowrap"),
    locations = cells_body(columns = !1:2)) %>%
  
  
  tab_style( # gene name column styling
    style = cell_text(size = px(12.5), style = "italic"),
    locations = cells_body(columns = gene)) %>%
  
  
  #______UP_____________
  tab_style( # UP & ADJ P SIG!
    style = list(cell_fill(color = "#fccad8"), cell_text(color = "#FF0099")),
    locations = up_sig_05_loc) %>%
  
  tab_style( 
    style = list(cell_fill(color = "#fddde6"), cell_text(color = "#d33682")),
    locations = up_sig_1_loc ) %>%
  
  tab_style( 
    style = list(cell_fill(color = "#FEEEF3"), cell_text(color = "#d33682")),
    locations = up_5_loc) %>%
  
  tab_style(style = list(cell_text(color = "#d33682")),
            locations = up_fc_loc) %>%
  
  #______DOWN_____________
  tab_style(
    style = list(cell_fill(color = "#C0DDF9"), cell_text(color = "#0080FF")),
    locations = down_sig_05_loc) %>%
  
  tab_style(
    style = list(cell_fill(color = "#DBECFC"),cell_text(color = "#2380c1")),
    locations = down_sig_1_loc) %>%
  
  tab_style( 
    style = list(cell_fill(color = "#EFF6FD"), cell_text(color = "#2380c1")),
    locations = down_5_loc) %>%
  
    tab_style(style = list(cell_text(color = "#2380c1")),
            locations = down_fc_loc) %>%
  
  
  #_______PAPER_ARROWS_________
  tab_style( # UP & ADJ P SIG!
    style = list(cell_fill(color = "#fddde6"), cell_text(color = "#d33682")),
    locations = list(cells_body(rows = sym == "\u2191 *", columns = c(sym)))) %>%
  
  tab_style( # Up!
        style = list(cell_fill(color = "#FEEEF3"), cell_text(color = "#d33682")),
    locations = list(cells_body(rows = sym == "\u2191", columns = c(sym)))) %>%
  
  tab_style( # Down & ADJ P SIG!
    style = list(cell_fill(color = "#DBECFC"),cell_text(color = "#2380c1")),
    locations = list(cells_body(rows = sym == "\u2193 *", columns = c(sym)))) %>%
  
  tab_style( # Down!
    style = list(cell_fill(color = "#EFF6FD"), cell_text(color = "#2380c1")),
    locations = list(cells_body(rows = sym == "\u2193", columns = c(sym)))) %>%
  
  
  #___SPANNER_COLORS____________
  tab_style(style = list(cell_fill(color = "#AFCC9F")),
      locations = cells_column_spanners(spanners = contains("H_EE_vs_SH"))) %>%
  # tab_style(style = list(cell_fill(color = "#D4E4CB"),
  #                        cell_borders(sides = c("left","right"),
  #                                     color = "#D3D3D3")),
  #       locations = cells_column_spanners(
  #         spanners = matches("HxA_EE_YNG_vs_SH_YNG|HxA_EE_OLD_vs_SH_OLD"))) %>%
  
  tab_style(style = cell_fill(color = "#BFBBD5"),
      locations = cells_column_spanners(spanners = contains("A_YNG_vs_OLD"))) %>%
  # tab_style(style = list(cell_fill(color = "#DDDCE9"),
  #                        cell_borders(sides = "left", color = "#D3D3D3")),
  #       locations = cells_column_spanners(
  #         spanners = matches("HxA_EE_YNG_vs_EE_OLD|HxA_SH_YNG_vs_SH_OLD"))) %>%
  
  
   tab_style(style = list(cell_fill(color = "#F9F7ED"), 
                          cell_borders(sides = "right", color = "#D3D3D3")),
      locations = cells_column_spanners(spanners = contains("Borgmeyer_2021"))) %>%
  
 #____NAs____Grey_out_NA_values________
    tab_style( # Down & ADJ P SIG!
    style = list(cell_text(color = "grey70", size = px(10))),
    locations = list(
      cells_body(rows = is.na(description),
        columns = c(description, sym,
           # H_EE_vs_SH.sym:HxA_SH_YNG_vs_SH_OLD.adjp,
            H_EE_vs_SH.sym:H_EE_vs_SH.adjp, A_YNG_vs_OLD.sym:A_YNG_vs_OLD.adjp
           ))
    )
  )
  
me_and_borgmeyer_gt  #print
```

### Perez & Wassouf - see above

```{r}
PKMzeta <- mapped_comp %>%
  filter(gene %in% c("Prkcz","Prkci", "Gria2", "Gria1", "Dlg4","Pick1", "Arc",
                     "Cebpb", "Creb1", "Igf2", "Rbm3")) 

# pp_calcineurin <-  mapped_comp %>%
# 
#     filter(str_detect(gene, regex("PPP|CABIN|CABN|CHP", ignore_case = TRUE)))
#   # pull(Gene_name) %>% #870
#   # unique() # 870
```

```{r}
cowplot::plot_grid(plotlist = Msig_HM_volcano_list[c(1,3,8)], ncol = 1,
          labels = c("a", "b", "c"),
          label_colour = "grey10", label_size = 11)


```

```{r}
cowplot::plot_grid(reactome_volcano_list$REACTOME_DISEASES_OF_SIGNAL_TRANSDUCTION_BY_GROWTH_FACTOR_RECEPTORS_AND_SECOND_MESSENGERS,
                   reactome_volcano_list$REACTOME_INTRACELLULAR_SIGNALING_BY_SECOND_MESSENGERS,
                   reactome_volcano_list$REACTOME_MAPK_FAMILY_SIGNALING_CASCADES,
                   reactome_volcano_list$REACTOME_PROLONGED_ERK_ACTIVATION_EVENTS,
                   reactome_volcano_list$REACTOME_REGULATION_OF_RAS_BY_GAPS,
                  
                   ncol = 1, labels = c("a","b", "c", "d", "e"), label_colour = "grey10", label_size = 11)
```

```{r}
cowplot::plot_grid( reactome_volcano_list$REACTOME_CREB1_PHOSPHORYLATION_THROUGH_THE_ACTIVATION_OF_ADENYLATE_CYCLASE,
                   reactome_volcano_list$REACTOME_CREB1_PHOSPHORYLATION_THROUGH_THE_ACTIVATION_OF_ADENYLATE_CYCLASE,
                   
                  
                   ncol = 1, labels = c("a"), label_colour = "grey10", label_size = 11)
```

```{r}
cowplot::plot_grid(reactome_volcano_list$REACTOME_NEUROTRANSMITTER_RECEPTORS_AND_POSTSYNAPTIC_SIGNAL_TRANSMISSION,
                   reactome_volcano_list$REACTOME_TRAFFICKING_OF_AMPA_RECEPTORS,
                   reactome_volcano_list$REACTOME_UNBLOCKING_OF_NMDA_RECEPTORS_GLUTAMATE_BINDING_AND_ACTIVATION,
                   reactome_volcano_list$REACTOME_ASSEMBLY_AND_CELL_SURFACE_PRESENTATION_OF_NMDA_RECEPTORS,
                   reactome_volcano_list$REACTOME_CREB1_PHOSPHORYLATION_THROUGH_THE_ACTIVATION_OF_ADENYLATE_CYCLASE,
                   
                  
                   ncol = 1, labels = c("a","b", "c", "d", "e"), label_colour = "grey10", label_size = 11)
```

```{r}

```
